### 2020-03-14 doc/Manuscript/article.tex
first iteration

\documentclass[preprint,12pt]{elsarticle}
\usepackage{amssymb}
\journal{Nuclear Instruments and Methods
in Physics Research Section A: Accelerators, Spectrometers,
Detectors and Associated Equipment}

\begin{document}

\begin{frontmatter}

%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \address{Address\fnref{label3}}
%% \fntext[label3]{}

\title{Latent Variable Machine Learning Algorithms and Classification of Events}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{}
%% \address[label1]{}
%% \address[label2]{}

\author{R.~Solli}
\address{Department of Physics, University of Oslo, POB 1048 Oslo, N-0316 Oslo, Norway}

\author{D.~Bazin}
\address{Department of Physics and Astronomy and Facility for Rare Ion Beams and National Superconducting Cyclotron Facility, Michigan State University, East Lansing, MI 48824, USA}
\author{M.P.~Kuchera}
\address{Physics Department, Davidson College, Davidson, North Carolina, USA}
\author{M.~Hjorth-Jensen}
\address{Department of Physics and Astronomy and Facility for Rare Ion Beams and National Superconducting Cyclotron Facility, Michigan State University, East Lansing, MI 48824, USA}
\address{Department of Physics and Center for Computing in Science Education, University of Oslo, POB 1048 Oslo, N-0316 Oslo, Norway}

\begin{abstract}
%% Text of abstract

\end{abstract}

%%Graphical abstract
\begin{graphicalabstract}
%\includegraphics{grabs}
\end{graphicalabstract}

%%Research highlights
\begin{highlights}
\item Research highlight 1
\item Research highlight 2
\end{highlights}

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword

%% PACS codes here, in the form: \PACS code \sep code

%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)

\end{keyword}

\end{frontmatter}

\section{Introduction}\label{sec:intro}

\section{Methods}\label{sec:methods}

\section{Results and Discussions}\label{sec:results}

\section{Conclusions and Perspectives}\label{sec{conclusion}

\appendix

\begin{thebibliography}{00}

%% \bibitem{label}
%% Text of bibliographic item

\bibitem{}

\end{thebibliography}
\end{document}

### 2020-03-14 doc/Manuscript/article.tex
Updates from Overleaf

@@
 \begin{abstract}
-%% Text of abstract
+In this work we introduce the application of convolutional autoencoder neural networks to the analysis of two-dimensional projections of particle tracks from a resonant proton scattering experiment on ${}^{46}$Ar.
+The data we analyze were recorded by an active target time-projection chamber (AT-TPC).
+Machine learning presents an interesting avenue for researchers operating an AT-TPC, as traditional analysis methods of AT-TPC data are both computationally expensive and fit all particle tracks against the event type of interest.
+The latter presents a considerable challenge when the space of reactions is not known prior to the analysis.
+
+We explore the performance of the autoencoder neural networks and a pre-trained VGG16 \cite{Simonyan2014} convolutional neural network on two tasks: a semi-supervised classification task and the unsupervised clustering of particle tracks.
+On the semi-supervised task, we find that a logistic regression classifier trained on small labelled subsets of the latent space of these models perform very well.
+On simulated data these classifiers achieve an $f1$ score \cite{Chinchor1992} of $f1>0.95$.
+The VGG16 latent classifier achieves this result with as few as $N=100$ samples, as does the convolutional autoencoder when trained on the VGG16 representations of the particle tracks.
+On real data, pre-processed with noise filtering, the same models achieve an $f1>0.7$.
+For unfiltered real data the models achieve an $f1$ value larger than $0.6$.
+Both of the previous results were found with the classifiers trained on $N=100$ samples.
+Furthermore, we found that the autoencoder model reduces the variability in the identification of proton events by $64\%$ from the benchmark logistic regression classifier trained on the VGG16 latent space on real experimental data.
+
+On the clustering task, we found that a $K$-means algorithm applied to the simulated data in the VGG16 latent space forms almost perfect clusters, with an adjusted rand index \cite{Hubert1985} ($ARI$) $ > 0.8$.
+Additionally, the VGG16+K-means approach finds high purity clusters of proton events for real experimental data.
+We also explore the application of neural networks to clustering by implementing a mixture of autoencoders algorithm.
+With this model we improved clustering performance on the real experimental data from an $ARI = 0.17$ to an $ARI = 0.40$.
+However, the neural network clustering suffers from stability issues necessitating further investigations into this approach.
 

### 2020-03-19 doc/Manuscript/article.tex
Updates from Overleaf
notes and files from thesis

@@
 \address{Department of Physics and Center for Computing in Science Education, University of Oslo, POB 1048 Oslo, N-0316 Oslo, Norway}
+\ead{hjensen@frib.msu.edu}
+\ead[url]{http://mhjgit.github.io/info/doc/web/}
 
@@
 \section{Methods}\label{sec:methods}
+\subsection{Why machine learning}
 
+\begin{itemize}
+    \item Traditional MC methods fall short in two principal ways:
+    \begin{itemize}
+        \item The computational cost per event is too large given the size of the data-sets
+        \item The broken tracks, noisy environment creates bad fit statistics for otherwise useful events. End result is $f1 \sim 0.7 $
+    \end{itemize}
+\end{itemize}
+
+\subsection{Machine Learning}
+
+\begin{itemize}
+    \item Reiterate the aim of unsupervised  clustering of events.
+    \item Challenges from a machine learning perspective. 
+    \begin{itemize}
+        \item Supervised vs. Unsupervised learning
+        \item Traditional unupervised - distances in high dimensional spaces etc. 
+    \end{itemize}
+    \item Digression here to explain some ML concepts? Link back to Conv part of Michelles paper? 
+    \item Opportunities presented by: 
+    \begin{itemize}
+        \item Transfer learning (train on VGG use on AT-TPC)
+        \item Compression is understanding (leverage autoencoders) 
+    \end{itemize}
+    \item Building on Kuchera et. al we investigate the VGG16 networks application. 
+    \item Novel contribution by applying clustering autoencoder networks. 
+\end{itemize}
 \section{Results and Discussions}\label{sec:results}
+
+\subsection{Transfer learning}
+
+\subsection{Autoencoder clustering}
 

### 2020-03-19 files_from_roberts_thesis/discussion/clustering/clustering.tex
notes and files from thesis

\section{Clustering of AT-TPC events}

From the semi-supervised results, we know that we can construct class separating latent spaces. The next challenge is then to expand that insight into the unsupervised clustering of events. Clustering is an intrinsically harder task, owing to the unsupervised nature of the problem. This leads to differing requirements on the representation of the data, as a linearly separable latent space is not necessarily suitable for clustering.

The most naive approach to clustering our data is to apply a simple clustering algorithm on the latent representation of an algorithm which we know separates our data. Using a pre-trained model is advantageous because they are generally well studied, but may offer less flexibility.  A more sophisticated approach is to use algorithms designed for clustering. These will invariably carry their own assumptions and pitfalls that we have to be wary of.

In this thesis, we have applied a K-means algorithm to the latent space of a pre-trained VGG16 model as our naive approach. Which means that the K-means assumptions bound this approach. These include that the clusters are distributed as Gaussians in the high-dimensional latent space and that we have a good idea of the number of clusters. We also investigate the more sophisticated mixture of autoencoder (MIXAE) and deep convolutional embedded clustering (DCEC) models, the latter of which fails to cluster AT-TPC data to any degree. However, the MIXAE model achieves promising results, but assumes uniformly distributed classes and requires a good idea of the number of clusters present in the data.

The assumption that we have a good idea of the number of clusters present is problematic for AT-TPC data. In the filtered and full datasets, the amorphous "other" class probably consist of a variety of event types. These were empirically placed in the same class for classification purposes when labelling the data, precisely because the researchers were unable to identify these events.

In this section, we will explore and discuss the clustering of AT-TPC data using a pre-trained network and the MIXAE model. We begin by considering the results from clustering on the pre-trained VGG16 model latent space.

\subsection{Clustering with a pre-trained network}

What is immediately clear from the results in table \ref{tab:clstr_vgg} is that the VGG16 latent space is able to capture difference between event classes, achieving an adjusted rand index (ARI) score of $>0.88$ on the simulated data against the ground-truth labels. Recall that the ARI is a measure of clustering similarities, adjusted for chance. A particularly interesting aspect of the results in table \ref{tab:clstr_vgg} is the ability to cluster despite the very high-dimensional latent space. As previously mentioned, Euclidean vector distances can be uninformative in high dimensional spaces, as shown by \cite{Aggarwal}. We can explain a part of this behaviour by the fact that a majority, on the order of $\sim 80\%$, of the entries in the latent vectors are zero-valued. However, that leaves some $\sim \num{1.5e3}$ non-zero elements per sample. This space is still very much a high dimensional representation. For the Euclidean distances to have discriminatory power, we must then infer that the space contains some very dense regions in-between regions of minimal density. In such a configuration, the ratios between distances will not tend to unity, and we find some success with K-means clustering. The results for the filtered and full data show worse performance than for the simulated data. However, these results are still promising for the application of clustering algorithms to pre-trained network latent spaces.

To further characterize the performance, we then consider the confusion matrices shown in figure \ref{fig:clster_confmat}. From these matrices, we can infer some interesting properties about the types of events placed in different clusters. It seems that the clustering of the filtered data identifies proton events that occur closer to the detector plane, and so have less developed spirals. We confirm this by visually inspecting the events in those clusters, as illustrated in figure \ref{fig:filtered_vgg_clster_repr}. Similarly, we infer that clustering of full events segments the proton class by the noisiness of the event, placing the more noisy proton events with the amorphous "other" category. We confirm this argument by inspecting the events contained in each cluster, illustrated in figure \ref{fig:full_vgg_clster_repr}.

One of the major strengths of this application is the consistency of the clustering algorithms. The K-means algorithm showed very little, if any, variation in performance when re-running the algorithm.

\subsection{Clustering with autoencoder based models}

In addition to the results from the VGG16+K-means algorithm, we implemented and applied the DCEC and MIXAE algorithms to the ${}^{46}$Ar data. The former failed to produce any salient clusterings of AT-TPC data. We are confident in the implementation of the algorithm as we are able to reproduce the original author's results on the MNIST handwritten digits dataset. Some value was gained from the experiments, as the mode of failure was informative for the choice to implement the MIXAE algorithm. The DCEC would consistently collapse all samples to one cluster. This exact mode of failure is addressed by the batch entropy loss term introduced in the MIXAE algorithm. Recall from equation \ref{eq:mixae_batch} that the batch entropy encourages the clustering probabilities to be diverse. This also explains the relative imbalance between the weighing of the losses that we found in section \ref{sec:convae_clustering}.

We then turn to the MIXAE performance shown in figures \ref{fig:mixae_sim}, \ref{fig:mixae_clean} and \ref{fig:mixae_full}. From these figures, we observe that while the MIXAE algorithm is able to achieve better performance than the VGG16+K-means algorithm, it suffers from severe stability problems. The results for the simulated data are noteworthy in this regard. To better understand the out of sample performance for the clustering algorithm, we segregate the training and test data entirely for the simulated data. Coupled with the high variability in the simulated performance, we infer that this method may not generalize well to unseen data. Being unsupervised, this is not as big a caveat for the MIXAE as it would be for a supervised algorithm.

Additionally, the best performing algorithms are systematically not those with the lowest loss-values. Indeed it seems that convergence to a perfect one-hot prediction of cluster belonging is indicative of the training having failed. This is especially clear for the filtered and full datasets whose performance are shown in figures \ref{fig:mixae_clean} and \ref{fig:mixae_full} respectively. A defining feature of the loss-plots is in the sample loss, as defined in equation \ref{eq:mixae_sample}. We observe that for the filtered and full datasets, the sample entropy is quite large for the models with the highest performance, with values $S_{sample} > 10$. Together with the observation that the batch-entropy is at a minimum, we conclude that the cluster assignments are near uniform. For comparison, a perfect one-hot representation has a sample loss of $\sim 10^{-1}$. The initially most plausible candidate for this behaviour is the class imbalance in the data, recall from table \ref{tab:class_distr} that the amorphous "other" class represents $\sim 60 \%$ of the full and filtered data. Indeed, we observe that the high performing models on the simulated data have an order of magnitude smaller sample loss compared to the models applied to full and filtered data.

To further characterize the performance, we inspect the properties of the clustering results achieved by the highest performing models on the filtered and full datasets. The confusion matrices listed in figure \ref{fig:mixae_confmat} initially displays some of the same properties we observed for the VGG16+K-means algorithm. To confirm this behavior, we display some select events from the proton class in figures \ref{fig:filtered_mixae_clster_repr} and \ref{fig:full_mixae_clster_repr}. From these plots, we confirm that while the clustering of the full dataset is in agreement with the VGG16+K-means clusters. That is, we observe noisy proton events being placed in the cluster with the "other" class. This is not the case for the filtered dataset. In the filtered data, the different clusters of proton events do not seem to be visually distinguishable.

\subsection{Comparing clustering methods}

Both the pre-trained and autoencoder based clustering methods hold promise. As is the case for classification, the ease of application of the pre-trained methods is a significant boon, but their static nature creates a hard to breach cap on their performance. We also observe an impressive purity in the proton clusters, as shown in figure \ref{fig:clster_confmat}, for both the filtered and full datasets. Notably, this purity is absent from the MIXAE clustering results. The increase in performance from the VGG16+K-means is then a result of a stronger segmentation of the "other" class of events.

We also observe a link between the clustering and classifying autoencoders in their performance on real AT-TPC data. Recall that for the real data, both the MIXAE and VGG16+K-means approaches showed confusion over very noisy proton events. We observe the same behaviour in the classifying latent spaces of the autoencoder and VGG16 models illustrated by their t-SNE projections shown in figures \ref{fig:vgg_tsne}, \ref{fig:ac_tnse} and \ref{fig:vgg_ac_tnse}.  In those figures, the majority of proton events are correctly segmented from the rest of the data. A portion of the proton events does, however, intermingle with the "other" class in a way that is familiar to us from the clustering results.

It remains to be seen if the algorithms explored in this thesis are capable of separating events in future experiments with more similar tracks. A related avenue for future research is then to combine the autoencoder methods discussed here with a duelling decoder objective on Hough transformed events. This representation is promising as it can explicitly encode the geometry in the event.

For on-going experiments with the AT-TPC, our recommendation for clustering is to employ a pre-trained model combined with K-means. Performance validation is, unfortunately, a necessary step. Some positive identifications of events are thus needed to validate the clustering.

In this thesis, we have demonstrated strong performance with the MIXAE algorithm. However, further inquiry is needed to investigate the stability of the algorithm. Other avenues of potential interest are the coupling of clustering with a duelling decoder objective, as well as other autoencoder based clustering algorithms. Lastly, there is a need for the coupling of unsupervised performance metrics to measures of performance against ground truth labels.

### 2020-03-19 files_from_roberts_thesis/results/clustering/clustering.tex
notes and files from thesis

\chapter{Clustering of AT-TPC events}\label{chap:clustering}

The principal challenge in the AT-TPC experiments that we are trying to solve is the reliance on labelled samples in the analysis as future experiments may not have as visually dissimilar reaction products  as we observe in the ${}^{46}$Ar experiment.  The  ${}^{46}$Ar experiment does, however, provide a useful example where we can then explore unsupervised techniques. In this chapter, we explore the application of clustering techniques to events represented in latent spaces.

We begin by exploring a naive K-means approach on the latent space of a pre-trained network. Subsequently, we investigate other clustering methods and two autoencoder based clustering algorithms, as outlined in section \ref{sec:deep_clustering}.

This chapter builds on the previous results from semi-supervised classification. We observe that we can construct high-quality latent spaces. These high-quality spaces facilitate an investigation of clustering techniques.

The approach for clustering of events is different from the semi-supervised approach in two meaningful ways. First, it is a harder task, as we will demonstrate. The clustering task thus necessitates a more exploratory approach to the problem. Second, as a consequence of the challenge, the focus will be a bit different than for the semi-supervised approach. We will still utilize the same architectures and models starting with a search over the parameter space over which we measure the performance using the adjusted rand score (ARS) and accuracy defined in section \ref{sec:unsupervised_perf} and \ref{sec:supervised_perf}, respectively.

As with chapter  \ref{chap:classification} where we explored the semi-supervised results, we begin this chapter by considering the VGG16 pre-trained model as a benchmark.

Lastly, we note that the focus of this work is mainly on discovering possible avenues for further research. This focus requires a broad scan of possible avenues rather than a rigorous analysis of one specific model.

### 2020-03-19 files_from_roberts_thesis/results/clustering/convae.tex
notes and files from thesis

\section{Deep clustering algorithms}\label{sec:convae_clustering}

In the previous section we demonstrated a potential for clustering of events from an AT-TPC experiment. To build on this result we will in this section explore the application of the mixture of autoencoders (MIXAE) algorithm introduced in section \ref{sec:deep_clustering}. The other algorithm introduced in section \ref{sec:deep_clustering}, deep clustering with convolutional autoencoders (DCEC), consistently collapsed to just one cluster for all datasets.

In the MIXAE algorithm the hyper-parameters to adjust are all the ordinary parameters that we introduced in table \ref{tab:convae_hyperparams}. In addition to those parameters we have the weighting of the loss terms: $\theta$, $\alpha$ and $\gamma$. These weighting parameters are attached to the reconstruction loss, sample entropy and batch-wise entropy respectively. \cite{Zhang} note that these parameters are very important for the model performance and so we focus primarily on these.

We empirically freeze the convolutional autoencoder hyperparameters to compress the original $128 \times 128$ to a $8 \times 8$ pixel grid using four convolutional layers. The parameters chosen for the autoencoders are listed in full in table \ref{tab:mixe_ae_hyperparams}.

\begin{table}[H]
\renewcommand*{\arraystretch}{0.5}
\centering
\caption{Hyperparameters selected for the autoencoder components of the MIXAE algorithm, see table \ref{tab:convae_hyperparams} for a full description of the parameters.}\label{tab:mixe_ae_hyperparams}
\setlength{\extrarowheight}{15pt}
\hspace*{-0.5in}
\begin{tabular}{ll}
\toprule
Hyperparameter & Value \\
\midrule
\multicolumn{2}{l}{Convolutional parameters: } \\
\midrule
Number of layers & $4$ \\
Kernels & $[3,\,3,\,3,\,3]$\\
Strides & $[2,\,2,\,2,\,2]$ \\
Filters & $[64,\, 32, \,16, \,8,]$ \\ 
\midrule
\multicolumn{2}{l}{Network parameters: } \\
\midrule
Activation & LReLu \\
Latent type & None \\
Latent dimension & 20  \\
$\beta$ & N/A \\
Batchnorm & False \\
\midrule
\multicolumn{2}{l}{Optimizer parameters: } \\
\midrule
$\eta$ & $10^{-3}$ \\
$\beta_1$ & $0.9$ \\
$\beta_2$ & $0.99$ \\
\bottomrule
\end{tabular}
\end{table}

Since there are then only three remaining hyperparameters we choose to perform a coarse grid-search as described in section \ref{sec:hyperparam_search_arch}.

\begin{table}
\centering
\caption{Hyperparameter grid for the MIXAE loss weighting terms. The grid is given as exponents for logarithmic scales.}\label{tab:mixae_loss_weights}
\begin{tabular}{lll}
\toprule
Parameter & Grid & Scale \\
\midrule 
$\theta$ & $[-1,\, 5]$ & Logarithmic \\
$\alpha$ & $[-5,\, -1]$ & Logarithmic \\
$\gamma$ & $[3,\, 5]$ & Logarithmic
\end{tabular}
\end{table}

\subsection{Simulated AT-TPC data}

To train the MIXAE clustering algorithm, we use the large simulated dataset with $M=80000$ points, evenly distributed between proton- and carbon-events. The algorithm is trained on a subset of $60000$ of these samples, and we track performance on the remaining $20000$ events.

The grids selected for the search are listed in table \ref{tab:mixae_loss_weights}. The search yielded an optimal configuration with

\begin{align}
\theta = 10^{-1}, \\
\alpha = 10^{-2}, \\
\gamma = 10^5.
\end{align}

Finally, for these parameters we re-ran the algorithm $N=10$ times to investigate the stability of the algorithm. The results are reported in figure \ref{fig:mixae_sim}. We observe that while the algorithm can achieve very strong performance, with an adjusted rand index (ARI) $ > 0.8$, it fluctuates strongly with repeated experiments. The $ARI$ is a clustering performance measure similar to accuracy, but adjusted for chance. We elaborate on the $ARI$ score in section \ref{sec:unsupervised_perf}. As mentioned in section \ref{sec:mixae} the batch-entropy has a second minimum when the cluster confidences are near equal. It is this behavior we observe in figure \ref{fig:mixae_sim}.

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{./plots/sim_mixae.pdf}
\caption[Clustering performance of MIXAE on simulated AT-TPC data]{Performance for the MIXAE model on simulated AT-TPC data. In the top row the loss components are plotted for each run, and in the bottom row the adjusted rand index (ARI) and clustering accuracy are shown. Each run is color-coded with the ARI achieved at the end of the run. }\label{fig:mixae_sim}
\end{figure}

\subsection{Filtered AT-TPC data}

We repeat the optimization steps in the previous section for the filtered AT-TPC data. The exception being that we allow the algorithm to train on the labelled samples. Beginning with a wide grid equal to the grid used for the simulated data we searched over all parameter configurations to find promising values. We then performed a local search around these values to pin-point the hyperparameter configuration. This search yielded the optimal hyper-parameters

\begin{align}
\theta &= 10^{1}, \\
\alpha &= 10^{-1}, \\
\gamma &= 3.162\times 10^3.
\end{align}

\noindent The results of the runs are included in figure \ref{fig:mixae_clean}. We observe ghat the highest performing models reach an $ARI > 0.5$, which is higher than the performance achieved by the K-means algorithm applied to the VGG16 latent space.

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{./plots/clean_mixae.pdf}
\caption[Clustering performance of MIXAE on filtered event AT-TPC data]{Performance for the MIXAE model on filtered AT-TPC data. In the top row the loss components are plotted for each run, and in the bottom row the adjusted rand index (ARI) and clustering accuracy are shown. Each run is color-coded with the ARI achieved at the end of the run.}\label{fig:mixae_clean}
\end{figure}

\subsection{Full AT-TPC data}

As in the two previous sections, we repeat the same procedure of iterative grid searches on the MIXAE loss-weights. Each configuration is re-run a total of $N=10$ times to capture fluctuations in the performance before a final selection is made on the hyperparameters. For the full dataset the MIXAE hyperparameters converge to the same values as for the clean data, i.e.

\begin{align}
\theta &= 10^{1}, \\
\alpha &= 10^{-1}, \\
\gamma &= 3.162\times 10^3.
\end{align}

\noindent As with the previous datasets we include a plot of the loss curves and performance measures for $N=10$ runs with the same loss-weight parameters, shown in figure \ref{fig:mixae_full}.

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{./plots/real_mixae.pdf}
\caption[Clustering performance of MIXAE on full event AT-TPC data]{Performance for the MIXAE model on un-filtered AT-TPC data. In the top row the loss components are plotted for each run, and in the bottom row the adjusted rand index (ARI) and clustering accuracy are shown. Each run is color-coded with the ARI achieved at the end of the run.}\label{fig:mixae_full}
\end{figure}

\subsection{Comparing performance}

It is also interesting to compare and contrast the clustering results from the MIXAE model with those of the VGG16$+$K-means outside the fairly abstract accuracies and rand scores. It is especially interesting to compare the cluster assignments, as they can inform further research efforts in the clustering of AT-TPC events. We illustrate the clustering with confusion matrices that are shown in figure \ref{fig:mixae_confmat}. From these matrices, we observe that the MIXAE applied to the clean data correctly clusters the noise events. Additionally, it identifies two proton clusters. We observe that these proton clusters are both less pure than the VGG16+K-means clusters and that there does not seem to be a visually meaningful difference between these clusters. The latter is inferred from figure \ref{fig:filtered_mixae_clster_repr}.

Applied to the real data the MIXAE correctly separates the proton class, however it is unable to separate the carbon events from the amorphous noise events or from the proton cluster.

\begin{figure}[H]
\centering

	\subfloat{
	\includegraphics[width=0.5\textwidth]{plots/filtered_mixae_conf_mat.pdf}
}
	\hspace{-1cm}
	\subfloat{
	\includegraphics[width=0.5\textwidth]{plots/full_mixae_conf_mat.pdf}
}
\caption[MIXAE - confusion matrices]{Confusion matrices for the MIXAE clustering algorithm on filtered and full AT-TPC events. The true labels indicate samples belonging to the p (proton), carbon (C), or other classes. }\label{fig:mixae_confmat}
\end{figure}

Lastly we wish to further investigate if there are systematic differences between proton events that were placed in different clusters for the clean and full data. From figure \ref{fig:mixae_confmat} we see that the MIXAE algorithm creates two proton clusters for the filtered data, and places about fifty per-cent of the proton events in a cluster with the amorphous "other" events. We extract some proton events from these clusters to inspect whether systematic differences occur.

\begin{figure}
\centering
\includegraphics[width=\textwidth]{plots/clean_cluster_repr.pdf}
\caption[Selection of carbon events in differing clusters]{Selection of carbon events belonging to different clusters. Each row represents one of the clusters for the filtered data as shown in figure \ref{fig:mixae_confmat}. There seems to be no clear distinction between these rows.}\label{fig:filtered_mixae_clster_repr}	
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\textwidth]{plots/real_cluster_repr.pdf}	
\caption[Selection of proton events in differing clusters]{Selection of proton events belonging to different clusters. Each row represents one of the clusters for the full data as shown in figure \ref{fig:mixae_confmat}. We observe a tendency for the more noisy events to be placed in the "other" cluster}\label{fig:full_mixae_clster_repr}	
\end{figure}

### 2020-03-19 files_from_roberts_thesis/results/clustering/vgg.tex
notes and files from thesis

% !TEX spellckeck=en_GB

\section{Clustering using a pre-trained model}\label{sec:vgg_cluster}

As in chapter \ref{ch:architectures}, we also use the VGG16 pre-trained network as a baseline for the clustering performance. We begin by considering a classical K-means approach to clustering. However, the output from the VGG16 network is very high dimensional with output vectors in $\R^{8192}$. One of the primary concerns is then the curse of dimensionality, where the ratio of distances goes to one with increasing dimensionality as shown by \citet{Aggarwal}. However, one of the central caveats to the authors  finding is that the elements are uniformly distributed in the latent space. It is then possible that all the class information lies in some sub-space of the latent data. To investigate this, we perform clustering analysis using the full representation and the $10^2$ first principal components only.

\subsection{K-means}

We begin by investigating the K-means clustering algorithm on the VGG16 latent space. As in chapter \ref{chap:classification} the VGG16 model is pre-trained on the imagenet dataset creating a set of vectors $\boldsymbol{x} \in \R^{8192}$. To cluster we use \lstinline{scikit-learn} implementation of the K-means algorithm, with default parameters \cite{Pedregosa2011}. The results of the clustering runs are included in table \ref{tab:clstr_vgg}. We observe that we are able to attain near-perfect clustering on simulated data and that there is a sharp decline in performance as we add noise by moving to the filtered and full datasets.

\begin{table}[H]
\centering 
\caption[K-means on pre-trained model]{K-means clustering results on AT-TPC event data. We observe that the performance predictably decreases with the amount of noise in the data.}\label{tab:clstr_vgg}
\input{./plots/kmeans_vgg.tex}
\end{table}

In addition to the performance measures reported in table \ref{tab:clstr_vgg}, it is interesting to observe which samples are being wrongly assigned. We achieve this by tabulating the assignments of samples relative to their ground truth labels. From these tables, we can infer which classes are more or less entangled with others. We tabulate the results for each dataset is in figure \ref{fig:clster_confmat}. We observe that the proton class is consistently assigned in a pure cluster. Purity is inferred by how much spread there is in the column between the ground truth labels. A high-quality cluster will, in addition to being pure, also capture most entries the class represented by the cluster. For example, consider the row corresponding to the proton class in figure \ref{fig:clster_confmat}. The column corresponding to the largest entry in the proton row has zero other predicted classes in it. From this, we conclude that the proton cluster is a high quality, high purity cluster.

\begin{figure}
\centering

	\subfloat{
	\includegraphics[width=0.35\textwidth]{./plots/Simulatedvgg_pca_conf_mat.pdf}

}
	\hspace{-1cm}
	\subfloat{
	\includegraphics[width=0.35\textwidth]{plots/Filteredvgg_pca_conf_mat.pdf}
}
	\hspace{-1cm}
	\subfloat{
	\includegraphics[width=0.35\textwidth]{plots/Fullvgg_pca_conf_mat.pdf}
}
\caption[Pre-trained network - confusion matrices]{Confusion matrices for the K-means clustering of simulated, filtered and full AT-TPC events. The true labels indicate samples belonging to the p (proton), carbon (C), or other classes. }\label{fig:clster_confmat}
\end{figure}

We repeat this analysis using a PCA dimensionality reduction on the latent space of the VGG16 model. This is done to estimate to what degree the class separating information is encoded in the entirety of the latent space, or in some select regions. The results from the PCA analysis were virtually identical to the results sans the PCA, and so we omit them for brevity.

Furthermore, we wish to characterize further the clusters presented in figure \ref{fig:clster_confmat}. To achieve this, we sample from the proton samples belonging to different clusters for the filtered and full data.

\begin{figure}
\includegraphics[width=\textwidth]{plots/filtered_vgg_cluster_repr.pdf}
\caption[Filtered proton samples by cluster belonging]{Illustrating a sample of proton events from different K-means clusters from the filtered dataset. Each row belongs to a single cluster corresponding to the filtered confusion matrix in figure \ref{fig:clster_confmat}}\label{fig:filtered_vgg_clster_repr}
\end{figure}

\begin{figure}
\includegraphics[width=\textwidth]{plots/full_vgg_cluster_repr.pdf}
\caption[Full proton samples by cluster belonging]{Illustrating a sample of proton events from different K-means clusters from the full dataset. Each row belongs to a single cluster corresponding to the full confusion matrix in figure \ref{fig:clster_confmat}}\label{fig:full_vgg_clster_repr}
\end{figure}

In addition to the results presented in this section, we performed clustering with a number of different algorithms included in the \lstinline{scikit-learn} package. None of them provided any notable differences from the K-means results or were significantly worse. Notably, the DBSCAN algorithm failed to provide any useful clustering results. We find this important as one of the significant drawbacks of K-means, and the deep clustering algorithms presented in section \ref{sec:deep_clustering}, is that they are all dependent on pre-determining the number of clusters. This is not the case for DBSCAN.

### 2020-03-19 files_from_roberts_thesis/results/results.tex
notes and files from thesis

\chapter{Experimental setup and design}

In this chapter, we describe the semi-supervised classification and clustering results obtained on the three active target time projection chamber (AT-TPC) datasets described in section \ref{sec:data}. The simulated data is used to provide a benchmark for the upper bound on the performance of a given algorithm. This chapter is structured by task first, and algorithm second. As such, we will begin with a consideration of the semi-supervised algorithms before continuing with the clustering task.

 We explore the models proposed in chapter \ref{ch:autoencoder} on two disparate tasks, one of semi-supervised classification and one of clustering. For each task, we evaluate the performance on each of the datasets using appropriate metrics, which were introduced in section \ref{sec:performance_val}.

 The primary objective for the ${}^{46}$Ar experiment was to identify resonant proton scattering events, but this thesis explores broader applications than this classification task. This broader picture is explicitly geared to the application of the models discussed in this thesis to other AT-TPC experiments. Following this argument, we measure individual class performance wherever appropriate.

The machine learning experiments conducted in this thesis was performed using the AI-Hub computational cluster at the University of Oslo.  This resource consists of three machines with four RTX $2080$ Nvidia GPU's (graphics processing unit) each. These cards have $\sim 10$GB of memory available for the allocation of models.

\subsection{Semi-supervised classification procedure}\label{sec:clf_procedure}

Our interest in the semi-supervised classification is three fold. Firstly, it provides a bound on the performance we can expect from the clustering algorithms, as the classifier problem is fundamentally an easier task. In addition to the bounding property, the semi-supervised task proved to be an excellent way to explore the types of algorithms we present in this thesis. Secondly, we wish to characterize the performance as a function of the number of available labelled samples. As previously mentioned, one of the principal challenges with the AT-TPC detector is the cost of finding examples of the positive class, if at all possible. Thirdly, we wish to lay the groundwork for a transfer-learning approach to bridging AT-TPC experiments. Take, for example, resonant proton scattering. This could also be a reaction of interest in a different experiment. Applying models trained to recognize these events in the ${}^{46}$Ar experiment could be a feasible approach. The results from the semi-supervised classification experiments are presented in chapter \ref{chap:classification}

The semi-supervised task is to train an autoencoder model on a large dataset, and evaluate the class separation of the latent space using a logistic regression classifier. We use a logistic regression classifier as we wish to measure qualities of the latent space, and not of the classification per se. To provide an additional benchmark for our algorithms, we measure the performance of a logistic regression classifier using the latent space of a pre-trained image classifier model, which has shown to be effective in the classification of events from the ${}^{46}$Ar experiment \cite{Kuchera2019}.

Intrinsic to the measurement of the semi-supervised performance is the budgeting of how many labelled samples one can feasibly extract. Moreover, the principal limitation of the semi-supervised approach is the assumption that the researchers can identify the event class(es) of interest positively. It is then interesting to quantify the change in model performance as a function of how many labelled samples the classification model has to train on. Bear in mind that the representation that the classification model sees is still trained on the full set of events for a given dataset.

For reference, the models are described in terms of their hyperparameters in table \ref{tab:convae_hyperparams} for the convolutional autoencoder and table \ref{tab:draw_hyperparams} for the DRAW-analogues. To determine the best hyperparameters, we perform on the order of $\sim 10^2$ runs, each taking on the order of $\sim 10^1$ minutes on a single Nvidia $2080$ GPU. For each configuration of the algorithms, we train a classifier on a subset of the labelled set. Subsequently, this classifier is evaluated on the remainder of the labelled data to estimate the out of sample (OOS) error. The best configuration is then lastly re-trained, and we evaluate the OOS error with k-fold cross validation as outlined in section \ref{sec:performance_val}.

\subsection{Clustering procedure}\label{sec:cster_procedure}

While the semi-supervised classification task provides context and understanding to the AT-TPC experiments, the clustering procedure is a direct attempt at solving challenges associated with these experiments. First and foremost is the challenge of acquiring labelled data of the event of interest. The clustering results are presented in chapter \ref{chap:clustering}

In contrast to the semi-supervised task, the clustering objective presumes that we have access to no labelled data. However, since the purpose of this thesis is in large part exploratory, we will measure performance on the labelled data provided. We explore the performance of the deep clustering algorithms described in section \ref{sec:deep_clustering} as well as a K-means approach using the latent space of a pre-trained image classifier model.

The clustering task is intrinsically a more challenging task than that of semi-supervised classification, and so more time was spent exploring algorithms to determine which were suitable for clustering AT-TPC data. Before training each of the algorithms in section \ref{sec:deep_clustering} was first configured to reproduce the results from their respective papers. Subsequently, they were applied to the AT-TPC data, with their autoencoder hyperparameters selected empirically from the semi-supervised task. This was done to enable a focus on the clustering components of the algorithm. In particular, we focus on the weighting parameters of the mixture of autoencoders (MIXAE) clustering algorithm.

### 2020-03-22 doc/Manuscript/article.tex
Updates from Overleaf
added discussion sections - cleanup to follow

@@
 \usepackage{amssymb}
+\usepackage{amsmath}
+\usepackage{float}
+\usepackage{listings}
+
+\newcommand{\inner}[2]{\langle #1 | #2 \rangle}
+\newcommand{\R}{\mathbb{R}}
+\newcommand{\wij}{W_{ij}}
+\newcommand{\loss}{\mathcal{L}}
+
 \journal{Nuclear Instruments and Methods
@@
 
-\title{Latent Variable Machine Learning Algorithms and Classification of Events}
+\title{Latent Variable Clustering for Classification of Events in the AT-TPC}
 
@@
 \section{Results and Discussions}\label{sec:results}
+The principal challenge in the AT-TPC experiments that we are trying to solve is the reliance on labelled samples in the analysis as future experiments may not have as visually dissimilar reaction products  as we observe in the ${}^{46}$Ar experiment.
+The  ${}^{46}$Ar experiment does, however, provide a useful example where we can then explore unsupervised techniques.
+In this chapter, we explore the application of clustering techniques to events represented in latent spaces.
 
+We begin by exploring a naive K-means approach on the latent space of a pre-trained network.
+Subsequently, we investigate other clustering methods and two autoencoder based clustering algorithms, as outlined in section \ref{sec:deep_clustering}.
+
+This chapter builds on the previous results from semi-supervised classification.
+We observe that we can construct high-quality latent spaces.
+These high-quality spaces facilitate an investigation of clustering techniques.
+
+The approach for clustering of events is different from the semi-supervised approach in two meaningful ways.
+First, it is a harder task, as we will demonstrate.
+The clustering task thus necessitates a more exploratory approach to the problem.
+Second, as a consequence of the challenge, the focus will be a bit different than for the semi-supervised approach.
+We will still utilize the same architectures and models starting with a search over the parameter space over which we measure the performance using the adjusted rand score (ARS) and accuracy defined in section \ref{sec:unsupervised_perf} and \ref{sec:supervised_perf}, respectively.
+
+As with chapter  \ref{chap:classification} where we explored the semi-supervised results, we begin this chapter by considering the VGG16 pre-trained model as a benchmark.
+
+Lastly, we note that the focus of this work is mainly on discovering possible avenues for further research.
+This focus requires a broad scan of possible avenues rather than a rigorous analysis of one specific model.
 \subsection{Transfer learning}
+As in chapter \ref{ch:architectures}, we also use the VGG16 pre-trained network as a baseline for the clustering performance.
+We begin by considering a classical K-means approach to clustering.
+However, the output from the VGG16 network is very high dimensional with output vectors in $\R^{8192}$.
+One of the primary concerns is then the curse of dimensionality, where the ratio of distances goes to one with increasing dimensionality as shown by \citet{Aggarwal}.
+However, one of the central caveats to the authors  finding is that the elements are uniformly distributed in the latent space.
+It is then possible that all the class information lies in some sub-space of the latent data.
+To investigate this, we perform clustering analysis using the full representation and the $10^2$ first principal components only.
 
+\subsection{K-means}
+
+We begin by investigating the K-means clustering algorithm on the VGG16 latent space.
+As in chapter \ref{chap:classification} the VGG16 model is pre-trained on the imagenet dataset creating a set of vectors $\boldsymbol{x} \in \R^{8192}$.
+To cluster we use \lstinline{scikit-learn} implementation of the K-means algorithm, with default parameters \cite{Pedregosa2011}.
+The results of the clustering runs are included in table \ref{tab:clstr_vgg}.
+We observe that we are able to attain near-perfect clustering on simulated data and that there is a sharp decline in performance as we add noise by moving to the filtered and full datasets.
+
+\begin{table}[H]
+\centering 
+\caption[K-means on pre-trained model]{K-means clustering results on AT-TPC event data.
+We observe that the performance predictably decreases with the amount of noise in the data.}\label{tab:clstr_vgg}
+\input{./plots/kmeans_vgg.tex}
+\end{table}
+
+In addition to the performance measures reported in table \ref{tab:clstr_vgg}, it is interesting to observe which samples are being wrongly assigned.
+We achieve this by tabulating the assignments of samples relative to their ground truth labels.
+From these tables, we can infer which classes are more or less entangled with others.
+We tabulate the results for each dataset is in figure \ref{fig:clster_confmat}.
+We observe that the proton class is consistently assigned in a pure cluster.
+Purity is inferred by how much spread there is in the column between the ground truth labels.
+A high-quality cluster will, in addition to being pure, also capture most entries the class represented by the cluster.
+For example, consider the row corresponding to the proton class in figure \ref{fig:clster_confmat}.
+The column corresponding to the largest entry in the proton row has zero other predicted classes in it.
+From this, we conclude that the proton cluster is a high quality, high purity cluster.
+
+\begin{figure}
+\centering
+
+	\subfloat{
+	\includegraphics[width=0.35\textwidth]{./plots/Simulatedvgg_pca_conf_mat.pdf}
+
+}
+	\hspace{-1cm}
+	\subfloat{
+	\includegraphics[width=0.35\textwidth]{plots/Filteredvgg_pca_conf_mat.pdf}
+}
+	\hspace{-1cm}
+	\subfloat{
+	\includegraphics[width=0.35\textwidth]{plots/Fullvgg_pca_conf_mat.pdf}
+}
+\caption[Pre-trained network - confusion matrices]{Confusion matrices for the K-means clustering of simulated, filtered and full AT-TPC events.
+The true labels indicate samples belonging to the p (proton), carbon (C), or other classes.
+}\label{fig:clster_confmat}
+\end{figure}
+
+We repeat this analysis using a PCA dimensionality reduction on the latent space of the VGG16 model.
+This is done to estimate to what degree the class separating information is encoded in the entirety of the latent space, or in some select regions.
+The results from the PCA analysis were virtually identical to the results sans the PCA, and so we omit them for brevity.
+
+Furthermore, we wish to characterize further the clusters presented in figure \ref{fig:clster_confmat}.
+To achieve this, we sample from the proton samples belonging to different clusters for the filtered and full data.
+
+\begin{figure}
+\includegraphics[width=\textwidth]{plots/filtered_vgg_cluster_repr.pdf}
+\caption[Filtered proton samples by cluster belonging]{Illustrating a sample of proton events from different K-means clusters from the filtered dataset.
+Each row belongs to a single cluster corresponding to the filtered confusion matrix in figure \ref{fig:clster_confmat}}\label{fig:filtered_vgg_clster_repr}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\textwidth]{plots/full_vgg_cluster_repr.pdf}
+\caption[Full proton samples by cluster belonging]{Illustrating a sample of proton events from different K-means clusters from the full dataset.
+Each row belongs to a single cluster corresponding to the full confusion matrix in figure \ref{fig:clster_confmat}}\label{fig:full_vgg_clster_repr}
+\end{figure}
+
+In addition to the results presented in this section, we performed clustering with a number of different algorithms included in the \lstinline{scikit-learn} package.
+None of them provided any notable differences from the K-means results or were significantly worse.
+Notably, the DBSCAN algorithm failed to provide any useful clustering results.
+We find this important as one of the significant drawbacks of K-means, and the deep clustering algorithms presented in section \ref{sec:deep_clustering}, is that they are all dependent on pre-determining the number of clusters.
+This is not the case for DBSCAN.
 \subsection{Autoencoder clustering}

### 2020-03-30 doc/Manuscript/article.tex
Updates from Overleaf

@@
 \usepackage{listings}
+\usepackage{amssymb}
+\usepackage{tabularx}
+\usepackage{tikz}
+\usepackage{float}
+\usepackage{mathtools}
+%\usepackage{minted}
+\usepackage[colorinlistoftodos]{todonotes}
+\usepackage[ruled,vlined]{algorithm2e}
+\usepackage{hyperref}
+\usepackage{enumerate}
+\usepackage[shortlabels]{enumitem}
+\usepackage{amssymb}
+\usepackage{subfig}
+\usepackage{makecell}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{mathpazo}
+%\usepackage{multicol}
+%\usepackage{siunits}
+\usepackage{siunitx}
+\usepackage{graphicx}
+\usepackage{longtable}
+\usepackage[toc,page]{appendix}
+\usepackage{booktabs}
+\usepackage{tabularx}
+\usepackage{adjustbox}
+\usepackage[section]{placeins}
+\usepackage{pdflscape}
+\usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc}
 
+\usepackage{natbib}
+\bibliographystyle{unsrtnat}
+
+\setlength{\marginparwidth}{4cm}
+\usepackage{todonotes}
 \newcommand{\inner}[2]{\langle #1 | #2 \rangle}
@@
 %% use optional labels to link authors explicitly to addresses:
-%% \author[label1,label2]{}
+%% \author[label1,label2]{}https://www.overleaf.com/project/5e73f990fcbd310001959b01
 %% \address[label1]{}
@@
 \end{itemize}
+
+\subsubsection{Pre-trained neural networks}
+
+Following the previous work of \citet{Kuchera2019}, we will consider representations of our events through the lens of a pre-trained network.
+In the Machine Learning community, it is not uncommon to publish packaged models with fitted parameters from image recognition contests.
+These models are trained on datasets with millions of images and classify between hundreds of distinct classes; one such is the imagenet dataset.
+In their work, \citet{Kuchera2019} use the VGG16 architecture trained on imagenet to classify AT-TPC events very successfully.
+We will build on the understanding of using these pre-trained networks in event classification by using VGG16 as an element in the end-to-end training of autoencoders.
+
+The VGG16 network is one of six analogous networks proposed by \citet{Simonyan2014}, and they were runners up in the ISLVRC(ImageNet large scale visual recognition competition) of 2014 \cite{Russakovsky2015}.
+The network architectures are fairly simple; for VGG16, there are sixteen layers in the network.
+The first thirteen of which are convolutional layers with exclusively $3 \times 3$ kernels.
+The choice of the kernel size is based on the fact that a stacked $3 \times 3$ kernel is equivalent to larger kernels in terms of the receptive field of the output.
+Three $3 \times 3$ kernels with stride $1$ have a $7 \times 7$ receptive field, but the larger kernel has $81\%$ more parameters and only one non-linearity \cite{Simonyan2014}.
+Stacking the smaller kernels then contributes to a lower computational cost.
+Additionally, there is a regularizing effect from the lowered number of parameters and increased explanatory power from the additional non-linearities.
+The full architecture is detailed in appendix \ref{tab:vgg}.
+
+\subsubsection{Deep clustering: Mixture of autoencoders}\label{sec:mixae}
+
+Another way of representing the clusters is by having multiple latent spaces representing the underlying manifolds that describe each class.
+This is the central idea in the Mixture of autoencoders (MIXAE) algorithm, introduced by  \cite{Zhang}.
+To ensure that each autoencoder represents a cluster, we attach a soft-max classifier to the set of latent samples.
+Recall from equation \ref{eq:softmax} that the soft-max function transforms a network output to a prediction of relative confidences.
+This classifier assigns a cluster to each autoencoder, coupling the latent space and the reconstructions.
+The soft-max classifier is trained to output cluster probabilities and penalized for collapsing to assigning one cluster only.
+The remaining task is then to connect the cluster assignments to the autoencoder reconstructions.
+This is achieved by multiplying by the cluster confidence with the reconstruction error of each autoencoder.
+
+\begin{figure}[tb]
+	\centering
+	\includegraphics[width=\textwidth]{plots/mixae.pdf}
+	\caption[Deep convolutional embedded clustering schematic]{Schematic of a DCEC model.
+A sample $\hat{\boldsymbol{x}}$ is compressed to set of lower-dimensional representations $\{\boldsymbol{z}^{(i)}\}$ by $N$ autoencoders.
+These samples are concatenated and passed through an auxiliary assignment network that predicts a confidence of cluster belonging for each autoencoder.
+For further details see the text.
+Figure adapted from \citet{Zhang}}
+	\label{fig:mixae}
+\end{figure}
+
+More formally let $\{\boldsymbol{z}_j^{(i)}\}^N$ be the set of $N$ latent samples from each of the $N$ auto-encoders from a single sample image $\hat{\boldsymbol{x}}^{(i)} \in \R ^{t\times v}$ with height $t$ and width $v$.
+Furthermore, let the soft cluster assignments be given as $\{p_j^{(i)}\}^N$ and the reconstructed samples be given as $\{\boldsymbol{x}_j^{(i)}\}$.
+The reconstruction loss is then the sum over each autoencoder multiplied with each cluster assignment, i.e.
+
+\begin{equation}\label{eq:mixae_reconst}
+\mathcal{L}_x = \sum_j p_j^{(i)} \mathcal{C}(\boldsymbol{x}_j^{(i)}, \hat{\boldsymbol{x}}^{(i)}),
+\end{equation}
+
+\noindent where $\mathcal{C}(\cdot)$ is a cost function like the mean squared error or a cross-entropy.
+
+We ensure that the soft cluster assignments encourage clustering by adding two terms to the total loss.
+The first is a simple entropy term which, when minimized, encourages the assignments to be one-hot vectors.
+Mathematically this loss can then be written as the entropy $S(\cdot)$ of a cluster assignment $\boldsymbol{p}$
+
+\begin{equation}\label{eq:mixae_sample}
+S(\boldsymbol{p}^{(i)})_{\text{sample}} = -\sum_j p_j^{(i)} \log p_j^{(i)}. 
+\end{equation}
+
+\noindent The last ingredient in the loss is then the term which discourages the trivial solution where only one cluster is being assigned.
+This loss is a batch-wise entropy term, given a mini-batch of $\beta$ samples the loss is computed as
+
+\begin{equation}\label{eq:mixae_batch}
+\begin{split}
+S(\{\boldsymbol{p}^{(i)}\}_i)_{\text{batch}}&= \left(-\sum_j \bar{p}_j \log \bar{p}_j \right)^{-1},\\
+\bar{\boldsymbol{p}} &= \frac{1}{\beta} \sum_i ^\beta \boldsymbol{p}^{(i)} .
+\end{split}
+\end{equation}
+
+\noindent The negative exponent in equation \ref{eq:mixae_batch} owes to the fact that we want to maximize the batch-wise entropy.
+
+From equations \ref{eq:mixae_reconst}, \ref{eq:mixae_sample}, and \ref{eq:mixae_batch} we can then compose the full loss for the MIXAE model.
+\cite{Zhang} note that the optimization depends heavily on the weighting of the different terms and so we introduce the weighting hyperparameters $\alpha$, $\gamma$ and $\theta$, and write the total loss over a mini-batch as
+
+\begin{equation}\label{eq:mixae_loss}
+\begin{split}
+\mathcal{L}_{\text{total}}(\{\hat{\boldsymbol{x}}^{(i)}\}^\beta_i) = \frac{1}{\beta}\sum_i \Big( &\frac{t\cdot v}{\theta} \mathcal{L}_x(\boldsymbol{p}^{(i)},\,\boldsymbol{x}^{(i)},\, \hat{\boldsymbol{x}}^{(i)} ) \\
+&+ \alpha S(\boldsymbol{p}^{(i)})_{\text{sample}} \Big) \\
+&+\gamma S(\{\boldsymbol{p}^{(i)}\}_i)_{\text{batch}}.
+\end{split}
+\end{equation}
+
+\noindent \citet{Zhang} note that this algorithm has a couple of shortcomings.
+The batch-entropy term encourages the cluster assignments to be uniformly distributed within the classes.
+However, it does have a second minimum when the assignments are all equally likely.
+For biased datasets with significant variations in the number of samples in each class, this can create problems for the algorithm.
+Additionally, the number of clusters are not dynamically computed by the algorithm but hard-coded by the researchers.
+This can be problematic if the number of cluster are not known prior to analysis.
+
 \section{Results and Discussions}\label{sec:results}
@@
 
+\todo{Drop this argument or build it in the ML section?}
 This chapter builds on the previous results from semi-supervised classification.
@@
 
+\todo{Reword this to not contrast against semi-supervised}
 The approach for clustering of events is different from the semi-supervised approach in two meaningful ways.
@@
 However, the output from the VGG16 network is very high dimensional with output vectors in $\R^{8192}$.
-One of the primary concerns is then the curse of dimensionality, where the ratio of distances goes to one with increasing dimensionality as shown by \citet{Aggarwal}.
+\todo{reword} One of the primary concerns is then the curse of dimensionality, where the ratio of distances goes to one with increasing dimensionality as shown by \citet{Aggarwal}.
 However, one of the central caveats to the authors  finding is that the elements are uniformly distributed in the latent space.
@@
 
-\section{Conclusions and Perspectives}\label{sec{conclusion}
+In the previous section we demonstrated a potential for clustering of events from an AT-TPC experiment.
+To build on this result we will in this section explore the application of the mixture of autoencoders (MIXAE) algorithm introduced in section \ref{sec:deep_clustering}.
+The other algorithm introduced in section \ref{sec:deep_clustering}, deep clustering with convolutional autoencoders (DCEC), consistently collapsed to just one cluster for all datasets.
+
+In the MIXAE algorithm the hyper-parameters to adjust are all the ordinary parameters that we introduced in table \ref{tab:convae_hyperparams}.
+In addition to those parameters we have the weighting of the loss terms: $\theta$, $\alpha$ and $\gamma$.
+These weighting parameters are attached to the reconstruction loss, sample entropy and batch-wise entropy respectively.
+\cite{Zhang} note that these parameters are very important for the model performance and so we focus primarily on these.
+
+\todo{but why}We empirically freeze the convolutional autoencoder hyperparameters to compress the original $128 \times 128$ to a $8 \times 8$ pixel grid using four convolutional layers.
+The parameters chosen for the autoencoders are listed in full in table \ref{tab:mixe_ae_hyperparams}.
+
+\begin{table}[H]
+\todo{prettify or remove}
+\renewcommand*{\arraystretch}{0.5}
+\centering
+\caption{Hyperparameters selected for the autoencoder components of the MIXAE algorithm, see table \ref{tab:convae_hyperparams} for a full description of the parameters.}\label{tab:mixe_ae_hyperparams}
+\setlength{\extrarowheight}{15pt}
+\hspace*{-0.5in}
+\begin{tabular}{ll}
+\toprule
+Hyperparameter & Value \\
+\midrule
+\multicolumn{2}{l}{Convolutional parameters: } \\
+\midrule
+Number of layers & $4$ \\
+Kernels & $[3,\,3,\,3,\,3]$\\
+Strides & $[2,\,2,\,2,\,2]$ \\
+Filters & $[64,\, 32, \,16, \,8,]$ \\ 
+\midrule
+\multicolumn{2}{l}{Network parameters: } \\
+\midrule
+Activation & LReLu \\
+Latent type & None \\
+Latent dimension & 20  \\
+$\beta$ & N/A \\
+Batchnorm & False \\
+\midrule
+\multicolumn{2}{l}{Optimizer parameters: } \\
+\midrule
+$\eta$ & $10^{-3}$ \\
+$\beta_1$ & $0.9$ \\
+$\beta_2$ & $0.99$ \\
+\bottomrule
+\end{tabular}
+\end{table}
+
+Since there are then only three remaining hyperparameters we choose to perform a coarse grid-search as described in section \ref{sec:hyperparam_search_arch}.
+
+\begin{table}
+\centering
+\caption{Hyperparameter grid for the MIXAE loss weighting terms. The grid is given as exponents for logarithmic scales.}\label{tab:mixae_loss_weights}
+\begin{tabular}{lll}
+\toprule
+Parameter & Grid & Scale \\
+\midrule 
+$\theta$ & $[-1,\, 5]$ & Logarithmic \\
+$\alpha$ & $[-5,\, -1]$ & Logarithmic \\
+$\gamma$ & $[3,\, 5]$ & Logarithmic
+\end{tabular}
+\end{table}
+
+\subsection{Simulated AT-TPC data}
+
+To train the MIXAE clustering algorithm, we use the large simulated dataset with $M=80000$ points, evenly distributed between proton- and carbon-events.
+The algorithm is trained on a subset of $60000$ of these samples, and we track performance on the remaining $20000$ events.
+
+The grids selected for the search are listed in table \ref{tab:mixae_loss_weights}. The search yielded an optimal configuration with
+
+\begin{align}
+\theta = 10^{-1}, \\
+\alpha = 10^{-2}, \\
+\gamma = 10^5.
+\end{align}
+
+Finally, for these parameters we re-ran the algorithm $N=10$ times to investigate the stability of the algorithm.
+The results are reported in figure \ref{fig:mixae_sim}.
+We observe that while the algorithm can achieve very strong performance, with an adjusted rand index (ARI) $ > 0.8$, it fluctuates strongly with repeated experiments.
+The $ARI$ is a clustering performance measure similar to accuracy, but adjusted for chance.
+We elaborate on the $ARI$ score in section \ref{sec:unsupervised_perf}.
+As mentioned in section \ref{sec:mixae} the batch-entropy has a second minimum when the cluster confidences are near equal.
+It is this behavior we observe in figure \ref{fig:mixae_sim}.
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=\textwidth]{./plots/sim_mixae.pdf}
+\caption[Clustering performance of MIXAE on simulated AT-TPC data]{Performance for the MIXAE model on simulated AT-TPC data.
+In the top row the loss components are plotted for each run, and in the bottom row the adjusted rand index (ARI) and clustering accuracy are shown.
+Each run is color-coded with the ARI achieved at the end of the run.
+}\label{fig:mixae_sim}
+\end{figure}
+
+\subsection{Filtered AT-TPC data}
+
+We repeat the optimization steps in the previous section for the filtered AT-TPC data.
+The exception being that we allow the algorithm to train on the labelled samples.
+Beginning with a wide grid equal to the grid used for the simulated data we searched over all parameter configurations to find promising values.
+We then performed a local search around these values to pin-point the hyperparameter configuration.
+This search yielded the optimal hyper-parameters
+
+\begin{align}
+\theta &= 10^{1}, \\
+\alpha &= 10^{-1}, \\
+\gamma &= 3.162\times 10^3.
+\end{align}
+
+\noindent The results of the runs are included in figure \ref{fig:mixae_clean}.
+We observe ghat the highest performing models reach an $ARI > 0.5$, which is higher than the performance achieved by the K-means algorithm applied to the VGG16 latent space.
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=\textwidth]{./plots/clean_mixae.pdf}
+\caption[Clustering performance of MIXAE on filtered event AT-TPC data]{Performance for the MIXAE model on filtered AT-TPC data.
+In the top row the loss components are plotted for each run, and in the bottom row the adjusted rand index (ARI) and clustering accuracy are shown.
+Each run is color-coded with the ARI achieved at the end of the run.}\label{fig:mixae_clean}
+\end{figure}
+
+\subsection{Full AT-TPC data}
+
+As in the two previous sections, we repeat the same procedure of iterative grid searches on the MIXAE loss-weights.
+Each configuration is re-run a total of $N=10$ times to capture fluctuations in the performance before a final selection is made on the hyperparameters.
+For the full dataset the MIXAE hyperparameters converge to the same values as for the clean data, i.e.
+
+\begin{align}
+\theta &= 10^{1}, \\
+\alpha &= 10^{-1}, \\
+\gamma &= 3.162\times 10^3.
+\end{align}
+
+\noindent As with the previous datasets we include a plot of the loss curves and performance measures for $N=10$ runs with the same loss-weight parameters, shown in figure \ref{fig:mixae_full}.
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=\textwidth]{./plots/real_mixae.pdf}
+\caption[Clustering performance of MIXAE on full event AT-TPC data]{Performance for the MIXAE model on un-filtered AT-TPC data.
+In the top row the loss components are plotted for each run, and in the bottom row the adjusted rand index (ARI) and clustering accuracy are shown.
+Each run is color-coded with the ARI achieved at the end of the run.}\label{fig:mixae_full}
+\end{figure}
+
+\subsection{Comparing performance}
+
+It is also interesting to compare and contrast the clustering results from the MIXAE model with those of the VGG16$+$K-means outside the fairly abstract accuracies and rand scores.
+It is especially interesting to compare the cluster assignments, as they can inform further research efforts in the clustering of AT-TPC events.
+We illustrate the clustering with confusion matrices that are shown in figure \ref{fig:mixae_confmat}.
+From these matrices, we observe that the MIXAE applied to the clean data correctly clusters the noise events.
+Additionally, it identifies two proton clusters.
+We observe that these proton clusters are both less pure than the VGG16+K-means clusters and that there does not seem to be a visually meaningful difference between these clusters.
+The latter is inferred from figure \ref{fig:filtered_mixae_clster_repr}.
+
+Applied to the real data the MIXAE correctly separates the proton class, however it is unable to separate the carbon events from the amorphous noise events or from the proton cluster.
+
+\begin{figure}[H]
+\centering
+
+	\subfloat{
+	\includegraphics[width=0.5\textwidth]{plots/filtered_mixae_conf_mat.pdf}
+}
+	\hspace{-1cm}
+	\subfloat{
+	\includegraphics[width=0.5\textwidth]{plots/full_mixae_conf_mat.pdf}
+}
+\caption[MIXAE - confusion matrices]{Confusion matrices for the MIXAE clustering algorithm on filtered and full AT-TPC events.
+The true labels indicate samples belonging to the p (proton), carbon (C), or other classes.
+}\label{fig:mixae_confmat}
+\end{figure}
+
+Lastly we wish to further investigate if there are systematic differences between proton events that were placed in different clusters for the clean and full data.
+From figure \ref{fig:mixae_confmat} we see that the MIXAE algorithm creates two proton clusters for the filtered data, and places about fifty per-cent of the proton events in a cluster with the amorphous "other" events.
+We extract some proton events from these clusters to inspect whether systematic differences occur.
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{plots/clean_cluster_repr.pdf}
+\caption[Selection of carbon events in differing clusters]{Selection of carbon events belonging to different clusters.
+Each row represents one of the clusters for the filtered data as shown in figure \ref{fig:mixae_confmat}.
+There seems to be no clear distinction between these rows.}\label{fig:filtered_mixae_clster_repr}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{plots/real_cluster_repr.pdf}	
+\caption[Selection of proton events in differing clusters]{Selection of proton events belonging to different clusters.
+Each row represents one of the clusters for the full data as shown in figure \ref{fig:mixae_confmat}.
+We observe a tendency for the more noisy events to be placed in the "other" cluster}\label{fig:full_mixae_clster_repr}
+\end{figure}
+
+\section{Conclusions and Perspectives}\label{sec{conclusion}}
 
 \appendix  
+\section{Neural network architectures}
 
-\begin{thebibliography}{00}
+\begin{table}
+\caption[vgg architectures]{Showing the details of the VGG network architectures.
+Network D trained on the ImageNet \cite{Russakovsky2015} dataset the network known as  VGG16 and is what we use in this thesis.}\label{tab:vgg}
+\includegraphics[width=\textwidth]{plots/vgg_architectures.png}
+\end{table}
 
-%% \bibitem{label}
-%% Text of bibliographic item
-
-\bibitem{}
-
-\end{thebibliography}
+\bibliography{bibliography}
 \end{document}

### 2020-04-01 doc/Manuscript/article.tex
Updates from Overleaf
Started writing the introduction

@@
 \section{Introduction}\label{sec:intro}
+\begin{itemize}
+    \item Overall goals for the AT-TPC
+    \item Nuclear physics motivation for clustering-based (unsupervised) methods
+    \item Motivation from ML approach
 
+\end{itemize}
+The AT-TPC \cite{Bradt2018} is a novel type of detector designed specifically for nuclear physics experiments where the energies of the recoiling particles are very low compared to the energy required to escape target material \cite{}.
+The luminosity of nuclear physics experiments performed with fixed targets is directly proportional to the amount of material encountered by the beam.
+On the other hand, for several classes of experiments the detection of recoil particles is paramount, therefore limiting the target thickness.
+In addition, the properties of the recoil particles are modified while traversing the target material, affecting the resolutions that can be achieved.
+This necessary balance between luminosity and resolution is particularly difficult when performing experiments with rare isotope beams, because of the low intensities available.
+
+The concept of active target aims at mitigating this compromise, by turning the target itself into a detector \cite{Beceiro2018}.
+Most active target detectors such as the AT-TPC are composed of a time projection chamber (TPC) where the detector gas is at the same time the target material.
+Recoil particles that originate from a nuclear reaction between a beam nucleus and a gas nucleus can be tracked from the vertex of the reaction to their final position inside the active volume of the target.
+Their properties can therefore be measured without any loss of resolution regardless of the amount of material traversed by the beam.
+At the same time, the detection efficiency is dramatically increased by the very large solid angle covered in this geometry.
+A direct consequence of this concept is the inclusiveness of the experimental data recorded by this type of detector: any nuclear reaction happening within the target is recorded.
+Although this sounds like an advantage from the scientific point of view, it poses great challenges during the analysis phase, that are reminiscent of bubble chamber times and on par with nowadays event classification issues in particle physics.
+More often than not, the reaction channel of interest has one of the lowest cross sections, therefore the physicist is faced with the task of sorting out the corresponding events from the ``background" of other reaction channels.
+
+Because TPCs produce 3-dimensional images of charged particle tracks, the event classification task is often akin to a visual inspection (bubble chamber times), which is not practical nowadays because of the large quantities of data.
+Machine learning techniques then appear as a promising prospect, in particular in the image recognition domain where much progress has been made recently.
+In addition, some ML algorithms offer new possibilities such as the potential discovery of unforeseen phenomena that would have been missed by more traditional analysis methods.
+The classification and clustering of events based on various ML algorithm is hereby examined, using experimental data recorded by the AT-TPC during its commissioning experiment from a radioactive $^{46}$Ar beam reacting of an isobutane target composed of proton and carbon nuclei.
+
+\section{Experimental Details} 
+(46Ar)
 \section{Methods}\label{sec:methods}

### 2020-04-03 doc/Manuscript/article.tex
Updates from Overleaf
Experimental details

@@
 In addition, some ML algorithms offer new possibilities such as the potential discovery of unforeseen phenomena that would have been missed by more traditional analysis methods.
-The classification and clustering of events based on various ML algorithm is hereby examined, using experimental data recorded by the AT-TPC during its commissioning experiment from a radioactive $^{46}$Ar beam reacting of an isobutane target composed of proton and carbon nuclei.
+The classification and clustering of events based on various ML algorithms is hereby examined, using experimental data recorded by the AT-TPC during its commissioning experiment from a radioactive $^{46}$Ar beam reacting of an isobutane target composed of proton and carbon nuclei.
 
 \section{Experimental Details} 
-(46Ar)
+The goal of the experiment was to measure the evolution of the elastic and inelastic cross sections between $^{46}$Ar and protons as a function of energy (also called excitation function), and observe resonances in the composite system $^{47}$K that correspond to analog states in the nucleus $^{47}$Ar.
+Spectroscopic information can then be obtained from the shape and amplitude of the observed resonances \cite{Bradt2018}.
+The experiment was performed at the National Superconducting Cyclotron Facility (NSCL) where a $^{46}$Ar beam was produced via fragmentation of a $^{48}$Ca beam on a $^9$Be target at about 140 MeV/u.
+The $^{46}$Ar isotopes were then filtered, thermalized, and finally re-accelerated to 4.6 MeV/u by a linear accelerator.
+This scheme was used to produce a low emittance beam necessary to guarantee a good energy resolution in the excitation function.
+Because of the energy loss of the $^{46}$Ar beam particles as they traverse the target gas volume, the position of the reaction vertex along the beam axis is directly related to the energy at which the reaction occurs.
+This allows the AT-TPC to measure the excitation function over a wide range of energies from a single beam energy.
 \section{Methods}\label{sec:methods}
@@
         \item Supervised vs. Unsupervised learning
+        \item Define Classifcation and Clustering 
+        \item Define latent space (refer)
         \item Traditional unupervised - distances in high dimensional spaces etc. 

### 2020-04-03 doc/Manuscript/article.tex
Updates from Overleaf

--- 2020-04-01 doc/Manuscript/article.tex
@@
 
-\subsection{Machine Learning}
+\subsection{Data}
+ In this section we give a brief overview of the data, for a more in-depth consideration we refer to \cite{Mittig2015}, \cite{Suzuki2012} and  \cite{Bradt2017a}.
+
+In this thesis we will explore the machine learning algorithms described in chapters \ref{ch:ml} and \ref{ch:autoencoder} on three datasets from the AT-TPC, listed in table \ref{tab:datasets}.
+The simulated data will serve as a baseline for performance, while the filtered and full data are real records from the ${}^{46}$Ar experiment.
+Each of these datasets have different distributions of the constituent classes.
+We list these differences in table \ref{tab:class_distr}.
+The filtered data differs from the full dataset as it has some post-processing applied to remove noise.
+This is not the case for the full data.
+The details of the post-processing are given in later sections.
+First, we consider the pipeline from raw data to images that the algorithms described in chapter \ref{ch:autoencoder} can process.
+
+\begin{table}
+\centering
+\caption{Descriptions of number of events in the data used for analysis in this thesis.
+In principle we can simulate infinite data, but it is both quite simple and not very interesting outside a case for a proof-of-concept}\label{tab:datasets}
+\begin{tabular}{lccc}
+\toprule
+{} & Simulated & Full & Filtered \\
+\midrule
+Total &  $8000$ & $51891$ & $49169$ \\
+labelled & $2400$ & $1774$ &  $1582$ \\ 
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{table}
+\centering
+\caption{Event class percentages for each of the datasets used in the analysis conducted in this thesis.
+The decrease in the "other" class of events from full to filtered data owes to the thresholding of events.
+If an event contains fewer than $20$ data-points it is discarded}\label{tab:class_distr}
+\begin{tabular}{lccc}
+\toprule
+{} & Simulated & Full & Filtered \\
+\midrule
+$\%$ Proton & $50$ & $25.3$ & $27.6$ \\ 
+$\%$ Carbon & $50$ & $12.0$ & $11.9$ \\
+$\%$ Other & $0$ & $62.8$ & $60.1$ \\
+\end{tabular}
+\end{table}
+
+\subsection{Data processing}
+
+Our data processing pipeline begins with localized point-cloud data in the two-dimensional detector coordinate system, with one time dimension and a corresponding charge measurement.
+The charge and time measurement are extracted as the peak of this signal over the event, resulting in a maximum of one measurement per pad in the sensor plane.
+The events range from fairly sparse $< 10^2$ records to being very populated, depending on where in the volume the reaction occurs as well as other noise-generating factors.
+We centre the charge data to values $>1$ by adding the lowest occurring record in the run and apply a log-transform to get values in $\R^+$.
+Subsequently, we scale by the maximum value in the run to get charge data in the interval $[0, 1]$.
+Lastly, the transformed charge data is saved in a two-dimensional $128px\, \times \,128px$ array using the \lstinline{matplotlib} package provided in the \lstinline{Python} programing language \cite{matplotlib}.
+The choice of scaling was made to accommodate a binary cross-entropy loss on a 2D projection, as it presumes the true values to be bounded as probabilities.
+
+We will begin by considering the simulated events, followed by subsequent considerations of the full and filtered experimental data.
+
+\subsection{Simulated \texorpdfstring{${}^{46}$Ar}{46Ar}  events}\label{sec:data_sim}
+
+The simulated AT-TPC tracks were simulated with the \lstinline{pytpc} package developed by \citet{Bradt2017a}.
+Using the same parameters as for the ${}^{46}$Ar$(p, p)$ experiment, a small set of $N=4000$ events were generated per class, as well as a larger set of $N=40000$ events per class.
+The events are generated as point-clouds, consisting of position data on the x-y plane, a time-stamp and an associated charge value.
+These point-clouds are transformed into pure x-y projections with charge intensity for the analysis in this thesis using the pipeline described in the previous section.
+
+More formally the events are originally composed of peak-only 4-tuples of $e_i = (x_i, y_i, t_i, c_i)$.
+The peak-only designation indicates that we use the recorded peak amplitude on each pad, the tuples then correspond to pads that recorded a signal for that event.
+Each event is then a set of these four-tuples: $\epsilon_j = \{e_i\}$ creating a track in three-dimensional space.
+
+To emulate the real-data case, we set a subset of the simulated data to be labelled and treat the rest as unlabelled data.
+We chose this partition to be $15\%$ of each class.
+We denote this subset and its associated labels as $\gamma_L=(\boldsymbol{X}_L, \boldsymbol{y}_L)$, the entire dataset which we will denote as $\boldsymbol{X}_F$.
+To clarify, please note that $\boldsymbol{X}_L \subset \boldsymbol{X}_F$.
+
+We display two simulated events in figure \ref{fig:sim_samples}. The top row illustrates a proton-event, and the bottom a carbon-event.
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=\textwidth]{plots/display_eventssimulated.pdf}
+\caption[Displaying simulated events in 2D and 3D]{Two- and three-dimensional representations of two events from a simulated ${}^{46}$Ar experiment.
+Each row is one event in two projections, where the color intensity of each point indicates higher charge values recorded by the detector.}\label{fig:sim_samples}
+\end{figure}
+
+\subsection{Full \texorpdfstring{${}^{46}$Ar}{46Ar}  events}\label{sec:data_real}
+
+The events analyzed in this section were retrieved from the ${}^{46}$Ar resonant proton scattering experiment recorded with the AT-TPC.
+
+The sensor plane in the AT-TPC is very sensitive, as such there is substantial noise in the ${}^{46}$Ar data.
+The noise can be attributed to structural noise from electronics cross-talk, and possible interactions with cosmic background radiation, as well as other sources of charged particles.
+Part of the challenge for this data then comes from understanding of the physics of the major contributing factors to this noise.
+
+We display two different events from the ${}^{46}$Ar experiment in figure \ref{fig:samples}.
+The top row illustrates an event with a large fraction of noise, while the bottom row shows an event nearly devoid of noise.
+The very clear spiral structure of the bottom row indicates that this is a proton-event.
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=\textwidth]{plots/display_eventsfull_.pdf}
+\caption[Displaying un-filtered events in 2D and 3D]{Two- and three-dimensional representations of two events from the ${}^{46}$Ar experiment.
+Each row is one event in two projections, where the color intensity of each point indicates higher charge values recorded by the detector.}\label{fig:samples}
+\end{figure}
+
+\subsection{Filtered \texorpdfstring{${}^{46}$Ar}{46Ar} events}\label{sec:filtered}
+
+As we saw in the previous section, the detector picks up a significant amount of noise.
+We split the noise broadly in two categories,  one being random-uncorrelated noise and the second is structured noise.
+The former can be quite trivially removed with a nearest-neighbour algorithm that checks if a point in the event is close to any other.
+To remove the correlated noise, researchers at the NSCL developed an algorithm based on the Hough transform \cite{Newman1972}.
+This transformation is a common technique in computer vision, used to identify common geometric shapes like lines and circles.
+Essentially, the algorithm draws many lines (of whatever desired geometry) through each data-point and checks whether these lines intersect with points in the dataset.
+Locations in parameter space that generate many intersections then become bright spots, allowing us to filter away points that are not close to these points.
+These algorithms remove a large amount of the unstructured noise and are computationally rather cheap.
+
+We illustrate two filtered events in figure \ref{fig:samples_filtered}. These are the same events as shown in figure \ref{fig:samples}, but with the Houghes' and nearest neighbours filtering applied.
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=0.9\textwidth, height=9cm]{plots/display_eventsclean_.pdf}
+\caption[Displaying filtered events in 2D and 3D]{Two- and three-dimensional representations of two events from the ${}^{46}$Ar experiment.
+Each row is one event in two projections, where the lightness of each point indicates higher charge values.
+These events have been filtered with a nearest neighbors algorithm and an algorithm based on the Hough transform, described in section \ref{sec:filtered}}\label{fig:samples_filtered}
+\end{figure}
+
+\section{Machine Learning Methods}
 
@@
 
-Another way of representing the clusters is by having multiple latent spaces representing the underlying manifolds that describe each class.
+\todo{re-word the intro-sentences} Another way of representing the clusters is by having multiple latent spaces representing the underlying manifolds that describe each class.
 This is the central idea in the Mixture of autoencoders (MIXAE) algorithm, introduced by  \cite{Zhang}.

### 2020-04-08 doc/Manuscript/article.tex
Updates from Overleaf
Continued on experiment and methods

@@
 This allows the AT-TPC to measure the excitation function over a wide range of energies from a single beam energy.
+
+The detector was placed inside the bore of a MRI solenoid energized to about 2 Tesla.
+This axial magnetic field served the purpose of bending the trajectories of the recoil particles in order to i) increase their length and ii) provide a measurement of their bending radius, directly related to their magnetic rigidity.
+Because the recoil particles travel in gas, they slow down and eventually stop, therefore their trajectories are described by 3-dimensional spirals (see \cite{Bradt2017}).
+One of the difficulties encountered in the analysis is that the shape of these spirals does not have any analytical form because it follows the energy-loss profile of the particles.
+It therefore needs to be simulated via an integration, which is numerically costly.
+Other difficulties are related to several experimental effects that deteriorate the quality of the data, namely saturation and cross-talk effects, as well as random noise.
+
+The method used in \cite{Bradt2018} to analyze the data followed a 3-phase sequence: cleaning, filtering and fitting.
+Traditional methods were used to perform each of these tasks, and ultimately extract the scientific information, but there were severe limitations and high computational costs that would have become prohibitive had the data set been larger by an order of magnitude.
+\todo{Should I give some details about how the 3 tasks were performed?}
+
 \section{Methods}\label{sec:methods}
+
+\subsection{Classifying  events} 
+\todo{Brute force serving as motivation for why we do unsupervised learning.}
+The filtering task relied on a set of $\chi^2$ distributions based on different criteria, however there was no easy way to quantify the effectiveness of the event selection from the thresholds applied to these distributions.
+In addition, the filtering task produced a binary result only, ``good" or ``bad", while the data clearly contained more classes of events, if only because of the presence of carbon in the target (from the isobutane gas).
+
 \subsection{Why machine learning}

### 2020-04-09 doc/Manuscript/article.tex
More material in chapter 3.

@@
 The filtering task relied on a set of $\chi^2$ distributions based on different criteria, however there was no easy way to quantify the effectiveness of the event selection from the thresholds applied to these distributions.
-In addition, the filtering task produced a binary result only, ``good" or ``bad", while the data clearly contained more classes of events, if only because of the presence of carbon in the target (from the isobutane gas).
+In addition, the filtering task produced a binary result only, "good" or "bad", while the data clearly contained more classes of events, if only because of the presence of carbon in the target (from the isobutane gas).
+In a broader perspective, an unsupervised classification algorithm would offer the possibility to "discover" rare events not anticipated (or overlooked) by the physicist.
+From a practical point of view, compared to supervised learning, it also avoids the necessary labeling task of the learning set events, which is error prone and time consuming.
 
@@
 \end{itemize}
+
+The $\chi^2$ approach used in the traditional analysis performed on the $^{46}$Ar data is very expensive computationally, because it involves the simulation of thousands of tracks for each recorded event, that need to be performed during the Monte-Carlo fitting sequence.
+Even though the reaction of interest in the $^{46}$Ar experiment had the largest cross section (elastic scattering), the sheer number of events to analyse (in the hundreds of thousands) necessitated the use of parallel processing on a high performance computer cluster.
+In the case of an experiment where the reaction of interest would represent less than a few percents of the total cross section, this procedure would become highly inefficient and prohibitive.
+Most of the CPU time would be spent on simulating events that are not of interest.
+A ML algorithm able to classify the data without a priori knowledge of the different types of events would increase the efficiency of the analysis tremendously, by concentrating the fitting efforts on events of interest only.
 

### 2020-04-10 doc/Manuscript/article.tex
reworded sections on pre-trained models and deep clustering.

@@
 
-Following the previous work of \citet{Kuchera2019}, we will consider representations of our events through the lens of a pre-trained network.
+To affect the clustering of AT-TPC events, we first considered the results from \citet{Kuchera2019}.
+In their work, the authors build a very successful classification approach using a pre-trained neural network to preprocess the events.
+From their results, it is clear that the pre-trained network captures salient information about the difference in event types.
+Our approach builds on previous work by considering the output from the pre-trained model as a representation on which we perform clustering with a simple K-means algorithm.
+
 In the Machine Learning community, it is not uncommon to publish packaged models with fitted parameters from image recognition contests.
 These models are trained on datasets with millions of images and classify between hundreds of distinct classes; one such is the imagenet dataset.
-In their work, \citet{Kuchera2019} use the VGG16 architecture trained on imagenet to classify AT-TPC events very successfully.
-We will build on the understanding of using these pre-trained networks in event classification by using VGG16 as an element in the end-to-end training of autoencoders.
+In their work, \citet{Kuchera2019} used the VGG16 architecture trained on imagenet to classify AT-TPC.
+Following on their work we will be applying the same VGG16 model to the clustering task.
 
-The VGG16 network is one of six analogous networks proposed by \citet{Simonyan2014}, and they were runners up in the ISLVRC(ImageNet large scale visual recognition competition) of 2014 \cite{Russakovsky2015}.
-The network architectures are fairly simple; for VGG16, there are sixteen layers in the network.
+The VGG16 network is one of six analogous networks proposed by \citet{Simonyan2014}.
+They were runners up in the ISLVRC(ImageNet large scale visual recognition competition) of 2014 \cite{Russakovsky2015}, narrowly being beat by GooLeNet.
+The network architectures are fairly straightforward; for VGG16, there are sixteen layers in the network.
 The first thirteen of which are convolutional layers with exclusively $3 \times 3$ kernels.
@@
 
-\todo{re-word the intro-sentences} Another way of representing the clusters is by having multiple latent spaces representing the underlying manifolds that describe each class.
-This is the central idea in the Mixture of autoencoders (MIXAE) algorithm, introduced by  \cite{Zhang}.
-To ensure that each autoencoder represents a cluster, we attach a soft-max classifier to the set of latent samples.
-Recall from equation \ref{eq:softmax} that the soft-max function transforms a network output to a prediction of relative confidences.
-This classifier assigns a cluster to each autoencoder, coupling the latent space and the reconstructions.
-The soft-max classifier is trained to output cluster probabilities and penalized for collapsing to assigning one cluster only.
-The remaining task is then to connect the cluster assignments to the autoencoder reconstructions.
-This is achieved by multiplying by the cluster confidence with the reconstruction error of each autoencoder.
+As an alternative to using a pre-trained model in conjunction with a clustering algorithm, we have also considered bespoke end-to-end clustering models.
+These models are optimized specifically to the AT-TPC data with the objective of clustering.
+For this paper, we considered the deep convolutional embedded clustering (DCEC) by \citet{Guo2017} as well as the mixture of autoencoders (MIXAE) method introduced by \citet{Zhang}.
+While we were able to reproduce the authors' results on their data, the DCEC algorithm proved unable to cluster AT-TPC in our implementation.
+However, this provides valuable insight as the seeds of the clusters are constructed by a K-means algorithm.
+This insight contrasts with our positive results from applying a pre-trained model with k-means and highlights potentially significant differences in models trained on a supervised or unsupervised objective for clustering tasks in nuclear physics.
+
+The MIXAE algorithm clusters data by pairing a set of encoder-decoder neural networks (autoencoder) with an auxiliary assignment network: each autoencoder constructs a representation of a given event while the assignment network assigns a set of cluster probabilities to the event.
+As noted by \citet{Zhang}, the MIXAe model assumes that the clusters are roughly similar in volume in the data and that one has some reasonable estimate on the number of clusters present.
+The architecture is portrayed in figure \ref{fig:mixae}, tapered boxes denote a direction of compression in the network components.
+
+As the MIXAE model does not have access to ground truth labels the training consists of applying simple heuristics that constitute a necessary, but not sufficient, fulfilment for a high-quality MIXAE model given the assumptions made of the model.
+The heuristics can be stated as:
+\begin{enumerate}
+	\item  Given a clustering assignment to a particular autoencoders' representation that autoencoders' reconstruction should be similar to the input event. 
+	\item Within a batch of events presented to the model, there should be a spread in the clustering assignment between all clusters. 
+	\item Each clustering prediction should be as close to a strong prediction as possible, i.e. assigning high probabilities is preferable to weak assignments.
+\end{enumerate}
+
+\noindent For a more formal consideration of these heuristics see \citet{Zhang}.
 
@@
 	\includegraphics[width=\textwidth]{plots/mixae.pdf}
-	\caption[Deep convolutional embedded clustering schematic]{Schematic of a DCEC model.
+	\caption[Mixture of autoencoders schematic]{Schematic of a MIXAE model.
 A sample $\hat{\boldsymbol{x}}$ is compressed to set of lower-dimensional representations $\{\boldsymbol{z}^{(i)}\}$ by $N$ autoencoders.
@@
 \end{figure}
-
-More formally let $\{\boldsymbol{z}_j^{(i)}\}^N$ be the set of $N$ latent samples from each of the $N$ auto-encoders from a single sample image $\hat{\boldsymbol{x}}^{(i)} \in \R ^{t\times v}$ with height $t$ and width $v$.
-Furthermore, let the soft cluster assignments be given as $\{p_j^{(i)}\}^N$ and the reconstructed samples be given as $\{\boldsymbol{x}_j^{(i)}\}$.
-The reconstruction loss is then the sum over each autoencoder multiplied with each cluster assignment, i.e.
-
-\begin{equation}\label{eq:mixae_reconst}
-\mathcal{L}_x = \sum_j p_j^{(i)} \mathcal{C}(\boldsymbol{x}_j^{(i)}, \hat{\boldsymbol{x}}^{(i)}),
-\end{equation}
-
-\noindent where $\mathcal{C}(\cdot)$ is a cost function like the mean squared error or a cross-entropy.
-
-We ensure that the soft cluster assignments encourage clustering by adding two terms to the total loss.
-The first is a simple entropy term which, when minimized, encourages the assignments to be one-hot vectors.
-Mathematically this loss can then be written as the entropy $S(\cdot)$ of a cluster assignment $\boldsymbol{p}$
-
-\begin{equation}\label{eq:mixae_sample}
-S(\boldsymbol{p}^{(i)})_{\text{sample}} = -\sum_j p_j^{(i)} \log p_j^{(i)}. 
-\end{equation}
-
-\noindent The last ingredient in the loss is then the term which discourages the trivial solution where only one cluster is being assigned.
-This loss is a batch-wise entropy term, given a mini-batch of $\beta$ samples the loss is computed as
-
-\begin{equation}\label{eq:mixae_batch}
-\begin{split}
-S(\{\boldsymbol{p}^{(i)}\}_i)_{\text{batch}}&= \left(-\sum_j \bar{p}_j \log \bar{p}_j \right)^{-1},\\
-\bar{\boldsymbol{p}} &= \frac{1}{\beta} \sum_i ^\beta \boldsymbol{p}^{(i)} .
-\end{split}
-\end{equation}
-
-\noindent The negative exponent in equation \ref{eq:mixae_batch} owes to the fact that we want to maximize the batch-wise entropy.
-
-From equations \ref{eq:mixae_reconst}, \ref{eq:mixae_sample}, and \ref{eq:mixae_batch} we can then compose the full loss for the MIXAE model.
-\cite{Zhang} note that the optimization depends heavily on the weighting of the different terms and so we introduce the weighting hyperparameters $\alpha$, $\gamma$ and $\theta$, and write the total loss over a mini-batch as
-
-\begin{equation}\label{eq:mixae_loss}
-\begin{split}
-\mathcal{L}_{\text{total}}(\{\hat{\boldsymbol{x}}^{(i)}\}^\beta_i) = \frac{1}{\beta}\sum_i \Big( &\frac{t\cdot v}{\theta} \mathcal{L}_x(\boldsymbol{p}^{(i)},\,\boldsymbol{x}^{(i)},\, \hat{\boldsymbol{x}}^{(i)} ) \\
-&+ \alpha S(\boldsymbol{p}^{(i)})_{\text{sample}} \Big) \\
-&+\gamma S(\{\boldsymbol{p}^{(i)}\}_i)_{\text{batch}}.
-\end{split}
-\end{equation}
-
-\noindent \citet{Zhang} note that this algorithm has a couple of shortcomings.
-The batch-entropy term encourages the cluster assignments to be uniformly distributed within the classes.
-However, it does have a second minimum when the assignments are all equally likely.
-For biased datasets with significant variations in the number of samples in each class, this can create problems for the algorithm.
-Additionally, the number of clusters are not dynamically computed by the algorithm but hard-coded by the researchers.
-This can be problematic if the number of cluster are not known prior to analysis.
 

### 2020-04-10 doc/Manuscript/article.tex
Updates from Overleaf

--- 2020-04-08 doc/Manuscript/article.tex
@@
 
-The detector was placed inside the bore of a MRI solenoid energized to about 2 Tesla.
-This axial magnetic field served the purpose of bending the trajectories of the recoil particles in order to i) increase their length and ii) provide a measurement of their bending radius, directly related to their magnetic rigidity.
-Because the recoil particles travel in gas, they slow down and eventually stop, therefore their trajectories are described by 3-dimensional spirals (see \cite{Bradt2017}).
-One of the difficulties encountered in the analysis is that the shape of these spirals does not have any analytical form because it follows the energy-loss profile of the particles.
-It therefore needs to be simulated via an integration, which is numerically costly.
-Other difficulties are related to several experimental effects that deteriorate the quality of the data, namely saturation and cross-talk effects, as well as random noise.
+\subsection{Standard approaches to analyze the data} 
+Brute force serving as motivation for why we do unsupervised learning.
+Supervised learning as well.
 
-The method used in \cite{Bradt2018} to analyze the data followed a 3-phase sequence: cleaning, filtering and fitting.
-Traditional methods were used to perform each of these tasks, and ultimately extract the scientific information, but there were severe limitations and high computational costs that would have become prohibitive had the data set been larger by an order of magnitude.
-\todo{Should I give some details about how the 3 tasks were performed?}
-
-\section{Methods}\label{sec:methods}
-
-\subsection{Classifying  events} 
-\todo{Brute force serving as motivation for why we do unsupervised learning.}
-The filtering task relied on a set of $\chi^2$ distributions based on different criteria, however there was no easy way to quantify the effectiveness of the event selection from the thresholds applied to these distributions.
-In addition, the filtering task produced a binary result only, ``good" or ``bad", while the data clearly contained more classes of events, if only because of the presence of carbon in the target (from the isobutane gas).
+\section{From supervised approaches to unsupervised methods to classify events}\label{sec:methods}
 

### 2020-04-17 doc/Manuscript/article.tex
context and motivation on pre-trained  + paragraph on clustering on latent

--- 2020-04-10 doc/Manuscript/article.tex
@@
 \subsection{Data}
+\todo{equalize to square plots - shorten down (maybe look at 3d)}
 In this section we give a brief overview of the data, for a more in-depth consideration we refer to \cite{Mittig2015}, \cite{Suzuki2012} and  \cite{Bradt2017a}.
@@
 
-\subsection{Data processing}
+\section{Data processing}
 
@@
 
-\subsubsection{Pre-trained neural networks}
+\subsection{Pre-trained neural networks}
 
 To affect the clustering of AT-TPC events, we first considered the results from \citet{Kuchera2019}.
-In their work, the authors build a very successful classification approach using a pre-trained neural network to preprocess the events.
-From their results, it is clear that the pre-trained network captures salient information about the difference in event types.
-Our approach builds on previous work by considering the output from the pre-trained model as a representation on which we perform clustering with a simple K-means algorithm.
+In their work, the authors consider the segmentation of reactions in the $^{46}$Ar experiment as a classification problem.
+Given ground-truth labels, they apply convolutional neural networks (CNN) to the problem.
+Using a CNN was in large part inspired by results from high-energy physics and, in particular, the work by \citet{Aurisano2016}.
+
+Using a CNN algorithm comes with the caveat that the data should be image-like.
+This caveat is especially stringent when applying pre-trained models to new data.
+For the $^{46}$Ar data we project the 3D point clouds in each event to a 2D matrix representation of the X-Y plane, summing over the Z-axis.
+Applying a pre-trained model is a common technique in computer science, and \citet{Kuchera2019} showed that a pre-trained CNN model outperforms bespoke models for the classification of $^{46}$Ar data.
+We build on the classification result in this work by using a pre-trained CNN model as a spring-board for our clustering models.
 
@@
 These models are trained on datasets with millions of images and classify between hundreds of distinct classes; one such is the imagenet dataset.
-In their work, \citet{Kuchera2019} used the VGG16 architecture trained on imagenet to classify AT-TPC.
-Following on their work we will be applying the same VGG16 model to the clustering task.
+In their work, \citet{Kuchera2019} used the VGG16 architecture trained on imagenet to classify ${}^{46}$Ar data.
+In our work we also found the VGG16 architecture to capture the strongest event-separating representation for clustering purposes.
 
 The VGG16 network is one of six analogous networks proposed by \citet{Simonyan2014}.
-They were runners up in the ISLVRC(ImageNet large scale visual recognition competition) of 2014 \cite{Russakovsky2015}, narrowly being beat by GooLeNet.
 The network architectures are fairly straightforward; for VGG16, there are sixteen layers in the network.
 The first thirteen of which are convolutional layers with exclusively $3 \times 3$ kernels.
-The choice of the kernel size is based on the fact that a stacked $3 \times 3$ kernel is equivalent to larger kernels in terms of the receptive field of the output.
-Three $3 \times 3$ kernels with stride $1$ have a $7 \times 7$ receptive field, but the larger kernel has $81\%$ more parameters and only one non-linearity \cite{Simonyan2014}.
-Stacking the smaller kernels then contributes to a lower computational cost.
-Additionally, there is a regularizing effect from the lowered number of parameters and increased explanatory power from the additional non-linearities.
-The full architecture is detailed in appendix \ref{tab:vgg}.
+The full architecture is detailed in appendix \ref{app:vgg}.
 
-\subsubsection{Deep clustering: Mixture of autoencoders}\label{sec:mixae}
+\subsubsection{Clustering on latent spaces}
+
+In contrast with the classification work by \citet{Kuchera2019}, we do not append a classifying layer to the VGG16 network given that we do not have access to the ground truth labels at training time in the clustering regime.
+We instead consider the output from the convolutional blocks of the VGG16 network as a representation of the events on which we will apply a clustering algorithm.
+
+The clustering algorithm we found to give the strongest result is the K-means algorithm.
+This algorithm is contingent on having a distance measurement between two datapoints.
+We found that the standard choice of a euclidian norm on the vector differences provided the best outcomes.
+
+\subsection{Deep clustering: Mixture of autoencoders}\label{sec:mixae}
 
@@
 \subsection{Autoencoder clustering}
+\todo{look at grayscale images or colorbar with propto charge}
 
@@
 \appendix  
-\section{Neural network architectures}
-
+\section{VGG16}\label{app:vgg}
+The choice of the kernel size is based on the fact that a stacked $3 \times 3$ kernel is equivalent to larger kernels in terms of the receptive field of the output.
+Three $3 \times 3$ kernels with stride $1$ have a $7 \times 7$ receptive field, but the larger kernel has $81\%$ more parameters and only one non-linearity \cite{Simonyan2014}.
+Stacking the smaller kernels then contributes to a lower computational cost.
+Additionally, there is a regularizing effect from the lowered number of parameters and increased explanatory power from the additional non-linearities.
 \begin{table}

### 2020-05-08 doc/Manuscript/article.tex
changes to data intro section

@@
 \subsection{Data}
-\todo{equalize to square plots - shorten down (maybe look at 3d)}
-In this section we give a brief overview of the data, for a more in-depth consideration we refer to \cite{Mittig2015}, \cite{Suzuki2012} and  \cite{Bradt2017a}.
-
-In this thesis we will explore the machine learning algorithms described in chapters \ref{ch:ml} and \ref{ch:autoencoder} on three datasets from the AT-TPC, listed in table \ref{tab:datasets}.
-The simulated data will serve as a baseline for performance, while the filtered and full data are real records from the ${}^{46}$Ar experiment.
-Each of these datasets have different distributions of the constituent classes.
-We list these differences in table \ref{tab:class_distr}.
-The filtered data differs from the full dataset as it has some post-processing applied to remove noise.
-This is not the case for the full data.
-The details of the post-processing are given in later sections.
-First, we consider the pipeline from raw data to images that the algorithms described in chapter \ref{ch:autoencoder} can process.
 
@@
 
-Our data processing pipeline begins with localized point-cloud data in the two-dimensional detector coordinate system, with one time dimension and a corresponding charge measurement.
-The charge and time measurement are extracted as the peak of this signal over the event, resulting in a maximum of one measurement per pad in the sensor plane.
-The events range from fairly sparse $< 10^2$ records to being very populated, depending on where in the volume the reaction occurs as well as other noise-generating factors.
-We centre the charge data to values $>1$ by adding the lowest occurring record in the run and apply a log-transform to get values in $\R^+$.
-Subsequently, we scale by the maximum value in the run to get charge data in the interval $[0, 1]$.
-Lastly, the transformed charge data is saved in a two-dimensional $128px\, \times \,128px$ array using the \lstinline{matplotlib} package provided in the \lstinline{Python} programing language \cite{matplotlib}.
-The choice of scaling was made to accommodate a binary cross-entropy loss on a 2D projection, as it presumes the true values to be bounded as probabilities.
+In this section we give a brief overview of the data, for a more in-depth consideration we refer to \cite{Mittig2015}, \cite{Suzuki2012} and  \cite{Bradt2017a}.
 
-We will begin by considering the simulated events, followed by subsequent considerations of the full and filtered experimental data.
+The AT-TPC data we studied for this work was originally recorded as charge time-series for each pad on the micromegas.
+In this representation an event is a record of 512 time-buckets for each of the $\sim10^4$ detector pads; however, in our analysis, we represent each event as a 2D slice.
+We chose to represent the data in 2D to facilitate the use of sophisticated image-recognition machine learning models.
+The 2D representation of events was computed by viewing the event as a point cloud of peaks in the charge time-series and summing over the time-axis.
+Additionally, we log-transform and normalize the data to  the interval $[0, 1]$ for each point.
+The final representation for each event is a $128 \times 128$ matrix with each element ranging from $0$ to $1$.
+
+One of the significant considerations for the analysis of AT-TPC data is to inject machine learning methods for track identification at the correct point in the analysis pipeline.
+Using raw data is advantageous as it provides an unbiased view of the event, but the data volume and noise levels might be prohibitive for analysis.
+On the opposite end of the pipeline, we will incrementally add bias to the analysis the more steps we take before applying the algorithm.
+With the benefit being that more pre-processing improves the signal-to-noise ratio, possibly improving model performance.
+To explore this trade-off between model performance and pre-processing bias, we performed analysis on simulated, cleaned and raw events.
 
@@
 Using the same parameters as for the ${}^{46}$Ar$(p, p)$ experiment, a small set of $N=4000$ events were generated per class, as well as a larger set of $N=40000$ events per class.
-The events are generated as point-clouds, consisting of position data on the x-y plane, a time-stamp and an associated charge value.
-These point-clouds are transformed into pure x-y projections with charge intensity for the analysis in this thesis using the pipeline described in the previous section.
-
-More formally the events are originally composed of peak-only 4-tuples of $e_i = (x_i, y_i, t_i, c_i)$.
-The peak-only designation indicates that we use the recorded peak amplitude on each pad, the tuples then correspond to pads that recorded a signal for that event.
-Each event is then a set of these four-tuples: $\epsilon_j = \{e_i\}$ creating a track in three-dimensional space.
 
@@
 We found that the standard choice of a euclidian norm on the vector differences provided the best outcomes.
+https://www.overleaf.com/project/5e73f990fcbd310001959b01
 

### 2020-05-08 doc/Manuscript/article.tex
Updates from Overleaf

--- 2020-04-17 doc/Manuscript/article.tex
@@
 \section{Introduction}\label{sec:intro}
-\begin{itemize}
-    \item Overall goals for the AT-TPC
-    \item Nuclear physics motivation for clustering-based (unsupervised) methods
-    \item Motivation from ML approach
+%\begin{itemize}
+%    \item Overall goals for the AT-TPC
+%   \item Nuclear physics motivation for clustering-based (unsupervised) methods
+%   \item Motivation from ML approach
 
-\end{itemize}
-The AT-TPC \cite{Bradt2018} is a novel type of detector designed specifically for nuclear physics experiments where the energies of the recoiling particles are very low compared to the energy required to escape target material \cite{}.
+%\end{itemize}
+The AT-TPC \cite{Bradt2018} is a novel type of detector designed specifically for nuclear physics experiments where the energies of the recoiling particles are very low compared to the energy required to escape target material.
 The luminosity of nuclear physics experiments performed with fixed targets is directly proportional to the amount of material encountered by the beam.
-On the other hand, for several classes of experiments the detection of recoil particles is paramount, therefore limiting the target thickness.
+On the other hand, for several classes of experiments, the detection of recoil particles is paramount, therefore limiting the target thickness.
 In addition, the properties of the recoil particles are modified while traversing the target material, affecting the resolutions that can be achieved.
@@
 
-The concept of active target aims at mitigating this compromise, by turning the target itself into a detector \cite{Beceiro2018}.
-Most active target detectors such as the AT-TPC are composed of a time projection chamber (TPC) where the detector gas is at the same time the target material.
+The concept of active target detectors aims at mitigating this compromise, by turning the target itself into a detector \cite{Beceiro2018}.
+Most active target detectors such as the AT-TPC are composed of a time projection chamber (TPC) where the detector material is at the same time the target material.
 Recoil particles that originate from a nuclear reaction between a beam nucleus and a gas nucleus can be tracked from the vertex of the reaction to their final position inside the active volume of the target.
 Their properties can therefore be measured without any loss of resolution regardless of the amount of material traversed by the beam.
-At the same time, the detection efficiency is dramatically increased by the very large solid angle covered in this geometry.
+At the same time, the detection efficiency is dramatically increased by the nearly $4\pi$ angle covered in this geometry.
 A direct consequence of this concept is the inclusiveness of the experimental data recorded by this type of detector: any nuclear reaction happening within the target is recorded.
-Although this sounds like an advantage from the scientific point of view, it poses great challenges during the analysis phase, that are reminiscent of bubble chamber times and on par with nowadays event classification issues in particle physics.
+Although this sounds like an advantage from the scientific point of view, it poses great challenges during the analysis phase, that are reminiscent of bubble chamber times and on par with event classification issues in today's particle physics experiments.
 More often than not, the reaction channel of interest has one of the lowest cross sections, therefore the physicist is faced with the task of sorting out the corresponding events from the ``background" of other reaction channels.
 
-Because TPCs produce 3-dimensional images of charged particle tracks, the event classification task is often akin to a visual inspection (bubble chamber times), which is not practical nowadays because of the large quantities of data.
+Because TPCs produce 3-dimensional images of charged particle tracks, the event classification task is often akin to a visual inspection (bubble chamber times), which is not tractable nowadays due tof the large quantities of data.
 Machine learning techniques then appear as a promising prospect, in particular in the image recognition domain where much progress has been made recently.
 In addition, some ML algorithms offer new possibilities such as the potential discovery of unforeseen phenomena that would have been missed by more traditional analysis methods.
-The classification and clustering of events based on various ML algorithms is hereby examined, using experimental data recorded by the AT-TPC during its commissioning experiment from a radioactive $^{46}$Ar beam reacting of an isobutane target composed of proton and carbon nuclei.
+In some cases, a set of labelled data can be utilized, which allows for the application of supervised learning methods, as previously demonstrated for the AT-TPC \cite{Kuchera2019}.
+In the cases where this in not feasible, unsupervised clustering methods must be applied.
+Here, clustering of events based on various ML algorithms are examined, using experimental data recorded by the AT-TPC during its commissioning experiment from a radioactive $^{46}$Ar beam reacting of an isobutane target composed of proton and carbon nuclei.
 
+%%% MPK here
 \section{Experimental Details} 
@@
 \subsection{Data}
-\todo{equalize to square plots - shorten down (maybe look at 3d)}
  In this section we give a brief overview of the data, for a more in-depth consideration we refer to \cite{Mittig2015}, \cite{Suzuki2012} and  \cite{Bradt2017a}.
@@
 
-\section{Data processing}
+\subsection{Data processing}
 
@@
 To affect the clustering of AT-TPC events, we first considered the results from \citet{Kuchera2019}.
-In their work, the authors consider the segmentation of reactions in the $^{46}$Ar experiment as a classification problem.
-Given ground-truth labels, they apply convolutional neural networks (CNN) to the problem.
-Using a CNN was in large part inspired by results from high-energy physics and, in particular, the work by \citet{Aurisano2016}.
+In their work, the authors build a very successful classification approach using a pre-trained neural network to preprocess the events.
+From their results, it is clear that the pre-trained network captures salient information about the difference in event types.
+Our approach builds on previous work by considering the output from the pre-trained model as a representation on which we perform clustering with a simple K-means algorithm.
 
-Using a CNN algorithm comes with the caveat that the data should be image-like.
-This caveat is especially stringent when applying pre-trained models to new data.
-For the $^{46}$Ar data we project the 3D point clouds in each event to a 2D matrix representation of the X-Y plane, summing over the Z-axis.
-Applying a pre-trained model is a common technique in computer science, and \citet{Kuchera2019} showed that a pre-trained CNN model outperforms bespoke models for the classification of $^{46}$Ar data.
-We build on the classification result in this work by using a pre-trained CNN model as a spring-board for our clustering models.
-
-In the Machine Learning community, it is not uncommon to publish packaged models with fitted parameters from image recognition contests.
+In the machine learning community, it is not uncommon to publish packaged models with fitted parameters from image recognition contests.
 These models are trained on datasets with millions of images and classify between hundreds of distinct classes; one such is the imagenet dataset.
-In their work, \citet{Kuchera2019} used the VGG16 architecture trained on imagenet to classify ${}^{46}$Ar data.
-In our work we also found the VGG16 architecture to capture the strongest event-separating representation for clustering purposes.
+In their work, \citet{Kuchera2019} used the VGG16 architecture trained on imagenet to classify AT-TPC.
+Following on their work we will be applying the same VGG16 model to the clustering task.
 
@@
 The first thirteen of which are convolutional layers with exclusively $3 \times 3$ kernels.
-The full architecture is detailed in appendix \ref{app:vgg}.
-
-\subsubsection{Clustering on latent spaces}
-
-In contrast with the classification work by \citet{Kuchera2019}, we do not append a classifying layer to the VGG16 network given that we do not have access to the ground truth labels at training time in the clustering regime.
-We instead consider the output from the convolutional blocks of the VGG16 network as a representation of the events on which we will apply a clustering algorithm.
-
-The clustering algorithm we found to give the strongest result is the K-means algorithm.
-This algorithm is contingent on having a distance measurement between two datapoints.
-We found that the standard choice of a euclidian norm on the vector differences provided the best outcomes.
+The choice of the kernel size is based on the fact that a stacked $3 \times 3$ kernel is equivalent to larger kernels in terms of the receptive field of the output.
+Three $3 \times 3$ kernels with stride $1$ have a $7 \times 7$ receptive field, but the larger kernel has $81\%$ more parameters and only one non-linearity \cite{Simonyan2014}.
+Stacking the smaller kernels then contributes to a lower computational cost.
+Additionally, there is a regularizing effect from the lowered number of parameters and increased explanatory power from the additional non-linearities.
+The full architecture is detailed in appendix \ref{tab:vgg}.
 
@@
 \subsection{Autoencoder clustering}
-\todo{look at grayscale images or colorbar with propto charge}
 
@@
 \appendix  
-\section{VGG16}\label{app:vgg}
-The choice of the kernel size is based on the fact that a stacked $3 \times 3$ kernel is equivalent to larger kernels in terms of the receptive field of the output.
-Three $3 \times 3$ kernels with stride $1$ have a $7 \times 7$ receptive field, but the larger kernel has $81\%$ more parameters and only one non-linearity \cite{Simonyan2014}.
-Stacking the smaller kernels then contributes to a lower computational cost.
-Additionally, there is a regularizing effect from the lowered number of parameters and increased explanatory power from the additional non-linearities.
+\section{Neural network architectures}
+
 \begin{table}

### 2020-05-10 doc/Manuscript/article.tex
Updates from Overleaf

--- 2020-05-08 doc/Manuscript/article.tex
@@
+%robert was here sup robert -MIchelle was also here  
 \documentclass[preprint,12pt]{elsarticle}
@@
 
-\subsection{Full \texorpdfstring{${}^{46}$Ar}{46Ar}  events}\label{sec:data_real}
+\subsection{Raw \texorpdfstring{${}^{46}$Ar}{46Ar}  events}\label{sec:data_real}
 
 The events analyzed in this section were retrieved from the ${}^{46}$Ar resonant proton scattering experiment recorded with the AT-TPC. 
+While we denote these events as raw, it is important to not that what we mean is a raw 2D representation of peak-only events.
 
@@
 Part of the challenge for this data then comes from understanding of the physics of the major contributing factors to this noise.
+\todo{I guess this paragraph can go.
+It reads very thesis-y, but I'm not sure what to do here to sensibly frame the raw data}
 
 We display two different events from the ${}^{46}$Ar experiment in figure \ref{fig:samples}.
-The top row illustrates an event with a large fraction of noise, while the bottom row shows an event nearly devoid of noise.
-The very clear spiral structure of the bottom row indicates that this is a proton-event.
+The top row illustrates a carbon event with a large fraction of noise, while the bottom row shows a proton event nearly devoid of noise.
 
@@
 These algorithms remove a large amount of the unstructured noise and are computationally rather cheap.
+\todo{Also pretty thesis-y language.
+Do not like}
 

### 2020-05-15 doc/Manuscript/article.tex
intro edits

@@
 \section{Introduction}\label{sec:intro}
-\begin{itemize}
-    \item Overall goals for the AT-TPC
-    \item Nuclear physics motivation for clustering-based (unsupervised) methods
-    \item Motivation from ML approach
 
-\end{itemize}
-The AT-TPC \cite{Bradt2018} is a novel type of detector designed specifically for nuclear physics experiments where the energies of the recoiling particles are very low compared to the energy required to escape target material \cite{}.
+%
+%\begin{itemize}
+%    \item Overall goals for the AT-TPC
+%    \item Nuclear physics motivation for clustering-based (unsupervised) %methods
+%   \item Motivation from ML approach
+%   
+%\end{itemize}
+
+The AT-TPC \cite{Bradt2018} is a novel type of detector designed specifically for nuclear physics experiments where the energies of the recoiling particles are very low compared to the energy required to escape target material.
 The luminosity of nuclear physics experiments performed with fixed targets is directly proportional to the amount of material encountered by the beam.
@@
 
-The concept of active target aims at mitigating this compromise, by turning the target itself into a detector \cite{Beceiro2018}.
+The concept of active targets aim at mitigating this compromise, by turning the target itself into a detector \cite{Beceiro2018}.
 Most active target detectors such as the AT-TPC are composed of a time projection chamber (TPC) where the detector gas is at the same time the target material.
@@
 A direct consequence of this concept is the inclusiveness of the experimental data recorded by this type of detector: any nuclear reaction happening within the target is recorded.
-Although this sounds like an advantage from the scientific point of view, it poses great challenges during the analysis phase, that are reminiscent of bubble chamber times and on par with nowadays event classification issues in particle physics.
-More often than not, the reaction channel of interest has one of the lowest cross sections, therefore the physicist is faced with the task of sorting out the corresponding events from the ``background" of other reaction channels.
+Although this sounds like an advantage from the scientific point of view, it poses great challenges during the analysis phase, that are reminiscent of bubble chamber times and on par with event classification challenges in particle physics today.
+{\color{blue} cite something HEP here} More often than not, the reaction channel of interest has one of the lowest cross sections, therefore the physicist is faced with the task of sorting out the corresponding events from the ``background" of other reaction channels.
 
-Because TPCs produce 3-dimensional images of charged particle tracks, the event classification task is often akin to a visual inspection (bubble chamber times), which is not practical nowadays because of the large quantities of data.
+Because TPCs produce 3-dimensional images of charged particle tracks, the event identification task is often akin to a visual inspection (akin to analysis in the bubble chamber era), which is not practical nowadays because of the large quantities of data.
 Machine learning techniques then appear as a promising prospect, in particular in the image recognition domain where much progress has been made recently.
-In addition, some ML algorithms offer new possibilities such as the potential discovery of unforeseen phenomena that would have been missed by more traditional analysis methods.
+{\color{blue} short discussion on prior work in nuclear/HEP/neutrino} In addition, some ML algorithms offer new possibilities such as the potential discovery of unforeseen phenomena that would have been missed by more traditional analysis methods.
+Prior work has demonstrated the ability to apply supervised classification machine learning methods to AT-TPC data when a labeled training set is available, whether through hand-labeled data or labeled simulated data (a {\em transfer learning} application) \cite{Kuchera2019}.
+In some experiments, a labeled dataset is unavailable.
+This could be  due to the inability to hand-label events, or the case where one must know a priori the types and behaviors of the reactions present in the detector in order to generate a simulated dataset.
+In the latter case, there must also exist a validation dataset of real data, still requiring the ability to label a subset of the real data.
+
 The classification and clustering of events based on various ML algorithms is hereby examined, using experimental data recorded by the AT-TPC during its commissioning experiment from a radioactive $^{46}$Ar beam reacting of an isobutane target composed of proton and carbon nuclei.

### 2020-05-15 doc/Manuscript/article.tex
Updates from Overleaf

@@
 
-The AT-TPC \cite{Bradt2018} is a novel type of detector designed specifically for nuclear physics experiments where the energies of the recoiling particles are very low compared to the energy required to escape target material.
+The AT-TPC \cite{Bradt2017} is a novel type of detector designed specifically for nuclear physics experiments where the energies of the recoiling particles are very low compared to the energy required to escape target material.
 The luminosity of nuclear physics experiments performed with fixed targets is directly proportional to the amount of material encountered by the beam.
@@
 
+% motivating ML work:
 Because TPCs produce 3-dimensional images of charged particle tracks, the event identification task is often akin to a visual inspection (akin to analysis in the bubble chamber era), which is not practical nowadays because of the large quantities of data.
@@
 In the latter case, there must also exist a validation dataset of real data, still requiring the ability to label a subset of the real data.
-
-The classification and clustering of events based on various ML algorithms is hereby examined, using experimental data recorded by the AT-TPC during its commissioning experiment from a radioactive $^{46}$Ar beam reacting of an isobutane target composed of proton and carbon nuclei.
+The unsupervised separation of event types, or {\em clustering}, based on various ML algorithms is hereby examined, using experimental data recorded by the AT-TPC during its commissioning experiment from a radioactive $^{46}$Ar beam reacting on an isobutane target composed of proton and carbon nuclei.
 
@@
 The goal of the experiment was to measure the evolution of the elastic and inelastic cross sections between $^{46}$Ar and protons as a function of energy (also called excitation function), and observe resonances in the composite system $^{47}$K that correspond to analog states in the nucleus $^{47}$Ar.
-Spectroscopic information can then be obtained from the shape and amplitude of the observed resonances \cite{Bradt2018}.
+Spectroscopic information can then be obtained from the shape and amplitude of the observed resonances \cite{Bradt2018} {\color{blue} check references: first should be Commisioning...
+this should be physics paper}.
 The experiment was performed at the National Superconducting Cyclotron Facility (NSCL) where a $^{46}$Ar beam was produced via fragmentation of a $^{48}$Ca beam on a $^9$Be target at about 140 MeV/u.
 The $^{46}$Ar isotopes were then filtered, thermalized, and finally re-accelerated to 4.6 MeV/u by a linear accelerator.
-This scheme was used to produce a low emittance beam necessary to guarantee a good energy resolution in the excitation function.
-Because of the energy loss of the $^{46}$Ar beam particles as they traverse the target gas volume, the position of the reaction vertex along the beam axis is directly related to the energy at which the reaction occurs.
+This scheme was used to produce a low-emittance beam.
+which is necessary to guarantee a good energy resolution in the excitation function.
+Because the  $^{46}$Ar beam particles lose energy as they traverse the target gas volume, the position of the reaction vertex along the beam axis is directly related to the energy at which the reaction occurs.
 This allows the AT-TPC to measure the excitation function over a wide range of energies from a single beam energy.
@@
 Traditional methods were used to perform each of these tasks, and ultimately extract the scientific information, but there were severe limitations and high computational costs that would have become prohibitive had the data set been larger by an order of magnitude.
-\todo{Should I give some details about how the 3 tasks were performed?}
+In addition, \cite{Kuchera2019} demonstrated that the purity and statistics of the data would have been improved with the use of ML.
+\todo{Should I give some details about how the 3 tasks were performed?}\todo{I think this is done sufficiently below}
 
@@
 \todo{Brute force serving as motivation for why we do unsupervised learning.}
-The filtering task relied on a set of $\chi^2$ distributions based on different criteria, however there was no easy way to quantify the effectiveness of the event selection from the thresholds applied to these distributions.
-In addition, the filtering task produced a binary result only, "good" or "bad", while the data clearly contained more classes of events, if only because of the presence of carbon in the target (from the isobutane gas).
-In a broader perspective, an unsupervised classification algorithm would offer the possibility to "discover" rare events not anticipated (or overlooked) by the physicist.
+The filtering task relied on a set of $\chi^2$ distributions based on different criteria, however there was no well-defined method to quantify the effectiveness of the event selection from the thresholds applied to these distributions.
+In addition, the filtering task produced a binary result only, ``good" or``bad", while the data clearly contained more classes of events, if only because of the presence of carbon in the target (from the isobutane gas).
+In a broader perspective, an unsupervised classification algorithm would offer the possibility to ``discover" rare events not anticipated (or overlooked) by the physicist.
+These events would likely be filtered out using the traditional methods.
 From a practical point of view, compared to supervised learning, it also avoids the necessary labeling task of the learning set events, which is error prone and time consuming.
@@
         \item The broken tracks, noisy environment creates bad fit statistics for otherwise useful events. End result is $f1 \sim 0.7 $
+        \item define a ``bad" fit as a non-proton event
     \end{itemize}

### 2020-05-21 doc/Manuscript/article.tex
Updates from Overleaf

@@
 \centering
-\caption{Descriptions of number of events in the data used for analysis in this thesis.
-In principle we can simulate infinite data, but it is both quite simple and not very interesting outside a case for a proof-of-concept}\label{tab:datasets}
+\caption{Descriptions of number of events in the data.
+In principle we can simulate an arbitrary amount of data, but the following quantities were shown to be sufficient for these studies.}\label{tab:datasets}
 \begin{tabular}{lccc}

### 2020-05-22 doc/Manuscript/article.tex
Few edits and comments

@@
 The $^{46}$Ar isotopes were then filtered, thermalized, and finally re-accelerated to 4.6 MeV/u by a linear accelerator.
-This scheme was used to produce a low-emittance beam.
-which is necessary to guarantee a good energy resolution in the excitation function.
+This scheme was used to produce a low-emittance beam, which is necessary to guarantee a good energy resolution in the excitation function.
 Because the  $^{46}$Ar beam particles lose energy as they traverse the target gas volume, the position of the reaction vertex along the beam axis is directly related to the energy at which the reaction occurs.
@@
 \subsection{Data}
+{\color{green}What is the intent of this section? Display some data examples to illustrate the points above?}
 
@@
 
-The AT-TPC data we studied for this work was originally recorded as charge time-series for each pad on the micromegas.
+The AT-TPC data we studied for this work was originally recorded as charge time-series for each pad on the micromegas {\color{green} We haven't defined what micromegas is}.
 In this representation an event is a record of 512 time-buckets for each of the $\sim10^4$ detector pads; however, in our analysis, we represent each event as a 2D slice.
 We chose to represent the data in 2D to facilitate the use of sophisticated image-recognition machine learning models.
-The 2D representation of events was computed by viewing the event as a point cloud of peaks in the charge time-series and summing over the time-axis.
+The 2D representation of events was computed by viewing the event as a point cloud in the charge time-series and summing over the time-axis.
 Additionally, we log-transform and normalize the data to  the interval $[0, 1]$ for each point.
@@
 
-One of the significant considerations for the analysis of AT-TPC data is to inject machine learning methods for track identification at the correct point in the analysis pipeline.
+One of the significant considerations for the analysis of AT-TPC data is to inject machine learning methods for track identification at the best point in the analysis pipeline.
 Using raw data is advantageous as it provides an unbiased view of the event, but the data volume and noise levels might be prohibitive for analysis.
@@
 
-The simulated AT-TPC tracks were simulated with the \lstinline{pytpc} package developed by \citet{Bradt2017a}.
-Using the same parameters as for the ${}^{46}$Ar$(p, p)$ experiment, a small set of $N=4000$ events were generated per class, as well as a larger set of $N=40000$ events per class.
+AT-TPC tracks were simulated with the \lstinline{pytpc} package developed by \citet{Bradt2017a}.
+Using the same parameters as for the ${}^{46}$Ar$(p, p)$ experiment, a small set of $N=4000$ events per class was generated, as well as a larger set of $N=40000$ events per class.
 
 To emulate the real-data case, we set a subset of the simulated data to be labelled and treat the rest as unlabelled data.
-We chose this partition to be $15\%$ of each class.
-We denote this subset and its associated labels as $\gamma_L=(\boldsymbol{X}_L, \boldsymbol{y}_L)$, the entire dataset which we will denote as $\boldsymbol{X}_F$.
-To clarify, please note that $\boldsymbol{X}_L \subset \boldsymbol{X}_F$.
+We choose this partition to be $15\%$ of each class.
+We denote this subset and its associated labels as $\gamma_L=(\boldsymbol{X}_L, \boldsymbol{y}_L)$, while the entire dataset which is identified as $\boldsymbol{X}_F$.
+Note that $\boldsymbol{X}_L \subset \boldsymbol{X}_F$.
 
-We display two simulated events in figure \ref{fig:sim_samples}. The top row illustrates a proton-event, and the bottom a carbon-event.
+Two simulated events are displayed in figure \ref{fig:sim_samples}. The top row illustrates a proton-event, and the bottom a carbon-event.
 
@@
 The events analyzed in this section were retrieved from the ${}^{46}$Ar resonant proton scattering experiment recorded with the AT-TPC. 
-While we denote these events as raw, it is important to not that what we mean is a raw 2D representation of peak-only events.
+While we denote these events as raw, it is important to note that what we mean is a raw 2D representation of peak-only events.
 
@@
 We begin by exploring a naive K-means approach on the latent space of a pre-trained network.
-Subsequently, we investigate other clustering methods and two autoencoder based clustering algorithms, as outlined in section \ref{sec:deep_clustering}.
+Subsequently, we investigate other clustering methods and the autoencoder based clustering algorithm outlined in section \ref{sec:mixae}.
 
-\todo{Drop this argument or build it in the ML section?}
-This chapter builds on the previous results from semi-supervised classification.
-We observe that we can construct high-quality latent spaces.
-These high-quality spaces facilitate an investigation of clustering techniques.
-
-\todo{Reword this to not contrast against semi-supervised}
-The approach for clustering of events is different from the semi-supervised approach in two meaningful ways.
-First, it is a harder task, as we will demonstrate.
-The clustering task thus necessitates a more exploratory approach to the problem.
-Second, as a consequence of the challenge, the focus will be a bit different than for the semi-supervised approach.
-We will still utilize the same architectures and models starting with a search over the parameter space over which we measure the performance using the adjusted rand score (ARS) and accuracy defined in section \ref{sec:unsupervised_perf} and \ref{sec:supervised_perf}, respectively.
-
-As with chapter  \ref{chap:classification} where we explored the semi-supervised results, we begin this chapter by considering the VGG16 pre-trained model as a benchmark.
-
-Lastly, we note that the focus of this work is mainly on discovering possible avenues for further research.
-This focus requires a broad scan of possible avenues rather than a rigorous analysis of one specific model.
 \subsection{Transfer learning}
-As in chapter \ref{ch:architectures}, we also use the VGG16 pre-trained network as a baseline for the clustering performance.
+Building on the work by \citet{Kuchera2019}, we also use the VGG16 pre-trained network as a baseline for the clustering performance.
 We begin by considering a classical K-means approach to clustering.
-However, the output from the VGG16 network is very high dimensional with output vectors in $\R^{8192}$.
-\todo{reword} One of the primary concerns is then the curse of dimensionality, where the ratio of distances goes to one with increasing dimensionality as shown by \citet{Aggarwal}.
-However, one of the central caveats to the authors  finding is that the elements are uniformly distributed in the latent space.
-It is then possible that all the class information lies in some sub-space of the latent data.
-To investigate this, we perform clustering analysis using the full representation and the $10^2$ first principal components only.
+However, the output from the VGG16 network is very high dimensional which traditionally conficts with good K-means results.
+With output vectors in $\R^{8192}$ the ratio between $L_p$ distances goes to one with increasing dimensionality, as shown by \citet{Aggarwal}.
+However, one of the central assumptions in the authors' finding is that the elements are uniformly distributed in the latent space.
+They note that it is then possible that if the class information lies in some sub-space of the latent data, $L_p$-norms may provide useful clusterings even in high dimensional spaces.
+To investigate this, we perform a $L_2$-norm based K-means clustering using the full VGG16 representation and the $10^2$ first principal components only.
 
@@
 We begin by investigating the K-means clustering algorithm on the VGG16 latent space.
-As in chapter \ref{chap:classification} the VGG16 model is pre-trained on the imagenet dataset creating a set of vectors $\boldsymbol{x} \in \R^{8192}$.
+The VGG16 model is pre-trained on the imagenet dataset, and subsequently applied to the AT-TPC events without the classification layers.
+This creates a set of vectors $\boldsymbol{x} \in \R^{8192}$ as our new representation of the events.
 To cluster we use \lstinline{scikit-learn} implementation of the K-means algorithm, with default parameters \cite{Pedregosa2011}.
 The results of the clustering runs are included in table \ref{tab:clstr_vgg}.
-We observe that we are able to attain near-perfect clustering on simulated data and that there is a sharp decline in performance as we add noise by moving to the filtered and full datasets.
+We observe that we are able to attain near-perfect clustering on simulated data and that there is a sharp decline in performance as we add noise by moving to the filtered and raw datasets.
 
@@
 
-We repeat this analysis using a PCA dimensionality reduction on the latent space of the VGG16 model.
+We repeat this analysis using a PCA dimensionality reduction {\color{green}(please define)} on the latent space of the VGG16 model.
 This is done to estimate to what degree the class separating information is encoded in the entirety of the latent space, or in some select regions.
@@
 None of them provided any notable differences from the K-means results or were significantly worse.
-Notably, the DBSCAN algorithm failed to provide any useful clustering results.
+Notably, the DBSCAN algorithm {\color{green} (add reference)} failed to provide any useful clustering results.
 We find this important as one of the significant drawbacks of K-means, and the deep clustering algorithms presented in section \ref{sec:deep_clustering}, is that they are all dependent on pre-determining the number of clusters.

### 2020-05-25 doc/Manuscript/article.tex
MPK edits Section 1

@@
 \author{R.~Solli}
+\address{Expert Analytics AS, Tordenskiolds gate 6, 0160, Oslo, Norway}
 \address{Department of Physics, University of Oslo, POB 1048 Oslo, N-0316 Oslo, Norway}
@@
 In some experiments, a labeled dataset is unavailable.
-This could be  due to the inability to hand-label events, or the case where one must know a priori the types and behaviors of the reactions present in the detector in order to generate a simulated dataset.
+This could be  due to the inability to hand-label events, or the case where one does not know a priori the types and behaviors of the reactions present in the detector in order to generate a labeled, simulated dataset.
 In the latter case, there must also exist a validation dataset of real data, still requiring the ability to label a subset of the real data.
-The unsupervised separation of event types, or {\em clustering}, based on various ML algorithms is hereby examined, using experimental data recorded by the AT-TPC during its commissioning experiment from a radioactive $^{46}$Ar beam reacting on an isobutane target composed of proton and carbon nuclei.
+The unsupervised separation of event types, or {\em clustering}, based on a set of ML algorithms is hereby examined, using experimental data recorded by the AT-TPC during its commissioning experiment from a radioactive $^{46}$Ar beam reacting on an isobutane target composed of proton and carbon nuclei.
 
 \section{Experimental Details} 
-The goal of the experiment was to measure the evolution of the elastic and inelastic cross sections between $^{46}$Ar and protons as a function of energy (also called excitation function), and observe resonances in the composite system $^{47}$K that correspond to analog states in the nucleus $^{47}$Ar.
-Spectroscopic information can then be obtained from the shape and amplitude of the observed resonances \cite{Bradt2018} {\color{blue} check references: first should be Commisioning...
+The goal of the experiment was to measure the evolution of the elastic and inelastic cross sections between $^{46}$Ar and protons as a function of energy (the {\em excitation function}), and observe resonances in the composite system $^{47}$K that correspond to analog states in the nucleus $^{47}$Ar.
+Spectroscopic information can then be obtained from the shape and amplitude of the observed resonances \cite{Bradt2018} {\color{blue} check references: first should be Commissioning...
 this should be physics paper}.
@@
 
-The detector was placed inside the bore of a MRI solenoid energized to about 2 Tesla.
+The detector was placed inside the bore of a MRI solenoid energized to $\sim 2$ Tesla.
 This axial magnetic field served the purpose of bending the trajectories of the recoil particles in order to i) increase their length and ii) provide a measurement of their bending radius, directly related to their magnetic rigidity.
 Because the recoil particles travel in gas, they slow down and eventually stop, therefore their trajectories are described by 3-dimensional spirals (see \cite{Bradt2017}).
-One of the difficulties encountered in the analysis is that the shape of these spirals does not have any analytical form because it follows the energy-loss profile of the particles.
+One of the difficulties encountered in the analysis is that the shape of these spirals does not have an analytical form because it follows the energy-loss profile of the particles.
 It therefore needs to be simulated via an integration, which is numerically costly.
@@
 The method used in \cite{Bradt2018} to analyze the data followed a 3-phase sequence: cleaning, filtering and fitting.
-Traditional methods were used to perform each of these tasks, and ultimately extract the scientific information, but there were severe limitations and high computational costs that would have become prohibitive had the data set been larger by an order of magnitude.
-In addition, \cite{Kuchera2019} demonstrated that the purity and statistics of the data would have been improved with the use of ML.
-\todo{Should I give some details about how the 3 tasks were performed?}\todo{I think this is done sufficiently below}
+Traditional methods were used to perform each of these tasks, and ultimately extract the scientific information, but there were severe limitations and high computational costs that become prohibitive in data sets larger by an order of magnitude.
+In addition, \cite{Kuchera2019} demonstrated that the purity and statistics of the data are improved with the use of ML.
 
@@
 {\color{green}What is the intent of this section? Display some data examples to illustrate the points above?}
+
+{\color{orange} Yes, but also this section is a necessary bridge for reproducibility. Maybe it's better as an appendix?}
 
@@
 
-We repeat this analysis using a PCA dimensionality reduction {\color{green}(please define)} on the latent space of the VGG16 model.
+We repeat this analysis using a Principle Component Analysis (PCA) dimensionality reduction \footnote{
+PCA is a common techinque to find the significant variations in data by projecting the data along a subset of its covariance matrix eigenvectors \cite{Marsland2009}
+}{\color{green}(please define)} {\color{orange} does this work?} on the latent space of the VGG16 model.
 This is done to estimate to what degree the class separating information is encoded in the entirety of the latent space, or in some select regions.
@@
 None of them provided any notable differences from the K-means results or were significantly worse.
-Notably, the DBSCAN algorithm {\color{green} (add reference)} failed to provide any useful clustering results.
-We find this important as one of the significant drawbacks of K-means, and the deep clustering algorithms presented in section \ref{sec:deep_clustering}, is that they are all dependent on pre-determining the number of clusters.
+Notably, the DBSCAN algorithm \cite{Ester96adensity-based}\cite{Bergstra2012} failed to provide any useful clustering results.
+We find this important as one of the significant drawbacks of K-means, and the deep clustering algorithm presented in section \ref{sec:mixae}, is that they are all dependent on pre-determining the number of clusters.
 This is not the case for DBSCAN.
@@
 In the previous section we demonstrated a potential for clustering of events from an AT-TPC experiment.
-To build on this result we will in this section explore the application of the mixture of autoencoders (MIXAE) algorithm introduced in section \ref{sec:deep_clustering}.
-The other algorithm introduced in section \ref{sec:deep_clustering}, deep clustering with convolutional autoencoders (DCEC), consistently collapsed to just one cluster for all datasets.
+To build on this result we will in this section explore the application of the mixture of autoencoders (MIXAE) algorithm introduced in section \ref{sec:mixae}.
+Other algorithms for deep clustering based on autoencoder architectures were also implemented.
+These did not succeed in the clustering task and were omitted for brevity\footnote{The algoritms we implemented were the Deep Embedded Convolutional Clustering \cite{Guo2017} and its predecessor Deep Embedded Clustering \cite{Xie2016}}.
 
@@
 In addition to those parameters we have the weighting of the loss terms: $\theta$, $\alpha$ and $\gamma$.
-These weighting parameters are attached to the reconstruction loss, sample entropy and batch-wise entropy respectively.
-\cite{Zhang} note that these parameters are very important for the model performance and so we focus primarily on these.
+These weighting parameters are attached to the reconstruction loss, sample entropy and batch-wise entropy respectively \cite{Zhang}.
+We focused on the tuning of the clustering hyper-parameters, and heuristically set the autoencoder hyper-parameters to a shallow $3\times3$ convolutional network.
 
-\todo{but why}We empirically freeze the convolutional autoencoder hyperparameters to compress the original $128 \times 128$ to a $8 \times 8$ pixel grid using four convolutional layers.
 The parameters chosen for the autoencoders are listed in full in table \ref{tab:mixe_ae_hyperparams}.
@@
 Activation & LReLu \\
-Latent type & None \\
 Latent dimension & 20  \\
-$\beta$ & N/A \\
 Batchnorm & False \\
@@
 
-Since there are then only three remaining hyperparameters we choose to perform a coarse grid-search as described in section \ref{sec:hyperparam_search_arch}.
+\noindent Since there are then only three remaining hyperparameters we choose to perform a coarse grid-search as described in section \ref{sec:hyperparam_search_arch}.
 

### 2020-05-25 doc/Manuscript/article.tex
Edits for Section 2

@@
 
-The $\chi^2$ approach used in the traditional analysis performed on the $^{46}$Ar data is very expensive computationally, because it involves the simulation of thousands of tracks for each recorded event, that need to be performed during the Monte-Carlo fitting sequence.
-Even though the reaction of interest in the $^{46}$Ar experiment had the largest cross section (elastic scattering), the sheer number of events to analyse (in the hundreds of thousands) necessitated the use of parallel processing on a high performance computer cluster.
-In the case of an experiment where the reaction of interest would represent less than a few percents of the total cross section, this procedure would become highly inefficient and prohibitive.
+The $\chi^2$ approach used in the traditional analysis performed on the $^{46}$Ar data is extremely computationally expensive because it involves the simulation of thousands of tracks for each recorded event, which are performed each iteration of the Monte-Carlo fitting sequence.
+Even though the reaction of interest in the $^{46}$Ar experiment had the largest cross section (elastic scattering), the sheer number of events to analyse (in the hundreds of thousands) necessitated the use of parallel processing on a high performance computing cluster.
+In the case of an experiment where the reaction of interest would represent less than a few percent of the total cross section, this procedure would become highly inefficient and prohibitive.
 Most of the CPU time would be spent on simulating events that are not of interest.
-A ML algorithm able to classify the data without a priori knowledge of the different types of events would increase the efficiency of the analysis tremendously, by concentrating the fitting efforts on events of interest only.
+An unsupervised ML algorithm able to separate the data without a priori knowledge of the different types of events would increase the efficiency of the analysis tremendously, by concentrating the fitting efforts only on events of interest.
+In addition, the clustering allows for more exploration of the data, potentially enabling new discovery of unexpected reaction types.
 

### 2020-05-25 doc/Manuscript/article.tex
Updates from Overleaf

@@
 The goal of the experiment was to measure the evolution of the elastic and inelastic cross sections between $^{46}$Ar and protons as a function of energy (the {\em excitation function}), and observe resonances in the composite system $^{47}$K that correspond to analog states in the nucleus $^{47}$Ar.
-Spectroscopic information can then be obtained from the shape and amplitude of the observed resonances \cite{Bradt2018} {\color{blue} check references: first should be Commissioning...
-this should be physics paper}.
+Spectroscopic information can then be obtained from the shape and amplitude of the observed resonances \cite{Bradt2018}.
 The experiment was performed at the National Superconducting Cyclotron Facility (NSCL) where a $^{46}$Ar beam was produced via fragmentation of a $^{48}$Ca beam on a $^9$Be target at about 140 MeV/u.
@@
 Most of the CPU time would be spent on simulating events that are not of interest.
-An unsupervised ML algorithm able to separate the data without a priori knowledge of the different types of events would increase the efficiency of the analysis tremendously, by concentrating the fitting efforts only on events of interest.
+An unsupervised ML algorithm able to separate the data without a priori knowledge of the different types of events increases the efficiency of the analysis tremendously, by concentrating the fitting efforts only on events of interest.
 In addition, the clustering allows for more exploration of the data, potentially enabling new discovery of unexpected reaction types.
@@
 
-In this section we give a brief overview of the data, for a more in-depth consideration we refer to \cite{Mittig2015}, \cite{Suzuki2012} and  \cite{Bradt2017a}.
+In this section we give a brief overview of the data, for a more in-depth consideration we refer the reader to \cite{Mittig2015}, \cite{Suzuki2012} and  \cite{Bradt2017a}.
 

### 2020-05-25 doc/Manuscript/article.tex
Updates from Overleaf

@@
 %robert was here sup robert -MIchelle was also here
-\documentclass[preprint,12pt]{elsarticle}
+
+% uncomment for submission to NIM A
+\documentclass[review,number,sort&compress]{elsarticle}
+
+%\documentclass[preprint,12pt]{elsarticle}
 \usepackage{amssymb}
@@
 \author{M.P.~Kuchera}
-\address{Physics Department, Davidson College, Davidson, North Carolina, USA}
+\address{Department of Physics, Davidson College, Davidson, North Carolina, USA}
+\author{R.R.~Strauss}
+\address{Department of Mathematics and Computer Science, Davidson College, Davidson, North Carolina, USA}
+
 \author{M.~Hjorth-Jensen}
@@
 
-\subsection{Data}
-{\color{green}What is the intent of this section? Display some data examples to illustrate the points above?}
+%\subsection{Data}
+%{\color{green}What is the intent of this section? Display some data examples to illustrate the points above?}
 
-{\color{orange} Yes, but also this section is a necessary bridge for reproducibility. Maybe it's better as an appendix?}
+%{\color{orange} Yes, but also this section is a necessary bridge for reproducibility. Maybe it's better as an appendix?}
 
-\begin{table}
+\begin{table}[hbtp]
 \centering
 \caption{Descriptions of number of events in the data.
-In principle we can simulate an arbitrary amount of data, but the following quantities were shown to be sufficient for these studies.}\label{tab:datasets}
+In principle we can simulate an arbitrary amount of data, but the following quantities were shown to be sufficient for these studies.{\color{blue} CITE THIS SOMEWHERE!}}\label{tab:datasets}
 \begin{tabular}{lccc}
@@
 
-\begin{table}
-\centering
-\caption{Event class percentages for each of the datasets used in the analysis conducted in this thesis.
-The decrease in the "other" class of events from full to filtered data owes to the thresholding of events.
-If an event contains fewer than $20$ data-points it is discarded}\label{tab:class_distr}
-\begin{tabular}{lccc}
-\toprule
-{} & Simulated & Full & Filtered \\
-\midrule
-$\%$ Proton & $50$ & $25.3$ & $27.6$ \\ 
-$\%$ Carbon & $50$ & $12.0$ & $11.9$ \\
-$\%$ Other & $0$ & $62.8$ & $60.1$ \\
-\end{tabular}
-\end{table}
-
-\section{Data processing}
-
+\section{Data Preparation}
+{\color{blue} MPK edit this section!}
 In this section we give a brief overview of the data, for a more in-depth consideration we refer the reader to \cite{Mittig2015}, \cite{Suzuki2012} and  \cite{Bradt2017a}.
 
-The AT-TPC data we studied for this work was originally recorded as charge time-series for each pad on the micromegas {\color{green} We haven't defined what micromegas is}.
-In this representation an event is a record of 512 time-buckets for each of the $\sim10^4$ detector pads; however, in our analysis, we represent each event as a 2D slice.
+The AT-TPC data we studied for this work was recorded as charge time-series for each of the  the $\sim10^4$ detector pads.
+In this representation an event is a record of 512 time-buckets for each of the detector pads. In our analysis, we represent each event as a downsampled 2D slice.
 We chose to represent the data in 2D to facilitate the use of sophisticated image-recognition machine learning models.
@@
 To explore this trade-off between model performance and pre-processing bias, we performed analysis on simulated, cleaned and raw events.
+
+\begin{table}
+\centering
+\caption{Event class percentages for each of the datasets used in the analysis conducted in this thesis.
+The decrease in the "other" class of events from full to filtered data owes to the thresholding of events.
+If an event contains fewer than $20$ data-points it is discarded }\label{tab:class_distr}
+\begin{tabular}{lccc}
+\toprule
+{} & Simulated & Full & Filtered \\
+\midrule
+$\%$ Proton & $50$ & $25.3$ & $27.6$ \\ 
+$\%$ Carbon & $50$ & $12.0$ & $11.9$ \\
+$\%$ Other & $0$ & $62.8$ & $60.1$ \\
+\end{tabular}
+\end{table}
 
@@
 
-The sensor plane in the AT-TPC is very sensitive, as such there is substantial noise in the ${}^{46}$Ar data.
+%The sensor plane in the AT-TPC is very sensitive, as such there is substantial noise in the ${}^{46}$Ar data.
 The noise can be attributed to structural noise from electronics cross-talk, and possible interactions with cosmic background radiation, as well as other sources of charged particles.

### 2020-05-26 doc/Manuscript/article.tex
Updates from Overleaf
syncing overleaf changes

@@
 
+%\documentclass[preprint,12pt]{elsarticle}
 % uncomment for submission to NIM A
 \documentclass[review,number,sort&compress]{elsarticle}
+\usepackage{lineno}
+\linenumbers
 
-%\documentclass[preprint,12pt]{elsarticle}
 \usepackage{amssymb}
@@
 We found that the standard choice of a euclidian norm on the vector differences provided the best outcomes.
-https://www.overleaf.com/project/5e73f990fcbd310001959b01
 
@@
 \centering
+	\includegraphics[width=\textwidth]{custom_work/Simulatedvgg_conf_mat.pdf} 
+	\caption[Pre-trained network - confusion matrices]{Confusion matrix for the K-means clustering of simulated AT-TPC events.
+The true labels indicate samples belonging to the p (proton), or the carbon (C) class}\label{fig:clster_confmat}
+\end{figure}
 
-	\subfloat{
-	\includegraphics[width=0.35\textwidth]{./plots/Simulatedvgg_pca_conf_mat.pdf}
+\begin{figure}
+\centering
+	\includegraphics[width=\textwidth]{custom_work/Filteredvgg_conf_mat.pdf}
+	\caption[Pre-trained network - confusion matrices]{Confusion matrix for the K-means clustering of filtered AT-TPC events.
+The true labels indicate samples belonging to the p (proton), carbon (C), or other classes.
+}\label{fig:clster_confmat}
+\end{figure}
 
-}
-	\hspace{-1cm}
-	\subfloat{
-	\includegraphics[width=0.35\textwidth]{plots/Filteredvgg_pca_conf_mat.pdf}
-}
-	\hspace{-1cm}
-	\subfloat{
-	\includegraphics[width=0.35\textwidth]{plots/Fullvgg_pca_conf_mat.pdf}
-}
-\caption[Pre-trained network - confusion matrices]{Confusion matrices for the K-means clustering of simulated, filtered and full AT-TPC events.
+\begin{figure}
+	\includegraphics[width=\textwidth]{custom_work/Filteredvgg_conf_mat.pdf}
+\caption[Pre-trained network - confusion matrices]{Confusion matrix for the K-means clustering of raw AT-TPC events.
 The true labels indicate samples belonging to the p (proton), carbon (C), or other classes.

### 2020-05-28 doc/Manuscript/article.tex
Comment out old content from 5.1 and add new outline for this section

@@
 
-To affect the clustering of AT-TPC events, we first considered the results from \citet{Kuchera2019}.
+% To affect the clustering of AT-TPC events, we first considered the results from \citet{Kuchera2019}.
 In their work, the authors consider the segmentation of reactions in the $^{46}$Ar experiment as a classification problem.
@@
 
-Using a CNN algorithm comes with the caveat that the data should be image-like.
+% Using a CNN algorithm comes with the caveat that the data should be image-like.
 This caveat is especially stringent when applying pre-trained models to new data.
@@
 
-In the Machine Learning community, it is not uncommon to publish packaged models with fitted parameters from image recognition contests.
+% In the Machine Learning community, it is not uncommon to publish packaged models with fitted parameters from image recognition contests.
 These models are trained on datasets with millions of images and classify between hundreds of distinct classes; one such is the imagenet dataset.
@@
 
-The VGG16 network is one of six analogous networks proposed by \citet{Simonyan2014}.
+% The VGG16 network is one of six analogous networks proposed by \citet{Simonyan2014}.
 The network architectures are fairly straightforward; for VGG16, there are sixteen layers in the network.
@@
 The full architecture is detailed in appendix \ref{app:vgg}.
+
+{\color{cyan} [Ryan:] The commented-out paragraphs above I think could be better organized and condensed to eliminate some off-topic stuff.
+The outline below is an idea for how this section might tell its story.
+
+\begin{itemize}
+    \item Training high-performing neural networks from scratch often requires enormous datasets and computation time, particularly for vision tasks.
+    \item However, it has been found that models which are trained at large scale will learn general features that are applicable to a variety of tasks.
+    \item For example, large neural networks which are trained on the ImageNet dataset --- a diverse image classification task --- learn how to identify lines, edges, and other common shapes which are useful for numerous problems.
+    \item Thus, it is common practice to initialize the convolutional layers of a network with the pre-trained weights learned from ImageNet (or some other large dataset).
+    \item The training process then only has to fine-tune the network for the specific task. Since we are building on prior knowledge, learning becomes far more efficient.
+    \item \citet{Kuchera2019} demonstrate the promise of this approach for classifying AT-TPC reaction products.
+\end{itemize}
+
+}
 

### 2020-06-01 doc/Manuscript/article.tex

@@
 
-\section{Methods}\label{sec:methods}
-
-\subsection{Classifying  events} 
-\todo{Brute force serving as motivation for why we do unsupervised learning.}
-The filtering task relied on a set of $\chi^2$ distributions based on different criteria, however there was no well-defined method to quantify the effectiveness of the event selection from the thresholds applied to these distributions.
-In addition, the filtering task produced a binary result only, ``good" or``bad", while the data clearly contained more classes of events, if only because of the presence of carbon in the target (from the isobutane gas).
-In a broader perspective, an unsupervised classification algorithm would offer the possibility to ``discover" rare events not anticipated (or overlooked) by the physicist.
-These events would likely be filtered out using the traditional methods.
-From a practical point of view, compared to supervised learning, it also avoids the necessary labeling task of the learning set events, which is error prone and time consuming.
-
-\subsection{Why machine learning}
-
-\begin{itemize}
-    \item Traditional MC methods fall short in two principal ways:
-    \begin{itemize}
-        \item The computational cost per event is too large given the size of the data-sets
-        \item The broken tracks, noisy environment creates bad fit statistics for otherwise useful events. End result is $f1 \sim 0.7 $
-        \item define a ``bad" fit as a non-proton event
-    \end{itemize}
-\end{itemize}
-
-The $\chi^2$ approach used in the traditional analysis performed on the $^{46}$Ar data is extremely computationally expensive because it involves the simulation of thousands of tracks for each recorded event, which are performed each iteration of the Monte-Carlo fitting sequence.
-Even though the reaction of interest in the $^{46}$Ar experiment had the largest cross section (elastic scattering), the sheer number of events to analyse (in the hundreds of thousands) necessitated the use of parallel processing on a high performance computing cluster.
-In the case of an experiment where the reaction of interest would represent less than a few percent of the total cross section, this procedure would become highly inefficient and prohibitive.
-Most of the CPU time would be spent on simulating events that are not of interest.
-An unsupervised ML algorithm able to separate the data without a priori knowledge of the different types of events increases the efficiency of the analysis tremendously, by concentrating the fitting efforts only on events of interest.
-In addition, the clustering allows for more exploration of the data, potentially enabling new discovery of unexpected reaction types.
-
-%\subsection{Data}
-%{\color{green}What is the intent of this section? Display some data examples to illustrate the points above?}
-
-%{\color{orange} Yes, but also this section is a necessary bridge for reproducibility. Maybe it's better as an appendix?}
-
-\begin{table}[hbtp]
-\centering
-\caption{Descriptions of number of events in the data.
-In principle we can simulate an arbitrary amount of data, but the following quantities were shown to be sufficient for these studies.{\color{blue} CITE THIS SOMEWHERE!}}\label{tab:datasets}
-\begin{tabular}{lccc}
-\toprule
-{} & Simulated & Full & Filtered \\
-\midrule
-Total &  $8000$ & $51891$ & $49169$ \\
-labelled & $2400$ & $1774$ &  $1582$ \\ 
-\bottomrule
-\end{tabular}
-\end{table}
-
 \section{Data Preparation}
@@
 
-\section{Machine Learning Methods}
+\section{Methods}\label{sec:methods}
+
+\subsection{Classifying  events} 
+\todo{Brute force serving as motivation for why we do unsupervised learning.}
+The filtering task relied on a set of $\chi^2$ distributions based on different criteria, however there was no well-defined method to quantify the effectiveness of the event selection from the thresholds applied to these distributions.
+In addition, the filtering task produced a binary result only, ``good" or``bad", while the data clearly contained more classes of events, if only because of the presence of carbon in the target (from the isobutane gas).
+In a broader perspective, an unsupervised classification algorithm would offer the possibility to ``discover" rare events not anticipated (or overlooked) by the physicist.
+These events would likely be filtered out using the traditional methods.
+From a practical point of view, compared to supervised learning, it also avoids the necessary labeling task of the learning set events, which is error prone and time consuming.
+
+\subsection{Why machine learning}
+
+\begin{itemize}
+    \item Traditional MC methods fall short in two principal ways:
+    \begin{itemize}
+        \item The computational cost per event is too large given the size of the data-sets
+        \item The broken tracks, noisy environment creates bad fit statistics for otherwise useful events. End result is $f1 \sim 0.7 $
+        \item define a ``bad" fit as a non-proton event
+    \end{itemize}
+\end{itemize}
+
+The $\chi^2$ approach used in the traditional analysis performed on the $^{46}$Ar data is extremely computationally expensive because it involves the simulation of thousands of tracks for each recorded event, which are performed each iteration of the Monte-Carlo fitting sequence.
+Even though the reaction of interest in the $^{46}$Ar experiment had the largest cross section (elastic scattering), the sheer number of events to analyse (in the hundreds of thousands) necessitated the use of parallel processing on a high performance computing cluster.
+In the case of an experiment where the reaction of interest would represent less than a few percent of the total cross section, this procedure would become highly inefficient and prohibitive.
+Most of the CPU time would be spent on simulating events that are not of interest.
+An unsupervised ML algorithm able to separate the data without a priori knowledge of the different types of events increases the efficiency of the analysis tremendously, by concentrating the fitting efforts only on events of interest.
+In addition, the clustering allows for more exploration of the data, potentially enabling new discovery of unexpected reaction types.
+
+%\subsection{Data}
+%{\color{green}What is the intent of this section? Display some data examples to illustrate the points above?}
+
+%{\color{orange} Yes, but also this section is a necessary bridge for reproducibility. Maybe it's better as an appendix?}
+
+\begin{table}[hbtp]
+\centering
+\caption{Descriptions of number of events in the data.
+In principle we can simulate an arbitrary amount of data, but the following quantities were shown to be sufficient for these studies.{\color{blue} CITE THIS SOMEWHERE!}}\label{tab:datasets}
+\begin{tabular}{lccc}
+\toprule
+{} & Simulated & Full & Filtered \\
+\midrule
+Total &  $8000$ & $51891$ & $49169$ \\
+labelled & $2400$ & $1774$ &  $1582$ \\ 
+\bottomrule
+\end{tabular}
+\end{table}
 
@@
         \item Supervised vs. Unsupervised learning
-        \item Define Classifcation and Clustering 
+        \item Define Classification and Clustering 
         \item Define latent space (refer)
-        \item Traditional unupervised - distances in high dimensional spaces etc. 
+        \item Traditional unsupervised - distances in high dimensional spaces etc. 
     \end{itemize}
@@
 	\caption[Pre-trained network - confusion matrices]{Confusion matrix for the K-means clustering of simulated AT-TPC events.
-The true labels indicate samples belonging to the p (proton), or the carbon (C) class}\label{fig:clster_confmat}
+The true labels indicate samples belonging to the p (proton), or the carbon (C) class}\label{fig:clster_confmat_sim}
 \end{figure}
@@
 The true labels indicate samples belonging to the p (proton), carbon (C), or other classes.
-}\label{fig:clster_confmat}
+}\label{fig:clster_confmati_filt}
 \end{figure}
@@
 \begin{figure}
-	\includegraphics[width=\textwidth]{custom_work/Filteredvgg_conf_mat.pdf}
+	\includegraphics[width=\textwidth]{custom_work/Rawvgg_conf_mat.pdf}
 \caption[Pre-trained network - confusion matrices]{Confusion matrix for the K-means clustering of raw AT-TPC events.
 The true labels indicate samples belonging to the p (proton), carbon (C), or other classes.
-}\label{fig:clster_confmat}
+}\label{fig:clster_confmati_raw}
 \end{figure}

### 2020-06-02 doc/Manuscript/article.tex
Updates from Overleaf

@@
 \section{Results and Discussions}\label{sec:results}
+\begin{itemize}
+\item Clear out the "methods" like parts.
+\item merge 5.1 and 5.2 - keeping point 1 in mind.
+\item merge 5.3 5.4 and 5.6 and create table from figure 10, 11 and 12. 
+\item go over 5.7 - mention instability in MIXAE results, peak performance of MIXAE is good - consistency is not. 
+\item remove fig. 14 and fig 15. 
+\end{itemize}
+
 The principal challenge in the AT-TPC experiments that we are trying to solve is the reliance on labelled samples in the analysis as future experiments may not have as visually dissimilar reaction products  as we observe in the ${}^{46}$Ar experiment.

### 2020-06-07 doc/Manuscript/article.tex
Updates from Overleaf
added table to mixae

@@
 
-{\color{cyan} [Ryan:] The commented-out paragraphs above I think could be better organized and condensed to eliminate some off-topic stuff.
-The outline below is an idea for how this section might tell its story.
-
-\begin{itemize}
-    \item Training high-performing neural networks from scratch often requires enormous datasets and computation time, particularly for vision tasks.
-    \item However, it has been found that models which are trained at large scale will learn general features that are applicable to a variety of tasks.
-    \item For example, large neural networks which are trained on the ImageNet dataset --- a diverse image classification task --- learn how to identify lines, edges, and other common shapes which are useful for numerous problems.
-    \item Thus, it is common practice to initialize the convolutional layers of a network with the pre-trained weights learned from ImageNet (or some other large dataset).
-    \item The training process then only has to fine-tune the network for the specific task. Since we are building on prior knowledge, learning becomes far more efficient.
-    \item \citet{Kuchera2019} demonstrate the promise of this approach for classifying AT-TPC reaction products.
-\end{itemize}
-
-}
+Training high-performing neural networks from scratch often requires enormous datasets and computation time.
+However, it has been found that models which are trained at large scale will learn general features that are applicable to a variety of tasks.
+For example, large neural networks which are trained on the ImageNet dataset \cite{Russakovsky2015} --- a diverse image classification task --- learn how to identify lines, edges, and other common shapes which are useful for numerous problems.
+Thus, it is common practice to initialize the convolutional layers of a network with the pre-trained weights learned from ImageNet (or some other large dataset).
+The training process then only has to fine-tune the network for the specific task.
+Since we are building on prior knowledge in this case, learning becomes far more efficient and better performance can often be achieved.
+\citet{Kuchera2019} used machine learning methods to classify the products of $^{46}$Ar reactions in the AT-TPC, and they found that a CNN initialized with ImageNet weights resulted in the most successful classification.
 
@@
 
-In contrast with the classification work by \citet{Kuchera2019}, we do not append a classifying layer to the VGG16 network given that we do not have access to the ground truth labels at training time in the clustering regime.
+% In contrast with the classification work of \citet{Kuchera2019}, we do not append a classifying layer to the VGG16 network given that we do not have access to the ground truth labels at training time in the clustering regime.
 We instead consider the output from the convolutional blocks of the VGG16 network as a representation of the events on which we will apply a clustering algorithm.
+
+% The clustering algorithm we found to give the strongest result is the K-means algorithm.
+This algorithm is contingent on having a distance measurement between two datapoints.
+We found that the standard choice of a euclidian norm on the vector differences provided the best outcomes.
+
+In contrast with the classification work of \citet{Kuchera2019}, we do not have access to ground truth labels and are trying to solve a fundamentally different learning problem.
+Thus, rather than fine-tuning a pre-trained network under the supervised learning regime, we extract the output of the pre-trained network's last convolutional layer as a latent representation of the events on which we will apply a clustering algorithm.
 
@@
 This is not the case for DBSCAN.
+
 \subsection{Autoencoder clustering}
-\todo{look at grayscale images or colorbar with propto charge}
+
+\begin{table}[H]
+\centering 
+\caption[MIXAE clustering performance]{MIXAE clustering performance on the ${}^{46} Ar$ experimental data.
+In contrast with the VGG-16 + K-means approach we observe  significant variations in performance.}\label{tab:clstr_vgg}
+\input{./plots/mixae_clustering.tex}
+\end{table}
 
@@
 
-\begin{figure}[H]
-\centering
-\includegraphics[width=\textwidth]{./plots/sim_mixae.pdf}
-\caption[Clustering performance of MIXAE on simulated AT-TPC data]{Performance for the MIXAE model on simulated AT-TPC data.
-In the top row the loss components are plotted for each run, and in the bottom row the adjusted rand index (ARI) and clustering accuracy are shown.
-Each run is color-coded with the ARI achieved at the end of the run.
-}\label{fig:mixae_sim}
-\end{figure}
-
 \subsection{Filtered AT-TPC data}
@@
 
-\begin{figure}[H]
-\centering
-\includegraphics[width=\textwidth]{./plots/clean_mixae.pdf}
-\caption[Clustering performance of MIXAE on filtered event AT-TPC data]{Performance for the MIXAE model on filtered AT-TPC data.
-In the top row the loss components are plotted for each run, and in the bottom row the adjusted rand index (ARI) and clustering accuracy are shown.
-Each run is color-coded with the ARI achieved at the end of the run.}\label{fig:mixae_clean}
-\end{figure}
-
 \subsection{Full AT-TPC data}
@@
 \noindent As with the previous datasets we include a plot of the loss curves and performance measures for $N=10$ runs with the same loss-weight parameters, shown in figure \ref{fig:mixae_full}.
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=\textwidth]{./plots/real_mixae.pdf}
-\caption[Clustering performance of MIXAE on full event AT-TPC data]{Performance for the MIXAE model on un-filtered AT-TPC data.
-In the top row the loss components are plotted for each run, and in the bottom row the adjusted rand index (ARI) and clustering accuracy are shown.
-Each run is color-coded with the ARI achieved at the end of the run.}\label{fig:mixae_full}
-\end{figure}
 
@@
 
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{plots/clean_cluster_repr.pdf}
-\caption[Selection of carbon events in differing clusters]{Selection of carbon events belonging to different clusters.
-Each row represents one of the clusters for the filtered data as shown in figure \ref{fig:mixae_confmat}.
-There seems to be no clear distinction between these rows.}\label{fig:filtered_mixae_clster_repr}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{plots/real_cluster_repr.pdf}	
-\caption[Selection of proton events in differing clusters]{Selection of proton events belonging to different clusters.
-Each row represents one of the clusters for the full data as shown in figure \ref{fig:mixae_confmat}.
-We observe a tendency for the more noisy events to be placed in the "other" cluster}\label{fig:full_mixae_clster_repr}
-\end{figure}
-
 \section{Conclusions and Perspectives}\label{sec{conclusion}}

### 2020-06-07 doc/Manuscript/plots/mixae_clustering.tex
added table to mixae

\begin{tabular}{lllll}
\toprule
{} & \multicolumn{2}{c}{Accuracy} &   \multicolumn{2}{c}{ARI} \\
\midrule
{} & Top 1 & $\mu \pm \sigma$ & Top 1 & $\mu \pm \sigma$ \\
Simulated &  0.96 &  $0.74 \pm 0.16 $  & 0.84 &  $0.33 \pm 0.32 $\\
Filtered  & 0.75 &  $0.71 \pm 0.04 $ & 0.52 &  $0.38 \pm 0.14 $\\
Full      & 0.71 &  $0.61 \pm 0.07 $ &  0.32 &  $0.09 \pm 0.10 $   \\
\bottomrule
\end{tabular}

### 2020-06-07 doc/Manuscript/article.tex
Updates from Overleaf

@@
 
-\begin{table}[H]
-\todo{prettify or remove}
-\renewcommand*{\arraystretch}{0.5}
-\centering
-\caption{Hyperparameters selected for the autoencoder components of the MIXAE algorithm, see table \ref{tab:convae_hyperparams} for a full description of the parameters.}\label{tab:mixe_ae_hyperparams}
-\setlength{\extrarowheight}{15pt}
-\hspace*{-0.5in}
-\begin{tabular}{ll}
-\toprule
-Hyperparameter & Value \\
-\midrule
-\multicolumn{2}{l}{Convolutional parameters: } \\
-\midrule
-Number of layers & $4$ \\
-Kernels & $[3,\,3,\,3,\,3]$\\
-Strides & $[2,\,2,\,2,\,2]$ \\
-Filters & $[64,\, 32, \,16, \,8,]$ \\ 
-\midrule
-\multicolumn{2}{l}{Network parameters: } \\
-\midrule
-Activation & LReLu \\
-Latent dimension & 20  \\
-Batchnorm & False \\
-\midrule
-\multicolumn{2}{l}{Optimizer parameters: } \\
-\midrule
-$\eta$ & $10^{-3}$ \\
-$\beta_1$ & $0.9$ \\
-$\beta_2$ & $0.99$ \\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\noindent Since there are then only three remaining hyperparameters we choose to perform a coarse grid-search as described in section \ref{sec:hyperparam_search_arch}.
-
-\begin{table}
-\centering
-\caption{Hyperparameter grid for the MIXAE loss weighting terms. The grid is given as exponents for logarithmic scales.}\label{tab:mixae_loss_weights}
-\begin{tabular}{lll}
-\toprule
-Parameter & Grid & Scale \\
-\midrule 
-$\theta$ & $[-1,\, 5]$ & Logarithmic \\
-$\alpha$ & $[-5,\, -1]$ & Logarithmic \\
-$\gamma$ & $[3,\, 5]$ & Logarithmic
-\end{tabular}
-\end{table}
-
 \subsection{Simulated AT-TPC data}
-
-To train the MIXAE clustering algorithm, we use the large simulated dataset with $M=80000$ points, evenly distributed between proton- and carbon-events.
-The algorithm is trained on a subset of $60000$ of these samples, and we track performance on the remaining $20000$ events.
-
-The grids selected for the search are listed in table \ref{tab:mixae_loss_weights}. The search yielded an optimal configuration with
-
-\begin{align}
-\theta = 10^{-1}, \\
-\alpha = 10^{-2}, \\
-\gamma = 10^5.
-\end{align}
 
@@
 We then performed a local search around these values to pin-point the hyperparameter configuration.
-This search yielded the optimal hyper-parameters
-
-\begin{align}
-\theta &= 10^{1}, \\
-\alpha &= 10^{-1}, \\
-\gamma &= 3.162\times 10^3.
-\end{align}
 
@@
 Each configuration is re-run a total of $N=10$ times to capture fluctuations in the performance before a final selection is made on the hyperparameters.
-For the full dataset the MIXAE hyperparameters converge to the same values as for the clean data, i.e.
-
-\begin{align}
-\theta &= 10^{1}, \\
-\alpha &= 10^{-1}, \\
-\gamma &= 3.162\times 10^3.
-\end{align}
 
@@
 
+\section{MIXAE hyper-parameter tuning}
+
+To train the MIXAE clustering algorithm, we use the large simulated dataset with $M=80000$ points, evenly distributed between proton- and carbon-events.
+The algorithm is trained on a subset of $60000$ of these samples, and we track performance on the remaining $20000$ events.
+Since there are then only three remaining hyperparameters we choose to perform a coarse grid-search as described in section \ref{sec:hyperparam_search_arch}.
+
+\begin{table}
+\centering
+\caption{Hyperparameter grid for the MIXAE loss weighting terms. The grid is given as exponents for logarithmic scales.}\label{tab:mixae_loss_weights}
+\begin{tabular}{lll}
+\toprule
+Parameter & Grid & Scale \\
+\midrule 
+$\theta$ & $[-1,\, 5]$ & Logarithmic \\
+$\alpha$ & $[-5,\, -1]$ & Logarithmic \\
+$\gamma$ & $[3,\, 5]$ & Logarithmic
+\end{tabular}
+\end{table}
+
+The grids selected for the search are listed in table \ref{tab:mixae_loss_weights}. The search yielded an optimal configuration with
+
+\begin{align}
+\theta = 10^{-1}, \\
+\alpha = 10^{-2}, \\
+\gamma = 10^5.
+\end{align}
+
+For the full dataset the MIXAE hyperparameters converge to the same values as for the clean data:
+
+\begin{align}
+\theta &= 10^{1}, \\
+\alpha &= 10^{-1}, \\
+\gamma &= 3.162\times 10^3.
+\end{align}
+
+Lastly we supply the configuration used for the individual convolutional autoencoder networks in table \ref{tab:mixe_ae_hyperparams}
+
+\begin{table}[H]
+\todo{prettify or remove}
+\renewcommand*{\arraystretch}{0.5}
+\centering
+\caption{Hyperparameters selected for the autoencoder components of the MIXAE algorithm, see table \ref{tab:convae_hyperparams} for a full description of the parameters.}\label{tab:mixe_ae_hyperparams}
+\setlength{\extrarowheight}{15pt}
+\hspace*{-0.5in}
+\begin{tabular}{ll}
+\toprule
+Hyperparameter & Value \\
+\midrule
+\multicolumn{2}{l}{Convolutional parameters: } \\
+\midrule
+Number of layers & $4$ \\
+Kernels & $[3,\,3,\,3,\,3]$\\
+Strides & $[2,\,2,\,2,\,2]$ \\
+Filters & $[64,\, 32, \,16, \,8,]$ \\ 
+\midrule
+\multicolumn{2}{l}{Network parameters: } \\
+\midrule
+Activation & LReLu \\
+Latent dimension & 20  \\
+Batchnorm & False \\
+\midrule
+\multicolumn{2}{l}{Optimizer parameters: } \\
+\midrule
+$\eta$ & $10^{-3}$ \\
+$\beta_1$ & $0.9$ \\
+$\beta_2$ & $0.99$ \\
+\bottomrule
+\end{tabular}
+\end{table}
+
 \bibliography{bibliography}

### 2020-06-07 doc/Manuscript/article.tex

@@
 \begin{figure}
-\includegraphics[width=\textwidth]{plots/filtered_vgg_cluster_repr.pdf}
+\includegraphics[width=\textwidth]{custom_work/Filteredexamples.pdf}
 \caption[Filtered proton samples by cluster belonging]{Illustrating a sample of proton events from different K-means clusters from the filtered dataset.
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{plots/full_vgg_cluster_repr.pdf}
-\caption[Full proton samples by cluster belonging]{Illustrating a sample of proton events from different K-means clusters from the full dataset.
+\includegraphics[width=\textwidth]{custom_work/Rawexamples.pdf}
+\caption[Full proton samples by cluster belonging]{Illustrating a sample of proton events from different K-means clusters from the raw dataset.
 Each row belongs to a single cluster corresponding to the full confusion matrix in figure \ref{fig:clster_confmat}}\label{fig:full_vgg_clster_repr}

### 2020-06-07 doc/Manuscript/article.tex
all plots new

@@
 
-\begin{figure}[ht]
-\centering
-\includegraphics[width=\textwidth]{plots/display_eventssimulated.pdf}
-\caption[Displaying simulated events in 2D and 3D]{Two- and three-dimensional representations of two events from a simulated ${}^{46}$Ar experiment.
-Each row is one event in two projections, where the color intensity of each point indicates higher charge values recorded by the detector.}\label{fig:sim_samples}
-\end{figure}
-
 \subsection{Raw \texorpdfstring{${}^{46}$Ar}{46Ar}  events}\label{sec:data_real}
@@
 \centering
-\includegraphics[width=\textwidth]{plots/display_eventsfull_.pdf}
+\includegraphics[width=\textwidth]{custom_work/examples_raw.pdf}
 \caption[Displaying un-filtered events in 2D and 3D]{Two- and three-dimensional representations of two events from the ${}^{46}$Ar experiment.
@@
 \centering
-\includegraphics[width=0.9\textwidth, height=9cm]{plots/display_eventsclean_.pdf}
+\includegraphics[width=0.9\textwidth, height=9cm]{custom_work/examples_filtered.pdf}
 \caption[Displaying filtered events in 2D and 3D]{Two- and three-dimensional representations of two events from the ${}^{46}$Ar experiment.
@@
 
-	\subfloat{
-	\includegraphics[width=0.5\textwidth]{plots/filtered_mixae_conf_mat.pdf}
-}
-	\hspace{-1cm}
-	\subfloat{
-	\includegraphics[width=0.5\textwidth]{plots/full_mixae_conf_mat.pdf}
-}
+	\includegraphics[width=\textwidth]{custom_work/Filtered_mixae_conf_mat.pdf}
+
+\end{figure}
+\begin{figure}[H]
+\centering
+
+	\includegraphics[width=\textwidth]{custom_work/Raw_mixae_conf_mat.pdf}
+
 \caption[MIXAE - confusion matrices]{Confusion matrices for the MIXAE clustering algorithm on filtered and full AT-TPC events.

### 2020-06-07 doc/Manuscript/article.tex
Data section edited

@@
 \section{Data Preparation}
-{\color{blue} MPK edit this section!}
+
 In this section we give a brief overview of the data, for a more in-depth consideration we refer the reader to \cite{Mittig2015}, \cite{Suzuki2012} and  \cite{Bradt2017a}.
@@
 The AT-TPC data we studied for this work was recorded as charge time-series for each of the  the $\sim10^4$ detector pads.
-In this representation an event is a record of 512 time-buckets for each of the detector pads. In our analysis, we represent each event as a downsampled 2D slice.
+In this representation, an event is a record of 512 time-buckets for each of the detector pads. In our analysis, we represent each event as a downsampled 2D projection.
 We chose to represent the data in 2D to facilitate the use of sophisticated image-recognition machine learning models.
@@
 Using raw data is advantageous as it provides an unbiased view of the event, but the data volume and noise levels might be prohibitive for analysis.
-On the opposite end of the pipeline, we will incrementally add bias to the analysis the more steps we take before applying the algorithm.
+%On the opposite end of the pipeline, 
+Therefore, we will incrementally add bias to the analysis by applying the algorithm further down the analysis pipeline..
 With the benefit being that more pre-processing improves the signal-to-noise ratio, possibly improving model performance.
@@
 \centering
-\caption{Event class percentages for each of the datasets used in the analysis conducted in this thesis.
+\caption{Event class percentages for each of the datasets used in this analysis.
 The decrease in the "other" class of events from full to filtered data owes to the thresholding of events.
@@
 AT-TPC tracks were simulated with the \lstinline{pytpc} package developed by \citet{Bradt2017a}.
-Using the same parameters as for the ${}^{46}$Ar$(p, p)$ experiment, a small set of $N=4000$ events per class was generated, as well as a larger set of $N=40000$ events per class.
+Using the same parameters {\color{green} what parameters are you refering to?
+-MPK} as for the ${}^{46}$Ar$(p, p)$ experiment, a small set of $N=4000$ events per class was generated, as well as a larger set of $N=40000$ events per class.
 
@@
 \includegraphics[width=\textwidth]{custom_work/examples_raw.pdf}
-\caption[Displaying un-filtered events in 2D and 3D]{Two- and three-dimensional representations of two events from the ${}^{46}$Ar experiment.
+\caption[Displaying unfiltered events in 2D and 3D]{Two- and three-dimensional representations of two events from the ${}^{46}$Ar experiment.
 Each row is one event in two projections, where the color intensity of each point indicates higher charge values recorded by the detector.}\label{fig:samples}
@@
 To remove the correlated noise, researchers at the NSCL developed an algorithm based on the Hough transform \cite{Newman1972}.
-This transformation is a common technique in computer vision, used to identify common geometric shapes like lines and circles.
-Essentially, the algorithm draws many lines (of whatever desired geometry) through each data-point and checks whether these lines intersect with points in the dataset.
+This transformation is a common technique in computer vision, used to identify common geometric shapes like lines and circles, and has been used extensively in high-energy particle physics since the bubble-chamber era \cite{Hough:1959}.
+%Essentially, the algorithm draws many lines (of whatever desired geometry) through each data-point and checks whether these lines intersect with points in the dataset.
 Locations in parameter space that generate many intersections then become bright spots, allowing us to filter away points that are not close to these points.
@@
 
-We illustrate two filtered events in figure \ref{fig:samples_filtered}. These are the same events as shown in figure \ref{fig:samples}, but with the Houghes' and nearest neighbours filtering applied.
+We illustrate two filtered events in figure \ref{fig:samples_filtered}. These are the same events as shown in figure \ref{fig:samples}, but with the Hough and nearest neighbours filtering applied.
 

### 2020-06-08 doc/Manuscript/article.tex
Edited section 4 -MPK

@@
 AT-TPC tracks were simulated with the \lstinline{pytpc} package developed by \citet{Bradt2017a}.
-Using the same parameters {\color{green} what parameters are you refering to?
+Using the same parameters {\color{blue} what parameters are you referring to?
 -MPK} as for the ${}^{46}$Ar$(p, p)$ experiment, a small set of $N=4000$ events per class was generated, as well as a larger set of $N=40000$ events per class.
@@
 \todo{Brute force serving as motivation for why we do unsupervised learning.}
-The filtering task relied on a set of $\chi^2$ distributions based on different criteria, however there was no well-defined method to quantify the effectiveness of the event selection from the thresholds applied to these distributions.
-In addition, the filtering task produced a binary result only, ``good" or``bad", while the data clearly contained more classes of events, if only because of the presence of carbon in the target (from the isobutane gas).
+%The filtering task relied on a set of $\chi^2$ distributions based on different criteria, 
+The traditional event selection process involved a Monte-Carlo $\chi^2$ minimization method, however there was no well-defined method to quantify the effectiveness of the event selection from the thresholds applied to these distributions.
+{\color{blue} I think we discussed this method earlier, but I need to verify. otherwise needs more explanation -MPK}
+In addition, the selection task produced a binary result only, ``good" or``bad", relative to the event of interest, while the data clearly contained more classes of events%, if only because of the presence of carbon in the target (from the isobutane gas).
 In a broader perspective, an unsupervised classification algorithm would offer the possibility to ``discover" rare events not anticipated (or overlooked) by the physicist.
@@
 The $\chi^2$ approach used in the traditional analysis performed on the $^{46}$Ar data is extremely computationally expensive because it involves the simulation of thousands of tracks for each recorded event, which are performed each iteration of the Monte-Carlo fitting sequence.
-Even though the reaction of interest in the $^{46}$Ar experiment had the largest cross section (elastic scattering), the sheer number of events to analyse (in the hundreds of thousands) necessitated the use of parallel processing on a high performance computing cluster.
+Even though the reaction of interest in the $^{46}$Ar experiment had the largest cross section (elastic scattering), the sheer number of events to analyse (in the hundreds of thousands) necessitated the use of parallel processing on a high performance computing cluster {\color{blue} is there a metric we can give here?
+-MPK}.
 In the case of an experiment where the reaction of interest would represent less than a few percent of the total cross section, this procedure would become highly inefficient and prohibitive.
-Most of the CPU time would be spent on simulating events that are not of interest.
-An unsupervised ML algorithm able to separate the data without a priori knowledge of the different types of events increases the efficiency of the analysis tremendously, by concentrating the fitting efforts only on events of interest.
+%Most of the CPU time would be spent on simulating events that are not of interest.
+The computationally expensive fitting procedure would be applied to every event, instead of the few percent of the events that are of interest for the analysis.
+An unsupervised ML algorithm able to separate the data without a priori knowledge of the different types of events increases the efficiency of the analysis tremendously, and allows the downstream analysis to concentrate on the fitting efforts only on events of interest.
 In addition, the clustering allows for more exploration of the data, potentially enabling new discovery of unexpected reaction types.
@@
 \caption{Descriptions of number of events in the data.
-In principle we can simulate an arbitrary amount of data, but the following quantities were shown to be sufficient for these studies.{\color{blue} CITE THIS SOMEWHERE!}}\label{tab:datasets}
+In principle we can simulate an arbitrary amount of data, but the following quantities were shown to be sufficient for these studies.{\color{blue} reference results?}}\label{tab:datasets}
 \begin{tabular}{lccc}
@@
 
-\begin{itemize}
-    \item Reiterate the aim of unsupervised  clustering of events.
-    \item Challenges from a machine learning perspective. 
-    \begin{itemize}
-        \item Supervised vs. Unsupervised learning
-        \item Define Classification and Clustering 
-        \item Define latent space (refer)
-        \item Traditional unsupervised - distances in high dimensional spaces etc. 
-    \end{itemize}
-    \item Digression here to explain some ML concepts? Link back to Conv part of Michelles paper? 
-    \item Opportunities presented by: 
-    \begin{itemize}
-        \item Transfer learning (train on VGG use on AT-TPC)
-        \item Compression is understanding (leverage autoencoders) 
-    \end{itemize}
-    \item Building on Kuchera et. al we investigate the VGG16 networks application. 
-    \item Novel contribution by applying clustering autoencoder networks. 
-\end{itemize}
+%\begin{itemize}
+%     \item Reiterate the aim of unsupervised  clustering of events.
+%     \item Challenges from a machine learning perspective. 
+%     \begin{itemize}
+%         \item Supervised vs. Unsupervised learning
+%         \item Define Classification and Clustering 
+%         \item Define latent space (refer)
+%         \item Traditional unsupervised - distances in high dimensional spaces etc. 
+%     \end{itemize}
+%     \item Digression here to explain some ML concepts? Link back to Conv part of Michelles paper? 
+%     \item Opportunities presented by: 
+%     \begin{itemize}
+%         \item Transfer learning (train on VGG use on AT-TPC)
+%         \item Compression is understanding (leverage autoencoders) 
+%     \end{itemize}
+%     \item Building on Kuchera et. al we investigate the VGG16 networks application. 
+%     \item Novel contribution by applying clustering autoencoder networks. 
+% \end{itemize}
 
@@
 
-In contrast with the classification work of \citet{Kuchera2019}, we do not have access to ground truth labels and are trying to solve a fundamentally different learning problem.
+In contrast with the classification work of \citet{Kuchera2019}, we do not assume access to ground truth labels and are trying to solve a fundamentally different learning problem.
 Thus, rather than fine-tuning a pre-trained network under the supervised learning regime, we extract the output of the pre-trained network's last convolutional layer as a latent representation of the events on which we will apply a clustering algorithm.
@@
 	\centering
-	\includegraphics[width=\textwidth]{plots/mixae.pdf}
+	\includegraphics[width=.8\textwidth]{plots/mixae.pdf}
 	\caption[Mixture of autoencoders schematic]{Schematic of a MIXAE model.
@@
 \end{figure}
-
+{\color{blue} This section ends quite abruptly -MPK}
 \section{Results and Discussions}\label{sec:results}

### 2020-06-08 doc/Manuscript/article.tex
Some edits to results, but I think it is still in progress -MPK

@@
 
-\subsection{Transfer learning}
+\subsection{Pre-trained model with K-means}
 Building on the work by \citet{Kuchera2019}, we also use the VGG16 pre-trained network as a baseline for the clustering performance.
 We begin by considering a classical K-means approach to clustering.
-However, the output from the VGG16 network is very high dimensional which traditionally conficts with good K-means results.
+However, the output from the VGG16 network is very high dimensional which traditionally conflicts with good K-means results.
 With output vectors in $\R^{8192}$ the ratio between $L_p$ distances goes to one with increasing dimensionality, as shown by \citet{Aggarwal}.
@@
 
-\subsection{K-means}
+%{\color{blue} Are there supposed to be results here? -MPK}
+%\subsection{K-means}
 
-We begin by investigating the K-means clustering algorithm on the VGG16 latent space.
+%We begin by investigating the K-means clustering algorithm on the VGG16 latent space. 
 The VGG16 model is pre-trained on the imagenet dataset, and subsequently applied to the AT-TPC events without the classification layers.

### 2020-06-08 doc/Manuscript/article.tex
Updates from Overleaf

@@
 \begin{abstract}
-In this work we introduce the application of convolutional autoencoder neural networks to the analysis of two-dimensional projections of particle tracks from a resonant proton scattering experiment on ${}^{46}$Ar.
-The data we analyze were recorded by an active target time-projection chamber (AT-TPC).
+This article presents novel applications of machine learning methods to the problem of event classification in an active target detector, the Active Target Time Projection Chamber (AT-TPC) \cite{Bradt2017}.
+Machine learning presents an interesting avenue for researchers operating the AT-TPC, as traditional analysis methods of AT-TPC data are computationally expensive and have to fit all tracks against the event type of interest.
+The latter presents a considerable challenge when either the space of reactions is not known prior to the analysis, or the event type of interest is not the most abundant.
+
+The application of various machine learning algorithms to the analysis of two-dimensional projections of particle tracks from a resonant proton scattering experiment on $^{46}$Ar is introduced.
+We explore the performance of autoencoder neural networks and a pre-trained VGG16 \cite{Simonyan2014} convolutional neural network on two tasks: a semi-supervised classification task and the unsupervised clustering of particle tracks.
+On the semi-supervised task, we find that a logistic regression classifier trained on small labelled subsets of the latent space of these models perform very well, both on simulated and real data.
+On the clustering task, we find that a $K$-means algorithm applied to the simulated data in the VGG16 latent space forms almost perfect clusters.
+Additionally, the VGG16+K-means approach finds high purity clusters of proton events for real experimental data.
+We also explore the application of neural networks to clustering by implementing a mixture of autoencoders algorithm, with improved results on real data.
+However, the neural network clustering suffers from stability issues necessitating further investigations into this approach.
+
+%In this work we introduce the application of convolutional autoencoder neural networks to the analysis of two-dimensional projections of particle tracks from a resonant proton scattering experiment on ${}^{46}$Ar.
+%The data we analyze were recorded by an active target time-projection chamber (AT-TPC).
 Machine learning presents an interesting avenue for researchers operating an AT-TPC, as traditional analysis methods of AT-TPC data are both computationally expensive and fit all particle tracks against the event type of interest.
@@
 
-We explore the performance of the autoencoder neural networks and a pre-trained VGG16 \cite{Simonyan2014} convolutional neural network on two tasks: a semi-supervised classification task and the unsupervised clustering of particle tracks.
+%We explore the performance of the autoencoder neural networks and a pre-trained VGG16 \cite{Simonyan2014} convolutional neural network on two tasks: a semi-supervised classification task and the unsupervised clustering of particle tracks.
 On the semi-supervised task, we find that a logistic regression classifier trained on small labelled subsets of the latent space of these models perform very well.
@@
 
-On the clustering task, we found that a $K$-means algorithm applied to the simulated data in the VGG16 latent space forms almost perfect clusters, with an adjusted rand index \cite{Hubert1985} ($ARI$) $ > 0.8$.
+%On the clustering task, we found that a $K$-means algorithm applied to the simulated data in the VGG16 latent space forms almost perfect clusters, with an adjusted rand index \cite{Hubert1985} ($ARI$) $ > 0.8$.
 Additionally, the VGG16+K-means approach finds high purity clusters of proton events for real experimental data.
@@
 
-The concept of active targets aim at mitigating this compromise, by turning the target itself into a detector \cite{Beceiro2018}.
+The concept of active targets aim at mitigating this compromise, by turning the target itself into a detector \cite{BECEIRONOVO2015}.
 Most active target detectors such as the AT-TPC are composed of a time projection chamber (TPC) where the detector gas is at the same time the target material.
@@
 A direct consequence of this concept is the inclusiveness of the experimental data recorded by this type of detector: any nuclear reaction happening within the target is recorded.
-Although this sounds like an advantage from the scientific point of view, it poses great challenges during the analysis phase, that are reminiscent of bubble chamber times and on par with event classification challenges in particle physics today.
-{\color{blue} cite something HEP here} More often than not, the reaction channel of interest has one of the lowest cross sections, therefore the physicist is faced with the task of sorting out the corresponding events from the ``background" of other reaction channels.
+Although this sounds like an advantage from the scientific point of view, it poses great challenges during the analysis phase, that are reminiscent of bubble chamber times and on par with event classification challenges in particle physics today, see for example the recent review of Mehta {\em et al.} \cite{mehta2019}.
+More often than not, the reaction channel of interest has one of the lowest cross sections, therefore the physicist is faced with the task of sorting out the corresponding events from the ``background" of other reaction channels.
 
@@
 Because TPCs produce 3-dimensional images of charged particle tracks, the event identification task is often akin to a visual inspection (akin to analysis in the bubble chamber era), which is not practical nowadays because of the large quantities of data.
-Machine learning techniques then appear as a promising prospect, in particular in the image recognition domain where much progress has been made recently.
-{\color{blue} short discussion on prior work in nuclear/HEP/neutrino} In addition, some ML algorithms offer new possibilities such as the potential discovery of unforeseen phenomena that would have been missed by more traditional analysis methods.
+Machine learning techniques then appear as a promising prospect, in particular in the image recognition domain where much progress has been made recently \cite{mehta2019}.
+In addition, some ML algorithms offer new possibilities such as the potential discovery of unforeseen phenomena that would have been missed by more traditional analysis methods.
 Prior work has demonstrated the ability to apply supervised classification machine learning methods to AT-TPC data when a labeled training set is available, whether through hand-labeled data or labeled simulated data (a {\em transfer learning} application) \cite{Kuchera2019}.
@@
 Subsequently, we investigate other clustering methods and the autoencoder based clustering algorithm outlined in section \ref{sec:mixae}.
-
+{\color{blue} Move methods to pervious section -MPK}
 \subsection{Pre-trained model with K-means}

### 2020-06-10 doc/Manuscript/article.tex
Updates from Overleaf

@@
 \begin{abstract}
-This article presents novel applications of machine learning methods to the problem of event classification in an active target detector, the Active Target Time Projection Chamber (AT-TPC) \cite{Bradt2017}.
+% Modified by DB (6/8/2020)
+This article presents novel applications of machine learning methods to the problem of event separation in an active target detector, the Active-Target Time Projection Chamber (AT-TPC) \cite{Bradt2017}.
 Machine learning presents an interesting avenue for researchers operating the AT-TPC, as traditional analysis methods of AT-TPC data are computationally expensive and have to fit all tracks against the event type of interest.
@@
 \centering
-\caption{Descriptions of number of events in the data.
-In principle we can simulate an arbitrary amount of data, but the following quantities were shown to be sufficient for these studies.{\color{blue} reference results?}}\label{tab:datasets}
+\caption{Descriptions of number of events in the data.}\label{tab:datasets}
 \begin{tabular}{lccc}
@@
 In contrast with the classification work of \citet{Kuchera2019}, we do not assume access to ground truth labels and are trying to solve a fundamentally different learning problem.
-Thus, rather than fine-tuning a pre-trained network under the supervised learning regime, we extract the output of the pre-trained network's last convolutional layer as a latent representation of the events on which we will apply a clustering algorithm.
-
-The clustering algorithm we found to give the strongest result is the K-means algorithm.
-This algorithm is contingent on having a distance measurement between two datapoints.
-We found that the standard choice of a euclidian norm on the vector differences provided the best outcomes.
+Thus, rather than fine-tuning a pre-trained network under the supervised learning regime, we extract the output of the pre-trained network's last convolutional layer as a latent representation of the events.
+We then cluster events based on this representation using the K-means++ algorithm with the standard choice of Euclidean distance.
 
@@
 These models are optimized specifically to the AT-TPC data with the objective of clustering.
-For this paper, we considered the deep convolutional embedded clustering (DCEC) by \citet{Guo2017} as well as the mixture of autoencoders (MIXAE) method introduced by \citet{Zhang}.
+We considered the deep convolutional embedded clustering (DCEC) by \citet{Guo2017} as well as the mixture of autoencoders (MIXAE) method introduced by \citet{Zhang}.
 While we were able to reproduce the authors' results on their data, the DCEC algorithm proved unable to cluster AT-TPC in our implementation.
@@
 {\color{blue} This section ends quite abruptly -MPK}
+
 \section{Results and Discussions}\label{sec:results}
@@
 
+From our analysis it is clear that we are a ways off from a fully automated unsupervised method for track classification.
+It is also clear that there is significant potential in both avenues explored in this work.
+In particular, the surprising ability of the K-means algorithm in picking out clear proton clusters from the VGG16 latent space lends itself well to an exploratory phase of analysis.
+The consistency of the predictive quality of the K-means model also means that
+
 \appendix  

### 2020-06-12 doc/Manuscript/article.tex
Updated abstract -MPK

@@
 This article presents novel applications of machine learning methods to the problem of event separation in an active target detector, the Active-Target Time Projection Chamber (AT-TPC) \cite{Bradt2017}.
-Machine learning presents an interesting avenue for researchers operating the AT-TPC, as traditional analysis methods of AT-TPC data are computationally expensive and have to fit all tracks against the event type of interest.
+%Machine learning presents an interesting avenue for researchers operating the AT-TPC, as traditional analysis methods of AT-TPC data are computationally expensive and have to fit all tracks against the event type of interest.
 The latter presents a considerable challenge when either the space of reactions is not known prior to the analysis, or the event type of interest is not the most abundant.
-
-The application of various machine learning algorithms to the analysis of two-dimensional projections of particle tracks from a resonant proton scattering experiment on $^{46}$Ar is introduced.
-We explore the performance of autoencoder neural networks and a pre-trained VGG16 \cite{Simonyan2014} convolutional neural network on two tasks: a semi-supervised classification task and the unsupervised clustering of particle tracks.
-On the semi-supervised task, we find that a logistic regression classifier trained on small labelled subsets of the latent space of these models perform very well, both on simulated and real data.
-On the clustering task, we find that a $K$-means algorithm applied to the simulated data in the VGG16 latent space forms almost perfect clusters.
+The application of unsupervised clustering  algorithms to the analysis of two-dimensional projections of particle tracks from a resonant proton scattering experiment on $^{46}$Ar is introduced.
+We explore the performance of autoencoder neural networks and a pre-trained VGG16 \cite{Simonyan2014} convolutional neural network.
+%on two tasks: a semi-supervised classification task and the unsupervised clustering of particle tracks. 
+%On the semi-supervised task, we find that a logistic regression classifier trained on small labelled subsets of the latent space of these models perform very well, both on simulated and real data.
+On the clustering task,
+We find that a $K$-means algorithm applied to the simulated data in the VGG16 latent space forms almost perfect clusters.
 Additionally, the VGG16+K-means approach finds high purity clusters of proton events for real experimental data.
 We also explore the application of neural networks to clustering by implementing a mixture of autoencoders algorithm, with improved results on real data.
-However, the neural network clustering suffers from stability issues necessitating further investigations into this approach.
+%
+%However, the neural network clustering suffers from stability issues. %necessitating further investigations into this approach.
 
@@
 Traditional methods were used to perform each of these tasks, and ultimately extract the scientific information, but there were severe limitations and high computational costs that become prohibitive in data sets larger by an order of magnitude.
+% Begin added by DB (06/10/2020)
+The cleaning was performed using a combination of linear and circular Hough transforms on a 2-dimensional projection of the tracks \cite{Bradt2017}.
+The following filtering and fitting phases were performed simultaneously, by defining the cost function to the fitting algorithm as a sum of 3 $\chi^2$ components based on i) the position of the track in space, ii) the energy deposited on each pixel of the sensor plane, and iii) the location of the vertex of the reaction.
+While various fitting algorithms were tested, the most accurate was a Monte-Carlo algorithm that explored the 6-dimensional phase space of the particle's kinematics parameters, reducing it progressively at each iteration step until the desired accuracy was reached \cite{Bradt2017}.
+Although this algorithm ended up being the most accurate, it is extremely costly computationally because of the very large number of simulated tracks needed for each event.
+The filtering was performed by setting limits to the $\chi^2$ distributions, below which the events were assigned as proton scattering.
+This is a very inefficient method because it requires performing the fitting for all events, including those that are not of interest.
+Pioneering work on event identification using ML methods, namely the use of a pre-trained CNN, later showed that the filtering phase would better be performed using this type of technique \cite{Kuchera2019}.
+% End added by DB (06/10/2020)
 In addition, \cite{Kuchera2019} demonstrated that the purity and statistics of the data are improved with the use of ML.
@@
 
+Building on the work by \citet{Kuchera2019}, we also use the VGG16 pre-trained network as a baseline for the clustering performance.
+We begin by considering a classical K-means approach to clustering.
+However, the output from the VGG16 network is very high dimensional which traditionally conflicts with good K-means results.
+With output vectors in $\R^{8192}$ the ratio between $L_p$ distances goes to one with increasing dimensionality, as shown by \citet{Aggarwal}.
+However, one of the central assumptions in the authors' finding is that the elements are uniformly distributed in the latent space.
+They note that it is then possible that if the class information lies in some sub-space of the latent data, $L_p$-norms may provide useful clusterings even in high dimensional spaces.
+To investigate this, we perform a $L_2$-norm based K-means clustering using the full VGG16 representation and the $10^2$ first principal components only.
+
+%{\color{blue} Are there supposed to be results here? -MPK}
+%\subsection{K-means}
+
+%We begin by investigating the K-means clustering algorithm on the VGG16 latent space. 
+The VGG16 model is pre-trained on the imagenet dataset, and subsequently applied to the AT-TPC events without the classification layers.
+This creates a set of vectors $\boldsymbol{x} \in \R^{8192}$ as our new representation of the events.
+To cluster we use \lstinline{scikit-learn} implementation of the K-means algorithm, with default parameters \cite{Pedregosa2011}.
+
 \subsection{Deep clustering: Mixture of autoencoders}\label{sec:mixae}
 
-As an alternative to using a pre-trained model in conjunction with a clustering algorithm, we have also considered bespoke end-to-end clustering models.
-These models are optimized specifically to the AT-TPC data with the objective of clustering.
-We considered the deep convolutional embedded clustering (DCEC) by \citet{Guo2017} as well as the mixture of autoencoders (MIXAE) method introduced by \citet{Zhang}.
+As an alternative to relying on a pre-trained model, we also consider bespoke end-to-end clustering models, which are specifically trained on the AT-TPC data.
+{\color{cyan} [Ryan: I think everything after this point in this paragraph should either be removed or go in a discussion at the end.
+Because it is entirely about DCEC, which isn't one of the two main methods we are presenting and more immediately isn't MIXAE, which is what this section is about.
+Or, this section just needs to be presented as something slightly more general, but right now, ``mixture of autoencoders'' is in the title.]} We considered the deep convolutional embedded clustering (DCEC) by \citet{Guo2017} as well as the mixture of autoencoders (MIXAE) method introduced by \citet{Zhang}.
 While we were able to reproduce the authors' results on their data, the DCEC algorithm proved unable to cluster AT-TPC in our implementation.
@@
 The MIXAE algorithm clusters data by pairing a set of encoder-decoder neural networks (autoencoder) with an auxiliary assignment network: each autoencoder constructs a representation of a given event while the assignment network assigns a set of cluster probabilities to the event.
-As noted by \citet{Zhang}, the MIXAe model assumes that the clusters are roughly similar in volume in the data and that one has some reasonable estimate on the number of clusters present.
+As noted by \citet{Zhang}, the MIXAe model assumes that the clusters are roughly similar in volume in the data and that one has some reasonable estimate on the number of clusters present {\color{cyan} [Ryan: Is this true of our data?
+I think if these assumptions are going to be mentioned, then we need to say that our data meets the criteria.}.
 The architecture is portrayed in figure \ref{fig:mixae}, tapered boxes denote a direction of compression in the network components.
@@
 \noindent For a more formal consideration of these heuristics see \citet{Zhang}.
+
+{\color{cyan} [Ryan: When I reach the end of this section, I'm still not quite sure its obvious what MIXAE is doing.
+I don't necessarily know how to reword things yet, but I think a more intuitive initial explanation of what it does could be helpful.}
 
@@
 The  ${}^{46}$Ar experiment does, however, provide a useful example where we can then explore unsupervised techniques.
-In this chapter, we explore the application of clustering techniques to events represented in latent spaces.
 
-We begin by exploring a naive K-means approach on the latent space of a pre-trained network.
-Subsequently, we investigate other clustering methods and the autoencoder based clustering algorithm outlined in section \ref{sec:mixae}.
-{\color{blue} Move methods to pervious section -MPK}
-\subsection{Pre-trained model with K-means}
-Building on the work by \citet{Kuchera2019}, we also use the VGG16 pre-trained network as a baseline for the clustering performance.
-We begin by considering a classical K-means approach to clustering.
-However, the output from the VGG16 network is very high dimensional which traditionally conflicts with good K-means results.
-With output vectors in $\R^{8192}$ the ratio between $L_p$ distances goes to one with increasing dimensionality, as shown by \citet{Aggarwal}.
-However, one of the central assumptions in the authors' finding is that the elements are uniformly distributed in the latent space.
-They note that it is then possible that if the class information lies in some sub-space of the latent data, $L_p$-norms may provide useful clusterings even in high dimensional spaces.
-To investigate this, we perform a $L_2$-norm based K-means clustering using the full VGG16 representation and the $10^2$ first principal components only.
+We begin by exploring the results of applying a K-means approach on the latent space of a pre-trained network.
+Subsequently, we investigate the performance of the MIXAE algorithm outlined in section \ref{sec:mixae}.
 
-%{\color{blue} Are there supposed to be results here? -MPK}
-%\subsection{K-means}
+\subsection{K-means clustering on the VGG16 latent space}
 
-%We begin by investigating the K-means clustering algorithm on the VGG16 latent space. 
-The VGG16 model is pre-trained on the imagenet dataset, and subsequently applied to the AT-TPC events without the classification layers.
-This creates a set of vectors $\boldsymbol{x} \in \R^{8192}$ as our new representation of the events.
-To cluster we use \lstinline{scikit-learn} implementation of the K-means algorithm, with default parameters \cite{Pedregosa2011}.
 The results of the clustering runs are included in table \ref{tab:clstr_vgg}.
@@
 A high-quality cluster will, in addition to being pure, also capture most entries the class represented by the cluster.
-For example, consider the row corresponding to the proton class in figure \ref{fig:clster_confmat}.
+For example, consider the row corresponding to the proton class in figure \ref{fig:clster_confmati_filt}.
 The column corresponding to the largest entry in the proton row has zero other predicted classes in it.
 From this, we conclude that the proton cluster is a high quality, high purity cluster.
+
+This high quality cluster also appears in the clustering of raw data.
+From figure \ref{fig:clster_confmati_raw} we observe that there is a high purity proton cluster.
+In contrast to the filtered data we observe that the deterioration in performance can largely be ascribed to the algorithm creating a proton + other cluster and a carbon + other cluster.
 
@@
 PCA is a common techinque to find the significant variations in data by projecting the data along a subset of its covariance matrix eigenvectors \cite{Marsland2009}
-}{\color{green}(please define)} {\color{orange} does this work?} on the latent space of the VGG16 model.
+} on the latent space of the VGG16 model.
 This is done to estimate to what degree the class separating information is encoded in the entirety of the latent space, or in some select regions.
@@
 
-Furthermore, we wish to characterize further the clusters presented in figure \ref{fig:clster_confmat}.
+Furthermore, we wish to characterize further the clusters presented in the matrix in figures \ref{fig:clster_confmati_filt} and \ref{fig:clster_confmati_raw}.
 To achieve this, we sample from the proton samples belonging to different clusters for the filtered and full data.
@@
 \caption[Filtered proton samples by cluster belonging]{Illustrating a sample of proton events from different K-means clusters from the filtered dataset.
-Each row belongs to a single cluster corresponding to the filtered confusion matrix in figure \ref{fig:clster_confmat}}\label{fig:filtered_vgg_clster_repr}
+Each row belongs to a single cluster corresponding to the filtered confusion matrix in figure \ref{fig:clster_confmati_filt}}\label{fig:filtered_vgg_clster_repr}
 \end{figure}
@@
 \caption[Full proton samples by cluster belonging]{Illustrating a sample of proton events from different K-means clusters from the raw dataset.
-Each row belongs to a single cluster corresponding to the full confusion matrix in figure \ref{fig:clster_confmat}}\label{fig:full_vgg_clster_repr}
+Each row belongs to a single cluster corresponding to the full confusion matrix in figure \ref{fig:clster_confmati_raw}}\label{fig:full_vgg_clster_repr}
 \end{figure}
@@
 
-\subsection{Autoencoder clustering}
+\subsection{MIXAE clustering results}
 
-\begin{table}[H]
-\centering 
-\caption[MIXAE clustering performance]{MIXAE clustering performance on the ${}^{46} Ar$ experimental data.
-In contrast with the VGG-16 + K-means approach we observe  significant variations in performance.}\label{tab:clstr_vgg}
-\input{./plots/mixae_clustering.tex}
-\end{table}
-
-In the previous section we demonstrated a potential for clustering of events from an AT-TPC experiment.
+In the previous section we demonstrated a rather naive clustering technique for AT-TPC event classification.
 To build on this result we will in this section explore the application of the mixture of autoencoders (MIXAE) algorithm introduced in section \ref{sec:mixae}.
@@
 
+\begin{table}[H]
+\centering 
+\caption[MIXAE clustering performance]{MIXAE clustering performance on the ${}^{46} Ar$ experimental data.
+In contrast with the VGG-16 + K-means approach we observe  significant variations in performance.}\label{tab:clstr_vgg}
+\input{./plots/mixae_clustering.tex}
+\end{table}
+
 \subsection{Simulated AT-TPC data}
@@
 We elaborate on the $ARI$ score in section \ref{sec:unsupervised_perf}.
-As mentioned in section \ref{sec:mixae} the batch-entropy has a second minimum when the cluster confidences are near equal.
-It is this behavior we observe in figure \ref{fig:mixae_sim}.
 
-\subsection{Filtered AT-TPC data}
-
-We repeat the optimization steps in the previous section for the filtered AT-TPC data.
+We repeat the optimization steps in the previous paragraph for the filtered AT-TPC data.
 The exception being that we allow the algorithm to train on the labelled samples.
@@
 \noindent The results of the runs are included in figure \ref{fig:mixae_clean}.
-We observe ghat the highest performing models reach an $ARI > 0.5$, which is higher than the performance achieved by the K-means algorithm applied to the VGG16 latent space.
+We observe that the highest performing models reach an $ARI > 0.5$, which is higher than the performance achieved by the K-means algorithm applied to the VGG16 latent space.
 
-\subsection{Full AT-TPC data}
-
-As in the two previous sections, we repeat the same procedure of iterative grid searches on the MIXAE loss-weights.
-Each configuration is re-run a total of $N=10$ times to capture fluctuations in the performance before a final selection is made on the hyperparameters.
-
-\noindent As with the previous datasets we include a plot of the loss curves and performance measures for $N=10$ runs with the same loss-weight parameters, shown in figure \ref{fig:mixae_full}.
+We repeated the same procedure described in the two previous paragraphs to the raw dataset.
+Each configuration was re-run a total of $N=10$ times to capture fluctuations in the performance before a final selection is made on the hyperparameters.
 
@@
 \centering
+	\includegraphics[width=\textwidth]{custom_work/Filtered_mixae_conf_mat.pdf}
+\end{figure}
 
-	\includegraphics[width=\textwidth]{custom_work/Filtered_mixae_conf_mat.pdf}
-
-\end{figure}
 \begin{figure}[H]
 \centering
-
 	\includegraphics[width=\textwidth]{custom_work/Raw_mixae_conf_mat.pdf}
-
 \caption[MIXAE - confusion matrices]{Confusion matrices for the MIXAE clustering algorithm on filtered and full AT-TPC events.
@@
 In particular, the surprising ability of the K-means algorithm in picking out clear proton clusters from the VGG16 latent space lends itself well to an exploratory phase of analysis.
-The consistency of the predictive quality of the K-means model also means that
+Another surprising facet of the K-means clustering is its consistent performance.
+As shown in table \ref{tab:clstr_vgg}, the variance is zero for the performance metrics.
+This result indicates that the clusters are very clearly defined in the VGG16 latent space.
+As we see from the non-proton clusters if figures \ref{fig:clster_confmati_filt} and \ref{fig:clster_confmati_raw}  that does not necessarily imply that the physical signals are correspondingly clear.
+
+Additionally, a clear disadvantage to the K-means method is that the researcher has to specify the number of clusters in the data themselves.
+Each experiment then has to be considered in light of possible reaction channels to determine a sensible number of clusters for this approach.
+
+The same weakness is present in the MIXAE implementation.
+Additionally, while it showed better optimal performance than the K-means method, it struggles with inconsistent results.
+This inconsistency was also notably not evident from the unsupervised training-objectives of the model.
+These two factors conspire to limit its immediate applicability.
+
+In summary, we recommend that researchers performing unsupervised track classification begin with an implementation of the VGG16 + K-means approach.
+For future work, it is worth investigating whether the adaptations of bespoke models like the MIXAE algorithm will allow better performance at no significant cost to consistency.
+Especially considering the flexibility to use 3D input representations or auxiliary physical goals in the optimization procedure.
+
+{\color{orange}[RS Another abrupt end to a paragraph. Could probably use some rounding out.]}
 

### 2020-06-12 doc/Manuscript/article.tex
Updates from Overleaf

@@
 
-\subsubsection{Clustering on latent spaces}
+\subsection{Clustering on latent spaces}
 
@@
 Because it is entirely about DCEC, which isn't one of the two main methods we are presenting and more immediately isn't MIXAE, which is what this section is about.
-Or, this section just needs to be presented as something slightly more general, but right now, ``mixture of autoencoders'' is in the title.]} We considered the deep convolutional embedded clustering (DCEC) by \citet{Guo2017} as well as the mixture of autoencoders (MIXAE) method introduced by \citet{Zhang}.
-While we were able to reproduce the authors' results on their data, the DCEC algorithm proved unable to cluster AT-TPC in our implementation.
-However, this provides valuable insight as the seeds of the clusters are constructed by a K-means algorithm.
-This insight contrasts with our positive results from applying a pre-trained model with k-means and highlights potentially significant differences in models trained on a supervised or unsupervised objective for clustering tasks in nuclear physics.
+Or, this section just needs to be presented as something slightly more general, but right now, ``mixture of autoencoders'' is in the title.]}
 
 The MIXAE algorithm clusters data by pairing a set of encoder-decoder neural networks (autoencoder) with an auxiliary assignment network: each autoencoder constructs a representation of a given event while the assignment network assigns a set of cluster probabilities to the event.
-As noted by \citet{Zhang}, the MIXAe model assumes that the clusters are roughly similar in volume in the data and that one has some reasonable estimate on the number of clusters present {\color{cyan} [Ryan: Is this true of our data?
+{\color[orange]{clear sentences about what the model does}} As noted by \citet{Zhang}, the MIXAE model assumes that the clusters are roughly similar in volume in the data and that one has some reasonable estimate on the number of clusters present {\color{cyan} [Ryan: Is this true of our data?
 I think if these assumptions are going to be mentioned, then we need to say that our data meets the criteria.}.
@@
 
-In addition to the results presented in this section, we performed clustering with a number of different algorithms included in the \lstinline{scikit-learn} package.
-None of them provided any notable differences from the K-means results or were significantly worse.
-Notably, the DBSCAN algorithm \cite{Ester96adensity-based}\cite{Bergstra2012} failed to provide any useful clustering results.
-We find this important as one of the significant drawbacks of K-means, and the deep clustering algorithm presented in section \ref{sec:mixae}, is that they are all dependent on pre-determining the number of clusters.
-This is not the case for DBSCAN.
-
 \subsection{MIXAE clustering results}
@@
 
+{\color{orange}[move to appendix]}
 In the MIXAE algorithm the hyper-parameters to adjust are all the ordinary parameters that we introduced in table \ref{tab:convae_hyperparams}.
@@
 Each configuration was re-run a total of $N=10$ times to capture fluctuations in the performance before a final selection is made on the hyperparameters.
+In contrast to the previous two results the MIXAE algorith very rarely converges to a high performig configuration.
+In fact, most runs are indistinguishable from random chance clusterings.
 
@@
 From figure \ref{fig:mixae_confmat} we see that the MIXAE algorithm creates two proton clusters for the filtered data, and places about fifty per-cent of the proton events in a cluster with the amorphous "other" events.
-We extract some proton events from these clusters to inspect whether systematic differences occur.
+
+\subsection{Alternative approaches}
+
+In addition to the results presented in this section, we performed clustering with a number of different algorithms included in the \lstinline{scikit-learn} package.
+None of them provided any notable differences from the K-means results or were significantly worse.
+Notably, the DBSCAN algorithm \cite{Ester96adensity-based}\cite{Bergstra2012} failed to provide any useful clustering results.
+We find this important as one of the significant drawbacks of K-means, and the deep clustering algorithm presented in section \ref{sec:mixae}, is that they are all dependent on pre-determining the number of clusters.
+This is not the case for DBSCAN.
+
+{\color{orange}[binding sentence]}
+We considered the deep convolutional embedded clustering (DCEC) by \citet{Guo2017} as well as the mixture of autoencoders (MIXAE) method introduced by \citet{Zhang}.
+While we were able to reproduce the authors' results on their data, the DCEC algorithm proved unable to cluster AT-TPC in our implementation.
+However, this provides valuable insight as the seeds of the clusters are constructed by a K-means algorithm.
+This insight contrasts with our positive results from applying a pre-trained model with k-means and highlights potentially significant differences in models trained on a supervised or unsupervised objective for clustering tasks in nuclear physics.
 
@@
 From our analysis it is clear that we are a ways off from a fully automated unsupervised method for track classification.
-It is also clear that there is significant potential in both avenues explored in this work.
+{\color{orange}[flip this to positive ]} It is also clear that there is significant potential in both avenues explored in this work.
 In particular, the surprising ability of the K-means algorithm in picking out clear proton clusters from the VGG16 latent space lends itself well to an exploratory phase of analysis.
@@
 
-Additionally, a clear disadvantage to the K-means method is that the researcher has to specify the number of clusters in the data themselves.
+%Additionally, a clear disadvantage to the K-means method is that the researcher has to specify the number of clusters in the data themselves.
 Each experiment then has to be considered in light of possible reaction channels to determine a sensible number of clusters for this approach.
 
-The same weakness is present in the MIXAE implementation.
+%The same weakness is present in the MIXAE implementation.
 Additionally, while it showed better optimal performance than the K-means method, it struggles with inconsistent results.

### 2020-06-12 doc/Manuscript/article.tex
Updates from Overleaf

@@
 This article presents novel applications of machine learning methods to the problem of event separation in an active target detector, the Active-Target Time Projection Chamber (AT-TPC) \cite{Bradt2017}.
+The overarching goal is to identify classes of events in the early stages of data processing, before 
 %Machine learning presents an interesting avenue for researchers operating the AT-TPC, as traditional analysis methods of AT-TPC data are computationally expensive and have to fit all tracks against the event type of interest.
@@
 We also explore the application of neural networks to clustering by implementing a mixture of autoencoders algorithm, with improved results on real data.
-%
-%However, the neural network clustering suffers from stability issues. %necessitating further investigations into this approach.
-
-%In this work we introduce the application of convolutional autoencoder neural networks to the analysis of two-dimensional projections of particle tracks from a resonant proton scattering experiment on ${}^{46}$Ar.
-%The data we analyze were recorded by an active target time-projection chamber (AT-TPC).
-Machine learning presents an interesting avenue for researchers operating an AT-TPC, as traditional analysis methods of AT-TPC data are both computationally expensive and fit all particle tracks against the event type of interest.
-The latter presents a considerable challenge when the space of reactions is not known prior to the analysis.
-
-%We explore the performance of the autoencoder neural networks and a pre-trained VGG16 \cite{Simonyan2014} convolutional neural network on two tasks: a semi-supervised classification task and the unsupervised clustering of particle tracks.
-On the semi-supervised task, we find that a logistic regression classifier trained on small labelled subsets of the latent space of these models perform very well.
-On simulated data these classifiers achieve an $f1$ score \cite{Chinchor1992} of $f1>0.95$.
-The VGG16 latent classifier achieves this result with as few as $N=100$ samples, as does the convolutional autoencoder when trained on the VGG16 representations of the particle tracks.
-On real data, pre-processed with noise filtering, the same models achieve an $f1>0.7$.
-For unfiltered real data the models achieve an $f1$ value larger than $0.6$.
-Both of the previous results were found with the classifiers trained on $N=100$ samples.
-Furthermore, we found that the autoencoder model reduces the variability in the identification of proton events by $64\%$ from the benchmark logistic regression classifier trained on the VGG16 latent space on real experimental data.
 
@@
 The MIXAE algorithm clusters data by pairing a set of encoder-decoder neural networks (autoencoder) with an auxiliary assignment network: each autoencoder constructs a representation of a given event while the assignment network assigns a set of cluster probabilities to the event.
-{\color[orange]{clear sentences about what the model does}} As noted by \citet{Zhang}, the MIXAE model assumes that the clusters are roughly similar in volume in the data and that one has some reasonable estimate on the number of clusters present {\color{cyan} [Ryan: Is this true of our data?
-I think if these assumptions are going to be mentioned, then we need to say that our data meets the criteria.}.
+{\color{orange}[clear sentences about what the model does]} .
 The architecture is portrayed in figure \ref{fig:mixae}, tapered boxes denote a direction of compression in the network components.
@@
 
-\noindent For a more formal consideration of these heuristics see \citet{Zhang}.
+\noindent For a more formal consideration of these heuristics and the assumptions made on the data by this model see \citet{Zhang}.
 
@@
 \section{Results and Discussions}\label{sec:results}
-\begin{itemize}
-\item Clear out the "methods" like parts.
-\item merge 5.1 and 5.2 - keeping point 1 in mind.
-\item merge 5.3 5.4 and 5.6 and create table from figure 10, 11 and 12. 
-\item go over 5.7 - mention instability in MIXAE results, peak performance of MIXAE is good - consistency is not. 
-\item remove fig. 14 and fig 15.
-\end{itemize}
 

### 2020-06-12 doc/Manuscript/article.tex
Updates from Overleaf

@@
 This article presents novel applications of machine learning methods to the problem of event separation in an active target detector, the Active-Target Time Projection Chamber (AT-TPC) \cite{Bradt2017}.
-The overarching goal is to identify classes of events in the early stages of data processing, before 
+The overarching goal is to identify classes of events in the early stages of the data analysis, thereby improving its efficiency by limiting the computationally expensive processing to events of interest only, which are often not the most abundant.
 %Machine learning presents an interesting avenue for researchers operating the AT-TPC, as traditional analysis methods of AT-TPC data are computationally expensive and have to fit all tracks against the event type of interest.

### 2020-06-15 doc/Manuscript/article.tex
Updates from Overleaf
Ryan: Rewording of section 4.5

@@
 
-\title{Latent Variable Clustering for Classification of Events in the AT-TPC}
+\title{Machine Learning Approaches  for Classification of Events in Active-Target Time Projection Chamber Experiments}
 
@@
 
-%On the clustering task, we found that a $K$-means algorithm applied to the simulated data in the VGG16 latent space forms almost perfect clusters, with an adjusted rand index \cite{Hubert1985} ($ARI$) $ > 0.8$.
-Additionally, the VGG16+K-means approach finds high purity clusters of proton events for real experimental data.
-We also explore the application of neural networks to clustering by implementing a mixture of autoencoders algorithm.
-With this model we improved clustering performance on the real experimental data from an $ARI = 0.17$ to an $ARI = 0.40$.
-However, the neural network clustering suffers from stability issues necessitating further investigations into this approach.
-
 \end{abstract}
@@
 \section{Introduction}\label{sec:intro}
-
-%
-%\begin{itemize}
-%    \item Overall goals for the AT-TPC
-%    \item Nuclear physics motivation for clustering-based (unsupervised) %methods
-%   \item Motivation from ML approach
-%   
-%\end{itemize}
 
@@
 
-Building on the work by \citet{Kuchera2019}, we also use the VGG16 pre-trained network as a baseline for the clustering performance.
+{\color{cyan} [Ryan: I think there are a lot of words in the following two paragraphs that don't need to exist.
+Essentially, these paragraphs are a repeat of the last paragraph and parts of the last subsection, but with one or two extra details added in (e.g.
+using the top 100 principle components).
+As such, I would remove them and add the necessary details to the previous paragraph (or replace them with a small paragraph).]} Building on the work by \citet{Kuchera2019}, we also use the VGG16 pre-trained network as a baseline for the clustering performance.
 We begin by considering a classical K-means approach to clustering.
@@
 
-As an alternative to relying on a pre-trained model, we also consider bespoke end-to-end clustering models, which are specifically trained on the AT-TPC data.
-{\color{cyan} [Ryan: I think everything after this point in this paragraph should either be removed or go in a discussion at the end.
-Because it is entirely about DCEC, which isn't one of the two main methods we are presenting and more immediately isn't MIXAE, which is what this section is about.
-Or, this section just needs to be presented as something slightly more general, but right now, ``mixture of autoencoders'' is in the title.]}
+As an alternative to relying on a pre-trained model, we also consider the MIXAE algorithm {\color{cyan} [Ryan: citation]}, a bespoke end-to-end clustering model which is specifically trained on the AT-TPC data.
 
-The MIXAE algorithm clusters data by pairing a set of encoder-decoder neural networks (autoencoder) with an auxiliary assignment network: each autoencoder constructs a representation of a given event while the assignment network assigns a set of cluster probabilities to the event.
+The MIXAE model comprises several autoencoders, each of which corresponds to a cluster.
+The number of autoencoders, and thus the number of clusters, has to be determined beforehand.
+Each autoencoder constructs a latent representation of a given example.
+Those representations are used as inputs to an auxiliary network which assigns probabilities the clusters, indicating the likelihood that the given example belongs to each cluster.
+Examples are then assigned to the cluster with the highest probability.
+{\color{cyan} [Ryan: A description of the objective functions may or may not be worth including here, but I don't know what they are.]} The architecture is portrayed in Figure \ref{fig:mixae}, tapered boxes denote a direction of compression in the network components.
+
+% The MIXAE algorithm clusters data by pairing a set of encoder-decoder neural networks (autoencoder) with an auxiliary assignment network: each autoencoder constructs a representation of a given event while the assignment network assigns a set of cluster probabilities to the event.
 {\color{orange}[clear sentences about what the model does]} .
@@
 
-As the MIXAE model does not have access to ground truth labels the training consists of applying simple heuristics that constitute a necessary, but not sufficient, fulfilment for a high-quality MIXAE model given the assumptions made of the model.
-The heuristics can be stated as:
+The MIXAE algorithm relies on a few simple assumptions which are necessary, but not sufficient, for producing a high-quality model. The assumptions can be stated as: 
 \begin{enumerate}
-	\item  Given a clustering assignment to a particular autoencoders' representation that autoencoders' reconstruction should be similar to the input event. 
-	\item Within a batch of events presented to the model, there should be a spread in the clustering assignment between all clusters. 
-	\item Each clustering prediction should be as close to a strong prediction as possible, i.e. assigning high probabilities is preferable to weak assignments.
+	\item  If an example is assigned to a particular cluster, the corresponding autoencoder's reconstruction should be accurate. 
+	\item Within a batch of examples presented to the model, assignments should spread across all clusters. 
+	\item Each clustering prediction should be as strong as possible, i.e. assigning high probabilities is preferable to weak assignments.
 \end{enumerate}
 
-\noindent For a more formal consideration of these heuristics and the assumptions made on the data by this model see \citet{Zhang}.
-
-{\color{cyan} [Ryan: When I reach the end of this section, I'm still not quite sure its obvious what MIXAE is doing.
-I don't necessarily know how to reword things yet, but I think a more intuitive initial explanation of what it does could be helpful.}
+\noindent {\color{cyan} [Ryan: Still needs a concluding sentence or two here.]} For a more formal consideration of these heuristics and the assumptions made on the data by this model see \citet{Zhang}.
 

### 2020-06-17 doc/Manuscript/article.tex
Updates from Overleaf

@@
-%robert was here sup robert -MIchelle was also here
-
 %\documentclass[preprint,12pt]{elsarticle}
@@
 The 2D representation of events was computed by viewing the event as a point cloud in the charge time-series and summing over the time-axis.
-Additionally, we log-transform and normalize the data to the interval $[0, 1]$ for each point.
+Additionally, we log-transform and normalize the data to the interval $[0, 1]$ for each point {\color{blue} are you sure we log transformed?
+were you given peaks data or the 2D projections?
+-MPK}.
 The final representation for each event is a $128 \times 128$ matrix with each element ranging from $0$ to $1$.
@@
 The traditional event selection process involved a Monte-Carlo $\chi^2$ minimization method, however there was no well-defined method to quantify the effectiveness of the event selection from the thresholds applied to these distributions.
-{\color{blue} I think we discussed this method earlier, but I need to verify. otherwise needs more explanation -MPK}
+{\color{blue} Needs more explanation earlier in paper -MPK}
 In addition, the selection task produced a binary result only, ``good" or``bad", relative to the event of interest, while the data clearly contained more classes of events%, if only because of the presence of carbon in the target (from the isobutane gas).
@@
 %We begin by investigating the K-means clustering algorithm on the VGG16 latent space. 
-The VGG16 model is pre-trained on the imagenet dataset, and subsequently applied to the AT-TPC events without the classification layers.
+The VGG16 model is pre-trained on the ImageNet dataset, and subsequently applied to the AT-TPC events without the classification layers.
 This creates a set of vectors $\boldsymbol{x} \in \R^{8192}$ as our new representation of the events.
@@
 
-As an alternative to relying on a pre-trained model, we also consider the MIXAE algorithm {\color{cyan} [Ryan: citation]}, a bespoke end-to-end clustering model which is specifically trained on the AT-TPC data.
+As an alternative to relying on a pre-trained model, we also consider the MIXAE algorithm \cite{Zhang}, a bespoke end-to-end clustering model which is specifically trained on the AT-TPC data.
 
@@
 Each autoencoder constructs a latent representation of a given example.
-Those representations are used as inputs to an auxiliary network which assigns probabilities the clusters, indicating the likelihood that the given example belongs to each cluster.
+Those representations are used as inputs to an auxiliary network which assigns probabilities to the clusters, indicating the likelihood that the given example belongs to each cluster.
 Examples are then assigned to the cluster with the highest probability.
-{\color{cyan} [Ryan: A description of the objective functions may or may not be worth including here, but I don't know what they are.]} The architecture is portrayed in Figure \ref{fig:mixae}, tapered boxes denote a direction of compression in the network components.
+The models' objective can then be de-composed in three parts:
+\begin{enumerate}
+    \item Each batch should contain different classes
+    \item Each predicted cluster assignment should be confident. 
+    \item The reconstructions should be good. But as they are weighted by the cluster assignment, the importance of the reconstruction is proportional to the confidence in the cluster predictions.
+\end{enumerate}
+
+{\color{cyan} [Ryan: A description of the objective functions may or may not be worth including here, but I don't know what they are.]} The architecture is portrayed in Figure \ref{fig:mixae}, wherein tapered boxes denote a direction of compression in the network components.
 
@@
 
-\noindent {\color{cyan} [Ryan: Still needs a concluding sentence or two here.]} For a more formal consideration of these heuristics and the assumptions made on the data by this model see \citet{Zhang}.
+\noindent {\color{cyan} [Ryan: Still needs a concluding sentence or two here.]} For a more formal consideration of these heuristics, the model objective aand the assumptions made on the data by this model see \citet{Zhang}.
 
@@
 We begin by exploring the results of applying a K-means approach on the latent space of a pre-trained network.
-Subsequently, we investigate the performance of the MIXAE algorithm outlined in section \ref{sec:mixae}.
+Subsequently, we investigate the performance of the MIXAE algorithm as outlined in section \ref{sec:mixae}.
 
@@
 In addition to the performance measures reported in table \ref{tab:clstr_vgg}, it is interesting to observe which samples are being wrongly assigned.
-We achieve this by tabulating the assignments of samples relative to their ground truth labels.
+To investigate this problem we tabulate the assignments of samples relative to their ground truth labels.
 From these tables, we can infer which classes are more or less entangled with others.
-We tabulate the results for each dataset is in figure \ref{fig:clster_confmat}.
+The results for each dataset is shown in figures \ref{fig:clster_confmat_sim} \ref{fig:clster_confmati_filt} and \ref{fig:clster_confmati_raw}.
 We observe that the proton class is consistently assigned in a pure cluster.
@@
 
-We repeat this analysis using a Principle Component Analysis (PCA) dimensionality reduction \footnote{
+We repeat this analysis using a Principle Component Analysis (PCA) dimensionality reduction\footnote{
 PCA is a common techinque to find the significant variations in data by projecting the data along a subset of its covariance matrix eigenvectors \cite{Marsland2009}
@@
 
-Furthermore, we wish to characterize further the clusters presented in the matrix in figures \ref{fig:clster_confmati_filt} and \ref{fig:clster_confmati_raw}.
-To achieve this, we sample from the proton samples belonging to different clusters for the filtered and full data.
+Furthermore, we wish to further characterize the clusters presented in the matrix in figures \ref{fig:clster_confmati_filt} and \ref{fig:clster_confmati_raw}.
+To achieve this, we sample examples from the proton samples belonging to different clusters for the filtered and full data.
+We display these examples in figures \ref{fig:filtered_vgg_clster_repr} and \ref{fig:full_vgg_clster_repr} respectively.
+Form these figures we gleam that the more noisy proton events are being clustered together with the amorphous other class.
+This is in line with the intuitive expectation from viewing the clustering tables in figures \ref{fig:clster_confmati_filt} and \ref{fig:clster_confmati_raw}.
 
@@
 
-In the previous section we demonstrated a rather naive clustering technique for AT-TPC event classification.
+In the previous section we demonstrated a powerful, but rather naive, clustering technique for AT-TPC track identification.
 To build on this result we will in this section explore the application of the mixture of autoencoders (MIXAE) algorithm introduced in section \ref{sec:mixae}.
-Other algorithms for deep clustering based on autoencoder architectures were also implemented.
-These did not succeed in the clustering task and were omitted for brevity\footnote{The algoritms we implemented were the Deep Embedded Convolutional Clustering \cite{Guo2017} and its predecessor Deep Embedded Clustering \cite{Xie2016}}.
+For details on hyper-parameter tuning and the experimental procedure for training the MIXAE algorithm see appendix \ref{ax: mixae}
 
-{\color{orange}[move to appendix]}
-In the MIXAE algorithm the hyper-parameters to adjust are all the ordinary parameters that we introduced in table \ref{tab:convae_hyperparams}.
-In addition to those parameters we have the weighting of the loss terms: $\theta$, $\alpha$ and $\gamma$.
-These weighting parameters are attached to the reconstruction loss, sample entropy and batch-wise entropy respectively \cite{Zhang}.
-We focused on the tuning of the clustering hyper-parameters, and heuristically set the autoencoder hyper-parameters to a shallow $3\times3$ convolutional network.
-
-The parameters chosen for the autoencoders are listed in full in table \ref{tab:mixe_ae_hyperparams}.
+After hyper parameter tuning each highest performing model is re-run $10$ times. The results are listed in table \ref{tab:clstr_mixae}'
+We observe that while the algorithm can achieve very strong performance, the performance varies.
+In some cases the MIXAE model converges to a seemingly good configuration, based on its unsupervised training goals.
+However, when inspecting its clustering performance against labeleled data that seemingly good model does no better than random.
+This happens more frequently with raw data, indicating an interaction with the noise levels present in the events.
 
@@
 \caption[MIXAE clustering performance]{MIXAE clustering performance on the ${}^{46} Ar$ experimental data.
-In contrast with the VGG-16 + K-means approach we observe  significant variations in performance.}\label{tab:clstr_vgg}
+In contrast with the VGG-16 + K-means approach we observe  significant variations in performance.}\label{tab:clstr_mixae}
 \input{./plots/mixae_clustering.tex}
@@
 
-\subsection{Simulated AT-TPC data}
-
-Finally, for these parameters we re-ran the algorithm $N=10$ times to investigate the stability of the algorithm.
-The results are reported in figure \ref{fig:mixae_sim}.
-We observe that while the algorithm can achieve very strong performance, with an adjusted rand index (ARI) $ > 0.8$, it fluctuates strongly with repeated experiments.
-The $ARI$ is a clustering performance measure similar to accuracy, but adjusted for chance.
-We elaborate on the $ARI$ score in section \ref{sec:unsupervised_perf}.
-
-We repeat the optimization steps in the previous paragraph for the filtered AT-TPC data.
-The exception being that we allow the algorithm to train on the labelled samples.
-Beginning with a wide grid equal to the grid used for the simulated data we searched over all parameter configurations to find promising values.
-We then performed a local search around these values to pin-point the hyperparameter configuration.
-
-\noindent The results of the runs are included in figure \ref{fig:mixae_clean}.
-We observe that the highest performing models reach an $ARI > 0.5$, which is higher than the performance achieved by the K-means algorithm applied to the VGG16 latent space.
-
-We repeated the same procedure described in the two previous paragraphs to the raw dataset.
-Each configuration was re-run a total of $N=10$ times to capture fluctuations in the performance before a final selection is made on the hyperparameters.
-In contrast to the previous two results the MIXAE algorith very rarely converges to a high performig configuration.
-In fact, most runs are indistinguishable from random chance clusterings.
+As with the VGG16 + KMeans approach we wish to further characterize the clustering results.
+Taking the best performing MIXAE model on filtered and raw data we tabulate the clusters against their labels.
+These tables are present in figures \ref{fig:mixae_confmat_filtered} and \ref{fig:mixae_confmat_raw} for filtered and raw data, respectively.
 
@@
 	\includegraphics[width=\textwidth]{custom_work/Filtered_mixae_conf_mat.pdf}
+\caption{}\label{fig:mixae_confmat_filtered}
 \end{figure}
@@
 	\includegraphics[width=\textwidth]{custom_work/Raw_mixae_conf_mat.pdf}
-\caption[MIXAE - confusion matrices]{Confusion matrices for the MIXAE clustering algorithm on filtered and full AT-TPC events.
+\caption[MIXAE - confusion matrices]{Confusion matrices for the MIXAE clustering algorithm on raw AT-TPC events.
 The true labels indicate samples belonging to the p (proton), carbon (C), or other classes.
-}\label{fig:mixae_confmat}
+We observe that in }\label{fig:mixae_confmat_raw}
 \end{figure}
@@
 
-\section{MIXAE hyper-parameter tuning}
+\section{MIXAE hyper-parameter tuning}\label{ax: mixae}
+
+In the MIXAE algorithm the hyper-parameters to adjust are all the ordinary parameters that we introduced in table \ref{tab:convae_hyperparams}.
+In addition to those parameters we have the weighting of the loss terms: $\theta$, $\alpha$ and $\gamma$.
+These weighting parameters are attached to the reconstruction loss, sample entropy and batch-wise entropy respectively \cite{Zhang}.
+We focused on the tuning of the clustering hyper-parameters, and heuristically set the autoencoder hyper-parameters to a shallow $3\times3$ convolutional network.
+
+The parameters chosen for the autoencoders are listed in full in table \ref{tab:mixe_ae_hyperparams}.
 
@@
 The algorithm is trained on a subset of $60000$ of these samples, and we track performance on the remaining $20000$ events.
-Since there are then only three remaining hyperparameters we choose to perform a coarse grid-search as described in section \ref{sec:hyperparam_search_arch}.
+Since there are then only three remaining hyperparameters we choose to perform a coarse grid-search as described in section \ref{sec:hyperparam_search_arch}.Finally, for these parameters we re-ran the algorithm $N=10$ times to investigate the stability of the algorithm.
 

### 2020-06-17 doc/Manuscript/article.tex
Updates from Overleaf

@@
 
-\title{Machine Learning Approaches  for Classification of Events in Active-Target Time Projection Chamber Experiments}
+\title{Unsupervised Learning  for Identifying  Events in Active Target  Experiments}
 
@@
 The AT-TPC data we studied for this work was recorded as charge time-series for each of the  the $\sim10^4$ detector pads.
-In this representation, an event is a record of 512 time-buckets for each of the detector pads. In our analysis, we represent each event as a downsampled 2D projection.
-We chose to represent the data in 2D to facilitate the use of sophisticated image-recognition machine learning models.
-The 2D representation of events was computed by viewing the event as a point cloud in the charge time-series and summing over the time-axis.
-Additionally, we log-transform and normalize the data to the interval $[0, 1]$ for each point {\color{blue} are you sure we log transformed?
-were you given peaks data or the 2D projections?
--MPK}.
+In this representation, an event is a record of $512$ time-buckets for each of the $10^4$ detector pads. In our analysis, we represent each event as a downsampled 2D projection.
+We chose to represent the data in 2D to facilitate the use of sophisticated image-recognition machine learning models, and this data representation was shown to successfully classify the events of this experiment in a supervised setting \cite{Kuchera2019}.
+First, the time-series data was represented as a 3D point cloud, where each point contains the maximum charge in the time-series trace.
+We log-transform to map the charge to the interval $[0, 1]$ for each point, then standardize the data.
+in the charge time-series and summing over the time-axis.
+%
+Additionally, we log-transform and normalize the data to the interval $[0, 1]$ for each point.
 The final representation for each event is a $128 \times 128$ matrix with each element ranging from $0$ to $1$.
@@
 \subsection{Why machine learning}
-
+{\bf we need to rewrite these bullet points, MHJ}
 \begin{itemize}
@@
         \item The computational cost per event is too large given the size of the data-sets
-        \item The broken tracks, noisy environment creates bad fit statistics for otherwise useful events. End result is $f1 \sim 0.7 $
+        \item The broken tracks with noisy environment create bad fit statistics for otherwise useful events. End result is f$_1 \sim 0.7 $
         \item define a ``bad" fit as a non-proton event
@@
 Each autoencoder constructs a latent representation of a given example.
-Those representations are used as inputs to an auxiliary network which assigns probabilities to the clusters, indicating the likelihood that the given example belongs to each cluster.
-Examples are then assigned to the cluster with the highest probability.
-The models' objective can then be de-composed in three parts:
-\begin{enumerate}
-    \item Each batch should contain different classes
-    \item Each predicted cluster assignment should be confident. 
-    \item The reconstructions should be good. But as they are weighted by the cluster assignment, the importance of the reconstruction is proportional to the confidence in the cluster predictions.
-\end{enumerate}
-
-{\color{cyan} [Ryan: A description of the objective functions may or may not be worth including here, but I don't know what they are.]} The architecture is portrayed in Figure \ref{fig:mixae}, wherein tapered boxes denote a direction of compression in the network components.
+Those representations are used as inputs to an auxiliary network which assigns scores to the clusters, indicating the likelihood that the given example belongs to each cluster.
+Examples are then assigned to the cluster with the highest score.
+The architecture is portrayed in Figure \ref{fig:mixae}, wherein tapered boxes denote a direction of compression in the network components.
 
@@
 
-\noindent {\color{cyan} [Ryan: Still needs a concluding sentence or two here.]} For a more formal consideration of these heuristics, the model objective aand the assumptions made on the data by this model see \citet{Zhang}.
+\noindent The objective during training is to enforce that these assumptions are met.
+For a more formal consideration of these heuristics, the model objective aand the assumptions made on the data by this model see \citet{Zhang}.
 
@@
 
-\begin{figure}[H]
+\begin{figure}
 \centering

### 2020-06-17 doc/Manuscript/article.tex
Updates from Overleaf

@@
 First, the time-series data was represented as a 3D point cloud, where each point contains the maximum charge in the time-series trace.
-We log-transform to map the charge to the interval $[0, 1]$ for each point, then standardize the data.
-in the charge time-series and summing over the time-axis.
-%
-Additionally, we log-transform and normalize the data to the interval $[0, 1]$ for each point.
+We log-transform then min-max scale the charge, mapping the charge to the interval $[0, 1]$.
+The data is projected into 2D space by summing over the time-axis, then downsampled into an $128\times128$ pixel image.
+%Additionally, we log-transform and normalize the data to the interval $[0, 1]$ for each point.
 The final representation for each event is a $128 \times 128$ matrix with each element ranging from $0$ to $1$.
@@
 
-AT-TPC tracks were simulated with the \lstinline{pytpc} package developed by \citet{Bradt2017a}.
-Using the same parameters {\color{blue} what parameters are you referring to?
--MPK} as for the ${}^{46}$Ar$(p, p)$ experiment, a small set of $N=4000$ events per class was generated, as well as a larger set of $N=40000$ events per class.
+A set of $N=40000$ simulated AT-TPC events per class were simulated from a ${}^{46}$Ar$(p, p)$ experiment with the \lstinline{pytpc} package developed by \citet{Bradt2017a}.
+%Using the same parameters {\color{blue} what parameters are you referring to?
+-MPK} as for the
+%Two simulated ${}^{46}$Ar$(p, p)$ experiment, a small set of $N=4000$ events per class was generated, as well as a larger set of $N=40000$ events per class.
 
@@
 
-{\color{cyan} [Ryan: I think there are a lot of words in the following two paragraphs that don't need to exist.
+%{\color{cyan} [Ryan: I think there are a lot of words in the following two paragraphs that don't need to exist.
 Essentially, these paragraphs are a repeat of the last paragraph and parts of the last subsection, but with one or two extra details added in (e.g.
@@
 \section{Conclusions and Perspectives}\label{sec{conclusion}}
-
+% Begin added by DB (6/17/2020)
+The purpose of this study is to explore the application of unsupervised learning algorithms to event identification from an active target detector.
+The necessity of identifying events from raw data prior to
+% End added by DB (6/17/2020)
 From our analysis it is clear that we are a ways off from a fully automated unsupervised method for track classification.

### 2020-06-17 doc/Manuscript/article.tex
Rewrote data prep + other smaller edits -MPK

@@
 %On the opposite end of the pipeline, 
-Therefore, we will incrementally add bias to the analysis by applying the algorithm further down the analysis pipeline..
+Therefore, incrementally add bias to the analysis by applying the algorithm further down the analysis pipeline..
 With the benefit being that more pre-processing improves the signal-to-noise ratio, possibly improving model performance.
@@
 
-To emulate the real-data case, we set a subset of the simulated data to be labelled and treat the rest as unlabelled data.
+For validation, we set a subset of the simulated data to be labelled and treat the rest as unlabelled data.
 We choose this partition to be $15\%$ of each class.
@@
 The purpose of this study is to explore the application of unsupervised learning algorithms to event identification from an active target detector.
-The necessity of identifying events from raw data prior to
+The necessity of identifying events from raw data prior to full processing is becoming a major issue in the data analysis
 % End added by DB (6/17/2020)

### 2020-06-20 doc/Manuscript/article.tex
Added statements in abstract, section 2, and reworked conclusion.

@@
 
+% From DB (06/19/2020)
+From the experimenter's point of view, it is clear that the method used in \cite{Bradt2018} to identify and filter events is not the most efficient computationally.
+The ML methods explored in \cite{Kuchera2019} are a step forward, but they still rely on supervised learning methods that require data labeling, a time-consuming and error-prone process.
+The aim of the present study is to investigate unsupervised learning methods that bypass the labeling step, and form classes of events independently from the experimenter's input.
+The task of labeling the different classes is then much less time-consuming and can potentially lead to the discovery of unforeseen types of events.
+
 \section{Data Preparation}
@@
 Note that $\boldsymbol{X}_L \subset \boldsymbol{X}_F$.
-
-Two simulated events are displayed in figure \ref{fig:sim_samples}. The top row illustrates a proton-event, and the bottom a carbon-event.
 
@@
 
-After hyper parameter tuning each highest performing model is re-run $10$ times. The results are listed in table \ref{tab:clstr_mixae}'
+After hyper parameter tuning each highest performing model is re-run $10$ times. The results are listed in table \ref{tab:clstr_mixae}.
 We observe that while the algorithm can achieve very strong performance, the performance varies.
@@
 
-\subsection{Comparing performance}
+Applied to raw data the MIXAE captures a proton cluster in a similar vein to the K-means approach.
+The MIXAE forms a proton-majority cluster, but with a significant portion of the more noisy proton events being clustered with the other-class.
+Additionally, we do not observe the carbon events being separated from either the amorphous noise events or from the proton cluster.
 
-It is also interesting to compare and contrast the clustering results from the MIXAE model with those of the VGG16$+$K-means outside the fairly abstract accuracies and rand scores.
-It is especially interesting to compare the cluster assignments, as they can inform further research efforts in the clustering of AT-TPC events.
-We illustrate the clustering with confusion matrices that are shown in figure \ref{fig:mixae_confmat}.
-From these matrices, we observe that the MIXAE applied to the clean data correctly clusters the noise events.
-Additionally, it identifies two proton clusters.
-We observe that these proton clusters are both less pure than the VGG16+K-means clusters and that there does not seem to be a visually meaningful difference between these clusters.
-The latter is inferred from figure \ref{fig:filtered_mixae_clster_repr}.
-
-Applied to the real data the MIXAE correctly separates the proton class, however it is unable to separate the carbon events from the amorphous noise events or from the proton cluster.
+On filtered data the highest performing MiXAE model achieves strong separation of the other-class, but curiously creates two proton-majority clusters.
 
@@
 	\includegraphics[width=\textwidth]{custom_work/Filtered_mixae_conf_mat.pdf}
-\caption{}\label{fig:mixae_confmat_filtered}
+\caption[MIXAE - confusion matrices]{Confusion matrix for the MIXAE clustering algorithm on filtered AT-TPC events.
+The true labels indicate samples belonging to the p (proton), carbon (C), or other classes.
+We observe that the algorithm forms two proton-majority clusters, and one clearly defined cluster of the other events.
+}\label{fig:mixae_confmat_filtered}
 \end{figure}
@@
 The true labels indicate samples belonging to the p (proton), carbon (C), or other classes.
-We observe that in }\label{fig:mixae_confmat_raw}
+We observe that the algorithm correctly caputers a majority proton-event cluster in cluster 0.
+However, in contrast with the K-means approach this cluster is contaminated to some extend with both carbon and other events.
+}\label{fig:mixae_confmat_raw}
 \end{figure}
 
-Lastly we wish to further investigate if there are systematic differences between proton events that were placed in different clusters for the clean and full data.
-From figure \ref{fig:mixae_confmat} we see that the MIXAE algorithm creates two proton clusters for the filtered data, and places about fifty per-cent of the proton events in a cluster with the amorphous "other" events.
+\subsection{Performance comparison and ?????}
+
+The most striking result we present in this work is the success of the K-means approach.
+As noted by \cite{Aggarwal} distance measures become less informative in higher dimensional spaces, but the K-means algorithm clusters events well in a very high-dimensional space.
+Another surprise is the stability of the K-means algorithm.
+A stability we attribute to the quality of the VGG16 latent space in creating class-separating sub-spaces.
+While the separations are not perfect, the stability and quality of the proton track identification create solid empirical grounding for applying this approach to other active target experiments.
+
+It is also interesting to compare and contrast the clustering results from the MIXAE model with those of the VGG16$+$K-means.
+In particular, the discrepancy in stability is worth noting.
+While the top performing MIXAE runs outperform the K-means approach, its reliability suffers.
+However, the high performance achieved shines a light on valuable potential research into more bespoke models for unsupervised track identification.
 
@@
 
-{\color{orange}[binding sentence]}
-We considered the deep convolutional embedded clustering (DCEC) by \citet{Guo2017} as well as the mixture of autoencoders (MIXAE) method introduced by \citet{Zhang}.
+Additionally, we considered the deep convolutional embedded clustering (DCEC) by \citet{Guo2017} as well as the MIXAE method introduced by \citet{Zhang}.
 While we were able to reproduce the authors' results on their data, the DCEC algorithm proved unable to cluster AT-TPC in our implementation.
@@
 \section{Conclusions and Perspectives}\label{sec{conclusion}}
-% Begin added by DB (6/17/2020)
+% Begin edit by DB (6/17/2020)
 The purpose of this study is to explore the application of unsupervised learning algorithms to event identification from an active target detector.
-The necessity of identifying events from raw data prior to full processing is becoming a major issue in the data analysis
-% End added by DB (6/17/2020)
-From our analysis it is clear that we are a ways off from a fully automated unsupervised method for track classification.
+The necessity to identify events from raw data prior to full processing is becoming a major issue in the data analysis of detectors with complex responses such as the AT-TPC.
+As shown by both avenues explored in this work, it is clear that there is significant potential to eventually achieve event classification using fully automated unsupervised methods.
+
+In particular, the ability of the K-means algorithm in picking out clear proton clusters from the VGG16 latent space lends itself well to an exploratory phase of analysis, where clusters of events corresponding to different reaction channels could be later identified by the experimenter.
+Another interesting facet of the K-means clustering is its consistent performance.
+As shown in table \ref{tab:clstr_vgg}, the variance is zero for the performance metrics.
+This result indicates that the clusters are very clearly defined in the VGG16 latent space.
+However, as can be seen from the non-proton clusters in figures \ref{fig:clster_confmati_filt} and \ref{fig:clster_confmati_raw},  this does not necessarily imply that the physical signals are correspondingly clear.
+A caveat to the K-means method is that the number of clusters has to be specified in advance.
+Each experiment then has to be considered in light of possible reaction channels to determine a sensible number of clusters for this approach.
+
+The same caveat is present in the MIXAE implementation.
+While it showed better optimal performance than the K-means method, some inconsistencies were observed, that were notably not evident from the unsupervised training-objectives of the model.
+These two factors currently conspire to limit its immediate applicability, and more developments are needed for this approach.
+
+In summary, our study shows that unsupervised track classification with an implementation of the VGG16 + K-means approach is a viable solution.
+For future work, it is worth investigating whether the adaptation of bespoke models like the MIXAE algorithm will allow better performance at no significant cost to consistency.
+%Especially considering the flexibility to use 3D input representations or auxiliary physical goals in the optimization procedure. 
+The two examples of unsupervised machine learning methods studied in this work are a first encouraging step towards automated classification of events that could not only greatly reduce the resource cost of analysis, but also eventually boost the efficiency of the experiment by allowing post-trigger decisions based on such algorithm implemented in hardware.
+Much progress remains to be done before achieving these goals.
+% End edit by DB (6/17/2020)
+%From our analysis it is clear that we are a ways off from a fully automated unsupervised method for track classification.
 {\color{orange}[flip this to positive ]} It is also clear that there is significant potential in both avenues explored in this work.
-In particular, the surprising ability of the K-means algorithm in picking out clear proton clusters from the VGG16 latent space lends itself well to an exploratory phase of analysis.
-Another surprising facet of the K-means clustering is its consistent performance.
+%n particular, the surprising ability of the K-means algorithm in picking out clear proton clusters from the VGG16 latent space lends itself well to an exploratory phase of analysis.
+%Another surprising facet of the K-means clustering is its consistent performance.
 As shown in table \ref{tab:clstr_vgg}, the variance is zero for the performance metrics.
@@
 
-In summary, we recommend that researchers performing unsupervised track classification begin with an implementation of the VGG16 + K-means approach.
+%In summary, we recommend that researchers performing unsupervised track classification begin with an implementation of the VGG16 + K-means approach.
 For future work, it is worth investigating whether the adaptations of bespoke models like the MIXAE algorithm will allow better performance at no significant cost to consistency.
@@
 
-{\color{orange}[RS Another abrupt end to a paragraph. Could probably use some rounding out.]}
+%{\color{orange}[RS Another abrupt end to a paragraph. Could probably use some rounding out.]}
 
@@
 \begin{table}[H]
-\todo{prettify or remove}
 \renewcommand*{\arraystretch}{0.5}

### 2020-06-26 doc/Manuscript/article.tex
Updates from Overleaf

@@
 {\color{blue} Needs more explanation earlier in paper -MPK}
-In addition, the selection task produced a binary result only, ``good" or``bad", relative to the event of interest, while the data clearly contained more classes of events%, if only because of the presence of carbon in the target (from the isobutane gas).
+In addition, the selection task produced a binary result only, ``good" or``bad", relative to the event of interest, while the data clearly contained more classes of events.%, if only because of the presence of carbon in the target (from the isobutane gas).
 In a broader perspective, an unsupervised classification algorithm would offer the possibility to ``discover" rare events not anticipated (or overlooked) by the physicist.
@@
 %We begin by investigating the K-means clustering algorithm on the VGG16 latent space. 
-The VGG16 model is pre-trained on the ImageNet dataset, and subsequently applied to the AT-TPC events without the classification layers.
+The user a VGG16 model that is pre-trained on the ImageNet dataset and subsequently applied to the AT-TPC events without the classification layers.
 This creates a set of vectors $\boldsymbol{x} \in \R^{8192}$ as our new representation of the events.
-To cluster we use \lstinline{scikit-learn} implementation of the K-means algorithm, with default parameters \cite{Pedregosa2011}.
+To cluster, we use the \lstinline{scikit-learn} implementation of the K-means++ algorithm with default parameters \cite{Pedregosa2011}.
 
@@
 Examples are then assigned to the cluster with the highest score.
-The architecture is portrayed in Figure \ref{fig:mixae}, wherein tapered boxes denote a direction of compression in the network components.
 
@@
 \noindent The objective during training is to enforce that these assumptions are met.
+The architecture is portrayed in Figure \ref{fig:mixae}, wherein tapered boxes denote a direction of compression in the network components.
+
 For a more formal consideration of these heuristics, the model objective aand the assumptions made on the data by this model see \citet{Zhang}.
@@
 
+{\color{blue} MPK edit this}
 \section{Results and Discussions}\label{sec:results}
@@
 The same caveat is present in the MIXAE implementation.
-While it showed better optimal performance than the K-means method, some inconsistencies were observed, that were notably not evident from the unsupervised training-objectives of the model.
+While it shows better optimal performance than the K-means method, some inconsistencies were observed, that were notably not evident from the unsupervised training-objectives of the model.
 These two factors currently conspire to limit its immediate applicability, and more developments are needed for this approach.

### 2020-07-06 doc/Manuscript/article.tex
Updates from Overleaf

@@
 \subsection{Why machine learning}
-{\bf we need to rewrite these bullet points, MHJ}
-\begin{itemize}
-    \item Traditional MC methods fall short in two principal ways:
-    \begin{itemize}
-        \item The computational cost per event is too large given the size of the data-sets
-        \item The broken tracks with noisy environment create bad fit statistics for otherwise useful events. End result is f$_1 \sim 0.7 $
-        \item define a ``bad" fit as a non-proton event
-    \end{itemize}
-\end{itemize}
+%{\bf we need to rewrite these bullet points, MHJ}
+%\begin{itemize}
+%    \item Traditional MC methods fall short in two principal ways:
+%    \begin{itemize}
+%        \item The computational cost per event is too large given the size of the data-sets
+%        \item The broken tracks with noisy environment create bad fit statistics for otherwise useful events. End result is f$_1 \sim 0.7 $
+%        \item define a ``bad" fit as a non-proton event
+%    \end{itemize}
+%\end{itemize}
 
@@
 However, it has been found that models which are trained at large scale will learn general features that are applicable to a variety of tasks.
-For example, large neural networks which are trained on the ImageNet dataset \cite{Russakovsky2015} --- a diverse image classification task --- learn how to identify lines, edges, and other common shapes which are useful for numerous problems.
+For example, large neural networks which are trained on the ImageNet dataset \cite{Russakovsky2015} --- a diverse image classification task --- learn how to identify lines, edges, and other common shapes that are useful for numerous problems.
 Thus, it is common practice to initialize the convolutional layers of a network with the pre-trained weights learned from ImageNet (or some other large dataset).
 The training process then only has to fine-tune the network for the specific task.
-Since we are building on prior knowledge in this case, learning becomes far more efficient and better performance can often be achieved.
+Since we are building on prior knowledge in this case, learning becomes far more efficient, and better performance can often be achieved.
 \citet{Kuchera2019} used machine learning methods to classify the products of $^{46}$Ar reactions in the AT-TPC, and they found that a CNN initialized with ImageNet weights resulted in the most successful classification.
@@
 In contrast with the classification work of \citet{Kuchera2019}, we do not assume access to ground truth labels and are trying to solve a fundamentally different learning problem.
-Thus, rather than fine-tuning a pre-trained network under the supervised learning regime, we extract the output of the pre-trained network's last convolutional layer as a latent representation of the events.
-We then cluster events based on this representation using the K-means++ algorithm with the standard choice of Euclidean distance.
+Thus, rather than fine-tuning a pre-trained network under the supervised learning regime, we extract the output of the pre-trained network's last convolutional layer as a latent representation of the events, where each event is represented as a vector in $\R^{8192}$.
+We then cluster events based on this representation using the Scikit-Learn implementation of the K-means++ algorithm with default parameters \cite{Pedregosa2011}.
 
@@
 %We begin by investigating the K-means clustering algorithm on the VGG16 latent space. 
-The user a VGG16 model that is pre-trained on the ImageNet dataset and subsequently applied to the AT-TPC events without the classification layers.
+% The user a VGG16 model that is pre-trained on the ImageNet dataset and subsequently applied to the AT-TPC events without the classification layers.
 This creates a set of vectors $\boldsymbol{x} \in \R^{8192}$ as our new representation of the events.
@@
 
-\noindent The objective during training is to enforce that these assumptions are met.
+\noindent The learning objective encourages these assumptions to be met.
 The architecture is portrayed in Figure \ref{fig:mixae}, wherein tapered boxes denote a direction of compression in the network components.
 
-For a more formal consideration of these heuristics, the model objective aand the assumptions made on the data by this model see \citet{Zhang}.
+For a more formal consideration of these heuristics, the model objective and the assumptions made on the data by this model see \citet{Zhang}.
 
@@
 
+\subsection{Measuring performance}
+
+We measure the performance of the clustering algorithms by two functions: clustering accuracy and the adjusted rand index (ARI). 
+Ordinarily, we define accuracy as the agreement between a set of class predictions and their ground truth labels. 
+However, with a clustering algorithm, there is no a priori knowledge of which cluster represents what ground-truth class. 
+Since, in this example, we have access to ground-truth labels, we can define a clustering accuracy by calculating the maximum accuracy achieved by permuting which cluster should indicate a given class.
+A measure of clustering accuracy falls in the $[0, 1]$ range where $1$ denotes perfect agreement.
+
+Since we have access to ground truth labels, we can also calculate the ARI, which  measures the agreement between two sets of labels but {\color{blue}adjusts for chance assignments:: define this mathematically} \cite{Hubert1985}.
+{\color{blue}  The effect is that a trivial clustering algorithm which predicts only one label will evaluate to ARI$=0$, but will have a nonzero accuracy:: need the math for this to be clear}.
+As with the clustering accuracy, an ARI score is bounded in the range $[0, 1]$ where $1$ denotes perfect agreement.
+%We refer to \citet{Hubert1985} for a detailed consideration of the ARI.
+
+The performance is measured by comparing model predictions on the labelled subset of data (see table \ref{tab:datasets}) of the ARI.
+
 {\color{blue} MPK edit this}
@@
 
-We begin by exploring the results of applying a K-means approach on the latent space of a pre-trained network.
+We first explore the results of applying a K-means approach on the latent space of a pre-trained network.
 Subsequently, we investigate the performance of the MIXAE algorithm as outlined in section \ref{sec:mixae}.
@@
 
-In addition to the performance measures reported in table \ref{tab:clstr_vgg}, it is interesting to observe which samples are being wrongly assigned.
-To investigate this problem we tabulate the assignments of samples relative to their ground truth labels.
+In addition to the performance measures reported in table \ref{tab:clstr_vgg}, it is interesting to observe which samples are wrongly assigned.
+To investigate this problem, we tabulate the assignments of samples relative to their ground truth labels.
 From these tables, we can infer which classes are more or less entangled with others.
@@
 
-This high quality cluster also appears in the clustering of raw data.
-From figure \ref{fig:clster_confmati_raw} we observe that there is a high purity proton cluster.
+This high-quality cluster also appears in the clustering of raw data.
+From figure \ref{fig:clster_confmati_raw}, we observe that there is a high purity proton cluster.
 In contrast to the filtered data we observe that the deterioration in performance can largely be ascribed to the algorithm creating a proton + other cluster and a carbon + other cluster.

### 2020-07-06 doc/Manuscript/article.tex
edited metric section + some other small edits -MPK

@@
 \section{Experimental Details} 
+\label{Sec:Exp}
 The goal of the experiment was to measure the evolution of the elastic and inelastic cross sections between $^{46}$Ar and protons as a function of energy (the {\em excitation function}), and observe resonances in the composite system $^{47}$K that correspond to analog states in the nucleus $^{47}$Ar.
@@
 \subsection{Classifying  events} 
-\todo{Brute force serving as motivation for why we do unsupervised learning.}
 %The filtering task relied on a set of $\chi^2$ distributions based on different criteria, 
-The traditional event selection process involved a Monte-Carlo $\chi^2$ minimization method, however there was no well-defined method to quantify the effectiveness of the event selection from the thresholds applied to these distributions.
-{\color{blue} Needs more explanation earlier in paper -MPK}
+The traditional Monte-Carlo event selection process, described in Section~\ref{Sec:Exp}, does not have a well-defined method to quantify the effectiveness of the event selection.
 In addition, the selection task produced a binary result only, ``good" or``bad", relative to the event of interest, while the data clearly contained more classes of events.%, if only because of the presence of carbon in the target (from the isobutane gas).
@@
 
-The principal challenge in the AT-TPC experiments that we are trying to solve is the reliance on labelled samples in the analysis as future experiments may not have as visually dissimilar reaction products as we observe in the ${}^{46}$Ar experiment.
+The principal challenge in the AT-TPC experiments that we are trying to address is the reliance on labelled samples in the analysis, as future experiments may not have as visually dissimilar reaction products as we observe in the ${}^{46}$Ar experiment.
 The  ${}^{46}$Ar experiment does, however, provide a useful example where we can then explore unsupervised techniques.

### 2020-07-06 doc/Manuscript/article.tex
Edited Section 5 -MPK

@@
 
-{\color{blue} MPK edit this}
 \section{Results and Discussions}\label{sec:results}
@@
 The principal challenge in the AT-TPC experiments that we are trying to address is the reliance on labelled samples in the analysis, as future experiments may not have as visually dissimilar reaction products as we observe in the ${}^{46}$Ar experiment.
-The  ${}^{46}$Ar experiment does, however, provide a useful example where we can then explore unsupervised techniques.
+The ability to label data in the the ${}^{46}$Ar experiment does, however, provide a useful example where we can then explore unsupervised techniques.
 
@@
 
+{\color{blue} This section may contain info that you can move to the metrics section??}
 In addition to the performance measures reported in table \ref{tab:clstr_vgg}, it is interesting to observe which samples are wrongly assigned.
@@
 This is done to estimate to what degree the class separating information is encoded in the entirety of the latent space, or in some select regions.
-The results from the PCA analysis were virtually identical to the results sans the PCA, and so we omit them for brevity.
+The results from the PCA analysis {\color{blue} using the top XX principal components} were virtually identical to the results sans the PCA, indicating that a lower-dimensional encoding of the data is sufficient for data clustering for this experiment.
+%and so we omit them for brevity.
 
-Furthermore, we wish to further characterize the clusters presented in the matrix in figures \ref{fig:clster_confmati_filt} and \ref{fig:clster_confmati_raw}.
-To achieve this, we sample examples from the proton samples belonging to different clusters for the filtered and full data.
+To further investigate the clusters presented in the matrix in figures \ref{fig:clster_confmati_filt} and \ref{fig:clster_confmati_raw}, we sample examples from the proton samples belonging to different clusters for the filtered and full data.
 We display these examples in figures \ref{fig:filtered_vgg_clster_repr} and \ref{fig:full_vgg_clster_repr} respectively.
@@
 
-After hyper parameter tuning each highest performing model is re-run $10$ times. The results are listed in table \ref{tab:clstr_mixae}.
-We observe that while the algorithm can achieve very strong performance, the performance varies.
+After hyperparameter tuning, each highest performing model is run $10$ times. The results are listed in table \ref{tab:clstr_mixae}.
+We observe that, while the algorithm can achieve very strong performance, the performance varies.
 In some cases the MIXAE model converges to a seemingly good configuration, based on its unsupervised training goals.
-However, when inspecting its clustering performance against labeleled data that seemingly good model does no better than random.
+However, when inspecting its clustering performance against labelled data that seemingly good model does no better than random.
 This happens more frequently with raw data, indicating an interaction with the noise levels present in the events.
@@
 
-As with the VGG16 + KMeans approach we wish to further characterize the clustering results.
+As with the VGG16 + KMeans approach we wish to further investigate the clustering results.
 Taking the best performing MIXAE model on filtered and raw data we tabulate the clusters against their labels.
@@
 \caption[MIXAE - confusion matrices]{Confusion matrix for the MIXAE clustering algorithm on filtered AT-TPC events.
-The true labels indicate samples belonging to the p (proton), carbon (C), or other classes.
+The true labels indicate samples belonging to the \texttt{p} (proton), \texttt{C} (Carbon), or \texttt{other} classes.
 We observe that the algorithm forms two proton-majority clusters, and one clearly defined cluster of the other events.
@@
 \caption[MIXAE - confusion matrices]{Confusion matrices for the MIXAE clustering algorithm on raw AT-TPC events.
-The true labels indicate samples belonging to the p (proton), carbon (C), or other classes.
+The true labels indicate samples belonging to the p (proton), carbon (C), or \texttt{other} classes.
 We observe that the algorithm correctly caputers a majority proton-event cluster in cluster 0.

### 2020-07-10 doc/Manuscript/article.tex
Updates from Overleaf

@@
 \subsection{Measuring performance}
+Unsupervised machine learning very typically goes hand-in-hand with a lack of ground-truth labelled data.
+In the face of such missing data, a model is ordinarily assessed by measures that do not depend on knowing the ground truth for any samples.
+However, as our work aims to explore the application of unsupervised methods to track identification, we chose the ${}^{46}$Ar data we have some ground truth labels to evaluate our models.
+The measures we use to evaluate the models we then chose to be ones which measure the similarities between two arbitrary sets of clustering assignments while holding one to be the ground truth.
 
-We measure the performance of the clustering algorithms by two functions: clustering accuracy and the adjusted rand index (ARI). 
-Ordinarily, we define accuracy as the agreement between a set of class predictions and their ground truth labels. 
-However, with a clustering algorithm, there is no a priori knowledge of which cluster represents what ground-truth class. 
-Since, in this example, we have access to ground-truth labels, we can define a clustering accuracy by calculating the maximum accuracy achieved by permuting which cluster should indicate a given class.
-A measure of clustering accuracy falls in the $[0, 1]$ range where $1$ denotes perfect agreement.
+We measure the performance of the clustering algorithms by two functions: clustering accuracy and the adjusted rand index (ARI)\cite{Hubert1985}.
+Both of these measurements fall in the range $ [0, \hspace{0.1cm}1]$, where $0$ denotes a complete disagreement and $1$ a complete agreement between the ground truth and predicted labels.
 
-Since we have access to ground truth labels, we can also calculate the ARI, which  measures the agreement between two sets of labels but {\color{blue}adjusts for chance assignments:: define this mathematically} \cite{Hubert1985}.
-{\color{blue}  The effect is that a trivial clustering algorithm which predicts only one label will evaluate to ARI$=0$, but will have a nonzero accuracy:: need the math for this to be clear}.
-As with the clustering accuracy, an ARI score is bounded in the range $[0, 1]$ where $1$ denotes perfect agreement.
+To compute the metrics we have to solve problem introduced by the arbitrary labels of a clustering algorithm.
+That is, we do not know which predicted assignment should correspond to a proton or carbon event.
+In short, we want to find the most reasonable correspondence between the clusters and the ground truth classes.
+To solve this problem we first define two sets; the ground truth labels $\mathbf{y} = [y_i, \hspace{0.2cm} y_{i +1}, \ldots]$ and the corresponding predictions $\mathbf{\hat{y}} = [\hat{y}_i, \hspace{0.2cm} \hat{y}_{i +1}, \ldots]$.
+Furthermore, let both $y_i$ and $\hat{y}_i$ be integer representations of an events ground truth, or predicted class, respectively.
+
+To compute both the ARI and the clustering accuracy we also have to construct the contingency table between two sets of clustering assignments.
+A contingency matrix defines the overlaps between classes in these sets, and its general form is shown in table \ref{tab:contingency}.
+
+\begin{table}[H]
+\centering
+\caption{General form of a contingency table.
+$y_n$ and $\hat{y}_n$ are the ground truth labels and clustering assignments respectively.
+The $w_{ij}$'s then describe how many samples are in clusters $y_i$ and simultaneously in $\hat{y}_j$.}\label{tab:contingency}
+\begin{tabular}{c|ccc|c}
+& $y_1$ &$y_2$ & $\cdots$ & sums \\
+\midrule
+$\hat{y}_1$ & $w_{11}$ & $w_{12}$ &  $\cdots$ & $e_1$ \\
+$\hat{y}_2$ & $w_{21}$ & $w_{22}$ &  $\cdots$ & $e_2$\\
+$\vdots$ & $\vdots$ & \vdots & $\ddots$ & \vdots \\
+\hline
+sums & $f_1$ & $f_2$ & $\hdots$ 
+\end{tabular}
+\end{table}
+
+The algorithm for finding the clustering accuracy can then be described in these steps:
+
+\begin{itemize}
+    \item Compute the matrix $\mathbf{W}$ from the contingency table between $\mathbf{y}$ and $\mathbf{\hat{y}}$
+    \item Subtract $\mathbf{W}$ from its maximum value to find the bipartite graph representation of the assignment problem.
+    \item Use an algorithm, like the Hungarian algorithm, to solve the assignment problem, and take the average of the values in $\mathbf{W}$ this solution indicates.
+This average is the clustering accuracy.
+\end{itemize}
+
+To compute the ARI we use the elements from the contingency table to evaluate the function introduced by \citet{Hubert1985}
+
+\begin{equation}\label{eq:ari}
+\text{ARI} = \frac{\sum  \binom{w_{ij}}{2} - \left[\sum  \binom{e_{i}}{2} \sum  \binom{f_{j}}{2}  \right]/\binom{n}{2}}{\frac{1}{2}\left[\sum  \binom{e_{i}}{2} + \sum  \binom{f_{j}}{2}  \right]- \left[\sum  \binom{e_{i}}{2} \sum  \binom{f_{j}}{2}  \right]/\binom{n}{2}}.
+\end{equation}
+
+The important distinction to make between the clustering accuracy and the ARI is that the clustering accuracy is a simple comparison that does not account for chance assignments, while the ARI does.
+In effect, this means that the accuracy is a good heuristic for the performance which is tempered by the ARI.
 %We refer to \citet{Hubert1985} for a detailed consideration of the ARI.
 
-The performance is measured by comparing model predictions on the labelled subset of data (see table \ref{tab:datasets}) of the ARI.
+The performance is measured by comparing model predictions on the labelled subset of data (see table \ref{tab:datasets}).
 
@@
 
-\subsection{Performance comparison and ?????}
+\subsection{Performance comparison}
 

### 2020-07-10 doc/Manuscript/article.tex
Updates from Overleaf

@@
 Additionally, the VGG16+K-means approach finds high purity clusters of proton events for real experimental data.
-We also explore the application of neural networks to clustering by implementing a mixture of autoencoders algorithm, with improved results on real data.
+We explore also the application of autoencoder neural networks to clustering by implementing.
+With autoencoder neural networks we find improved descriptions of data from experiments.
 
@@
 \begin{highlights}
-\item Research highlight 1
-\item Research highlight 2
+\item We present  novel applications of unsupervised machine learning methods to the analysis of event separations in active target experiments. 
+\item Autoencoder neural networks represent a promising path to the analysis of experimental data. 
 \end{highlights}
@@
 \begin{keyword}
-%% keywords here, in the form: keyword \sep keyword
-
+Active target experiments \sep Machine Learning \sep Unsupervised Learning \sep Autoencoder Neural Networks
+\PACS
 %% PACS codes here, in the form: \PACS code \sep code
-
-%% MSC codes here, in the form: \MSC code \sep code
-%% or \MSC[2008] code \sep code (2000 is the default)
 
@@
 % motivating ML work:
-Because TPCs produce 3-dimensional images of charged particle tracks, the event identification task is often akin to a visual inspection (akin to analysis in the bubble chamber era), which is not practical nowadays because of the large quantities of data.
+Because TPCs produce three-dimensional images of charged particle tracks, the event identification task is often akin to a visual inspection (comparable to the analyses made in the bubble chamber era), which is not practical nowadays because of the large quantities of data.
 Machine learning techniques then appear as a promising prospect, in particular in the image recognition domain where much progress has been made recently \cite{mehta2019}.
-In addition, some ML algorithms offer new possibilities such as the potential discovery of unforeseen phenomena that would have been missed by more traditional analysis methods.
+In addition, machine learning (ML) algorithms offer new possibilities such as the potential discovery of unforeseen phenomena that would have been missed by more traditional analysis methods.
 Prior work has demonstrated the ability to apply supervised classification machine learning methods to AT-TPC data when a labeled training set is available, whether through hand-labeled data or labeled simulated data (a {\em transfer learning} application) \cite{Kuchera2019}.
 In some experiments, a labeled dataset is unavailable.
-This could be  due to the inability to hand-label events, or the case where one does not know a priori the types and behaviors of the reactions present in the detector in order to generate a labeled, simulated dataset.
+This could be  due to the inability to hand-label events, or the case where one does not know {\em a priori} the types and behaviors of the reactions present in the detector in order to generate a labeled, simulated dataset.
 In the latter case, there must also exist a validation dataset of real data, still requiring the ability to label a subset of the real data.
@@
 This axial magnetic field served the purpose of bending the trajectories of the recoil particles in order to i) increase their length and ii) provide a measurement of their bending radius, directly related to their magnetic rigidity.
-Because the recoil particles travel in gas, they slow down and eventually stop, therefore their trajectories are described by 3-dimensional spirals (see \cite{Bradt2017}).
+Because the recoil particles travel in gas, they slow down and eventually stop, therefore their trajectories are described by three-dimensional spirals (see \cite{Bradt2017}).
 One of the difficulties encountered in the analysis is that the shape of these spirals does not have an analytical form because it follows the energy-loss profile of the particles.
@@
 
-The method used in \cite{Bradt2018} to analyze the data followed a 3-phase sequence: cleaning, filtering and fitting.
+The method used in \cite{Bradt2018} to analyze the data followed a three-phase sequence: cleaning, filtering and fitting.
 Traditional methods were used to perform each of these tasks, and ultimately extract the scientific information, but there were severe limitations and high computational costs that become prohibitive in data sets larger by an order of magnitude.
@@
 The cleaning was performed using a combination of linear and circular Hough transforms on a 2-dimensional projection of the tracks \cite{Bradt2017}.
-The following filtering and fitting phases were performed simultaneously, by defining the cost function to the fitting algorithm as a sum of 3 $\chi^2$ components based on i) the position of the track in space, ii) the energy deposited on each pixel of the sensor plane, and iii) the location of the vertex of the reaction.
-While various fitting algorithms were tested, the most accurate was a Monte-Carlo algorithm that explored the 6-dimensional phase space of the particle's kinematics parameters, reducing it progressively at each iteration step until the desired accuracy was reached \cite{Bradt2017}.
+The following filtering and fitting phases were performed simultaneously, by defining the cost function to the fitting algorithm as a sum of three $\chi^2$ components based on i) the position of the track in space, ii) the energy deposited on each pixel of the sensor plane, and iii) the location of the vertex of the reaction.
+While various fitting algorithms were tested, the most accurate was a Monte-Carlo algorithm that explored the six-dimensional phase space of the particle's kinematics parameters, reducing it progressively at each iteration step until the desired accuracy was reached \cite{Bradt2017}.
 Although this algorithm ended up being the most accurate, it is extremely costly computationally because of the very large number of simulated tracks needed for each event.
@@
 This is a very inefficient method because it requires performing the fitting for all events, including those that are not of interest.
-Pioneering work on event identification using ML methods, namely the use of a pre-trained CNN, later showed that the filtering phase would better be performed using this type of technique \cite{Kuchera2019}.
+Pioneering work on event identification using machine learning methods, namely the use of a pre-trained convolutional neural networks (CNN), later showed that the filtering phase would better be performed using this type of technique \cite{Kuchera2019}.
 % End added by DB (06/10/2020)

### 2020-07-10 doc/Manuscript/article.tex
Updates from Overleaf

@@
 %on two tasks: a semi-supervised classification task and the unsupervised clustering of particle tracks. 
-%On the semi-supervised task, we find that a logistic regression classifier trained on small labelled subsets of the latent space of these models perform very well, both on simulated and real data.
+%On the semi-supervised task, we find that a logistic regression classifier trained on small labeled subsets of the latent space of these models perform very well, both on simulated and real data.
 On the clustering task,
@@
 Prior work has demonstrated the ability to apply supervised classification machine learning methods to AT-TPC data when a labeled training set is available, whether through hand-labeled data or labeled simulated data (a {\em transfer learning} application) \cite{Kuchera2019}.
-In some experiments, a labeled dataset is unavailable.
-This could be  due to the inability to hand-label events, or the case where one does not know {\em a priori} the types and behaviors of the reactions present in the detector in order to generate a labeled, simulated dataset.
-In the latter case, there must also exist a validation dataset of real data, still requiring the ability to label a subset of the real data.
+In some experiments, a labeled data set is unavailable.
+This could be  due to the inability to hand-label events, or the case where one does not know {\em a priori} the types and behaviors of the reactions present in the detector in order to generate a labeled, simulated data set.
+In the latter case, there must also exist a validation data set of real data, still requiring the ability to label a subset of the real data.
 The unsupervised separation of event types, or {\em clustering}, based on a set of ML algorithms is hereby examined, using experimental data recorded by the AT-TPC during its commissioning experiment from a radioactive $^{46}$Ar beam reacting on an isobutane target composed of proton and carbon nuclei.
@@
 Traditional methods were used to perform each of these tasks, and ultimately extract the scientific information, but there were severe limitations and high computational costs that become prohibitive in data sets larger by an order of magnitude.
+In particular, as for the data presented and analyzed here, with data sets in the terabyte region or larger, these more traditional methods become prohibitively expensive from a computational stand.
 % Begin added by DB (06/10/2020)
@@
 % End added by DB (06/10/2020)
-In addition, \cite{Kuchera2019} demonstrated that the purity and statistics of the data are improved with the use of ML.
+In addition, the authors of Ref.~\cite{Kuchera2019} demonstrated that the purity and statistics of the data are improved with the use of ML.
 
@@
 The AT-TPC data we studied for this work was recorded as charge time-series for each of the  the $\sim10^4$ detector pads.
-In this representation, an event is a record of $512$ time-buckets for each of the $10^4$ detector pads. In our analysis, we represent each event as a downsampled 2D projection.
-We chose to represent the data in 2D to facilitate the use of sophisticated image-recognition machine learning models, and this data representation was shown to successfully classify the events of this experiment in a supervised setting \cite{Kuchera2019}.
-First, the time-series data was represented as a 3D point cloud, where each point contains the maximum charge in the time-series trace.
+In this representation, an event is a record of $512$ time-buckets for each of the $10^4$ detector pads. In our analysis, we represent each event as a down-sampled two-dimensional  projection.
+We chose to represent the data in two dimensions to facilitate the use of sophisticated image-recognition machine learning models, and this data representation was shown to successfully classify the events of this experiment in a supervised setting \cite{Kuchera2019}.
+First, the time-series data was represented as a three-dimensional  point cloud, where each point contains the maximum charge in the time-series trace.
 We log-transform then min-max scale the charge, mapping the charge to the interval $[0, 1]$.
-The data is projected into 2D space by summing over the time-axis, then downsampled into an $128\times128$ pixel image.
+The data is projected into two-dimensional  space by summing over the time-axis, then down-sampled into a $128\times128$ pixel image.
 %Additionally, we log-transform and normalize the data to the interval $[0, 1]$ for each point.
@@
 %On the opposite end of the pipeline, 
-Therefore, incrementally add bias to the analysis by applying the algorithm further down the analysis pipeline..
-With the benefit being that more pre-processing improves the signal-to-noise ratio, possibly improving model performance.
-To explore this trade-off between model performance and pre-processing bias, we performed analysis on simulated, cleaned and raw events.
+Therefore, we incrementally add bias to the analysis by applying the algorithm further down the analysis pipeline, with the benefit being that more preprocessing improves the signal-to-noise ratio, possibly improving model performance.
+To explore this trade-off between model performance and preprocessing bias, we performed our analysis on simulated, cleaned and raw events.
 
@@
 \centering
-\caption{Event class percentages for each of the datasets used in this analysis.
+\caption{Event class percentages for each of the data sets used in this analysis.
 The decrease in the "other" class of events from full to filtered data owes to the thresholding of events.
-If an event contains fewer than $20$ data-points it is discarded }\label{tab:class_distr}
+If an event contains fewer than $20$ data points it is discarded }\label{tab:class_distr}
 \begin{tabular}{lccc}
@@
 
-A set of $N=40000$ simulated AT-TPC events per class were simulated from a ${}^{46}$Ar$(p, p)$ experiment with the \lstinline{pytpc} package developed by \citet{Bradt2017a}.
+A set of $N=40000$ simulated AT-TPC events per class was obtained  from a ${}^{46}$Ar$(p, p)$ experiment with the \lstinline{pytpc} package developed by \citet{Bradt2017a}.
 %Using the same parameters {\color{blue} what parameters are you referring to?
@@
 
-For validation, we set a subset of the simulated data to be labelled and treat the rest as unlabelled data.
-We choose this partition to be $15\%$ of each class.
-We denote this subset and its associated labels as $\gamma_L=(\boldsymbol{X}_L, \boldsymbol{y}_L)$, while the entire dataset which is identified as $\boldsymbol{X}_F$.
+For validation, we set a subset of the simulated data to be labeled and treat the rest as unlabeled data.
+We chose this partition to consist of  $15\%$ of the data.
+We denote this subset and its associated labels as $\gamma_L=(\boldsymbol{X}_L, \boldsymbol{y}_L)$, while the entire data set is identified as $\boldsymbol{X}_F$.
 Note that $\boldsymbol{X}_L \subset \boldsymbol{X}_F$.
@@
 The events analyzed in this section were retrieved from the ${}^{46}$Ar resonant proton scattering experiment recorded with the AT-TPC. 
-While we denote these events as raw, it is important to note that what we mean is a raw 2D representation of peak-only events.
+While we denote these events as raw, it is important to note that what we mean is a raw two-dimensional representation of peak-only events.
 
@@
 We display two different events from the ${}^{46}$Ar experiment in figure \ref{fig:samples}.
-The top row illustrates a carbon event with a large fraction of noise, while the bottom row shows a proton event nearly devoid of noise.
+The top row illustrates a carbon event with a large fraction of noise, while the bottom row shows a proton event almost free of noise.
 
@@
 \includegraphics[width=\textwidth]{custom_work/examples_raw.pdf}
-\caption[Displaying unfiltered events in 2D and 3D]{Two- and three-dimensional representations of two events from the ${}^{46}$Ar experiment.
+\caption[Displaying unfiltered events in two and three dimensions]{Two- and three-dimensional representations of two events from the ${}^{46}$Ar experiment.
 Each row is one event in two projections, where the color intensity of each point indicates higher charge values recorded by the detector.}\label{fig:samples}
@@
 This transformation is a common technique in computer vision, used to identify common geometric shapes like lines and circles, and has been used extensively in high-energy particle physics since the bubble-chamber era \cite{Hough:1959}.
-%Essentially, the algorithm draws many lines (of whatever desired geometry) through each data-point and checks whether these lines intersect with points in the dataset.
+%Essentially, the algorithm draws many lines (of whatever desired geometry) through each data-point and checks whether these lines intersect with points in the data set.
 Locations in parameter space that generate many intersections then become bright spots, allowing us to filter away points that are not close to these points.
@@
 The traditional Monte-Carlo event selection process, described in Section~\ref{Sec:Exp}, does not have a well-defined method to quantify the effectiveness of the event selection.
-In addition, the selection task produced a binary result only, ``good" or``bad", relative to the event of interest, while the data clearly contained more classes of events.%, if only because of the presence of carbon in the target (from the isobutane gas).
-In a broader perspective, an unsupervised classification algorithm would offer the possibility to ``discover" rare events not anticipated (or overlooked) by the physicist.
+In addition, the selection task produced a binary result only, either ``good" or``bad" relative to the event of interest.
+The data contain however many more classes of events.%, if only because of the presence of carbon in the target (from the isobutane gas).
+In a broader perspective, an unsupervised classification algorithm would offer the possibility to ``discover" rare events which may not be expected or are overlooked.
 These events would likely be filtered out using the traditional methods.
@@
 The computationally expensive fitting procedure would be applied to every event, instead of the few percent of the events that are of interest for the analysis.
-An unsupervised ML algorithm able to separate the data without a priori knowledge of the different types of events increases the efficiency of the analysis tremendously, and allows the downstream analysis to concentrate on the fitting efforts only on events of interest.
+An unsupervised ML algorithm able to separate the data without {\em a priori} knowledge of the different types of events increases the efficiency of the analysis tremendously, and allows the downstream analysis to concentrate on the fitting efforts only on events of interest.
 In addition, the clustering allows for more exploration of the data, potentially enabling new discovery of unexpected reaction types.
@@
 \centering
-\caption{Descriptions of number of events in the data.}\label{tab:datasets}
+\caption{Descriptions of number of events in the data.}\label{tab:data sets}
 \begin{tabular}{lccc}
@@
 % In the Machine Learning community, it is not uncommon to publish packaged models with fitted parameters from image recognition contests.
-These models are trained on datasets with millions of images and classify between hundreds of distinct classes; one such is the imagenet dataset.
+These models are trained on data sets with millions of images and classify between hundreds of distinct classes; one such is the imagenet data set.
 In their work, \citet{Kuchera2019} used the VGG16 architecture trained on imagenet to classify ${}^{46}$Ar data.
@@
 
-Training high-performing neural networks from scratch often requires enormous datasets and computation time.
+Training high-performing neural networks from scratch often requires enormous data sets and computation time.
 However, it has been found that models which are trained at large scale will learn general features that are applicable to a variety of tasks.
-For example, large neural networks which are trained on the ImageNet dataset \cite{Russakovsky2015} --- a diverse image classification task --- learn how to identify lines, edges, and other common shapes that are useful for numerous problems.
-Thus, it is common practice to initialize the convolutional layers of a network with the pre-trained weights learned from ImageNet (or some other large dataset).
+For example, large neural networks which are trained on the ImageNet data set \cite{Russakovsky2015} --- a diverse image classification task --- learn how to identify lines, edges, and other common shapes that are useful for numerous problems.
+Thus, it is common practice to initialize the convolutional layers of a network with the pre-trained weights learned from ImageNet (or some other large data set).
 The training process then only has to fine-tune the network for the specific task.
@@
 %We begin by investigating the K-means clustering algorithm on the VGG16 latent space. 
-% The user a VGG16 model that is pre-trained on the ImageNet dataset and subsequently applied to the AT-TPC events without the classification layers.
+% The user a VGG16 model that is pre-trained on the ImageNet data set and subsequently applied to the AT-TPC events without the classification layers.
 This creates a set of vectors $\boldsymbol{x} \in \R^{8192}$ as our new representation of the events.
@@
 
-The performance is measured by comparing model predictions on the labelled subset of data (see table \ref{tab:datasets}).
+The performance is measured by comparing model predictions on the labelled subset of data (see table \ref{tab:data sets}).
 
@@
 The results of the clustering runs are included in table \ref{tab:clstr_vgg}.
-We observe that we are able to attain near-perfect clustering on simulated data and that there is a sharp decline in performance as we add noise by moving to the filtered and raw datasets.
+We observe that we are able to attain near-perfect clustering on simulated data and that there is a sharp decline in performance as we add noise by moving to the filtered and raw data sets.
 
@@
 From these tables, we can infer which classes are more or less entangled with others.
-The results for each dataset is shown in figures \ref{fig:clster_confmat_sim} \ref{fig:clster_confmati_filt} and \ref{fig:clster_confmati_raw}.
+The results for each data set is shown in figures \ref{fig:clster_confmat_sim} \ref{fig:clster_confmati_filt} and \ref{fig:clster_confmati_raw}.
 We observe that the proton class is consistently assigned in a pure cluster.
@@
 \includegraphics[width=\textwidth]{custom_work/Filteredexamples.pdf}
-\caption[Filtered proton samples by cluster belonging]{Illustrating a sample of proton events from different K-means clusters from the filtered dataset.
+\caption[Filtered proton samples by cluster belonging]{Illustrating a sample of proton events from different K-means clusters from the filtered data set.
 Each row belongs to a single cluster corresponding to the filtered confusion matrix in figure \ref{fig:clster_confmati_filt}}\label{fig:filtered_vgg_clster_repr}
@@
 \includegraphics[width=\textwidth]{custom_work/Rawexamples.pdf}
-\caption[Full proton samples by cluster belonging]{Illustrating a sample of proton events from different K-means clusters from the raw dataset.
+\caption[Full proton samples by cluster belonging]{Illustrating a sample of proton events from different K-means clusters from the raw data set.
 Each row belongs to a single cluster corresponding to the full confusion matrix in figure \ref{fig:clster_confmati_raw}}\label{fig:full_vgg_clster_repr}
@@
 % Begin edit by DB (6/17/2020)
-The purpose of this study is to explore the application of unsupervised learning algorithms to event identification from an active target detector.
+The purpose of this study has been to explore the application of unsupervised learning algorithms to event identification from an active target detector.
 The necessity to identify events from raw data prior to full processing is becoming a major issue in the data analysis of detectors with complex responses such as the AT-TPC.
@@
 In summary, our study shows that unsupervised track classification with an implementation of the VGG16 + K-means approach is a viable solution.
-For future work, it is worth investigating whether the adaptation of bespoke models like the MIXAE algorithm will allow better performance at no significant cost to consistency.
+For future work, it is worth investigating whether the adaptation of  models like the MIXAE algorithm will allow better performance at no significant cost to consistency.
 %Especially considering the flexibility to use 3D input representations or auxiliary physical goals in the optimization procedure. 
@@
 \caption[vgg architectures]{Showing the details of the VGG network architectures.
-Network D trained on the ImageNet \cite{Russakovsky2015} dataset the network known as  VGG16 and is what we use in this thesis.}\label{tab:vgg}
+Network D trained on the ImageNet \cite{Russakovsky2015} data set the network known as  VGG16 and is what we use in this thesis.}\label{tab:vgg}
 \includegraphics[width=\textwidth]{plots/vgg_architectures.png}
@@
 
-To train the MIXAE clustering algorithm, we use the large simulated dataset with $M=80000$ points, evenly distributed between proton- and carbon-events.
+To train the MIXAE clustering algorithm, we use the large simulated data set with $M=80000$ points, evenly distributed between proton- and carbon-events.
 The algorithm is trained on a subset of $60000$ of these samples, and we track performance on the remaining $20000$ events.
@@
 
-For the full dataset the MIXAE hyperparameters converge to the same values as for the clean data:
+For the full data set the MIXAE hyperparameters converge to the same values as for the clean data:
 

### 2020-07-10 doc/Manuscript/article.tex
Updates from Overleaf

@@
 
-As an alternative to relying on a pre-trained model, we also consider the MIXAE algorithm \cite{Zhang}, a bespoke end-to-end clustering model which is specifically trained on the AT-TPC data.
+As an alternative to relying on a pre-trained model, we also consider the MIXAE algorithm \cite{Zhang}, which is an end-to-end clustering model specifically trained on the AT-TPC data.
 
@@
 While the top performing MIXAE runs outperform the K-means approach, its reliability suffers.
-However, the high performance achieved shines a light on valuable potential research into more bespoke models for unsupervised track identification.
+However, the high performance achieved shines a light on valuable potential research into more tailored models for unsupervised track identification.
 

### 2020-07-14 doc/Manuscript/article.tex
Updates from Overleaf

@@
 The important distinction to make between the clustering accuracy and the ARI is that the clustering accuracy is a simple comparison that does not account for chance assignments, while the ARI does.
-In effect, this means that the accuracy is a good heuristic for the performance which is tempered by the ARI.
+In effect, this means that the accuracy is a good heuristic for the performance, tempered by the ARI.
 %We refer to \citet{Hubert1985} for a detailed consideration of the ARI.
@@
 In addition to the performance measures reported in table \ref{tab:clstr_vgg}, it is interesting to observe which samples are wrongly assigned.
-To investigate this problem, we tabulate the assignments of samples relative to their ground truth labels.
+To investigate this problem, we compute the matrices $\mathbf{W}$ as shown in table \ref{tab:contingency}.
 From these tables, we can infer which classes are more or less entangled with others.
@@
 This is done to estimate to what degree the class separating information is encoded in the entirety of the latent space, or in some select regions.
-The results from the PCA analysis {\color{blue} using the top XX principal components} were virtually identical to the results sans the PCA, indicating that a lower-dimensional encoding of the data is sufficient for data clustering for this experiment.
+The results from the PCA analysis, using the top 100 principal components , were virtually identical to the results sans the PCA, indicating that a lower-dimensional encoding of the data is sufficient for data clustering for this experiment.
 %and so we omit them for brevity.

### 2020-07-16 doc/Manuscript/article.tex
Modified motivation for unsupervised ML -MPK

@@
 The $\chi^2$ approach used in the traditional analysis performed on the $^{46}$Ar data is extremely computationally expensive because it involves the simulation of thousands of tracks for each recorded event, which are performed each iteration of the Monte-Carlo fitting sequence.
-Even though the reaction of interest in the $^{46}$Ar experiment had the largest cross section (elastic scattering), the sheer number of events to analyse (in the hundreds of thousands) necessitated the use of parallel processing on a high performance computing cluster {\color{blue} is there a metric we can give here?
--MPK}.
+Even though the reaction of interest in the $^{46}$Ar experiment had the largest cross section (elastic scattering), the time spent on Monte Carlo fitting of {\em all} of the events produced in the experiment was the largest computational bottleneck in the analysis.
 In the case of an experiment where the reaction of interest would represent less than a few percent of the total cross section, this procedure would become highly inefficient and prohibitive.
@@
 However, it has been found that models which are trained at large scale will learn general features that are applicable to a variety of tasks.
-For example, large neural networks which are trained on the ImageNet data set \cite{Russakovsky2015} --- a diverse image classification task --- learn how to identify lines, edges, and other common shapes that are useful for numerous problems.
+For example, large neural networks which are trained on the ImageNet data set \cite{deng2009imagenet} --- a diverse image classification task --- learn how to identify lines, edges, and other common shapes that are useful for numerous problems.
 Thus, it is common practice to initialize the convolutional layers of a network with the pre-trained weights learned from ImageNet (or some other large data set).
@@
 Since we are building on prior knowledge in this case, learning becomes far more efficient, and better performance can often be achieved.
-\citet{Kuchera2019} used machine learning methods to classify the products of $^{46}$Ar reactions in the AT-TPC, and they found that a CNN initialized with ImageNet weights resulted in the most successful classification.
+\citet{Kuchera2019} used machine learning methods to classify the products of $^{46}$Ar reactions in the AT-TPC, and they found that a CNN initialized with weights trained on ImageNet data resulted in the most successful classification.
 
@@
 The MIXAE model comprises several autoencoders, each of which corresponds to a cluster.
-The number of autoencoders, and thus the number of clusters, has to be determined beforehand.
 Each autoencoder constructs a latent representation of a given example.
@@
 Examples are then assigned to the cluster with the highest score.
-
+The number of autoencoders, and thus the number of clusters, has to be determined beforehand.
 % The MIXAE algorithm clusters data by pairing a set of encoder-decoder neural networks (autoencoder) with an auxiliary assignment network: each autoencoder constructs a representation of a given event while the assignment network assigns a set of cluster probabilities to the event.
@@
 
-\noindent The learning objective encourages these assumptions to be met.
+\noindent The learning objective encourages these assumptions to be met. For a more formal consideration of the model objective and the assumptions made on the data by this model see \citet{Zhang}.
+
 The architecture is portrayed in Figure \ref{fig:mixae}, wherein tapered boxes denote a direction of compression in the network components.
-
-For a more formal consideration of these heuristics, the model objective and the assumptions made on the data by this model see \citet{Zhang}.
+In the figure each encoder and decoder pair constitute an autoencoder.
+The assignment of the cluster belonging of a sample, $\mathbf{\hat{x}}$, is taken to be the index of the maximal element in the vector $\mathbf{p}$ as shown in the right-most part of the figure.
+Finally, the model is trained end-to-end with backpropagation \cite{Linnainmaa1976}, as implemented in the machine learning library TensorFlow \cite{tensorflow}.
 
@@
 \end{figure}
-{\color{blue} This section ends quite abruptly -MPK}
 
@@
 To solve this problem we first define two sets; the ground truth labels $\mathbf{y} = [y_i, \hspace{0.2cm} y_{i +1}, \ldots]$ and the corresponding predictions $\mathbf{\hat{y}} = [\hat{y}_i, \hspace{0.2cm} \hat{y}_{i +1}, \ldots]$.
-Furthermore, let both $y_i$ and $\hat{y}_i$ be integer representations of an events ground truth, or predicted class, respectively.
+Furthermore, let both $y_i$ and $\hat{y}_i$ be integer representations of an events ground truth, and predicted class, respectively.
 
@@
 In effect, this means that the accuracy is a good heuristic for the performance, tempered by the ARI.
+
+Lastly, we introduce two terms to describe cluster quality: purity and quality.
+Purity is inferred by how much spread there is in the column between the ground truth labels in the matrix $\mathbf{W}$.
+A high-quality cluster will, in addition to being pure, also capture most entries the class represented by the cluster.
 %We refer to \citet{Hubert1985} for a detailed consideration of the ARI.
@@
 
-{\color{blue} This section may contain info that you can move to the metrics section??}
 In addition to the performance measures reported in table \ref{tab:clstr_vgg}, it is interesting to observe which samples are wrongly assigned.
@@
 The results for each data set is shown in figures \ref{fig:clster_confmat_sim} \ref{fig:clster_confmati_filt} and \ref{fig:clster_confmati_raw}.
-We observe that the proton class is consistently assigned in a pure cluster.
-Purity is inferred by how much spread there is in the column between the ground truth labels.
-A high-quality cluster will, in addition to being pure, also capture most entries the class represented by the cluster.
+We observe that the proton class is consistently assigned in a pure cluster..
 For example, consider the row corresponding to the proton class in figure \ref{fig:clster_confmati_filt}.
 The column corresponding to the largest entry in the proton row has zero other predicted classes in it.
-From this, we conclude that the proton cluster is a high quality, high purity cluster.
 
@@
 This is done to estimate to what degree the class separating information is encoded in the entirety of the latent space, or in some select regions.
-The results from the PCA analysis, using the top 100 principal components , were virtually identical to the results sans the PCA, indicating that a lower-dimensional encoding of the data is sufficient for data clustering for this experiment.
+The results from the PCA analysis, using the top 100 principal components , were virtually identical to the results sans the PCA.
+This result indicates that the class-encoding information is contained in a minority of the axes of variation in the data.
 %and so we omit them for brevity.
@@
 Additionally, we considered the deep convolutional embedded clustering (DCEC) by \citet{Guo2017} as well as the MIXAE method introduced by \citet{Zhang}.
-While we were able to reproduce the authors' results on their data, the DCEC algorithm proved unable to cluster AT-TPC in our implementation.
+While we were able to reproduce the authors' results on their data, the DCEC algorithm proved unable to cluster AT-TPC data in our implementation.
 However, this provides valuable insight as the seeds of the clusters are constructed by a K-means algorithm.
-This insight contrasts with our positive results from applying a pre-trained model with k-means and highlights potentially significant differences in models trained on a supervised or unsupervised objective for clustering tasks in nuclear physics.
+This insight contrasts with our positive results from applying a pre-trained classification model with k-means and highlights potentially significant differences in models trained on a supervised or unsupervised objective for clustering tasks in nuclear physics.
 

### 2020-07-20 doc/Manuscript/article.tex
Updates from Overleaf

@@
 
-The results of the clustering runs are included in table \ref{tab:clstr_vgg}.
-We observe that we are able to attain near-perfect clustering on simulated data and that there is a sharp decline in performance as we add noise by moving to the filtered and raw data sets.
+The results of the clustering runs are included in table \ref{tab:clstr_vgg}, we ran the k-means algorithm $N=1000$ times to assess the variability in the performance.
+We observe that the clustering on simulated data attains the highest performance, and that there is a decline in performance as we add noise by moving to the filtered and raw data sets.
+We observe that the k-means algorithm for the simulated data converges to either to a very good solution, or to a very poor one.
+This fluctuation is ordinarily suppressed with the convention for re-running the fit several times and picking the strongest performing one for prediction.
 
@@
 \centering 
-\caption[K-means on pre-trained model]{K-means clustering results on AT-TPC event data.
+\caption[K-means on pre-trained model]{K-means clustering results on AT-TPC event data, for $N=1000$ runs of the k-means algorithm.
 We observe that the performance predictably decreases with the amount of noise in the data.}\label{tab:clstr_vgg}

### 2020-07-29 doc/Manuscript/article.tex
Updates from Overleaf

@@
 The traditional Monte-Carlo event selection process, described in Section~\ref{Sec:Exp}, does not have a well-defined method to quantify the effectiveness of the event selection.
-In addition, the selection task produced a binary result only, either ``good" or``bad" relative to the event of interest.
-The data contain however many more classes of events.%, if only because of the presence of carbon in the target (from the isobutane gas).
+In addition, the selection task produced a binary result only, either a ``good" or``bad" fit to the event of interest.
+A ``bad'' fit is then assumed to be a different event type, and is removed from the analysis.
+%, if only because of the presence of carbon in the target (from the isobutane gas).
 In a broader perspective, an unsupervised classification algorithm would offer the possibility to ``discover" rare events which may not be expected or are overlooked.
@@
 \centering 
-\caption[K-means on pre-trained model]{K-means clustering results on AT-TPC event data, for $N=1000$ runs of the k-means algorithm.
+\caption[K-means on pre-trained model]{K-means clustering results on AT-TPC event data in the VGG-16 latent space, for $N=1000$ runs of the k-means algorithm.
 We observe that the performance predictably decreases with the amount of noise in the data.}\label{tab:clstr_vgg}

### 2020-08-04 doc/Manuscript/article.tex
Updates from Overleaf

@@
 
-\subsection{K-means clustering on the VGG16 latent space}
+\subsection{K-means clustering on the VGG16 latent space}\label{sec:kmeans_results}
 
@@
 \caption[MIXAE clustering performance]{MIXAE clustering performance on the ${}^{46} Ar$ experimental data.
-In contrast with the VGG-16 + K-means approach we observe  significant variations in performance.}\label{tab:clstr_mixae}
+In contrast with the VGG-16 + K-means approach we observe  significant variations in performance.
+{\color{blue}, however, the KMeans++ algorithm still clusters by reaction type, as evidenced when initializing $10$ random centroid seeds, and selecting the most separated clusters without supervision.
+The mean and standard deviations come from $N = 1000$ runs.
+}}\label{tab:clstr_mixae}
 \input{./plots/mixae_clustering.tex}
@@
 \section{VGG16}\label{app:vgg}
+
+The VGGNet models are a family of high-performing image classification, and object localization networks.
+In the VGGNet architecture a small kernel size is leveraged to increase expressive power in a very deep convolutional network.
+
 The choice of the kernel size is based on the fact that a stacked $3 \times 3$ kernel is equivalent to larger kernels in terms of the receptive field of the output.
@@
 Additionally, there is a regularizing effect from the lowered number of parameters and increased explanatory power from the additional non-linearities.
+The architecture is shown in detail in table \ref{tab:vgg}.
+
+VGGnet models are distributed freely with weights trained on the imagenet \cite{deng2009imagenet} image classification task.
+For the results in section \ref{sec:kmeans_results} we used a VGG16 model pre trained on imagenet data.
+
 \begin{table}
@@
 
-In the MIXAE algorithm the hyper-parameters to adjust are all the ordinary parameters that we introduced in table \ref{tab:convae_hyperparams}.
+In the MIXAE algorithm the hyper-parameters to adjust are all the ordinary parameters associated with a neueral network.
+We chose to base our neural network parameter choices on the VGG16 architecture.
+The parameters chosen for the autoencoders are listed in full in table \ref{tab:mixe_ae_hyperparams}.
+
 In addition to those parameters we have the weighting of the loss terms: $\theta$, $\alpha$ and $\gamma$.
 These weighting parameters are attached to the reconstruction loss, sample entropy and batch-wise entropy respectively \cite{Zhang}.
-We focused on the tuning of the clustering hyper-parameters, and heuristically set the autoencoder hyper-parameters to a shallow $3\times3$ convolutional network.
-
-The parameters chosen for the autoencoders are listed in full in table \ref{tab:mixe_ae_hyperparams}.
+We focused on the tuning of the clustering hyper-parameters, and defined the autoencoder hyper-parameters to be a shallow $3\times3$ convolutional network as detailed in the previous paragraph.
 
@@
 The algorithm is trained on a subset of $60000$ of these samples, and we track performance on the remaining $20000$ events.
-Since there are then only three remaining hyperparameters we choose to perform a coarse grid-search as described in section \ref{sec:hyperparam_search_arch}.Finally, for these parameters we re-ran the algorithm $N=10$ times to investigate the stability of the algorithm.
+On real data the algorithm is trained on an unlabelled set of data, and evaluated on an unlabelled subset\footnote{see table \ref{tab:data sets} for details}.
+Since there are then only three remaining hyperparameters we choose to perform a coarse grid-search for the optimal configuration.
+Finally, for the best parameters we re-ran the algorithm $N=10$ times to investigate the stability of the algorithm.
 

### 2020-08-04 doc/Manuscript/plots/mixae_clustering.tex
Updates from Overleaf

@@
 \midrule
-{} & Top 1 & $\mu \pm \sigma$ & Top 1 & $\mu \pm \sigma$ \\
-Simulated &  0.96 &  $0.74 \pm 0.16 $  & 0.84 &  $0.33 \pm 0.32 $\\
-Filtered  & 0.75 &  $0.71 \pm 0.04 $ & 0.52 &  $0.38 \pm 0.14 $\\
-Full      & 0.71 &  $0.61 \pm 0.07 $ &  0.32 &  $0.09 \pm 0.10 $   \\
+{} & {\color{red}Top 1}{\color{blue} init $= 10$} &  {\color{red}$\mu \pm \sigma$}{\color{blue} init $= 1$} & Top 1 & $\mu \pm \sigma$ \\
+Simulated &  0.96{\color{blue} $\pm 0.00$} &  $0.74 \pm 0.16 $  & 0.84 &  $0.33 \pm 0.32 $\\
+Filtered  & 0.75{\color{blue} $\pm 0.00$} &  $0.71 \pm 0.04 $ & 0.52 &  $0.38 \pm 0.14 $\\
+Raw & 0.71{\color{blue} $\pm 0.00$} &  $0.61 \pm 0.07 $ &  0.32 &  $0.09 \pm 0.10 $   \\
 \bottomrule

### 2020-08-05 doc/Manuscript/article.tex
Updates from Overleaf

@@
 We find that a $K$-means algorithm applied to the simulated data in the VGG16 latent space forms almost perfect clusters.
-Additionally, the VGG16+K-means approach finds high purity clusters of proton events for real experimental data.
+Additionally, the VGG16+$K$-means approach finds high purity clusters of proton events for real experimental data.
 We explore also the application of autoencoder neural networks to clustering by implementing.
@@
 
-The AT-TPC \cite{Bradt2017} is a novel type of detector designed specifically for nuclear physics experiments where the energies of the recoiling particles are very low compared to the energy required to escape target material.
+The Active-Target Time Projection Chamber (AT-TPC) \cite{Bradt2017} is a novel type of detector designed specifically for nuclear physics experiments where the energies of the recoiling particles are very low compared to the energy required to escape target material.
 The luminosity of nuclear physics experiments performed with fixed targets is directly proportional to the amount of material encountered by the beam.
@@
 Although this sounds like an advantage from the scientific point of view, it poses great challenges during the analysis phase, that are reminiscent of bubble chamber times and on par with event classification challenges in particle physics today, see for example the recent review of Mehta {\em et al.} \cite{mehta2019}.
-More often than not, the reaction channel of interest has one of the lowest cross sections, therefore the physicist is faced with the task of sorting out the corresponding events from the ``background" of other reaction channels.
+More often than not, the reaction channel of interest has one of the lowest cross sections.
+When analyzing the data one is therefore faced with the task of sorting out the corresponding events from the {\em background} of other reaction channels.
 
@@
 In some experiments, a labeled data set is unavailable.
-This could be  due to the inability to hand-label events, or the case where one does not know {\em a priori} the types and behaviors of the reactions present in the detector in order to generate a labeled, simulated data set.
+This could be  due to the inability to hand-label events, or the case where one does not know {\em a priori} the types and behaviors of the reactions present in the detector in order to generate a labeled and simulated data set.
 In the latter case, there must also exist a validation data set of real data, still requiring the ability to label a subset of the real data.
@@
 
-The detector was placed inside the bore of a MRI solenoid energized to $\sim 2$ Tesla.
+The detector was placed inside the bore of an MRI solenoid energized to $\sim 2$ Tesla.
 This axial magnetic field served the purpose of bending the trajectories of the recoil particles in order to i) increase their length and ii) provide a measurement of their bending radius, directly related to their magnetic rigidity.
@@
 % End added by DB (06/10/2020)
-In addition, the authors of Ref.~\cite{Kuchera2019} demonstrated that the purity and statistics of the data are improved with the use of ML.
+In addition, the authors of Ref.~\cite{Kuchera2019} demonstrated that the purity and statistics of the data are improved with the use of machine learning methods.
 
@@
 The task of labeling the different classes is then much less time-consuming and can potentially lead to the discovery of unforeseen types of events.
+Furthermore, it allows us to process larger amounts of data in a much shorter time.
 
@@
 
-In this section we give a brief overview of the data, for a more in-depth consideration we refer the reader to \cite{Mittig2015}, \cite{Suzuki2012} and  \cite{Bradt2017a}.
+In this section we give a brief overview of the data, for a more in-depth consideration we refer the reader to Refs.~\cite{Mittig2015,Suzuki2012,Bradt2017a}.
 

### 2020-08-05 doc/Manuscript/article.tex
Updates from Overleaf

@@
 A direct consequence of this concept is the inclusiveness of the experimental data recorded by this type of detector: any nuclear reaction happening within the target is recorded.
-Although this sounds like an advantage from the scientific point of view, it poses great challenges during the analysis phase, that are reminiscent of bubble chamber times and on par with event classification challenges in particle physics today, see for example the recent review of Mehta {\em et al.} \cite{mehta2019}.
+Although this sounds like an advantage from the scientific point of view, it poses great challenges during the analysis phase, that are reminiscent
+of bubble-chamber times \cite{Hough:1959} and on par with event classification challenges in particle physics today, see for example the recent review of Mehta {\em et al.} \cite{mehta2019}.
 More often than not, the reaction channel of interest has one of the lowest cross sections.
@@
 % motivating ML work:
-Because TPCs produce three-dimensional images of charged particle tracks, the event identification task is often akin to a visual inspection (comparable to the analyses made in the bubble chamber era), which is not practical nowadays because of the large quantities of data.
+Because TPCs produce three-dimensional images of charged particle tracks, the event identification task is often akin to a visual inspection (comparable to the analyses made in the bubble-chamber era \cite{Hough:1959}), which is not practical nowadays because of the large quantities of data.
 Machine learning techniques then appear as a promising prospect, in particular in the image recognition domain where much progress has been made recently \cite{mehta2019}.
@@
 In this representation, an event is a record of $512$ time-buckets for each of the $10^4$ detector pads. In our analysis, we represent each event as a down-sampled two-dimensional  projection.
-We chose to represent the data in two dimensions to facilitate the use of sophisticated image-recognition machine learning models, and this data representation was shown to successfully classify the events of this experiment in a supervised setting \cite{Kuchera2019}.
+We chose to represent the data in two dimensions to facilitate the use of advanced image-recognition machine learning models, and this data representation was shown to successfully classify the events of this experiment in a supervised setting \cite{Kuchera2019}.
 First, the time-series data was represented as a three-dimensional  point cloud, where each point contains the maximum charge in the time-series trace.
-We log-transform then min-max scale the charge, mapping the charge to the interval $[0, 1]$.
-The data is projected into two-dimensional  space by summing over the time-axis, then down-sampled into a $128\times128$ pixel image.
+We log-transform and map the data to the interval $[0, 1]$.
+The data is projected into a two-dimensional  space by summing over the time-axis.
+Thereafter the data is down-sampled into a $128\times128$ pixel image.
 %Additionally, we log-transform and normalize the data to the interval $[0, 1]$ for each point.
@@
 One of the significant considerations for the analysis of AT-TPC data is to inject machine learning methods for track identification at the best point in the analysis pipeline.
-Using raw data is advantageous as it provides an unbiased view of the event, but the data volume and noise levels might be prohibitive for analysis.
+Using raw data is advantageous as it provides an unbiased view of the event, but the data volume and noise levels might be prohibitive for the analysis.
 %On the opposite end of the pipeline, 
 Therefore, we incrementally add bias to the analysis by applying the algorithm further down the analysis pipeline, with the benefit being that more preprocessing improves the signal-to-noise ratio, possibly improving model performance.
-To explore this trade-off between model performance and preprocessing bias, we performed our analysis on simulated, cleaned and raw events.
+To explore this trade-off between model performance and preprocessing bias, we performed our analysis on simulated, cleaned and raw events as discussed below.
 
@@
 \midrule
-$\%$ Proton & $50$ & $25.3$ & $27.6$ \\ 
-$\%$ Carbon & $50$ & $12.0$ & $11.9$ \\
-$\%$ Other & $0$ & $62.8$ & $60.1$ \\
+Percentage Proton & $50$ & $25.3$ & $27.6$ \\ 
+Percentage Carbon & $50$ & $12.0$ & $11.9$ \\
+Percentage Other & $0$ & $62.8$ & $60.1$ \\
 \end{tabular}
@@
 
-A set of $N=40000$ simulated AT-TPC events per class was obtained  from a ${}^{46}$Ar$(p, p)$ experiment with the \lstinline{pytpc} package developed by \citet{Bradt2017a}.
+A set of $N=40000$ simulated AT-TPC events per class was obtained  from a ${}^{46}$Ar$(p, p)$ experiment with the \lstinline{pytpc} package developed by \citet{Bradt2017a}, see table \ref{tab:class_distr}.
 %Using the same parameters {\color{blue} what parameters are you referring to?

### 2020-08-05 doc/Manuscript/article.tex
Updates from Overleaf

@@
 
+% MHJ comment: this table is barely described in the text, do we need it? If so, we need to add more text.
+
 \begin{table}
@@
 The events analyzed in this section were retrieved from the ${}^{46}$Ar resonant proton scattering experiment recorded with the AT-TPC. 
-While we denote these events as raw, it is important to note that what we mean is a raw two-dimensional representation of peak-only events.
+While we denote these events as raw, it is important to note that what we intend is a raw two-dimensional representation of peak-only events.
 
@@
 As we saw in the previous section, the detector picks up a significant amount of noise.
-We split the noise broadly in two categories,  one being random-uncorrelated noise and the second is structured noise.
+We split the noise broadly in two categories,  one being randomly uncorrelated noise and the second one being structured noise.
 The former can be quite trivially removed with a nearest-neighbour algorithm that checks if a point in the event is close to any other.
@@
 
-We illustrate two filtered events in figure \ref{fig:samples_filtered}. These are the same events as shown in figure \ref{fig:samples}, but with the Hough and nearest neighbours filtering applied.
+We illustrate two filtered events in figure \ref{fig:samples_filtered}.
+These are the same events as shown in figure \ref{fig:samples}, but with the Hough transform and nearest neighbours filtering applied.
 
@@
 The traditional Monte-Carlo event selection process, described in Section~\ref{Sec:Exp}, does not have a well-defined method to quantify the effectiveness of the event selection.
-In addition, the selection task produced a binary result only, either a ``good" or``bad" fit to the event of interest.
-A ``bad'' fit is then assumed to be a different event type, and is removed from the analysis.
+In addition, the selection task produced a binary result only, either a {\em good} or {\em bad} fit to the event of interest.
+A {\em bad} fit is then assumed to be a different event type, and is removed from the analysis.
 %, if only because of the presence of carbon in the target (from the isobutane gas).
-In a broader perspective, an unsupervised classification algorithm would offer the possibility to ``discover" rare events which may not be expected or are overlooked.
+In a broader perspective, an unsupervised classification algorithm would offer the possibility to {\em discover} rare events which may not be expected or are overlooked.
 These events would likely be filtered out using the traditional methods.
@@
 
-The $\chi^2$ approach used in the traditional analysis performed on the $^{46}$Ar data is extremely computationally expensive because it involves the simulation of thousands of tracks for each recorded event, which are performed each iteration of the Monte-Carlo fitting sequence.
+The $\chi^2$ approach used in the traditional analysis performed on the $^{46}$Ar data is extremely expensive from a computational stand because it involves the simulation of thousands of tracks for each recorded event.
+These events are in turn simulated for each iteration of the Monte Carlo fitting sequence.
 Even though the reaction of interest in the $^{46}$Ar experiment had the largest cross section (elastic scattering), the time spent on Monte Carlo fitting of {\em all} of the events produced in the experiment was the largest computational bottleneck in the analysis.
 In the case of an experiment where the reaction of interest would represent less than a few percent of the total cross section, this procedure would become highly inefficient and prohibitive.
+Adding to this the large amount of data produced in this experiment (with even larger data sets expected in future experiments), simply begs for more efficient analysis tools.
 %Most of the CPU time would be spent on simulating events that are not of interest.
@@
 %{\color{orange} Yes, but also this section is a necessary bridge for reproducibility. Maybe it's better as an appendix?}
+
+% MHJ comment: his table is barely mentioned in the text, do we need it?
 
@@
 
-% The clustering algorithm we found to give the strongest result is the K-means algorithm.
+% The clustering algorithm we found to give the strongest result is the $K$-means algorithm.
 This algorithm is contingent on having a distance measurement between two datapoints.
@@
 
-In contrast with the classification work of \citet{Kuchera2019}, we do not assume access to ground truth labels and are trying to solve a fundamentally different learning problem.
+In contrast with the classification work of \citet{Kuchera2019}, we do not assume access to ground truth labels and we are trying to solve a fundamentally different learning problem.
 Thus, rather than fine-tuning a pre-trained network under the supervised learning regime, we extract the output of the pre-trained network's last convolutional layer as a latent representation of the events, where each event is represented as a vector in $\R^{8192}$.
-We then cluster events based on this representation using the Scikit-Learn implementation of the K-means++ algorithm with default parameters \cite{Pedregosa2011}.
+We then cluster events based on this representation using the Scikit-Learn implementation of the $K$-means algorithm with default parameters \cite{Pedregosa2011}.
 
@@
 As such, I would remove them and add the necessary details to the previous paragraph (or replace them with a small paragraph).]} Building on the work by \citet{Kuchera2019}, we also use the VGG16 pre-trained network as a baseline for the clustering performance.
-We begin by considering a classical K-means approach to clustering.
-However, the output from the VGG16 network is very high dimensional which traditionally conflicts with good K-means results.
+We begin by considering a classical $K$-means approach to clustering.
+However, the output from the VGG16 network is very high dimensional which traditionally conflicts with good $K$-means results.
 With output vectors in $\R^{8192}$ the ratio between $L_p$ distances goes to one with increasing dimensionality, as shown by \citet{Aggarwal}.
@@
 They note that it is then possible that if the class information lies in some sub-space of the latent data, $L_p$-norms may provide useful clusterings even in high dimensional spaces.
-To investigate this, we perform a $L_2$-norm based K-means clustering using the full VGG16 representation and the $10^2$ first principal components only.
+To investigate this, we perform a $L_2$-norm based $K$-means clustering using the full VGG16 representation and the $10^2$ first principal components only.
 
 %{\color{blue} Are there supposed to be results here? -MPK}
-%\subsection{K-means}
+%\subsection{$K$-means}
 
-%We begin by investigating the K-means clustering algorithm on the VGG16 latent space. 
+%We begin by investigating the $K$-means clustering algorithm on the VGG16 latent space. 
 % The user a VGG16 model that is pre-trained on the ImageNet data set and subsequently applied to the AT-TPC events without the classification layers.
 This creates a set of vectors $\boldsymbol{x} \in \R^{8192}$ as our new representation of the events.
-To cluster, we use the \lstinline{scikit-learn} implementation of the K-means++ algorithm with default parameters \cite{Pedregosa2011}.
+To cluster, we use the \lstinline{scikit-learn} implementation of the $K$-means++ algorithm with default parameters \cite{Pedregosa2011}.
 
@@
 In the figure each encoder and decoder pair constitute an autoencoder.
-The assignment of the cluster belonging of a sample, $\mathbf{\hat{x}}$, is taken to be the index of the maximal element in the vector $\mathbf{p}$ as shown in the right-most part of the figure.
-Finally, the model is trained end-to-end with backpropagation \cite{Linnainmaa1976}, as implemented in the machine learning library TensorFlow \cite{tensorflow}.
+The assignment of the cluster belonging to a sample, $\mathbf{\hat{x}}$, is taken to be the index of the maximal element in the vector $\mathbf{p}$ as shown in the right-most part of the figure.
+Finally, the model is trained end-to-end with back propagation \cite{Linnainmaa1976}, as implemented in the machine learning library TensorFlow \cite{tensorflow}.
 
@@
 In the face of such missing data, a model is ordinarily assessed by measures that do not depend on knowing the ground truth for any samples.
-However, as our work aims to explore the application of unsupervised methods to track identification, we chose the ${}^{46}$Ar data we have some ground truth labels to evaluate our models.
+However, as our work aims to explore the application of unsupervised methods to track identification, we chose the ${}^{46}$Ar data since we have some ground truth labels to evaluate our models.
 The measures we use to evaluate the models we then chose to be ones which measure the similarities between two arbitrary sets of clustering assignments while holding one to be the ground truth.
@@
 
-To compute the metrics we have to solve problem introduced by the arbitrary labels of a clustering algorithm.
+To compute the metrics we have to solve problems introduced by the arbitrary labels of a clustering algorithm.
 That is, we do not know which predicted assignment should correspond to a proton or carbon event.
@@
 To solve this problem we first define two sets; the ground truth labels $\mathbf{y} = [y_i, \hspace{0.2cm} y_{i +1}, \ldots]$ and the corresponding predictions $\mathbf{\hat{y}} = [\hat{y}_i, \hspace{0.2cm} \hat{y}_{i +1}, \ldots]$.
-Furthermore, let both $y_i$ and $\hat{y}_i$ be integer representations of an events ground truth, and predicted class, respectively.
+Furthermore, we let both $y_i$ and $\hat{y}_i$ be integer representations of an event's ground truth, and predicted class, respectively.
 
@@
 \caption{General form of a contingency table.
-$y_n$ and $\hat{y}_n$ are the ground truth labels and clustering assignments respectively.
+Here $y_n$ and $\hat{y}_n$ are the ground truth labels and clustering assignments respectively.
 The $w_{ij}$'s then describe how many samples are in clusters $y_i$ and simultaneously in $\hat{y}_j$.}\label{tab:contingency}
@@
     \item Subtract $\mathbf{W}$ from its maximum value to find the bipartite graph representation of the assignment problem.
-    \item Use an algorithm, like the Hungarian algorithm, to solve the assignment problem, and take the average of the values in $\mathbf{W}$ this solution indicates.
+    \item Use an algorithm, like the Hungarian algorithm \cite{burkard2012}, to solve the assignment problem, and take the average of the values in $\mathbf{W}$ this solution indicates.
 This average is the clustering accuracy.
@@
 
-We first explore the results of applying a K-means approach on the latent space of a pre-trained network.
+We first explore the results of applying a $K$-means approach on the latent space of a pre-trained network.
 Subsequently, we investigate the performance of the MIXAE algorithm as outlined in section \ref{sec:mixae}.
 
-\subsection{K-means clustering on the VGG16 latent space}\label{sec:kmeans_results}
+\subsection{$K$-means clustering on the VGG16 latent space}\label{sec:kmeans_results}
 
-The results of the clustering runs are included in table \ref{tab:clstr_vgg}, we ran the k-means algorithm $N=1000$ times to assess the variability in the performance.
+The results of the clustering runs are included in table \ref{tab:clstr_vgg}, we ran the $K$-means algorithm $N=1000$ times to assess the variability in the performance.
 We observe that the clustering on simulated data attains the highest performance, and that there is a decline in performance as we add noise by moving to the filtered and raw data sets.
-We observe that the k-means algorithm for the simulated data converges to either to a very good solution, or to a very poor one.
+We observe that the $K$-means algorithm for the simulated data converges to either to a very good solution, or to a very poor one.
 This fluctuation is ordinarily suppressed with the convention for re-running the fit several times and picking the strongest performing one for prediction.
@@
 \centering 
-\caption[K-means on pre-trained model]{K-means clustering results on AT-TPC event data in the VGG-16 latent space, for $N=1000$ runs of the k-means algorithm.
+\caption[$K$-means on pre-trained model]{$K$-means clustering results on AT-TPC event data in the VGG-16 latent space, for $N=1000$ runs of the $K$-means algorithm.
 We observe that the performance predictably decreases with the amount of noise in the data.}\label{tab:clstr_vgg}
@@
 	\includegraphics[width=\textwidth]{custom_work/Simulatedvgg_conf_mat.pdf} 
-	\caption[Pre-trained network - confusion matrices]{Confusion matrix for the K-means clustering of simulated AT-TPC events.
+	\caption[Pre-trained network - confusion matrices]{Confusion matrix for the $K$-means clustering of simulated AT-TPC events.
 The true labels indicate samples belonging to the p (proton), or the carbon (C) class}\label{fig:clster_confmat_sim}
@@
 	\includegraphics[width=\textwidth]{custom_work/Filteredvgg_conf_mat.pdf}
-	\caption[Pre-trained network - confusion matrices]{Confusion matrix for the K-means clustering of filtered AT-TPC events.
+	\caption[Pre-trained network - confusion matrices]{Confusion matrix for the $K$-means clustering of filtered AT-TPC events.
 The true labels indicate samples belonging to the p (proton), carbon (C), or other classes.
@@
 	\includegraphics[width=\textwidth]{custom_work/Rawvgg_conf_mat.pdf}
-\caption[Pre-trained network - confusion matrices]{Confusion matrix for the K-means clustering of raw AT-TPC events.
+\caption[Pre-trained network - confusion matrices]{Confusion matrix for the $K$-means clustering of raw AT-TPC events.
 The true labels indicate samples belonging to the p (proton), carbon (C), or other classes.
@@
 \includegraphics[width=\textwidth]{custom_work/Filteredexamples.pdf}
-\caption[Filtered proton samples by cluster belonging]{Illustrating a sample of proton events from different K-means clusters from the filtered data set.
+\caption[Filtered proton samples by cluster belonging]{Illustrating a sample of proton events from different $K$-means clusters from the filtered data set.
 Each row belongs to a single cluster corresponding to the filtered confusion matrix in figure \ref{fig:clster_confmati_filt}}\label{fig:filtered_vgg_clster_repr}
@@
 \includegraphics[width=\textwidth]{custom_work/Rawexamples.pdf}
-\caption[Full proton samples by cluster belonging]{Illustrating a sample of proton events from different K-means clusters from the raw data set.
+\caption[Full proton samples by cluster belonging]{Illustrating a sample of proton events from different $K$-means clusters from the raw data set.
 Each row belongs to a single cluster corresponding to the full confusion matrix in figure \ref{fig:clster_confmati_raw}}\label{fig:full_vgg_clster_repr}
@@
 \caption[MIXAE clustering performance]{MIXAE clustering performance on the ${}^{46} Ar$ experimental data.
-In contrast with the VGG-16 + K-means approach we observe  significant variations in performance.
+In contrast with the VGG-16 + $K$-means approach we observe  significant variations in performance.
 {\color{blue}, however, the KMeans++ algorithm still clusters by reaction type, as evidenced when initializing $10$ random centroid seeds, and selecting the most separated clusters without supervision.
@@
 
-Applied to raw data the MIXAE captures a proton cluster in a similar vein to the K-means approach.
+Applied to raw data the MIXAE captures a proton cluster in a similar vein to the $K$-means approach.
 The MIXAE forms a proton-majority cluster, but with a significant portion of the more noisy proton events being clustered with the other-class.
@@
 We observe that the algorithm correctly caputers a majority proton-event cluster in cluster 0.
-However, in contrast with the K-means approach this cluster is contaminated to some extend with both carbon and other events.
+However, in contrast with the $K$-means approach this cluster is contaminated to some extend with both carbon and other events.
 }\label{fig:mixae_confmat_raw}
@@
 
-The most striking result we present in this work is the success of the K-means approach.
-As noted by \cite{Aggarwal} distance measures become less informative in higher dimensional spaces, but the K-means algorithm clusters events well in a very high-dimensional space.
-Another surprise is the stability of the K-means algorithm.
+The most striking result we present in this work is the success of the $K$-means approach.
+As noted by \cite{Aggarwal} distance measures become less informative in higher dimensional spaces, but the $K$-means algorithm clusters events well in a very high-dimensional space.
+Another surprise is the stability of the $K$-means algorithm.
 A stability we attribute to the quality of the VGG16 latent space in creating class-separating sub-spaces.
@@
 
-It is also interesting to compare and contrast the clustering results from the MIXAE model with those of the VGG16$+$K-means.
+It is also interesting to compare and contrast the clustering results from the MIXAE model with those of the VGG16$+$$K$-means.
 In particular, the discrepancy in stability is worth noting.
-While the top performing MIXAE runs outperform the K-means approach, its reliability suffers.
+While the top performing MIXAE runs outperform the $K$-means approach, its reliability suffers.
 However, the high performance achieved shines a light on valuable potential research into more tailored models for unsupervised track identification.
@@
 In addition to the results presented in this section, we performed clustering with a number of different algorithms included in the \lstinline{scikit-learn} package.
-None of them provided any notable differences from the K-means results or were significantly worse.
+None of them provided any notable differences from the $K$-means results or were significantly worse.
 Notably, the DBSCAN algorithm \cite{Ester96adensity-based}\cite{Bergstra2012} failed to provide any useful clustering results.
-We find this important as one of the significant drawbacks of K-means, and the deep clustering algorithm presented in section \ref{sec:mixae}, is that they are all dependent on pre-determining the number of clusters.
+We find this important as one of the significant drawbacks of $K$-means, and the deep clustering algorithm presented in section \ref{sec:mixae}, is that they are all dependent on pre-determining the number of clusters.
 This is not the case for DBSCAN.
@@
 While we were able to reproduce the authors' results on their data, the DCEC algorithm proved unable to cluster AT-TPC data in our implementation.
-However, this provides valuable insight as the seeds of the clusters are constructed by a K-means algorithm.
-This insight contrasts with our positive results from applying a pre-trained classification model with k-means and highlights potentially significant differences in models trained on a supervised or unsupervised objective for clustering tasks in nuclear physics.
+However, this provides valuable insight as the seeds of the clusters are constructed by a $K$-means algorithm.
+This insight contrasts with our positive results from applying a pre-trained classification model with $K$-means and highlights potentially significant differences in models trained on a supervised or unsupervised objective for clustering tasks in nuclear physics.
 
@@
 
-In particular, the ability of the K-means algorithm in picking out clear proton clusters from the VGG16 latent space lends itself well to an exploratory phase of analysis, where clusters of events corresponding to different reaction channels could be later identified by the experimenter.
-Another interesting facet of the K-means clustering is its consistent performance.
+In particular, the ability of the $K$-means algorithm in picking out clear proton clusters from the VGG16 latent space lends itself well to an exploratory phase of analysis, where clusters of events corresponding to different reaction channels could be later identified by the experimenter.
+Another interesting facet of the $K$-means clustering is its consistent performance.
 As shown in table \ref{tab:clstr_vgg}, the variance is zero for the performance metrics.
@@
 However, as can be seen from the non-proton clusters in figures \ref{fig:clster_confmati_filt} and \ref{fig:clster_confmati_raw},  this does not necessarily imply that the physical signals are correspondingly clear.
-A caveat to the K-means method is that the number of clusters has to be specified in advance.
+A caveat to the $K$-means method is that the number of clusters has to be specified in advance.
 Each experiment then has to be considered in light of possible reaction channels to determine a sensible number of clusters for this approach.
@@
 The same caveat is present in the MIXAE implementation.
-While it shows better optimal performance than the K-means method, some inconsistencies were observed, that were notably not evident from the unsupervised training-objectives of the model.
+While it shows better optimal performance than the $K$-means method, some inconsistencies were observed, that were notably not evident from the unsupervised training-objectives of the model.
 These two factors currently conspire to limit its immediate applicability, and more developments are needed for this approach.
 
-In summary, our study shows that unsupervised track classification with an implementation of the VGG16 + K-means approach is a viable solution.
+In summary, our study shows that unsupervised track classification with an implementation of the VGG16 + $K$-means approach is a viable solution.
 For future work, it is worth investigating whether the adaptation of  models like the MIXAE algorithm will allow better performance at no significant cost to consistency.
@@
 {\color{orange}[flip this to positive ]} It is also clear that there is significant potential in both avenues explored in this work.
-%n particular, the surprising ability of the K-means algorithm in picking out clear proton clusters from the VGG16 latent space lends itself well to an exploratory phase of analysis.
-%Another surprising facet of the K-means clustering is its consistent performance.
+%n particular, the surprising ability of the $K$-means algorithm in picking out clear proton clusters from the VGG16 latent space lends itself well to an exploratory phase of analysis.
+%Another surprising facet of the $K$-means clustering is its consistent performance.
 As shown in table \ref{tab:clstr_vgg}, the variance is zero for the performance metrics.
@@
 
-%Additionally, a clear disadvantage to the K-means method is that the researcher has to specify the number of clusters in the data themselves.
+%Additionally, a clear disadvantage to the $K$-means method is that the researcher has to specify the number of clusters in the data themselves.
 Each experiment then has to be considered in light of possible reaction channels to determine a sensible number of clusters for this approach.
@@
 %The same weakness is present in the MIXAE implementation.
-Additionally, while it showed better optimal performance than the K-means method, it struggles with inconsistent results.
+Additionally, while it showed better optimal performance than the $K$-means method, it struggles with inconsistent results.
 This inconsistency was also notably not evident from the unsupervised training-objectives of the model.
@@
 
-%In summary, we recommend that researchers performing unsupervised track classification begin with an implementation of the VGG16 + K-means approach.
+%In summary, we recommend that researchers performing unsupervised track classification begin with an implementation of the VGG16 + $K$-means approach.
 For future work, it is worth investigating whether the adaptations of bespoke models like the MIXAE algorithm will allow better performance at no significant cost to consistency.

### 2020-08-05 doc/Manuscript/article.tex
Updates from Overleaf

@@
 The results for each data set is shown in figures \ref{fig:clster_confmat_sim} \ref{fig:clster_confmati_filt} and \ref{fig:clster_confmati_raw}.
-We observe that the proton class is consistently assigned in a pure cluster..
-For example, consider the row corresponding to the proton class in figure \ref{fig:clster_confmati_filt}.
+We observe that the proton class is consistently assigned in a pure cluster.
+For example, consider the row corresponding to the proton class in Fig.~\ref{fig:clster_confmati_filt}.
 The column corresponding to the largest entry in the proton row has zero other predicted classes in it.
@@
 This high-quality cluster also appears in the clustering of raw data.
-From figure \ref{fig:clster_confmati_raw}, we observe that there is a high purity proton cluster.
-In contrast to the filtered data we observe that the deterioration in performance can largely be ascribed to the algorithm creating a proton + other cluster and a carbon + other cluster.
+From Fig.~\ref{fig:clster_confmati_raw}, we observe that there is a high purity proton cluster.
+In contrast to the filtered data we observe that the deterioration in performance can largely be ascribed to the algorithm creating a proton plus another cluster and a carbon plus another cluster.
 
@@
 We repeat this analysis using a Principle Component Analysis (PCA) dimensionality reduction\footnote{
-PCA is a common techinque to find the significant variations in data by projecting the data along a subset of its covariance matrix eigenvectors \cite{Marsland2009}
+PCA is a common techinque to find the significant variations in data by projecting the data along a subset of its covariance matrix eigenvectors \cite{vidal2016,Marsland2009}
 } on the latent space of the VGG16 model.
 This is done to estimate to what degree the class separating information is encoded in the entirety of the latent space, or in some select regions.
-The results from the PCA analysis, using the top 100 principal components , were virtually identical to the results sans the PCA.
-This result indicates that the class-encoding information is contained in a minority of the axes of variation in the data.
+The results from the PCA analysis, using the top 100 principal components, were virtually identical to our previous results not containing  the PCA analysis.
+This in an interesting observation which indicates that the class-encoding information is contained in a minority of the axes of variation in the data.
 %and so we omit them for brevity.
 
-To further investigate the clusters presented in the matrix in figures \ref{fig:clster_confmati_filt} and \ref{fig:clster_confmati_raw}, we sample examples from the proton samples belonging to different clusters for the filtered and full data.
-We display these examples in figures \ref{fig:filtered_vgg_clster_repr} and \ref{fig:full_vgg_clster_repr} respectively.
-Form these figures we gleam that the more noisy proton events are being clustered together with the amorphous other class.
-This is in line with the intuitive expectation from viewing the clustering tables in figures \ref{fig:clster_confmati_filt} and \ref{fig:clster_confmati_raw}.
+To further investigate the clusters presented in the matrix in figures \ref{fig:clster_confmati_filt} and \ref{fig:clster_confmati_raw}, we study examples from the proton samples belonging to different clusters for the filtered and full data.
+We display these examples in Figs.~\ref{fig:filtered_vgg_clster_repr} and \ref{fig:full_vgg_clster_repr}, respectively.
+Form these figures we observe that the more noisy proton events are being clustered together with the amorphous other class events.
+This is in line with the intuitive expectation from viewing the clustering results in Figs.~\ref{fig:clster_confmati_filt} and \ref{fig:clster_confmati_raw}.
 
@@
 \includegraphics[width=\textwidth]{custom_work/Filteredexamples.pdf}
-\caption[Filtered proton samples by cluster belonging]{Illustrating a sample of proton events from different $K$-means clusters from the filtered data set.
-Each row belongs to a single cluster corresponding to the filtered confusion matrix in figure \ref{fig:clster_confmati_filt}}\label{fig:filtered_vgg_clster_repr}
+\caption[Filtered proton samples by cluster belonging]{Illustration of  a sample of proton events from different $K$-means clusters originating from the filtered data set.
+Each row belongs to a single cluster corresponding to the filtered confusion matrix in Fig.~\ref{fig:clster_confmati_filt}}\label{fig:filtered_vgg_clster_repr}
 \end{figure}
@@
 \includegraphics[width=\textwidth]{custom_work/Rawexamples.pdf}
-\caption[Full proton samples by cluster belonging]{Illustrating a sample of proton events from different $K$-means clusters from the raw data set.
-Each row belongs to a single cluster corresponding to the full confusion matrix in figure \ref{fig:clster_confmati_raw}}\label{fig:full_vgg_clster_repr}
+\caption[Full proton samples by cluster belonging]{Illustration of  a sample of proton events from different $K$-means clusters originating from the raw data set.
+Each row belongs to a single cluster corresponding to the full confusion matrix in Fig.~\ref{fig:clster_confmati_raw}}\label{fig:full_vgg_clster_repr}
 \end{figure}
@@
 To build on this result we will in this section explore the application of the mixture of autoencoders (MIXAE) algorithm introduced in section \ref{sec:mixae}.
-For details on hyper-parameter tuning and the experimental procedure for training the MIXAE algorithm see appendix \ref{ax: mixae}
+For details on hyper-parameter tuning and the experimental procedure for training the MIXAE algorithm see the appendix \ref{ax: mixae}.
 
-After hyperparameter tuning, each highest performing model is run $10$ times. The results are listed in table \ref{tab:clstr_mixae}.
-We observe that, while the algorithm can achieve very strong performance, the performance varies.
+With the best set of hyperparameters, each highest performing model is thereafter run $10$ times. The results are listed in table \ref{tab:clstr_mixae}.
+We observe that, while the algorithm can achieve a very strong performance, the performance varies.
 In some cases the MIXAE model converges to a seemingly good configuration, based on its unsupervised training goals.
-However, when inspecting its clustering performance against labelled data that seemingly good model does no better than random.
-This happens more frequently with raw data, indicating an interaction with the noise levels present in the events.
+However, when inspecting its clustering performance against labelled data, the seemingly good model does no better than a model based on random selection.
+This happens more frequently with the raw data, indicating an interaction with the noise levels present in the events.
 
@@
 
-As with the VGG16 + KMeans approach we wish to further investigate the clustering results.
+As with the VGG16 + $K$-Means approach we wish to further investigate the clustering results.
 Taking the best performing MIXAE model on filtered and raw data we tabulate the clusters against their labels.
@@
 
-On filtered data the highest performing MiXAE model achieves strong separation of the other-class, but curiously creates two proton-majority clusters.
+On filtered data the highest performing MIXAE model achieves strong separation of the other-class, but curiously creates two proton-majority clusters.
 
@@
 The most striking result we present in this work is the success of the $K$-means approach.
-As noted by \cite{Aggarwal} distance measures become less informative in higher dimensional spaces, but the $K$-means algorithm clusters events well in a very high-dimensional space.
+As noted by \cite{Aggarwal}, distance measures become less informative in higher dimensional spaces, but the $K$-means algorithm clusters events well in a very high-dimensional space.
 Another surprise is the stability of the $K$-means algorithm.
-A stability we attribute to the quality of the VGG16 latent space in creating class-separating sub-spaces.
+We attribute this stability to the quality of the VGG16 latent space in creating class-separating sub-spaces.
 While the separations are not perfect, the stability and quality of the proton track identification create solid empirical grounding for applying this approach to other active target experiments.
 
-It is also interesting to compare and contrast the clustering results from the MIXAE model with those of the VGG16$+$$K$-means.
+It is also interesting to compare and contrast the clustering results from the MIXAE model with those of the VGG16 with the  $K$-means approach.
 In particular, the discrepancy in stability is worth noting.
 While the top performing MIXAE runs outperform the $K$-means approach, its reliability suffers.
-However, the high performance achieved shines a light on valuable potential research into more tailored models for unsupervised track identification.
+However, the high performance achieved indicates that this may represent a valuable potential research path into more tailored models for unsupervised track identification.
 
@@
 Notably, the DBSCAN algorithm \cite{Ester96adensity-based}\cite{Bergstra2012} failed to provide any useful clustering results.
-We find this important as one of the significant drawbacks of $K$-means, and the deep clustering algorithm presented in section \ref{sec:mixae}, is that they are all dependent on pre-determining the number of clusters.
+We find this important as one of the significant drawbacks of $K$-means, and the deep clustering algorithm presented in section \ref{sec:mixae}, is that they depend  on pre-determining the number of clusters.
 This is not the case for DBSCAN.
 
-Additionally, we considered the deep convolutional embedded clustering (DCEC) by \citet{Guo2017} as well as the MIXAE method introduced by \citet{Zhang}.
+Additionally, we considered the deep convolutional embedded clustering (DCEC) approach by \citet{Guo2017} as well as the MIXAE method introduced by \citet{Zhang}.
 While we were able to reproduce the authors' results on their data, the DCEC algorithm proved unable to cluster AT-TPC data in our implementation.

### 2020-08-05 doc/Manuscript/article.tex
Updates from Overleaf

@@
 % uncomment for submission to NIM A
-\documentclass[review,number,sort&compress]{elsarticle}
+\documentclass[review,sort&compress]{elsarticle}
 \usepackage{lineno}
-\linenumbers
+%\linenumbers
 
@@
 Active target experiments \sep Machine Learning \sep Unsupervised Learning \sep Autoencoder Neural Networks
-\PACS
+%\PACS
 %% PACS codes here, in the form: \PACS code \sep code
@@
 
-We display two different events from the ${}^{46}$Ar experiment in figure \ref{fig:samples}.
+We display two different events from the ${}^{46}$Ar experiment in Fig.~\ref{fig:samples}.
 The top row illustrates a carbon event with a large fraction of noise, while the bottom row shows a proton event almost free of noise.
@@
 
-We illustrate two filtered events in figure \ref{fig:samples_filtered}.
-These are the same events as shown in figure \ref{fig:samples}, but with the Hough transform and nearest neighbours filtering applied.
+We illustrate two filtered events in Fig.~\ref{fig:samples_filtered}.
+These are the same events as shown in Fig.~\ref{fig:samples}, but with the Hough transform and nearest neighbours filtering applied.
 
@@
 
-The architecture is portrayed in Figure \ref{fig:mixae}, wherein tapered boxes denote a direction of compression in the network components.
+The architecture is portrayed in Fig.~\ref{fig:mixae}, wherein tapered boxes denote a direction of compression in the network components.
 In the figure each encoder and decoder pair constitute an autoencoder.
@@
 From these tables, we can infer which classes are more or less entangled with others.
-The results for each data set is shown in figures \ref{fig:clster_confmat_sim} \ref{fig:clster_confmati_filt} and \ref{fig:clster_confmati_raw}.
+The results for each data set is shown in  Figs.~\ref{fig:clster_confmat_sim} \ref{fig:clster_confmati_filt} and \ref{fig:clster_confmati_raw}.
 We observe that the proton class is consistently assigned in a pure cluster.
@@
 
-To further investigate the clusters presented in the matrix in figures \ref{fig:clster_confmati_filt} and \ref{fig:clster_confmati_raw}, we study examples from the proton samples belonging to different clusters for the filtered and full data.
+To further investigate the clusters presented in the matrix in Figs.~\ref{fig:clster_confmati_filt} and \ref{fig:clster_confmati_raw}, we study examples from the proton samples belonging to different clusters for the filtered and full data.
 We display these examples in Figs.~\ref{fig:filtered_vgg_clster_repr} and \ref{fig:full_vgg_clster_repr}, respectively.
@@
 
+% This table, 5,  is not properly explained, what is init=10? and what is top 1? 
 \begin{table}[H]
@@
 In contrast with the VGG-16 + $K$-means approach we observe  significant variations in performance.
-{\color{blue}, however, the KMeans++ algorithm still clusters by reaction type, as evidenced when initializing $10$ random centroid seeds, and selecting the most separated clusters without supervision.
+The $K$-means algorithm still clusters by reaction type, as evidenced when initializing $10$ random centroid seeds, and selecting the most separated clusters without supervision.
 The mean and standard deviations come from $N = 1000$ runs.
-}}\label{tab:clstr_mixae}
+}\label{tab:clstr_mixae}
 \input{./plots/mixae_clustering.tex}
@@
 Taking the best performing MIXAE model on filtered and raw data we tabulate the clusters against their labels.
-These tables are present in figures \ref{fig:mixae_confmat_filtered} and \ref{fig:mixae_confmat_raw} for filtered and raw data, respectively.
+These tables are present in Figs.~\ref{fig:mixae_confmat_filtered} and \ref{fig:mixae_confmat_raw} for filtered and raw data, respectively.
 
@@
 This result indicates that the clusters are very clearly defined in the VGG16 latent space.
-However, as can be seen from the non-proton clusters in figures \ref{fig:clster_confmati_filt} and \ref{fig:clster_confmati_raw},  this does not necessarily imply that the physical signals are correspondingly clear.
+However, as can be seen from the non-proton clusters in Figs.~\ref{fig:clster_confmati_filt} and \ref{fig:clster_confmati_raw},  this does not necessarily imply that the physical signals are correspondingly clear.
 A caveat to the $K$-means method is that the number of clusters has to be specified in advance.
@@
 The same caveat is present in the MIXAE implementation.
-While it shows better optimal performance than the $K$-means method, some inconsistencies were observed, that were notably not evident from the unsupervised training-objectives of the model.
+While it shows better optimal performance than the $K$-means method, some inconsistencies were observed that were notably not evident from the unsupervised training-objectives of the model.
 These two factors currently conspire to limit its immediate applicability, and more developments are needed for this approach.
 
-In summary, our study shows that unsupervised track classification with an implementation of the VGG16 + $K$-means approach is a viable solution.
+In summary, our study shows that unsupervised track classification with an implementation of the VGG16 and the  $K$-means approach is a viable solution.
 For future work, it is worth investigating whether the adaptation of  models like the MIXAE algorithm will allow better performance at no significant cost to consistency.

### 2020-08-06 doc/Manuscript/article.tex
Updates from Overleaf

@@
 First, the time-series data was represented as a three-dimensional  point cloud, where each point contains the maximum charge in the time-series trace.
-We log-transform and map the data to the interval $[0, 1]$.
-The data is projected into a two-dimensional  space by summing over the time-axis.
+We log-transform the data and perform a min-max scaling in order to  map the data to the interval $[0, 1]$.
+The data is projected onto a two-dimensional  space by summing over the time-axis.
 Thereafter the data is down-sampled into a $128\times128$ pixel image.
@@
 %{\color{orange}[RS Another abrupt end to a paragraph. Could probably use some rounding out.]} 
+\subsection*{Acknowledgements}
+MHJ’s work is supported by U.S. National Science Foundation Grants No. PHY-1404159 and PHY-2013047.
 

### 2020-08-06 doc/Manuscript/article.tex
Updates from Overleaf

@@
 
-The Active-Target Time Projection Chamber (AT-TPC) \cite{Bradt2017} is a novel type of detector designed specifically for nuclear physics experiments where the energies of the recoiling particles are very low compared to the energy required to escape target material.
+The Active-Target Time Projection Chamber (AT-TPC) \cite{Bradt2017} is a novel type of detector designed specifically for nuclear physics experiments where the energies of the recoiling particles are very low compared to the energy required to escape the target material.
 The luminosity of nuclear physics experiments performed with fixed targets is directly proportional to the amount of material encountered by the beam.
@@
 In addition, the properties of the recoil particles are modified while traversing the target material, affecting the resolutions that can be achieved.
-This necessary balance between luminosity and resolution is particularly difficult when performing experiments with rare isotope beams, because of the low intensities available.
+This necessary balance between luminosity and resolution is particularly difficult when performing experiments with rare isotope beams, chiefly because of the low intensities available.
 
@@
 A direct consequence of this concept is the inclusiveness of the experimental data recorded by this type of detector: any nuclear reaction happening within the target is recorded.
-Although this sounds like an advantage from the scientific point of view, it poses great challenges during the analysis phase, that are reminiscent
-of bubble-chamber times \cite{Hough:1959} and on par with event classification challenges in particle physics today, see for example the recent review of Mehta {\em et al.} \cite{mehta2019}.
+Although this sounds like an advantage from the scientific point of view, it poses great challenges during the analysis phase that are reminiscent
+of bubble-chamber times and on par with event classification challenges in particle physics today, see for example the recent review of Mehta {\em et al.} \cite{mehta2019}.
 More often than not, the reaction channel of interest has one of the lowest cross sections.
@@
 
-\begin{table}
-\centering
-\caption{Event class percentages for each of the data sets used in this analysis.
-The decrease in the "other" class of events from full to filtered data owes to the thresholding of events.
-If an event contains fewer than $20$ data points it is discarded }\label{tab:class_distr}
-\begin{tabular}{lccc}
-\toprule
-{} & Simulated & Full & Filtered \\
-\midrule
-Percentage Proton & $50$ & $25.3$ & $27.6$ \\ 
-Percentage Carbon & $50$ & $12.0$ & $11.9$ \\
-Percentage Other & $0$ & $62.8$ & $60.1$ \\
-\end{tabular}
-\end{table}
+%\begin{table}
+%\centering
+%\caption{Event class percentages for each of the data sets used in this %analysis.
+The decrease in the "other" class of events from full to filtered %data owes to the thresholding of events.
+If an event contains fewer than $20$ %data points it is discarded }\label{tab:class_distr}
+%\begin{tabular}{lccc}
+%\toprule
+%{} & Simulated & Full & Filtered \\
+%\midrule
+%Percentage Proton & $50$ & $25.3$ & $27.6$ \\ 
+%Percentage Carbon & $50$ & $12.0$ & $11.9$ \\
+%Percentage Other & $0$ & $62.8$ & $60.1$ \\
+%\end{tabular}
+%\end{table}
 
@@
 
-A set of $N=40000$ simulated AT-TPC events per class was obtained  from a ${}^{46}$Ar$(p, p)$ experiment with the \lstinline{pytpc} package developed by \citet{Bradt2017a}, see table \ref{tab:class_distr}.
+A set of $N=40000$ simulated AT-TPC events per class was obtained  from a ${}^{46}$Ar$(p, p)$ experiment with the {\em pytpc} package developed by \citet{Bradt2017a}.
+%, see table \ref{tab:class_distr}.
 %Using the same parameters {\color{blue} what parameters are you referring to?
@@
 \caption[Displaying unfiltered events in two and three dimensions]{Two- and three-dimensional representations of two events from the ${}^{46}$Ar experiment.
-Each row is one event in two projections, where the color intensity of each point indicates higher charge values recorded by the detector.}\label{fig:samples}
+Each row is one event in two projections, where the color intensity of each point indicates higher charge values recorded by the detector.
+The top row illustrates a carbon event with a large fraction of noise, while the bottom row shows a proton event almost free of noise.}\label{fig:samples}
 \end{figure}
@@
 We split the noise broadly in two categories,  one being randomly uncorrelated noise and the second one being structured noise.
-The former can be quite trivially removed with a nearest-neighbour algorithm that checks if a point in the event is close to any other.
-To remove the correlated noise, researchers at the NSCL developed an algorithm based on the Hough transform \cite{Newman1972}.
+The former can be quite trivially removed with a nearest-neighbor algorithm, see for example \cite{hastie2009}, that checks if a point in the event is close to any other.
+To remove the correlated noise, researchers at the National Superconducting Cyclotron Laboratory of Michigan State University, developed an algorithm based on the Hough transform \cite{Newman1972}.
 This transformation is a common technique in computer vision, used to identify common geometric shapes like lines and circles, and has been used extensively in high-energy particle physics since the bubble-chamber era \cite{Hough:1959}.
@@
 
-We illustrate two filtered events in Fig.~\ref{fig:samples_filtered}.
-These are the same events as shown in Fig.~\ref{fig:samples}, but with the Hough transform and nearest neighbours filtering applied.
+We illustrate two filtered events in Fig.~\ref{fig:samples_filtered}. These are the same events as shown in Fig.~\ref{fig:samples}, but with the Hough transform and nearest-neighbor filtering applied.
 
@@
 Each row is one event in two projections, where the lightness of each point indicates higher charge values.
-These events have been filtered with a nearest neighbors algorithm and an algorithm based on the Hough transform, described in section \ref{sec:filtered}}\label{fig:samples_filtered}
+These events have been filtered with a nearest neighbors algorithm and an algorithm based on the Hough transform \cite{Newman1972}.}
 \end{figure}
@@
 %The filtering task relied on a set of $\chi^2$ distributions based on different criteria, 
-The traditional Monte-Carlo event selection process, described in Section~\ref{Sec:Exp}, does not have a well-defined method to quantify the effectiveness of the event selection.
+The traditional Monte Carlo event selection process, described in Section~\ref{Sec:Exp}, does not have a well-defined method to quantify the effectiveness of the event selection.
 In addition, the selection task produced a binary result only, either a {\em good} or {\em bad} fit to the event of interest.
@@
 In the case of an experiment where the reaction of interest would represent less than a few percent of the total cross section, this procedure would become highly inefficient and prohibitive.
-Adding to this the large amount of data produced in this experiment (with even larger data sets expected in future experiments), simply begs for more efficient analysis tools.
+Adding to this the large amount of data produced in this experiment (with even larger data sets expected in future experiments), the analysis simply begs for more efficient analysis tools.
 %Most of the CPU time would be spent on simulating events that are not of interest.
@@
 % MHJ comment: his table is barely mentioned in the text, do we need it?
-
-\begin{table}[hbtp]
-\centering
-\caption{Descriptions of number of events in the data.}\label{tab:data sets}
-\begin{tabular}{lccc}
-\toprule
-{} & Simulated & Full & Filtered \\
-\midrule
-Total &  $8000$ & $51891$ & $49169$ \\
-labelled & $2400$ & $1774$ &  $1582$ \\ 
-\bottomrule
-\end{tabular}
-\end{table}
 
@@
 
-\noindent The learning objective encourages these assumptions to be met. For a more formal consideration of the model objective and the assumptions made on the data by this model see \citet{Zhang}.
+The learning objective encourages these assumptions to be met. For a more formal consideration of the model objective and the assumptions made on the data by this model see \citet{Zhang}.
 
 The architecture is portrayed in Fig.~\ref{fig:mixae}, wherein tapered boxes denote a direction of compression in the network components.
-In the figure each encoder and decoder pair constitute an autoencoder.
+In the figure each encoder and decoder pair makes up an autoencoder.
 The assignment of the cluster belonging to a sample, $\mathbf{\hat{x}}$, is taken to be the index of the maximal element in the vector $\mathbf{p}$ as shown in the right-most part of the figure.
@@
 
-\begin{figure}[tb]
+\begin{figure}[tbh]
 	\centering
@@
 \subsection{Measuring performance}
-Unsupervised machine learning very typically goes hand-in-hand with a lack of ground-truth labelled data.
+Unsupervised machine learning  is often accompanied with a  lack of ground-truth labelled data.
 In the face of such missing data, a model is ordinarily assessed by measures that do not depend on knowing the ground truth for any samples.
 However, as our work aims to explore the application of unsupervised methods to track identification, we chose the ${}^{46}$Ar data since we have some ground truth labels to evaluate our models.
-The measures we use to evaluate the models we then chose to be ones which measure the similarities between two arbitrary sets of clustering assignments while holding one to be the ground truth.
+The measures we use to evaluate the models we then chose to be those which measure the similarities between two arbitrary sets of clustering assignments, while holding one to be the ground truth.
 
-We measure the performance of the clustering algorithms by two functions: clustering accuracy and the adjusted rand index (ARI)\cite{Hubert1985}.
+We measure the performance of the clustering algorithms by two functions: the clustering accuracy and the adjusted rand index (ARI)\cite{Hubert1985}.
 Both of these measurements fall in the range $ [0, \hspace{0.1cm}1]$, where $0$ denotes a complete disagreement and $1$ a complete agreement between the ground truth and predicted labels.
@@
 \end{equation}
-
+The quantities $f_i$ and $e_i$ are defined in table {tab:contingency}.
 The important distinction to make between the clustering accuracy and the ARI is that the clustering accuracy is a simple comparison that does not account for chance assignments, while the ARI does.
@@
 
-The performance is measured by comparing model predictions on the labelled subset of data (see table \ref{tab:data sets}).
+The performance is measured by comparing model predictions on the labelled subset of data (see table \ref{tab:data sets} in  appendix B for more details).
 
@@
 
-The results of the clustering runs are included in table \ref{tab:clstr_vgg}, we ran the $K$-means algorithm $N=1000$ times to assess the variability in the performance.
+The results of the clustering runs are included in table \ref{tab:clstr_vgg1}.
+We ran the $K$-means algorithm $N=100$ times with $M=10$ initializations, per run, of the cluster centroids to assess the variability in the performance.
+To quantify the results we report the best performing model (Top 1), and the mean and standard deviation for the result.
+
 We observe that the clustering on simulated data attains the highest performance, and that there is a decline in performance as we add noise by moving to the filtered and raw data sets.
-We observe that the $K$-means algorithm for the simulated data converges to either to a very good solution, or to a very poor one.
-This fluctuation is ordinarily suppressed with the convention for re-running the fit several times and picking the strongest performing one for prediction.
+The results are shown in table \ref{tab:clstr_vgg10}.
 
@@
 \centering 
-\caption[$K$-means on pre-trained model]{$K$-means clustering results on AT-TPC event data in the VGG-16 latent space, for $N=1000$ runs of the $K$-means algorithm.
-We observe that the performance predictably decreases with the amount of noise in the data.}\label{tab:clstr_vgg}
-\input{./plots/kmeans_vgg.tex}
+\caption[$K$-means on pre-trained model]{$K$-means clustering results on AT-TPC event data in the VGG-16 latent space, for $N=100$ runs of the $K$-means algorithm with $M=10$ initializations.
+We observe that the performance predictably decreases with the amount of noise in the data.}\label{tab:clstr_vgg10}
+\input{./plots/kmeans_vgg_init10.tex}
 \end{table}
 
-In addition to the performance measures reported in table \ref{tab:clstr_vgg}, it is interesting to observe which samples are wrongly assigned.
+\noindent The lack of variability is explained by the number of initializations.
+As can be seen from table \ref{tab:clstr_vgg1} where we run the $K$-means algorithm $N=1000$ times with $M=1$ initializations of the centroids.
+
+\begin{table}[H]
+\centering 
+\caption[$K$-means on pre-trained model]{$K$-means clustering results on AT-TPC event data in the VGG-16 latent space, for $N=1000$ runs of the $K$-means algorithm with $M=1$ initializations.
+We observe that there is significant variability in the results, which is ordinarily masked by picking the best-performing model by re-running the algorithm.
+}\label{tab:clstr_vgg1}
+\input{./plots/kmeans_vgg_init1.tex}
+\end{table}
+
+In addition to the performance measures reported in table \ref{tab:clstr_vgg1}, it is interesting to observe which samples are wrongly assigned.
 To investigate this problem, we compute the matrices $\mathbf{W}$ as shown in table \ref{tab:contingency}.
@@
 
-% This table, 5,  is not properly explained, what is init=10? and what is top 1? 
+% This table is not properly explained, what is init=10? and what is top 1? 
 \begin{table}[H]
@@
 In contrast with the VGG-16 + $K$-means approach we observe  significant variations in performance.
-The $K$-means algorithm still clusters by reaction type, as evidenced when initializing $10$ random centroid seeds, and selecting the most separated clusters without supervision.
-The mean and standard deviations come from $N = 1000$ runs.
 }\label{tab:clstr_mixae}
@@
 The true labels indicate samples belonging to the p (proton), carbon (C), or \texttt{other} classes.
-We observe that the algorithm correctly caputers a majority proton-event cluster in cluster 0.
-However, in contrast with the $K$-means approach this cluster is contaminated to some extend with both carbon and other events.
+We observe that the algorithm correctly captures a majority proton-event cluster in cluster 0.
+However, in contrast with the $K$-means approach this cluster is contaminated to some extent with both carbon events and other events.
 }\label{fig:mixae_confmat_raw}
 \end{figure}
-
-\subsection{Performance comparison}
 
@@
 None of them provided any notable differences from the $K$-means results or were significantly worse.
-Notably, the DBSCAN algorithm \cite{Ester96adensity-based}\cite{Bergstra2012} failed to provide any useful clustering results.
+Notably, the DBSCAN algorithm \cite{Ester96adensity-based,Bergstra2012} failed to provide any useful clustering results.
 We find this important as one of the significant drawbacks of $K$-means, and the deep clustering algorithm presented in section \ref{sec:mixae}, is that they depend  on pre-determining the number of clusters.
@@
 Another interesting facet of the $K$-means clustering is its consistent performance.
-As shown in table \ref{tab:clstr_vgg}, the variance is zero for the performance metrics.
+As shown in table \ref{tab:clstr_vgg1}, the variance is zero for the performance metrics.
 This result indicates that the clusters are very clearly defined in the VGG16 latent space.
 However, as can be seen from the non-proton clusters in Figs.~\ref{fig:clster_confmati_filt} and \ref{fig:clster_confmati_raw},  this does not necessarily imply that the physical signals are correspondingly clear.
+{\color{orange} Furthermore, the unsupervised metric that decides which of the $M$ $K$-means initializations perform the best does not necessarily coincide perfectly with separating the event classes.
+This is evidenced by the highest performing model for the raw data showing up in table \ref{tab:clstr_vgg1}, and being filtered out in table \ref{tab:clstr_vgg10}}
+
 A caveat to the $K$-means method is that the number of clusters has to be specified in advance.
@@
 
+\begin{table}[hbtp]
+\centering
+\caption{Descriptions of number of events in the data.}\label{tab:data sets}
+\begin{tabular}{lccc}
+\toprule
+{} & Simulated & Full & Filtered \\
+\midrule
+Total &  $8000$ & $51891$ & $49169$ \\
+labelled & $2400$ & $1774$ &  $1582$ \\ 
+\bottomrule
+\end{tabular}
+\end{table}
+
 \bibliography{bibliography}
 \end{document}
+
+{tab:contingency}

### 2020-08-06 doc/Manuscript/article.tex
small fixes on references and changes to 5.1

@@
 
+Within the context of machine learning methods applied to the analysis of nuclear physics experiments, 
+the purpose of this work is thus to explore the application of unsupervised learning algorithms to event identification from an active target detector.
+The necessity to identify events from raw data prior to full processing is becoming a major issue in the data analysis of detectors with complex responses such as the AT-TPC.
+After these introductory words, we review the experimental information in the next section.
+In section \ref{sec:datadescript} we describe the set up of the data while sections \ref{sec:me
+
 \section{Experimental Details} 
@@
 
-\section{Data Preparation}
+\section{Data Preparation}\label{sec:datadescript}
 
@@
 Each row is one event in two projections, where the lightness of each point indicates higher charge values.
-These events have been filtered with a nearest neighbors algorithm and an algorithm based on the Hough transform \cite{Newman1972}.}
+These events have been filtered with a nearest neighbors algorithm and an algorithm based on the Hough transform \cite{Newman1972}.}\label{fig:samples_filtered}
 \end{figure}
@@
 \end{equation}
-The quantities $f_i$ and $e_i$ are defined in table {tab:contingency}.
+The quantities $f_i$ and $e_i$ are defined in table \ref{tab:contingency}.
 The important distinction to make between the clustering accuracy and the ARI is that the clustering accuracy is a simple comparison that does not account for chance assignments, while the ARI does.
@@
 
-With the best set of hyperparameters, each highest performing model is thereafter run $10$ times. The results are listed in table \ref{tab:clstr_mixae}.
+With the best set of hyperparameters, each highest performing model is thereafter run $10$ times\footnote{The MIXAE model is significantly more computationally expensive to train than a $K$-means model.
+Reulting in the skew in number of runs in the two cases.}.
+The results are listed in table \ref{tab:clstr_mixae}.
 We observe that, while the algorithm can achieve a very strong performance, the performance varies.
@@
 \centering 
-\caption[MIXAE clustering performance]{MIXAE clustering performance on the ${}^{46} Ar$ experimental data.
-In contrast with the VGG-16 + $K$-means approach we observe  significant variations in performance.
-}\label{tab:clstr_mixae}
+\caption[MIXAE clustering performance]{MIXAE clustering performance on the ${}^{46} Ar$ experimental data with $N=10$ runs of the algorithm.
+In contrast with the VGG-16 + $K$-means approach we observe  significant variations in performance.}\label{tab:clstr_mixae}
 \input{./plots/mixae_clustering.tex}
@@
 
-In addition to the results presented in this section, we performed clustering with a number of different algorithms included in the \lstinline{scikit-learn} package.
+In addition to the results presented in this section, we performed clustering with a number of different algorithms included in the Scikit-learn package of \citet{Pedregosa2011}.
 None of them provided any notable differences from the $K$-means results or were significantly worse.
@@
 
-\section{Conclusions and Perspectives}\label{sec{conclusion}}
+\section{Conclusions and Perspectives}\label{sec:conclusion}
 % Begin edit by DB (6/17/2020)
@@
 \subsection*{Acknowledgements}
-MHJ’s work is supported by U.S. National Science Foundation Grants No. PHY-1404159 and PHY-2013047.
+MHJ’s work is supported by U.S.
+National Science Foundation Grants No.
+PHY-1404159 and PHY-2013047.
+This project has also received support from the INTPART project of the Research Council of Norway (Grant No.
+288125).
 

### 2020-08-06 doc/Manuscript/article.tex
Updates from Overleaf

@@
 After these introductory words, we review the experimental information in the next section.
-In section \ref{sec:datadescript} we describe the set up of the data while sections \ref{sec:me
+In section \ref{sec:datadescript} we describe the set up of the data while sections \ref{sec:methods} and \ref{sec:results}
+present the methods applied and our results and discussions, respectively. Finally, our conclusions and perspectives for future work are presented in section \ref{sec:conclusion}.
 
@@
 The algorithm is trained on a subset of $60000$ of these samples, and we track performance on the remaining $20000$ events.
-On real data the algorithm is trained on an unlabelled set of data, and evaluated on an unlabelled subset\footnote{see table \ref{tab:data sets} for details}.
+On real data the algorithm is trained on an unlabelled set of data, and evaluated on a labelled subset\footnote{see table \ref{tab:data sets} for details}.
 Since there are then only three remaining hyperparameters we choose to perform a coarse grid-search for the optimal configuration.

### 2020-08-06 doc/Manuscript/article.tex
Updates from Overleaf

@@
 \caption[$K$-means on pre-trained model]{$K$-means clustering results on AT-TPC event data in the VGG-16 latent space, for $N=100$ runs of the $K$-means algorithm with $M=10$ initializations.
+To quantify the results we report the best performing model (Top 1), and the mean and standard deviation for the result ($\mu\pm \sigma$).
 We observe that the performance predictably decreases with the amount of noise in the data.}\label{tab:clstr_vgg10}
@@
 \caption[$K$-means on pre-trained model]{$K$-means clustering results on AT-TPC event data in the VGG-16 latent space, for $N=1000$ runs of the $K$-means algorithm with $M=1$ initializations.
+To quantify the results we report the best performing model (Top 1), and the mean and standard deviation for the result ($\mu\pm \sigma$).
 We observe that there is significant variability in the results, which is ordinarily masked by picking the best-performing model by re-running the algorithm.
@@
 \caption[MIXAE clustering performance]{MIXAE clustering performance on the ${}^{46} Ar$ experimental data with $N=10$ runs of the algorithm.
+To quantify the results we report the best performing model (Top 1), and the mean and standard deviation for the result ($\mu\pm \sigma$).
 In contrast with the VGG-16 + $K$-means approach we observe  significant variations in performance.}\label{tab:clstr_mixae}
@@
 However, as can be seen from the non-proton clusters in Figs.~\ref{fig:clster_confmati_filt} and \ref{fig:clster_confmati_raw},  this does not necessarily imply that the physical signals are correspondingly clear.
-{\color{orange} Furthermore, the unsupervised metric that decides which of the $M$ $K$-means initializations perform the best does not necessarily coincide perfectly with separating the event classes.
-This is evidenced by the highest performing model for the raw data showing up in table \ref{tab:clstr_vgg1}, and being filtered out in table \ref{tab:clstr_vgg10}}
+Furthermore, the unsupervised metric that decides which of the $M$ $K$-means initializations perform the best does not necessarily coincide perfectly with separating the event classes.
+This is evidenced by the highest performing model for the raw data showing up in table \ref{tab:clstr_vgg1}, and being filtered out from table \ref{tab:clstr_vgg10}
 
@@
 \centering
-\caption{Hyperparameters selected for the autoencoder components of the MIXAE algorithm, see table \ref{tab:convae_hyperparams} for a full description of the parameters.}\label{tab:mixe_ae_hyperparams}
+\caption{Hyperparameters selected for the autoencoder components of the MIXAE algorithm}\label{tab:mixe_ae_hyperparams}
 \setlength{\extrarowheight}{15pt}
@@
 
+\section{Data}
+
+The data used  for the analysis in this work were partitioned as shown in table \ref{tab:data sets}. 
 \begin{table}[hbtp]

### 2020-08-06 doc/Manuscript/article.tex
Updates from Overleaf

@@
 	\caption[Mixture of autoencoders schematic]{Schematic of a MIXAE model.
-A sample $\hat{\boldsymbol{x}}$ is compressed to set of lower-dimensional representations $\{\boldsymbol{z}^{(i)}\}$ by $N$ autoencoders.
-These samples are concatenated and passed through an auxiliary assignment network that predicts a confidence of cluster belonging for each autoencoder.
-For further details see the text.
-Figure adapted from \citet{Zhang}}
+A sample $\hat{\boldsymbol{x}}$ is compressed to set of lower-dimensional representations $\{\boldsymbol{z}^{(i)}\}$ by $N$ autoencoders \cite{Zhang}.
+These samples are concatenated and passed through an auxiliary assignment network that predicts a confidence of cluster belonging for each autoencoder.}
+	%For further details see the text. Figure adapted from \citet{Zhang}}
 	\label{fig:mixae}
@@
 With the best set of hyperparameters, each highest performing model is thereafter run $10$ times\footnote{The MIXAE model is significantly more computationally expensive to train than a $K$-means model.
-Reulting in the skew in number of runs in the two cases.}.
+Resulting in the skew in number of runs in the two cases.}.
 The results are listed in table \ref{tab:clstr_mixae}.
@@
 
-VGGnet models are distributed freely with weights trained on the imagenet \cite{deng2009imagenet} image classification task.
-For the results in section \ref{sec:kmeans_results} we used a VGG16 model pre trained on imagenet data.
+VGGnet models are distributed freely with weights trained on the ImageNet \cite{deng2009imagenet} image classification task.
+For the results in section \ref{sec:kmeans_results} we used a VGG16 model pre trained on ImageNet data.
 
@@
 \caption[vgg architectures]{Showing the details of the VGG network architectures.
-Network D trained on the ImageNet \cite{Russakovsky2015} data set the network known as  VGG16 and is what we use in this thesis.}\label{tab:vgg}
+Network D trained on the ImageNet \cite{Russakovsky2015} data set the network known as  VGG16 and is what we use in this thesis.
+Table  }\label{tab:vgg}
 \includegraphics[width=\textwidth]{plots/vgg_architectures.png}
@@
 
-In the MIXAE algorithm the hyper-parameters to adjust are all the ordinary parameters associated with a neueral network.
+In the MIXAE algorithm the hyper-parameters to adjust are all the ordinary parameters associated with a neural network.
 We chose to base our neural network parameter choices on the VGG16 architecture.

### 2020-08-06 doc/Manuscript/article.tex
small editorial

@@
 \journal{Nuclear Instruments and Methods
-in Physics Research Section A: Accelerators, Spectrometers,
-Detectors and Associated Equipment}
+in Physics Research Section A}
 
@@
 % Modified by DB (6/8/2020)
-This article presents novel applications of machine learning methods to the problem of event separation in an active target detector, the Active-Target Time Projection Chamber (AT-TPC) \cite{Bradt2017}.
-The overarching goal is to identify classes of events in the early stages of the data analysis, thereby improving its efficiency by limiting the computationally expensive processing to events of interest only, which are often not the most abundant.
+This article presents novel applications of unsupervised machine learning methods to the problem of event separation in an active target detector, the Active-Target Time Projection Chamber (AT-TPC) \cite{Bradt2017}.
+The overarching goal is to group similar events in the early stages of the data analysis, thereby improving efficiency by limiting the computationally expensive processing of unnecessary events.
 %Machine learning presents an interesting avenue for researchers operating the AT-TPC, as traditional analysis methods of AT-TPC data are computationally expensive and have to fit all tracks against the event type of interest.
@@
 %%Graphical abstract
-\begin{graphicalabstract}
+%\begin{graphicalabstract}
 %\includegraphics{grabs}
-\end{graphicalabstract}
+%\end{graphicalabstract}
 
@@
 
-A set of $N=40000$ simulated AT-TPC events per class was obtained  from a ${}^{46}$Ar$(p, p)$ experiment with the {\em pytpc} package developed by \citet{Bradt2017a}.
+A set of $N=40000$ simulated AT-TPC events per class was obtained  from a ${}^{46}$Ar$(p, p)$ experiment with the \texttt{pytpc} package developed by \citet{Bradt2017a}.
 %, see table \ref{tab:class_distr}.
@@
 
-For validation, we set a subset of the simulated data to be labeled and treat the rest as unlabeled data.
+For validation, we select a subset of the simulated data to be labeled and treat the rest as unlabeled data.
 We chose this partition to consist of  $15\%$ of the data.
@@
 The events analyzed in this section were retrieved from the ${}^{46}$Ar resonant proton scattering experiment recorded with the AT-TPC. 
-While we denote these events as raw, it is important to note that what we intend is a raw two-dimensional representation of peak-only events.
+While we denote these events as raw, it is important to note that what we intend is a raw two-dimensional projection of events.
 
@@
 The assignment of the cluster belonging to a sample, $\mathbf{\hat{x}}$, is taken to be the index of the maximal element in the vector $\mathbf{p}$ as shown in the right-most part of the figure.
-Finally, the model is trained end-to-end with back propagation \cite{Linnainmaa1976}, as implemented in the machine learning library TensorFlow \cite{tensorflow}.
+Finally, the model is trained end-to-end with back propagation \cite{Linnainmaa1976}, as implemented in the machine learning package TensorFlow \cite{tensorflow}.
 
@@
 The results of the clustering runs are included in table \ref{tab:clstr_vgg1}.
-We ran the $K$-means algorithm $N=100$ times with $M=10$ initializations, per run, of the cluster centroids to assess the variability in the performance.
-To quantify the results we report the best performing model (Top 1), and the mean and standard deviation for the result.
+We ran the $K$-means algorithm $N=100$ times with $M=10$ initializations per run, of the cluster centroids to assess the variability in the performance.
+The K-means algorithm returns the best performing model of the $M$ initializations on the unsupervised objective.
+We report performance on the labelled subset of data, using the labels to identify the top-performing model (Top 1).
+Additionally, we report the mean and standard deviation of the algorithm on the $N$ trials, which indicate unsupervised performance.
+
+%The mean and standard deviation are metrics reported using no label information.
+%To quantify the results we report the best performing model (Top 1), and the mean and standard deviation for the result.
 
@@
 \caption[$K$-means on pre-trained model]{$K$-means clustering results on AT-TPC event data in the VGG-16 latent space, for $N=100$ runs of the $K$-means algorithm with $M=10$ initializations.
-To quantify the results we report the best performing model (Top 1), and the mean and standard deviation for the result ($\mu\pm \sigma$).
 We observe that the performance predictably decreases with the amount of noise in the data.}\label{tab:clstr_vgg10}
@@
 \caption[$K$-means on pre-trained model]{$K$-means clustering results on AT-TPC event data in the VGG-16 latent space, for $N=1000$ runs of the $K$-means algorithm with $M=1$ initializations.
-To quantify the results we report the best performing model (Top 1), and the mean and standard deviation for the result ($\mu\pm \sigma$).
-We observe that there is significant variability in the results, which is ordinarily masked by picking the best-performing model by re-running the algorithm.
+We observe that there is significant variability in the results, which is ordinarily masked by $M$ re-initializations that avoid local minima.
 }\label{tab:clstr_vgg1}
@@
 We repeat this analysis using a Principle Component Analysis (PCA) dimensionality reduction\footnote{
-PCA is a common techinque to find the significant variations in data by projecting the data along a subset of its covariance matrix eigenvectors \cite{vidal2016,Marsland2009}
+PCA is a common technique to find the significant variations in data by projecting the data along a subset of its covariance matrix eigenvectors \cite{vidal2016,Marsland2009}
 } on the latent space of the VGG16 model.
@@
 Furthermore, the unsupervised metric that decides which of the $M$ $K$-means initializations perform the best does not necessarily coincide perfectly with separating the event classes.
-This is evidenced by the highest performing model for the raw data showing up in table \ref{tab:clstr_vgg1}, and being filtered out from table \ref{tab:clstr_vgg10}
+This is evidenced by the highest performing model measured on labelled data (Top 1) for the raw data showing up in table \ref{tab:clstr_vgg1}, and being filtered out from table \ref{tab:clstr_vgg10}.
 
@@
 This project has also received support from the INTPART project of the Research Council of Norway (Grant No.
-288125).
+288125) and the Davidson Research Initiative.
 
@@
 In the VGGNet architecture a small kernel size is leveraged to increase expressive power in a very deep convolutional network.
+A tabulated view of the VGG models can be seen in \cite{Simonyan2014}.
 
@@
 Additionally, there is a regularizing effect from the lowered number of parameters and increased explanatory power from the additional non-linearities.
-The architecture is shown in detail in table \ref{tab:vgg}.
 
@@
 For the results in section \ref{sec:kmeans_results} we used a VGG16 model pre trained on ImageNet data.
-
-\begin{table}
-\caption[vgg architectures]{Showing the details of the VGG network architectures.
-Network D trained on the ImageNet \cite{Russakovsky2015} data set the network known as  VGG16 and is what we use in this thesis.
-Table  }\label{tab:vgg}
-\includegraphics[width=\textwidth]{plots/vgg_architectures.png}
-\end{table}
 

### 2020-08-06 doc/Manuscript/article.tex
Updates from Overleaf

@@
 \address{Department of Physics, University of Oslo, POB 1048 Oslo, N-0316 Oslo, Norway}
+\ead{robert@xal.no}
 
@@
 \address{Department of Physics and Center for Computing in Science Education, University of Oslo, POB 1048 Oslo, N-0316 Oslo, Norway}
-\ead{hjensen@frib.msu.edu}
-\ead[url]{http://mhjgit.github.io/info/doc/web/}
+%\ead[url]{http://mhjgit.github.io/info/doc/web/}
 
@@
 Additionally, the VGG16+$K$-means approach finds high purity clusters of proton events for real experimental data.
-We explore also the application of autoencoder neural networks to clustering by implementing.
-With autoencoder neural networks we find improved descriptions of data from experiments.
+We also explore the application of clustering the latent space of autoencoder neural networks for event separation.
+While these networks show strong performance, they suffer from high variability in their results.
+%With autoencoder neural networks we find improved descriptions of data from experiments.
 
@@
 %%Research highlights
-\begin{highlights}
-\item We present  novel applications of unsupervised machine learning methods to the analysis of event separations in active target experiments. 
-\item Autoencoder neural networks represent a promising path to the analysis of experimental data. 
-\end{highlights}
+%\begin{highlights}
+%\item We present  novel applications of unsupervised machine learning methods to the analysis %of event separations in active target experiments. 
+%\item Autoencoder neural networks represent a promising path to the analysis of experimental %data. 
+%\end{highlights}
 
@@
 From these tables, we can infer which classes are more or less entangled with others.
-The results for each data set is shown in  Figs.~\ref{fig:clster_confmat_sim} \ref{fig:clster_confmati_filt} and \ref{fig:clster_confmati_raw}.
+The results for each data set is shown in  Figs.~\ref{fig:clster_confmat_sim}, \ref{fig:clster_confmati_filt} and \ref{fig:clster_confmati_raw}.
 We observe that the proton class is consistently assigned in a pure cluster.
@@
 To build on this result we will in this section explore the application of the mixture of autoencoders (MIXAE) algorithm introduced in section \ref{sec:mixae}.
-For details on hyper-parameter tuning and the experimental procedure for training the MIXAE algorithm see the appendix \ref{ax: mixae}.
+For details on hyper-parameter tuning and the experimental procedure for training the MIXAE algorithm see the \ref{ax: mixae}.
 
@@
 While we were able to reproduce the authors' results on their data, the DCEC algorithm proved unable to cluster AT-TPC data in our implementation.
+Further details on these experiments are presented in the thesis by \citet{SolliThesis}.
+
 However, this provides valuable insight as the seeds of the clusters are constructed by a $K$-means algorithm.
@@
 Another interesting facet of the $K$-means clustering is its consistent performance.
-As shown in table \ref{tab:clstr_vgg1}, the variance is zero for the performance metrics.
+As shown in table \ref{tab:clstr_vgg10}, the variance is zero for the performance metrics.
 This result indicates that the clusters are very clearly defined in the VGG16 latent space.
@@
 The two examples of unsupervised machine learning methods studied in this work are a first encouraging step towards automated classification of events that could not only greatly reduce the resource cost of analysis, but also eventually boost the efficiency of the experiment by allowing post-trigger decisions based on such algorithm implemented in hardware.
-Much progress remains to be done before achieving these goals.
+%Much progress remains to be done before achieving these goals.
 % End edit by DB (6/17/2020)
@@
 PHY-1404159 and PHY-2013047.
+DB's work is supported by the U.S.
+Department of Energy, Office of Science, Office of Nuclear Physics, under Grant No.
+DE-SC0020451.
 This project has also received support from the INTPART project of the Research Council of Norway (Grant No.
@@
 
-\bibliography{bibliography}
+\bibliography{article}
 \end{document}

### 2020-11-30 doc/Manuscript/article.tex
Updates from Overleaf

@@
 \ead{robert@xal.no}
-
 \author{D.~Bazin}
 \address{Department of Physics and Astronomy and Facility for Rare Ion Beams and National Superconducting Cyclotron Facility, Michigan State University, East Lansing, MI 48824, USA}
+\author{M.~Hjorth-Jensen}
+\address{Department of Physics and Astronomy and Facility for Rare Ion Beams and National Superconducting Cyclotron Facility, Michigan State University, East Lansing, MI 48824, USA}
+\address{Department of Physics and Center for Computing in Science Education, University of Oslo, POB 1048 Oslo, N-0316 Oslo, Norway}
 \author{M.P.~Kuchera}
@@
 \address{Department of Mathematics and Computer Science, Davidson College, Davidson, North Carolina, USA}
-
-\author{M.~Hjorth-Jensen}
-\address{Department of Physics and Astronomy and Facility for Rare Ion Beams and National Superconducting Cyclotron Facility, Michigan State University, East Lansing, MI 48824, USA}
-\address{Department of Physics and Center for Computing in Science Education, University of Oslo, POB 1048 Oslo, N-0316 Oslo, Norway}
-%\ead[url]{http://mhjgit.github.io/info/doc/web/}
+\address{Department of Computer Science, University of North Carolina, Chapel Hill, North Carolina, USA}
 
@@
 % Modified by DB (6/8/2020)
-This article presents novel applications of unsupervised machine learning methods to the problem of event separation in an active target detector, the Active-Target Time Projection Chamber (AT-TPC) \cite{Bradt2017}.
-The overarching goal is to group similar events in the early stages of the data analysis, thereby improving efficiency by limiting the computationally expensive processing of unnecessary events.
 %Machine learning presents an interesting avenue for researchers operating the AT-TPC, as traditional analysis methods of AT-TPC data are computationally expensive and have to fit all tracks against the event type of interest.
 The latter presents a considerable challenge when either the space of reactions is not known prior to the analysis, or the event type of interest is not the most abundant.
-The application of unsupervised clustering algorithms to the analysis of two-dimensional projections of particle tracks from a resonant proton scattering experiment on $^{46}$Ar is introduced.
-We explore the performance of autoencoder neural networks and a pre-trained VGG16 \cite{Simonyan2014} convolutional neural network.
 %on two tasks: a semi-supervised classification task and the unsupervised clustering of particle tracks. 
@@
 On the clustering task,
+%With autoencoder neural networks we find improved descriptions of data from experiments.
+This article presents novel applications of unsupervised machine learning methods to the problem of event separation in an active target detector, the Active-Target Time Projection Chamber (AT-TPC) \cite{Bradt2017}.
+The overarching goal is to group similar events in the early stages of the data analysis, thereby improving efficiency by limiting the computationally expensive processing of unnecessary events.
+The application of unsupervised clustering algorithms to the analysis of two-dimensional projections of particle tracks from a resonant proton scattering experiment on $^{46}$Ar is introduced.
+We explore the performance of autoencoder neural networks and a pre-trained VGG16 \cite{Simonyan2014} convolutional neural network.
 We find that a $K$-means algorithm applied to the simulated data in the VGG16 latent space forms almost perfect clusters.
@@
 While these networks show strong performance, they suffer from high variability in their results.
-%With autoencoder neural networks we find improved descriptions of data from experiments.
 
@@
 
-A set of $N=40000$ simulated AT-TPC events per class was obtained  from a ${}^{46}$Ar$(p, p)$ experiment with the \texttt{pytpc} package developed by \citet{Bradt2017a}.
+A set of $N=8000$ simulated AT-TPC events per class was obtained  from a ${}^{46}$Ar$(p, p)$ experiment with the \texttt{pytpc} package developed by \citet{Bradt2017a}.
 %, see table \ref{tab:class_distr}.
@@
 
-To train the MIXAE clustering algorithm, we use the large simulated data set with $M=80000$ points, evenly distributed between proton- and carbon-events.
+To train the MIXAE clustering algorithm, we use a large simulated data set with $M=80000$ points, evenly distributed between proton- and carbon-events.
 The algorithm is trained on a subset of $60000$ of these samples, and we track performance on the remaining $20000$ events.
@@
 Total &  $8000$ & $51891$ & $49169$ \\
-labelled & $2400$ & $1774$ &  $1582$ \\ 
+Labelled & $2400$ & $1774$ &  $1582$ \\ 
 \bottomrule

### 2021-01-20 reviewer_responses.tex
added file for reviewer responses -MPK

\documentclass[12pt]{article}

\begin{document}

This article presents unsupervised machine learning methods and results for event separation in
the Active-Target Time Projection Chamber (AT-TPC). This work extends a previous related
study [Ref. 6, published in NIMA 2019] from supervised machine learning to unsupervised
learning. Some of the potentially attractive feature of unsupervised learning could include
bypassing the labeling step, forming classes of events independently from the experimenter’s input,
and processing larger amounts (~ TB) of data efficiently. By filtering out un-interesting events, an
unsupervised classification algorithm would also offer the possibility to discover rare events which
may not be expected or are overlooked. These rare events would likely be filtered out using the
traditional methods. The c2 approach used in the traditional analysis performed on the 46Ar data
could be computationally expensive because it involves the simulation of thousands of tracks for
each recorded event.

I am in favor of the publication of the paper in NIMA, subject to clarifications of the terms used
and data method descriptions in the manuscript. Furthermore, the discussions about the results
have mainly been focusing on data method metrics. It would be very helpful to connect or
extrapolate the results to the physics models and understanding or traditional instrumentation
metrics such as ground truth derived from physics models, signal to noise, noise reduction, etc.,
since NIMA is not a data science journal, at least not yet.

Below are some examples.
\begin{itemize}
\item Line 29, Simulated data? Line 193 simulated events… from a Ar(p,p) experiment. Are these
simulation data or experimental data? The use of word simulated is confusing. Based on the context,
it appears that only experimental data is used for this work.
\item Line 124-125, simulated via an integration, which is numerically costly. Rephrase and clarify how
costly quantitatively. Similar clarifications needed in Line 131, Line 145 and Table 2.
\item Figure 1. Add some discussions about the experimental information and physics behind the
observed patterns. For example, which direction is the incoming beam? Beam energy? Vertex at
(0,0)? The direction of the magnetic field?
\item Line 306, Sample $\hat{x}$, presumably each of these is a 2D image x-y? what are the sample
dimensions?
\item Lines 330 \& 331, define Ground truth labels y and predictions $\hat{y}$, how are they derived from
the raw data (2D projections of the AT-TPC)? Or the latent layer outputs of a pre-trained neural
network? Or else?
\item Figures 4, Figure 5, Figure 6, Figure 9, Figure 10, are data science metric results. Needs update or
move some of them to the appendix.
\item Figure 7, Figure 8, elucidate the physics and understanding behind the pattern would be helpful.
Similar to the comments to Figure 1 above.
\end{itemize}
\end{document}

### 2021-01-25 reviewer_responses.tex
Updates from Overleaf

@@
 \documentclass[12pt]{article}
-
+\usepackage{xcolor}
 \begin{document}
@@
 simulation data or experimental data? The use of word simulated is confusing. Based on the context,
-it appears that only experimental data is used for this work.
+it appears that only experimental data is used for this work.\\
+\textcolor{blue}{We apologize for the confusion.
+The events referred to in section 3.1 are simulated, not real.
+They were generated from an event simulator written for the purpose of analysing the Ar(p,p) reaction.
+The phrasing of this sentence is modified to clarify the information.}
 \item Line 124-125, simulated via an integration, which is numerically costly. Rephrase and clarify how
-costly quantitatively. Similar clarifications needed in Line 131, Line 145 and Table 2.
+costly quantitatively. Similar clarifications needed in Line 131, Line 145 and Table 2.\\
+\textcolor{blue}{More quantitative information about the computational cost of traditional analysis methods was added in the paragraphs mentioned by the referee.}
 \item Figure 1. Add some discussions about the experimental information and physics behind the
 observed patterns. For example, which direction is the incoming beam? Beam energy? Vertex at
-(0,0)? The direction of the magnetic field?
+(0,0)? The direction of the magnetic field?\\
+
+\textcolor{blue}{Additional text has been added in the figure caption to better explain the experimental conditions of the displayed data.}
 \item Line 306, Sample $\hat{x}$, presumably each of these is a 2D image x-y? what are the sample
 dimensions?
+
+\textcolor{orange}{The text has been modified to clarify that the samples are images encoded as a square matrix with 128 elements along one axis.}
+
 \item Lines 330 \& 331, define Ground truth labels y and predictions $\hat{y}$, how are they derived from
@@
 network? Or else?
+\textcolor{blue}{Additional text in each subsection of Section 3: Data Preparation was added to indicate where the ground truth labels were derived from.
+In fact, we used this dataset {\it because} we had rare access to ground truth labels for experimental data.}
 \item Figures 4, Figure 5, Figure 6, Figure 9, Figure 10, are data science metric results. Needs update or
 move some of them to the appendix.
+\textcolor{blue}{The confusion matrices have been moved to the appendix, and language was added to clarify their meaning.}
 \item Figure 7, Figure 8, elucidate the physics and understanding behind the pattern would be helpful.
 Similar to the comments to Figure 1 above.
+\textcoloe{blue}{Now Figures 4 and 5.}
 \end{itemize}

### 2021-01-25 doc/Manuscript/article.tex
added file for reviewer responses -MPK
Updates from Overleaf

@@
 \usepackage{lineno}
-%\linenumbers
+\linenumbers
 
@@
 \usepackage[T1]{fontenc}
+\usepackage{xcolor}
 
@@
 It therefore needs to be simulated via an integration, which is numerically costly.
+Common integration methods such as Runge-Kutta have a computational cost typically one of magnitude larger than an analytical calculation.
 Other difficulties are related to several experimental effects that deteriorate the quality of the data, namely saturation and cross-talk effects, as well as random noise.
@@
 In particular, as for the data presented and analyzed here, with data sets in the terabyte region or larger, these more traditional methods become prohibitively expensive from a computational stand.
+This is particularly the case when using the Monte-Carlo fitting procedure (explained later in text), because of the thousands of calculated tracks that are required for each individual event.
+For instance, the analysis of the $^{46}$Ar(p,p) data required several seconds per event on a CPU cluster composed of about 100 nodes.
+
 % Begin added by DB (06/10/2020)
@@
 The data is projected onto a two-dimensional  space by summing over the time-axis.
-Thereafter the data is down-sampled into a $128\times128$ pixel image.
+The two-dimensional data is then down-sampled into a $128\times128$ pixel image by discretizing the space in the $x-y$ plane and summing all charge values that fall within the bounds for each element location.
+%{\bf via sum-pooling}.
 %Additionally, we log-transform and normalize the data to the interval $[0, 1]$ for each point.
@@
 Therefore, we incrementally add bias to the analysis by applying the algorithm further down the analysis pipeline, with the benefit being that more preprocessing improves the signal-to-noise ratio, possibly improving model performance.
-To explore this trade-off between model performance and preprocessing bias, we performed our analysis on simulated, cleaned and raw events as discussed below.
+To explore this trade-off between model performance and preprocessing bias, we performed our analysis on simulated, raw and cleaned events as discussed below.
 
@@
 
-A set of $N=8000$ simulated AT-TPC events per class was obtained  from a ${}^{46}$Ar$(p, p)$ experiment with the \texttt{pytpc} package developed by \citet{Bradt2017a}.
-%, see table \ref{tab:class_distr}.
+A set of $N=8000$ simulated AT-TPC events per class was generated from the \texttt{pytpc} package developed by \citet{Bradt2017a} for the analysis of the  ${}^{46}$Ar$(p, p)$ experiment.
+%A set of $N=8000$ simulated AT-TPC events per class was obtained  from a ${}^{46}$Ar$(p, p)$ experiment with the \texttt{pytpc} package developed by \citet{Bradt2017a}.%, see table \ref{tab:class_distr}.
 %Using the same parameters {\color{blue} what parameters are you referring to?
@@
 
+{\bf As this dataset is generated via simulation, the true labels are known for this dataset.}
+
 \subsection{Raw \texorpdfstring{${}^{46}$Ar}{46Ar}  events}\label{sec:data_real}
@@
 The events analyzed in this section were retrieved from the ${}^{46}$Ar resonant proton scattering experiment recorded with the AT-TPC. 
-While we denote these events as raw, it is important to note that what we intend is a raw two-dimensional projection of events.
+While we denote these events as raw, it is important to note that what we intend is a raw two-dimensional projection of events {\bf without any post-processing done to clean the data}.
 
@@
 We display two different events from the ${}^{46}$Ar experiment in Fig.~\ref{fig:samples}.
-The top row illustrates a carbon event with a large fraction of noise, while the bottom row shows a proton event almost free of noise.
+The top row illustrates a proton event with a distinctive spiral-shaped track, while the bottom row shows a carbon event with a large fraction of noise.
 
@@
 Each row is one event in two projections, where the color intensity of each point indicates higher charge values recorded by the detector.
-The top row illustrates a carbon event with a large fraction of noise, while the bottom row shows a proton event almost free of noise.}\label{fig:samples}
+The top row illustrates a proton event with a distinctive spiral-shaped track, while the bottom row shows a carbon event with a large fraction of noise.
+The beam direction is oriented along the z-axis (from positive to negative), close to the (x,y) origin.
+The magnetic field is oriented on the same axis as the beam, causing the bending of the charged particle's trajectories.
+The spiral shape is caused by the slowing down of the particles inside the gas volume.
+The bending radius of the tracks depends not only on the energy, but also on the mass-to-charge ratio of the particles, hence the large difference between proton and carbon events (see also section \ref{Sec:Exp}).}
+\label{fig:samples}
 \end{figure}
 
+{\bf A subset of this data was hand-labeled for validation.}
 \subsection{Filtered \texorpdfstring{${}^{46}$Ar}{46Ar} events}\label{sec:filtered}
@@
 \end{figure}
+{\bf The same subset of hand-labeled data from Section~\ref{sec:data_real} was used for validation.}
 
@@
 In the figure each encoder and decoder pair makes up an autoencoder.
-The assignment of the cluster belonging to a sample, $\mathbf{\hat{x}}$, is taken to be the index of the maximal element in the vector $\mathbf{p}$ as shown in the right-most part of the figure.
+The assignment of the cluster belonging to a sample image, $\mathbf{\hat{x}} \in \mathcal{R}^{128 \times \hspace{0.01in} 128} $, is taken to be the index of the maximal element in the vector $\mathbf{p}$ as shown in the right-most part of the figure.
 Finally, the model is trained end-to-end with back propagation \cite{Linnainmaa1976}, as implemented in the machine learning package TensorFlow \cite{tensorflow}.
@@
 
-\begin{figure}
-\centering
-	\includegraphics[width=\textwidth]{custom_work/Simulatedvgg_conf_mat.pdf} 
-	\caption[Pre-trained network - confusion matrices]{Confusion matrix for the $K$-means clustering of simulated AT-TPC events.
-The true labels indicate samples belonging to the p (proton), or the carbon (C) class}\label{fig:clster_confmat_sim}
-\end{figure}
-
-\begin{figure}
-\centering
-	\includegraphics[width=\textwidth]{custom_work/Filteredvgg_conf_mat.pdf}
-	\caption[Pre-trained network - confusion matrices]{Confusion matrix for the $K$-means clustering of filtered AT-TPC events.
-The true labels indicate samples belonging to the p (proton), carbon (C), or other classes.
-}\label{fig:clster_confmati_filt}
-\end{figure}
-
-\begin{figure}
-	\includegraphics[width=\textwidth]{custom_work/Rawvgg_conf_mat.pdf}
-\caption[Pre-trained network - confusion matrices]{Confusion matrix for the $K$-means clustering of raw AT-TPC events.
-The true labels indicate samples belonging to the p (proton), carbon (C), or other classes.
-}\label{fig:clster_confmati_raw}
-\end{figure}
-
 We repeat this analysis using a Principle Component Analysis (PCA) dimensionality reduction\footnote{
@@
 On filtered data the highest performing MIXAE model achieves strong separation of the other-class, but curiously creates two proton-majority clusters.
-
-\begin{figure}
-\centering
-	\includegraphics[width=\textwidth]{custom_work/Filtered_mixae_conf_mat.pdf}
-\caption[MIXAE - confusion matrices]{Confusion matrix for the MIXAE clustering algorithm on filtered AT-TPC events.
-The true labels indicate samples belonging to the \texttt{p} (proton), \texttt{C} (Carbon), or \texttt{other} classes.
-We observe that the algorithm forms two proton-majority clusters, and one clearly defined cluster of the other events.
-}\label{fig:mixae_confmat_filtered}
-\end{figure}
-
-\begin{figure}
-\centering
-	\includegraphics[width=\textwidth]{custom_work/Raw_mixae_conf_mat.pdf}
-\caption[MIXAE - confusion matrices]{Confusion matrices for the MIXAE clustering algorithm on raw AT-TPC events.
-The true labels indicate samples belonging to the p (proton), carbon (C), or \texttt{other} classes.
-We observe that the algorithm correctly captures a majority proton-event cluster in cluster 0.
-However, in contrast with the $K$-means approach this cluster is contaminated to some extent with both carbon events and other events.
-}\label{fig:mixae_confmat_raw}
-\end{figure}
 
@@
 \appendix
+
 \section{VGG16}\label{app:vgg}
@@
 
+\section{Clustering confusion tables}
+
+To elucidate the results presented in tables \ref{tab:clstr_mixae} and \ref{tab:clstr_vgg1} we computed clustering confusion-matrices.
+These matrices show a more detailed picture of intermingled classes in a clustering or classification task where ground truth labels are available.
+In the figures below we tabulate the clusters, as predicted by the algorithm, along the x-axis.
+Each cluster is decomposed in its ground-truth members along the y-axis.
+
+In this view, a perfect clustering algorithm will produce a confusion matrix which only has nonzero elements in the primary diagonal under free permutation of its columns.
+
+\begin{figure}
+\centering
+	\includegraphics[width=\textwidth]{custom_work/Simulatedvgg_conf_mat.pdf} 
+	\caption[Pre-trained network - confusion matrices]{Confusion matrix for the $K$-means clustering of simulated AT-TPC events.
+The true labels indicate samples belonging to the p (proton), or the carbon (C) class.
+We observe very high quality separation between the proton and carbon classes.}\label{fig:clster_confmat_sim}
+\end{figure}
+
+\begin{figure}
+\centering
+	\includegraphics[width=\textwidth]{custom_work/Filteredvgg_conf_mat.pdf}
+	\caption[Pre-trained network - confusion matrices]{Confusion matrix for the $K$-means clustering of filtered AT-TPC events.
+The true labels indicate samples belonging to the p (proton), carbon (C), or other classes.
+Each column denotes a cluster, with each cell in the column denoting the count of that rows' class in the cluster.
+We observe that cluster 2 is a high quality proton event cluster }\label{fig:clster_confmati_filt}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=\textwidth]{custom_work/Rawvgg_conf_mat.pdf}
+\caption[Pre-trained network - confusion matrices]{Confusion matrix for the $K$-means clustering of raw AT-TPC events.
+The true labels indicate samples belonging to the p (proton), carbon (C), or other classes.
+Each column denotes a cluster, with each cell in the column denoting the count of that rows' class in the cluster.
+We observe that cluster 0 is a high quality proton event cluster }\label{fig:clster_confmati_raw}
+\end{figure}
+
+\begin{figure}
+\centering
+	\includegraphics[width=\textwidth]{custom_work/Filtered_mixae_conf_mat.pdf}
+\caption[MIXAE - confusion matrices]{Confusion matrix for the MIXAE clustering algorithm on filtered AT-TPC events.
+The true labels indicate samples belonging to the \texttt{p} (proton), \texttt{C} (Carbon), or \texttt{other} classes.
+We observe that the algorithm forms two proton-majority clusters, and one clearly defined cluster of the other events.
+}\label{fig:mixae_confmat_filtered}
+\end{figure}
+
+\begin{figure}
+\centering
+	\includegraphics[width=\textwidth]{custom_work/Raw_mixae_conf_mat.pdf}
+\caption[MIXAE - confusion matrices]{Confusion matrices for the MIXAE clustering algorithm on raw AT-TPC events.
+The true labels indicate samples belonging to the p (proton), carbon (C), or \texttt{other} classes.
+We observe that the algorithm correctly captures a majority proton-event cluster in cluster 0.
+However, in contrast with the $K$-means approach this cluster is contaminated to some extent with both carbon events and other events.
+}\label{fig:mixae_confmat_raw}
+\end{figure}
+
 \bibliography{article}

### 2021-01-25 reviewer_responses.tex
Reviewer address in progress

@@
 since NIMA is not a data science journal, at least not yet.
+
+\textcolor{blue}{We thank the reviewer for a thorough review of our work. We have }
 
@@
 Similar to the comments to Figure 1 above.
-\textcoloe{blue}{Now Figures 4 and 5.}
+\textcolor{blue}{Now Figures 4 and 5.
+Since this is unsupervised clustering, we could not conjecture as to the physical meaning behind the separation.
+However, the plots visualize a sampling of proton events that were attributed to a noisy and pure cluster (each cluster), allowing a reader to infer the behavior of the unsupervised learning on these events.
+The figure caption texts have been updated to clarify this point.
+}
 \end{itemize}

### 2021-01-25 doc/Manuscript/article.tex
Reviewer address in progress

@@
 In short, we want to find the most reasonable correspondence between the clusters and the ground truth classes.
-To solve this problem we first define two sets; the ground truth labels $\mathbf{y} = [y_i, \hspace{0.2cm} y_{i +1}, \ldots]$ and the corresponding predictions $\mathbf{\hat{y}} = [\hat{y}_i, \hspace{0.2cm} \hat{y}_{i +1}, \ldots]$.
+To solve this problem we first define two sets; the ground truth labels $\mathbf{y} = [y_i, \hspace{0.2cm} y_{i +1}, \ldots]$ as determined in section \ref{sec:datadescript}, and the corresponding predictions $\mathbf{\hat{y}} = [\hat{y}_i, \hspace{0.2cm} \hat{y}_{i +1}, \ldots]$.
 Furthermore, we let both $y_i$ and $\hat{y}_i$ be integer representations of an event's ground truth, and predicted class, respectively.
@@
 \includegraphics[width=\textwidth]{custom_work/Filteredexamples.pdf}
-\caption[Filtered proton samples by cluster belonging]{Illustration of  a sample of proton events from different $K$-means clusters originating from the filtered data set.
+\caption[Filtered proton samples by cluster belonging]{A sample of proton events from different $K$-means clusters from the filtered data set.
+The top row shows proton samples that are intermingled with noise events in the models' predictions, and the bottom row belongs to a cluster that contains proton events almost exclusively.
 Each row belongs to a single cluster corresponding to the filtered confusion matrix in Fig.~\ref{fig:clster_confmati_filt}}\label{fig:filtered_vgg_clster_repr}

### 2021-01-31 reviewer_responses.tex
Updates from Overleaf

--- 2021-01-20 reviewer_responses.tex
@@
 \documentclass[12pt]{article}
-
+\usepackage{xcolor}
 \begin{document}
 
-This article presents unsupervised machine learning methods and results for event separation in
-the Active-Target Time Projection Chamber (AT-TPC). This work extends a previous related
-study [Ref. 6, published in NIMA 2019] from supervised machine learning to unsupervised
-learning. Some of the potentially attractive feature of unsupervised learning could include
-bypassing the labeling step, forming classes of events independently from the experimenter’s input,
-and processing larger amounts (~ TB) of data efficiently. By filtering out un-interesting events, an
-unsupervised classification algorithm would also offer the possibility to discover rare events which
-may not be expected or are overlooked. These rare events would likely be filtered out using the
-traditional methods. The c2 approach used in the traditional analysis performed on the 46Ar data
-could be computationally expensive because it involves the simulation of thousands of tracks for
-each recorded event.
+\section*{Dear Editors,}
 
-I am in favor of the publication of the paper in NIMA, subject to clarifications of the terms used
-and data method descriptions in the manuscript. Furthermore, the discussions about the results
-have mainly been focusing on data method metrics. It would be very helpful to connect or
-extrapolate the results to the physics models and understanding or traditional instrumentation
-metrics such as ground truth derived from physics models, signal to noise, noise reduction, etc.,
-since NIMA is not a data science journal, at least not yet.
+we are much indebted for the positive feedback from the reviewer and your willingness to consider our manuscript for publication.
+We have tried as faithfully as possible to answer the referee and we hope the revised version meets your standards for publications in NIMA.\newline
+Our answers to the referee are included below here. The revised version is included as well.\newline
+Yours sincerely,\newline
+Daniel Bazin, Morten Hjorth-Jensen, Michelle Kuchera, Robert Solli, and Ryan Strauss
 
-Below are some examples.
+\section*{Answers to the referee}
+
+We are very much indebted for the positive comments on our manuscript and we would like to  thank the reviewer for a careful read of our work.
+We have tried as faithfully as possible to answer the various comments.
+We have included them here for the sake of completeness, together with our responses
+(highlighted in blue). We hope our responses  aid in improving the quality of the manuscript and thank the reviewer again for her/his willingness for considering our manuscript again.
+
 \begin{itemize}
-\item Line 29, Simulated data? Line 193 simulated events… from a Ar(p,p) experiment. Are these
+\item Reviewer comment: Line 29, Simulated data? Line 193 simulated events… from a Ar(p,p) experiment. Are these
 simulation data or experimental data? The use of word simulated is confusing. Based on the context,
-it appears that only experimental data is used for this work.
-\item Line 124-125, simulated via an integration, which is numerically costly. Rephrase and clarify how
-costly quantitatively. Similar clarifications needed in Line 131, Line 145 and Table 2.
-\item Figure 1. Add some discussions about the experimental information and physics behind the
+it appears that only experimental data is used for this work.\\
+\textcolor{blue}{Our answer: We apologize for the confusion.
+The events referred to in section 3.1 are simulated, not real.
+They were generated from an event simulator written for the purpose of analysing the Ar(p,p) reaction.
+The phrasing of this sentence is modified to clarify the information.}
+\item Reviewer comment: Line 124-125, simulated via an integration, which is numerically costly. Rephrase and clarify how
+costly quantitatively. Similar clarifications needed in Line 131, Line 145 and Table 2.\\
+\textcolor{blue}{Our answer: More quantitative information about the computational cost of traditional analysis methods was added in the paragraphs mentioned by the referee.}
+\item Reviewer comment: Figure 1. Add some discussions about the experimental information and physics behind the
 observed patterns. For example, which direction is the incoming beam? Beam energy? Vertex at
-(0,0)? The direction of the magnetic field?
-\item Line 306, Sample $\hat{x}$, presumably each of these is a 2D image x-y? what are the sample
+(0,0)? The direction of the magnetic field?\\
+
+\textcolor{blue}{Our answer: Additional text has been added in the figure caption to better explain the experimental conditions of the displayed data.}
+\item Reviewer comment: Line 306, Sample $\hat{x}$, presumably each of these is a 2D image x-y? what are the sample
 dimensions?
-\item Lines 330 \& 331, define Ground truth labels y and predictions $\hat{y}$, how are they derived from
+
+\textcolor{blue}{Our answer: The text has been modified to clarify that the samples are images encoded as a square matrix with 128 elements along one axis.}
+
+\item Reviewer comment: Lines 330 \& 331, define Ground truth labels y and predictions $\hat{y}$, how are they derived from
 the raw data (2D projections of the AT-TPC)? Or the latent layer outputs of a pre-trained neural
 network? Or else?
-\item Figures 4, Figure 5, Figure 6, Figure 9, Figure 10, are data science metric results. Needs update or
+\textcolor{blue}{Our answer: We have added more text in each subsection of Section 3: Data Preparation was added to indicate where the ground truth labels were derived from.
+In fact, we used this data set {\it because} we had rare access to ground truth labels for experimental data.}
+\item Reviewer comment: Figures 4, Figure 5, Figure 6, Figure 9, Figure 10, are data science metric results. Needs update or
 move some of them to the appendix.
-\item Figure 7, Figure 8, elucidate the physics and understanding behind the pattern would be helpful.
+\textcolor{blue}{Our answer: We have followed the reviewer's suggestions and have moved the confusion matrix figures to the appendix, and we added text to clarify their meaning.}
+\item Reviewer comment: Figure 7, Figure 8, elucidate the physics and understanding behind the pattern would be helpful.
 Similar to the comments to Figure 1 above.
+\textcolor{blue}{Our answer: These figures are wow the new Figures 4 and 5.
+Since this is unsupervised clustering, it is normally difficult to  conjecture a clear   physical meaning behind the separation.
+However, the plots visualize a sampling of proton events that were attributed to a noisy and pure cluster (each cluster), allowing a reader to infer the behavior of the unsupervised learning on these events.
+The figure caption texts have been updated to clarify this point.
+}
 \end{itemize}
+
+We hope our answers have addressed properly the comments of the reviewer. \newline
+Yours sincerely,\newline
+The Authors
+
 \end{document}

### 2021-03-09 doc/Manuscript/article.tex
Updates from Overleaf

@@
 \subsection*{Acknowledgements}
-MHJ’s work is supported by U.S.
+MHJ’s work is supported by the U.S.
+Department of Energy, Office of Science, office of Nuclear Physics under grant No.
+DE-SC0021152 and U.S.
 National Science Foundation Grants No.
@@
 \centering
-	\includegraphics[width=\textwidth]{custom_work/Filteredvgg_conf_mat.pdf}
+	\includegraphics[width=\textwidth]{custom_work/Filteredcache_conf_mat.pdf}
 	\caption[Pre-trained network - confusion matrices]{Confusion matrix for the $K$-means clustering of filtered AT-TPC events.
@@
 \begin{figure}
-	\includegraphics[width=\textwidth]{custom_work/Rawvgg_conf_mat.pdf}
+	\includegraphics[width=\textwidth]{custom_work/Rawcache_conf_mat.pdf}
 \caption[Pre-trained network - confusion matrices]{Confusion matrix for the $K$-means clustering of raw AT-TPC events.

### 2021-03-11 reviewer_responses.tex
Updates from Overleaf

@@
 \textcolor{blue}{Our answer: We apologize for the confusion.
-The events referred to in section 3.1 are simulated, not real.
-They were generated from an event simulator written for the purpose of analysing the Ar(p,p) reaction.
-The phrasing of this sentence is modified to clarify the information.}
+Both simulated events and records from the AT-TPC were studied.
+Simulated events were generated from an event simulator written for the purpose of analysing the Ar(p,p) reaction.
+We added a sentence in the abstract to clarify the circumstance.
+}
 \item Reviewer comment: Line 124-125, simulated via an integration, which is numerically costly. Rephrase and clarify how
@@
 Similar to the comments to Figure 1 above.
-\textcolor{blue}{Our answer: These figures are wow the new Figures 4 and 5.
+\textcolor{blue}{Our answer: These figures are now the new Figures 4 and 5.
 Since this is unsupervised clustering, it is normally difficult to  conjecture a clear   physical meaning behind the separation.

### 2021-03-11 doc/Manuscript/article.tex
Updates from Overleaf

@@
 We explore the performance of autoencoder neural networks and a pre-trained VGG16 \cite{Simonyan2014} convolutional neural network.
-We find that a $K$-means algorithm applied to the simulated data in the VGG16 latent space forms almost perfect clusters.
+We study clustering performance on data from a simulated $^{46}$Ar experiment, and records from the AT-TPC detector.
+We find that a $K$-means algorithm applied to simulated data in the VGG16 latent space forms almost perfect clusters.
 Additionally, the VGG16+$K$-means approach finds high purity clusters of proton events for real experimental data.
@@
 In the figure each encoder and decoder pair makes up an autoencoder.
-The assignment of the cluster belonging to a sample image, $\mathbf{\hat{x}} \in \mathcal{R}^{128 \times \hspace{0.01in} 128} $, is taken to be the index of the maximal element in the vector $\mathbf{p}$ as shown in the right-most part of the figure.
+The assignment of the cluster for a sample x-y event image, $\mathbf{\hat{x}} \in \mathcal{R}^{128 \times \hspace{0.01in} 128} $, is taken to be the index of the maximal element in the vector $\mathbf{p}$ as shown in the right-most part of the figure.
 Finally, the model is trained end-to-end with back propagation \cite{Linnainmaa1976}, as implemented in the machine learning package TensorFlow \cite{tensorflow}.
@@
 \caption[Filtered proton samples by cluster belonging]{A sample of proton events from different $K$-means clusters from the filtered data set.
-The top row shows proton samples that are intermingled with noise events in the models' predictions, and the bottom row belongs to a cluster that contains proton events almost exclusively.
+The bottom row shows proton samples that are intermingled with noise events in the models' predictions, and the top row belongs to a cluster that contains proton events almost exclusively.
 Each row belongs to a single cluster corresponding to the filtered confusion matrix in Fig.~\ref{fig:clster_confmati_filt}}\label{fig:filtered_vgg_clster_repr}
@@
 \centering
-	\includegraphics[width=\textwidth]{custom_work/Simulatedvgg_conf_mat.pdf} 
+	\includegraphics[width=\textwidth]{custom_work/Simulatedcache_conf_mat.pdf} 
 	\caption[Pre-trained network - confusion matrices]{Confusion matrix for the $K$-means clustering of simulated AT-TPC events.
@@
 Each column denotes a cluster, with each cell in the column denoting the count of that rows' class in the cluster.
-We observe that cluster 0 is a high quality proton event cluster }\label{fig:clster_confmati_raw}
+We observe that cluster 2 is a high quality proton event cluster }\label{fig:clster_confmati_raw}
 \end{figure}

### 2021-03-13 reviewer_responses.tex
v2 - revision submission

@@
 
-we are much indebted for the positive feedback from the reviewer and your willingness to consider our manuscript for publication.
+We are much indebted for the positive feedback from the reviewer and your willingness to consider our manuscript for publication.
 We have tried as faithfully as possible to answer the referee and we hope the revised version meets your standards for publications in NIMA.\newline
@@
 However, the plots visualize a sampling of proton events that were attributed to a noisy and pure cluster (each cluster), allowing a reader to infer the behavior of the unsupervised learning on these events.
-The figure caption texts have been updated to clarify this point.
+The text has been updated with intuitive physics deductions from the plots.
+The figure caption texts have also been updated for clarity.
 }

### 2021-03-13 credit_statement.tex
v2 - revision submission

\documentclass[12pt]{article}
\usepackage{xcolor}
\begin{document}
{\bf Robert Solli:} Conceptualization, Methodology, Software, Validation, Formal Analysis, Writing - Original Draft, Writing - Review \& Editing, Visualization,
{\bf Daniel Bazin:} Conceptualization, Investigation, Writing - Original Draft, Writing - Review \& Editing, Supervision,
{\bf Morten Hjorth-Jensen:} Resources, Writing - Original Draft, Writing - Review \& Editing, Supervision
{\bf Michelle Kuchera:} Conceptualization, Methodology, Resources, Writing - Original Draft, Writing - Review \& Editing, Supervision,
{\bf Ryan Strauss:} Methodology, Software, Validation, Formal Analysis, Writing - Original Draft, Writing - Review \& Editing,

\end{document}

### 2021-03-13 doc/Manuscript/article.tex
v2 - revision submission

@@
-%\documentclass[preprint,12pt]{elsarticle}
+\documentclass[preprint,12pt]{elsarticle}
 % uncomment for submission to NIM A
-\documentclass[review,sort&compress]{elsarticle}
-\usepackage{lineno}
-\linenumbers
+%\documentclass[review,sort&compress]{elsarticle}
+%\usepackage{lineno}
+%\linenumbers
 
@@
 We explore the performance of autoencoder neural networks and a pre-trained VGG16 \cite{Simonyan2014} convolutional neural network.
-We study clustering performance on data from a simulated $^{46}$Ar experiment, and records from the AT-TPC detector.
-We find that a $K$-means algorithm applied to simulated data in the VGG16 latent space forms almost perfect clusters.
-Additionally, the VGG16+$K$-means approach finds high purity clusters of proton events for real experimental data.
+We study clustering performance on both data from a simulated $^{46}$Ar experiment, and real events from the AT-TPC detector.
+We find that a $k$-means algorithm applied to simulated data in the VGG16 latent space forms almost perfect clusters.
+Additionally, the VGG16+$k$-means approach finds high purity clusters of proton events for real experimental data.
 We also explore the application of clustering the latent space of autoencoder neural networks for event separation.
@@
 A set of $N=8000$ simulated AT-TPC events per class was generated from the \texttt{pytpc} package developed by \citet{Bradt2017a} for the analysis of the  ${}^{46}$Ar$(p, p)$ experiment.
-%A set of $N=8000$ simulated AT-TPC events per class was obtained  from a ${}^{46}$Ar$(p, p)$ experiment with the \texttt{pytpc} package developed by \citet{Bradt2017a}.%, see table \ref{tab:class_distr}.
+%A set of $N=8000$ simulated AT-TPC events per class was obtained  from a ${}^{46}$Ar$(p, p)$ experiment with the \texttt{pytpc} package developed by \citet{Bradt2017a}.%, see Table \ref{tab:class_distr}.
 %Using the same parameters {\color{blue} what parameters are you referring to?
@@
 Note that $\boldsymbol{X}_L \subset \boldsymbol{X}_F$.
-
-{\bf As this dataset is generated via simulation, the true labels are known for this dataset.}
+As this dataset is generated via simulation, the true labels are known for this dataset.
 
@@
 The events analyzed in this section were retrieved from the ${}^{46}$Ar resonant proton scattering experiment recorded with the AT-TPC. 
-While we denote these events as raw, it is important to note that what we intend is a raw two-dimensional projection of events {\bf without any post-processing done to clean the data}.
+While we denote these events as raw, it is important to note that what we intend is a raw two-dimensional projection of events without any post-processing done to clean the data.
 
@@
 The top row illustrates a proton event with a distinctive spiral-shaped track, while the bottom row shows a carbon event with a large fraction of noise.
-
+A subset of this data was hand-labeled for validation, as discussed in \cite{Kuchera2019}.
 \begin{figure}[ht]
@@
 
-{\bf A subset of this data was hand-labeled for validation.}
 \subsection{Filtered \texorpdfstring{${}^{46}$Ar}{46Ar} events}\label{sec:filtered}
@@
 
-We illustrate two filtered events in Fig.~\ref{fig:samples_filtered}. These are the same events as shown in Fig.~\ref{fig:samples}, but with the Hough transform and nearest-neighbor filtering applied.
+We illustrate two filtered events in Fig.~\ref{fig:samples_filtered}.
+These are the same events as shown in Fig.~\ref{fig:samples}, but with the Hough transform and nearest-neighbor filtering applied.
+The same subset of hand-labeled data from Section~\ref{sec:data_real} was used for validation.
 
@@
 \end{figure}
-{\bf The same subset of hand-labeled data from Section~\ref{sec:data_real} was used for validation.}
 
@@
 The traditional Monte Carlo event selection process, described in Section~\ref{Sec:Exp}, does not have a well-defined method to quantify the effectiveness of the event selection.
-In addition, the selection task produced a binary result only, either a {\em good} or {\em bad} fit to the event of interest.
+In addition, the selection task produced a binary, cut-based result: either a {\em good} or {\em bad} fit to the event of interest.
 A {\em bad} fit is then assumed to be a different event type, and is removed from the analysis.
@@
 
-The $\chi^2$ approach used in the traditional analysis performed on the $^{46}$Ar data is extremely expensive from a computational stand because it involves the simulation of thousands of tracks for each recorded event.
+The $\chi^2$ approach used in the traditional analysis performed on the $^{46}$Ar data is extremely expensive computationally because it requires the simulation of thousands of tracks for each recorded event.
 These events are in turn simulated for each iteration of the Monte Carlo fitting sequence.
@@
 
-% The clustering algorithm we found to give the strongest result is the $K$-means algorithm.
+% The clustering algorithm we found to give the strongest result is the $k$-means algorithm.
 This algorithm is contingent on having a distance measurement between two datapoints.
@@
 Thus, rather than fine-tuning a pre-trained network under the supervised learning regime, we extract the output of the pre-trained network's last convolutional layer as a latent representation of the events, where each event is represented as a vector in $\R^{8192}$.
-We then cluster events based on this representation using the Scikit-Learn implementation of the $K$-means algorithm with default parameters \cite{Pedregosa2011}.
+We then cluster events based on this representation using the \texttt{scikit-learn} implementation of the $k$-means algorithm with default parameters \cite{Pedregosa2011}.
 
@@
 As such, I would remove them and add the necessary details to the previous paragraph (or replace them with a small paragraph).]} Building on the work by \citet{Kuchera2019}, we also use the VGG16 pre-trained network as a baseline for the clustering performance.
-We begin by considering a classical $K$-means approach to clustering.
-However, the output from the VGG16 network is very high dimensional which traditionally conflicts with good $K$-means results.
+We begin by considering a classical $k$-means approach to clustering.
+However, the output from the VGG16 network is very high dimensional which traditionally conflicts with good $k$-means results.
 With output vectors in $\R^{8192}$ the ratio between $L_p$ distances goes to one with increasing dimensionality, as shown by \citet{Aggarwal}.
@@
 They note that it is then possible that if the class information lies in some sub-space of the latent data, $L_p$-norms may provide useful clusterings even in high dimensional spaces.
-To investigate this, we perform a $L_2$-norm based $K$-means clustering using the full VGG16 representation and the $10^2$ first principal components only.
+To investigate this, we perform a $L_2$-norm based $k$-means clustering using the full VGG16 representation and the $10^2$ first principal components only.
 
 %{\color{blue} Are there supposed to be results here? -MPK}
-%\subsection{$K$-means}
+%\subsection{$k$-means}
 
-%We begin by investigating the $K$-means clustering algorithm on the VGG16 latent space. 
+%We begin by investigating the $k$-means clustering algorithm on the VGG16 latent space. 
 % The user a VGG16 model that is pre-trained on the ImageNet data set and subsequently applied to the AT-TPC events without the classification layers.
 This creates a set of vectors $\boldsymbol{x} \in \R^{8192}$ as our new representation of the events.
-To cluster, we use the \lstinline{scikit-learn} implementation of the $K$-means++ algorithm with default parameters \cite{Pedregosa2011}.
+To cluster, we use the \lstinline{scikit-learn} implementation of the $k$-means++ algorithm with default parameters \cite{Pedregosa2011}.
 
@@
 To compute both the ARI and the clustering accuracy we also have to construct the contingency table between two sets of clustering assignments.
-A contingency matrix defines the overlaps between classes in these sets, and its general form is shown in table \ref{tab:contingency}.
+A contingency matrix defines the overlaps between classes in these sets, and its general form is shown in Table \ref{tab:contingency}.
 
@@
 \end{equation}
-The quantities $f_i$ and $e_i$ are defined in table \ref{tab:contingency}.
+The quantities $f_i$ and $e_i$ are defined in Table \ref{tab:contingency}.
 The important distinction to make between the clustering accuracy and the ARI is that the clustering accuracy is a simple comparison that does not account for chance assignments, while the ARI does.
@@
 
-The performance is measured by comparing model predictions on the labelled subset of data (see table \ref{tab:data sets} in  appendix B for more details).
+The performance is measured by comparing model predictions on the labelled subset of data (see Table \ref{tab:data sets} in  appendix B for more details).
 
@@
 
-We first explore the results of applying a $K$-means approach on the latent space of a pre-trained network.
+We first explore the results of applying a $k$-means approach on the latent space of a pre-trained network.
 Subsequently, we investigate the performance of the MIXAE algorithm as outlined in section \ref{sec:mixae}.
 
-\subsection{$K$-means clustering on the VGG16 latent space}\label{sec:kmeans_results}
+\subsection{$k$-means clustering on the VGG16 latent space}\label{sec:kmeans_results}
 
-The results of the clustering runs are included in table \ref{tab:clstr_vgg1}.
-We ran the $K$-means algorithm $N=100$ times with $M=10$ initializations per run, of the cluster centroids to assess the variability in the performance.
-The K-means algorithm returns the best performing model of the $M$ initializations on the unsupervised objective.
+The results of the clustering runs are included in Table \ref{tab:clstr_vgg1}.
+We ran the $k$-means algorithm $N=100$ times with $M=10$ initializations per run, of the cluster centroids to assess the variability in the performance.
+The $k$-means algorithm returns the best performing model of the $M$ initializations on the unsupervised objective.
 We report performance on the labelled subset of data, using the labels to identify the top-performing model (Top 1).
@@
 We observe that the clustering on simulated data attains the highest performance, and that there is a decline in performance as we add noise by moving to the filtered and raw data sets.
-The results are shown in table \ref{tab:clstr_vgg10}.
+The results are shown in Table \ref{tab:clstr_vgg10}.
 
@@
 \centering 
-\caption[$K$-means on pre-trained model]{$K$-means clustering results on AT-TPC event data in the VGG-16 latent space, for $N=100$ runs of the $K$-means algorithm with $M=10$ initializations.
+\caption[$k$-means on pre-trained model]{$k$-means clustering results on AT-TPC event data in the VGG-16 latent space, for $N=100$ runs of the $k$-means algorithm with $M=10$ initializations.
 We observe that the performance predictably decreases with the amount of noise in the data.}\label{tab:clstr_vgg10}
@@
 \noindent The lack of variability is explained by the number of initializations.
-As can be seen from table \ref{tab:clstr_vgg1} where we run the $K$-means algorithm $N=1000$ times with $M=1$ initializations of the centroids.
+As can be seen from Table \ref{tab:clstr_vgg1} where we run the $k$-means algorithm $N=1000$ times with $M=1$ initializations of the centroids.
 
@@
 \centering 
-\caption[$K$-means on pre-trained model]{$K$-means clustering results on AT-TPC event data in the VGG-16 latent space, for $N=1000$ runs of the $K$-means algorithm with $M=1$ initializations.
+\caption[$k$-means on pre-trained model]{$k$-means clustering results on AT-TPC event data in the VGG-16 latent space, for $N=1000$ runs of the $k$-means algorithm with $M=1$ initializations.
 We observe that there is significant variability in the results, which is ordinarily masked by $M$ re-initializations that avoid local minima.
@@
 
-In addition to the performance measures reported in table \ref{tab:clstr_vgg1}, it is interesting to observe which samples are wrongly assigned.
-To investigate this problem, we compute the matrices $\mathbf{W}$ as shown in table \ref{tab:contingency}.
+In addition to the performance measures reported in Table \ref{tab:clstr_vgg1}, it is interesting to observe which samples are wrongly assigned.
+To investigate this problem, we compute the matrices $\mathbf{W}$ as shown in Table \ref{tab:contingency}.
 From these tables, we can infer which classes are more or less entangled with others.
@@
 
-To further investigate the clusters presented in the matrix in Figs.~\ref{fig:clster_confmati_filt} and \ref{fig:clster_confmati_raw}, we study examples from the proton samples belonging to different clusters for the filtered and full data.
-We display these examples in Figs.~\ref{fig:filtered_vgg_clster_repr} and \ref{fig:full_vgg_clster_repr}, respectively.
-Form these figures we observe that the more noisy proton events are being clustered together with the amorphous other class events.
+To further investigate the clusters presented in the matrix in Figs.~\ref{fig:clster_confmati_filt} and \ref{fig:clster_confmati_raw}, we visualize a random subset of examples from the proton events belonging to different clusters for the filtered and full data in Figs.~\ref{fig:filtered_vgg_clster_repr} and \ref{fig:full_vgg_clster_repr}, respectively.
+We look at random samples of proton events in two clusters to infer an intuition on the track properties that are considered similar.
+The figures indicate that shorter tracks, therefore low-energy or small scattering angle protons, are more likely to appear similar to other event types.
+%From these figures, we observe that the more noisy proton events are being clustered together with the amorphous other class events.
 This is in line with the intuitive expectation from viewing the clustering results in Figs.~\ref{fig:clster_confmati_filt} and \ref{fig:clster_confmati_raw}.
+A visual inspection of the cluster components serves as a sanity-check for our clustering approach.
+While we have demonstrated some level of success in clustering events, the algorithm is agnostic to the underlying physics.
+    %a major caveat remains in the fundamental aphysical nature of the approach. 
+    %The aforementioned figures are then a way to better understand how the algorithm mislabels events of interest. 
+    %In lieu of a physics-grounded analysis, we wish to infer the differences which places proton events in other clusters than the high-purity clusters we observe in Figs ~\ref{fig:clster_confmati_filt} and \ref{fig:clster_confmati_raw}.
+   % \item While we have achieved some level of success in our clustering methods, we lack the ability to interpret what aspects of the data guided the cluster assignments.
+To better understand how the algorithm mislabels events of interest, we used visualization of clustered events to infer proton events that the algorithm considers similar.
+%\end{itemize}
 
@@
 \includegraphics[width=\textwidth]{custom_work/Filteredexamples.pdf}
-\caption[Filtered proton samples by cluster belonging]{A sample of proton events from different $K$-means clusters from the filtered data set.
+\caption[Filtered proton samples by cluster belonging]{A sample of proton events from different $k$-means clusters from the filtered data set.
 The bottom row shows proton samples that are intermingled with noise events in the models' predictions, and the top row belongs to a cluster that contains proton events almost exclusively.
-Each row belongs to a single cluster corresponding to the filtered confusion matrix in Fig.~\ref{fig:clster_confmati_filt}}\label{fig:filtered_vgg_clster_repr}
+Each row belongs to a single cluster corresponding to the filtered confusion matrix in Fig.~\ref{fig:clster_confmati_filt}.
+While the events in the pure proton cluster are visually distinct from the impure cluster, the difference is less pronounced than what we observe for the full data-set in figure \ref{fig:full_vgg_clster_repr}.
+}\label{fig:filtered_vgg_clster_repr}.
 \end{figure}
@@
 \includegraphics[width=\textwidth]{custom_work/Rawexamples.pdf}
-\caption[Full proton samples by cluster belonging]{Illustration of  a sample of proton events from different $K$-means clusters originating from the raw data set.
-Each row belongs to a single cluster corresponding to the full confusion matrix in Fig.~\ref{fig:clster_confmati_raw}}\label{fig:full_vgg_clster_repr}
+\caption[Full proton samples by cluster belonging]{A random sample of proton events from different $k$-means clusters from the raw data set.
+Each row belongs to a single cluster corresponding to the full confusion matrix in Fig.~\ref{fig:clster_confmati_raw}.
+The proton events in the cluster corresponding to the first row appear more visually distinct from other reaction types while the proton events corresponding to the cluster in the bottom row appear visually more similar to other reaction types in our dataset.
+}\label{fig:full_vgg_clster_repr}
 \end{figure}
@@
 
-With the best set of hyperparameters, each highest performing model is thereafter run $10$ times\footnote{The MIXAE model is significantly more computationally expensive to train than a $K$-means model.
+With the best set of hyperparameters, each highest performing model is thereafter run $10$ times\footnote{The MIXAE model is significantly more computationally expensive to train than a $k$-means model.
 Resulting in the skew in number of runs in the two cases.}.
-The results are listed in table \ref{tab:clstr_mixae}.
+The results are listed in Table \ref{tab:clstr_mixae}.
 We observe that, while the algorithm can achieve a very strong performance, the performance varies.
@@
 To quantify the results we report the best performing model (Top 1), and the mean and standard deviation for the result ($\mu\pm \sigma$).
-In contrast with the VGG-16 + $K$-means approach we observe  significant variations in performance.}\label{tab:clstr_mixae}
+In contrast with the VGG-16 + $k$-means approach we observe  significant variations in performance.}\label{tab:clstr_mixae}
 \input{./plots/mixae_clustering.tex}
@@
 
-Applied to raw data the MIXAE captures a proton cluster in a similar vein to the $K$-means approach.
+Applied to raw data the MIXAE captures a proton cluster in a similar vein to the $k$-means approach.
 The MIXAE forms a proton-majority cluster, but with a significant portion of the more noisy proton events being clustered with the other-class.
@@
 
-The most striking result we present in this work is the success of the $K$-means approach.
-As noted by \cite{Aggarwal}, distance measures become less informative in higher dimensional spaces, but the $K$-means algorithm clusters events well in a very high-dimensional space.
-Another surprise is the stability of the $K$-means algorithm.
+The most striking result we present in this work is the success of the $k$-means approach.
+As noted by \cite{Aggarwal}, distance measures become less informative in higher dimensional spaces, but the $k$-means algorithm clusters events well in our fairly high-dimensional VGG latent spaces.
+Another surprise is the stability of the $k$-means algorithm.
 We attribute this stability to the quality of the VGG16 latent space in creating class-separating sub-spaces.
@@
 
-It is also interesting to compare and contrast the clustering results from the MIXAE model with those of the VGG16 with the  $K$-means approach.
+It is also interesting to compare and contrast the clustering results from the MIXAE model with those of the VGG16 with the  $k$-means approach.
 In particular, the discrepancy in stability is worth noting.
-While the top performing MIXAE runs outperform the $K$-means approach, its reliability suffers.
+While the top performing MIXAE runs outperform the $k$-means approach, its reliability suffers.
 However, the high performance achieved indicates that this may represent a valuable potential research path into more tailored models for unsupervised track identification.
@@
 
-In addition to the results presented in this section, we performed clustering with a number of different algorithms included in the Scikit-learn package of \citet{Pedregosa2011}.
-None of them provided any notable differences from the $K$-means results or were significantly worse.
+In addition to the results presented in this section, we performed clustering with a number of different algorithms included in the \texttt{scikit-learn} package of \citet{Pedregosa2011}.
+None of them provided any notable differences from the $k$-means results or were significantly worse.
 Notably, the DBSCAN algorithm \cite{Ester96adensity-based,Bergstra2012} failed to provide any useful clustering results.
-We find this important as one of the significant drawbacks of $K$-means, and the deep clustering algorithm presented in section \ref{sec:mixae}, is that they depend  on pre-determining the number of clusters.
+We find this important as one of the significant drawbacks of $k$-means, and the deep clustering algorithm presented in section \ref{sec:mixae}, is that they depend  on pre-determining the number of clusters.
 This is not the case for DBSCAN.
@@
 
-However, this provides valuable insight as the seeds of the clusters are constructed by a $K$-means algorithm.
-This insight contrasts with our positive results from applying a pre-trained classification model with $K$-means and highlights potentially significant differences in models trained on a supervised or unsupervised objective for clustering tasks in nuclear physics.
+However, this provides valuable insight as the seeds of the clusters are constructed by a $k$-means algorithm.
+This insight contrasts with our positive results from applying a pre-trained classification model with $k$-means and highlights potentially significant differences in models trained on a supervised or unsupervised objective for clustering tasks in nuclear physics.
 
@@
 
-In particular, the ability of the $K$-means algorithm in picking out clear proton clusters from the VGG16 latent space lends itself well to an exploratory phase of analysis, where clusters of events corresponding to different reaction channels could be later identified by the experimenter.
-Another interesting facet of the $K$-means clustering is its consistent performance.
-As shown in table \ref{tab:clstr_vgg10}, the variance is zero for the performance metrics.
+In particular, the ability of the $k$-means algorithm in picking out clear proton clusters from the VGG16 latent space lends itself well to an exploratory phase of analysis, where clusters of events corresponding to different reaction channels could be later identified by the experimenter.
+Another interesting facet of the $k$-means clustering is its consistent performance.
+As shown in Table \ref{tab:clstr_vgg10}, the variance is zero for the performance metrics.
 This result indicates that the clusters are very clearly defined in the VGG16 latent space.
 However, as can be seen from the non-proton clusters in Figs.~\ref{fig:clster_confmati_filt} and \ref{fig:clster_confmati_raw},  this does not necessarily imply that the physical signals are correspondingly clear.
-Furthermore, the unsupervised metric that decides which of the $M$ $K$-means initializations perform the best does not necessarily coincide perfectly with separating the event classes.
-This is evidenced by the highest performing model measured on labelled data (Top 1) for the raw data showing up in table \ref{tab:clstr_vgg1}, and being filtered out from table \ref{tab:clstr_vgg10}.
+Furthermore, the unsupervised metric that decides which of the $M$ $k$-means initializations perform the best does not necessarily coincide perfectly with separating the event classes.
+This is evidenced by the highest performing model measured on labelled data (Top 1) for the raw data showing up in Table \ref{tab:clstr_vgg1}, and being filtered out from Table \ref{tab:clstr_vgg10}.
 
-A caveat to the $K$-means method is that the number of clusters has to be specified in advance.
+A caveat to the $k$-means method is that the number of clusters has to be specified in advance.
 Each experiment then has to be considered in light of possible reaction channels to determine a sensible number of clusters for this approach.
@@
 The same caveat is present in the MIXAE implementation.
-While it shows better optimal performance than the $K$-means method, some inconsistencies were observed that were notably not evident from the unsupervised training-objectives of the model.
+While it shows better optimal performance than the $k$-means method, some inconsistencies were observed that were notably not evident from the unsupervised training-objectives of the model.
 These two factors currently conspire to limit its immediate applicability, and more developments are needed for this approach.
 
-In summary, our study shows that unsupervised track classification with an implementation of the VGG16 and the  $K$-means approach is a viable solution.
+In summary, our study shows that unsupervised track classification with an implementation of the VGG16 and the  $k$-means approach is a viable solution.
 For future work, it is worth investigating whether the adaptation of  models like the MIXAE algorithm will allow better performance at no significant cost to consistency.
 %Especially considering the flexibility to use 3D input representations or auxiliary physical goals in the optimization procedure. 
-The two examples of unsupervised machine learning methods studied in this work are a first encouraging step towards automated classification of events that could not only greatly reduce the resource cost of analysis, but also eventually boost the efficiency of the experiment by allowing post-trigger decisions based on such algorithm implemented in hardware.
+The two examples of unsupervised machine learning methods studied in this work are a first encouraging step towards automated selection of similar events that could  greatly reduce the resource cost of analysis.
+Selection and classification algorithms have the potential to eventually boost the efficiency of the experiment by allowing post-trigger decisions based on such algorithm implemented in hardware.
 %Much progress remains to be done before achieving these goals.
@@
 {\color{orange}[flip this to positive ]} It is also clear that there is significant potential in both avenues explored in this work.
-%n particular, the surprising ability of the $K$-means algorithm in picking out clear proton clusters from the VGG16 latent space lends itself well to an exploratory phase of analysis.
-%Another surprising facet of the $K$-means clustering is its consistent performance.
-As shown in table \ref{tab:clstr_vgg}, the variance is zero for the performance metrics.
+%n particular, the surprising ability of the $k$-means algorithm in picking out clear proton clusters from the VGG16 latent space lends itself well to an exploratory phase of analysis.
+%Another surprising facet of the $k$-means clustering is its consistent performance.
+As shown in Table \ref{tab:clstr_vgg}, the variance is zero for the performance metrics.
 This result indicates that the clusters are very clearly defined in the VGG16 latent space.
@@
 
-%Additionally, a clear disadvantage to the $K$-means method is that the researcher has to specify the number of clusters in the data themselves.
+%Additionally, a clear disadvantage to the $k$-means method is that the researcher has to specify the number of clusters in the data themselves.
 Each experiment then has to be considered in light of possible reaction channels to determine a sensible number of clusters for this approach.
@@
 %The same weakness is present in the MIXAE implementation.
-Additionally, while it showed better optimal performance than the $K$-means method, it struggles with inconsistent results.
+Additionally, while it showed better optimal performance than the $k$-means method, it struggles with inconsistent results.
 This inconsistency was also notably not evident from the unsupervised training-objectives of the model.
@@
 
-%In summary, we recommend that researchers performing unsupervised track classification begin with an implementation of the VGG16 + $K$-means approach.
+%In summary, we recommend that researchers performing unsupervised track classification begin with an implementation of the VGG16 + $k$-means approach.
 For future work, it is worth investigating whether the adaptations of bespoke models like the MIXAE algorithm will allow better performance at no significant cost to consistency.
@@
 DE-SC0020451.
+This material is based upon work supported by the National Science Foundation under Grant No.
+PHY-2012865  (MPK).
 This project has also received support from the INTPART project of the Research Council of Norway (Grant No.
@@
 We chose to base our neural network parameter choices on the VGG16 architecture.
-The parameters chosen for the autoencoders are listed in full in table \ref{tab:mixe_ae_hyperparams}.
+The parameters chosen for the autoencoders are listed in full in Table \ref{tab:mixe_ae_hyperparams}.
 
@@
 The algorithm is trained on a subset of $60000$ of these samples, and we track performance on the remaining $20000$ events.
-On real data the algorithm is trained on an unlabelled set of data, and evaluated on a labelled subset\footnote{see table \ref{tab:data sets} for details}.
+On real data the algorithm is trained on an unlabelled set of data, and evaluated on a labelled subset\footnote{see Table \ref{tab:data sets} for details}.
 Since there are then only three remaining hyperparameters we choose to perform a coarse grid-search for the optimal configuration.
@@
 
-The grids selected for the search are listed in table \ref{tab:mixae_loss_weights}. The search yielded an optimal configuration with
+The grids selected for the search are listed in Table \ref{tab:mixae_loss_weights}. The search yielded an optimal configuration with
 
@@
 
-Lastly we supply the configuration used for the individual convolutional autoencoder networks in table \ref{tab:mixe_ae_hyperparams}
+Lastly we supply the configuration used for the individual convolutional autoencoder networks in Table \ref{tab:mixe_ae_hyperparams}
 
@@
 
-The data used  for the analysis in this work were partitioned as shown in table \ref{tab:data sets}. 
+The data used  for the analysis in this work were partitioned as shown in Table \ref{tab:data sets}. 
 \begin{table}[hbtp]
@@
 
-To elucidate the results presented in tables \ref{tab:clstr_mixae} and \ref{tab:clstr_vgg1} we computed clustering confusion-matrices.
+To elucidate the results presented in Tables \ref{tab:clstr_mixae} and \ref{tab:clstr_vgg1} we computed clustering confusion-matrices.
 These matrices show a more detailed picture of intermingled classes in a clustering or classification task where ground truth labels are available.
@@
 	\includegraphics[width=\textwidth]{custom_work/Simulatedcache_conf_mat.pdf} 
-	\caption[Pre-trained network - confusion matrices]{Confusion matrix for the $K$-means clustering of simulated AT-TPC events.
+	\caption[Pre-trained network - confusion matrices]{Confusion matrix for the $k$-means clustering of simulated AT-TPC events.
 The true labels indicate samples belonging to the p (proton), or the carbon (C) class.
@@
 	\includegraphics[width=\textwidth]{custom_work/Filteredcache_conf_mat.pdf}
-	\caption[Pre-trained network - confusion matrices]{Confusion matrix for the $K$-means clustering of filtered AT-TPC events.
+	\caption[Pre-trained network - confusion matrices]{Confusion matrix for the $k$-means clustering of filtered AT-TPC events.
 The true labels indicate samples belonging to the p (proton), carbon (C), or other classes.
@@
 	\includegraphics[width=\textwidth]{custom_work/Rawcache_conf_mat.pdf}
-\caption[Pre-trained network - confusion matrices]{Confusion matrix for the $K$-means clustering of raw AT-TPC events.
+\caption[Pre-trained network - confusion matrices]{Confusion matrix for the $k$-means clustering of raw AT-TPC events.
 The true labels indicate samples belonging to the p (proton), carbon (C), or other classes.
@@
 We observe that the algorithm correctly captures a majority proton-event cluster in cluster 0.
-However, in contrast with the $K$-means approach this cluster is contaminated to some extent with both carbon events and other events.
+However, in contrast with the $k$-means approach this cluster is contaminated to some extent with both carbon events and other events.
 }\label{fig:mixae_confmat_raw}
